### 2015-06-01 analyses/ResTime.Rmd

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide'}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/QuantitativeBiodiversity/Assignments/Alpha") 
```

## INTRODUCTION

Here, we explore the constraining influence of ...  alpha ($\alpha$) diversity. 
We will quantify general components of $\alpha$-diversity: **richness** and **evenness**.
From there, we ...  **species-abundance distributions (SAD)**.

```{r, results = 'hide', message=FALSE}
#install.packages("vegan") 
#instead, use try("vegan")
require("vegan") 
```

## 2) LOADING DATA

```{r}
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
```

```{r}
par(mfrow=c(2, 2))
x <- log(sim.data$TracerParticle_ResTime)
N <- log(sim.data$TotalAbundance)
plot(x, N, xlab="Particle Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='SteelBlue')

x <- log(sim.data$MicrobeCell_ResTime)
plot(x, N, xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='darkorchid4')

x <- log(sim.data$TracerParticle_ResTime)
S <- log(sim.data$SpeciesRichness)
plot(x, S, xlab="Particle Residence Time", 
     ylab="Total Abundance", 
     main = "S vs. Tau", col='SteelBlue')

x <- log(sim.data$MicrobeCell_ResTime)
plot(x, S, xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "S vs. Tau", col='darkorchid4')
```

```{r}
rads <- c()
rad.list <- readLines("~/GitHub/hydrobide/results/simulated_data/RADs.csv")  
for (rad in rad.list) {
  rad <- unlist(strsplit(rad, split = ","))  
  rad <- as.numeric(rad)
  rads <- c(rads, rad)
  }
```

### 2015-06-01 analyses/ResTime.Rmd

@@
 getwd() 
-setwd("~/GitHub/QuantitativeBiodiversity/Assignments/Alpha") 
+setwd("~/GitHub/hydrobide") 
 ```
@@
 ```{r}
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
+sim.data <- read.csv("./results/simulated_data/SimData.csv")  
 ```

### 2015-06-01 models/Hydro/DataGenerator/DataGenerator.pyx
edited ResTime.md and fluid dynamics simulation model DataGenerator.pyx -KL

@@ -21,10 +21,10 @@ import metrics
 
 
 def get_mrrmax():
-    m = int(choice(range(1, 2))) # number of individuals immigrating per time step
-    r = int(choice(range(1, 2))) # number of resource particles flowing in per time step
+    m = int(choice(range(1, 10))) # number of individuals immigrating per time step
+    r = int(choice(range(1, 10))) # number of resource particles flowing in per time step
     nr = int(choice(range(1, 10))) # max number of resources types
-    rmax = int(choice(range(10, 20))) # max value of resource particle sizes
+    rmax = int(choice(range(10, 100))) # max value of resource particle sizes
     return [m, r, nr, rmax]
 
 #######################  MICROBE COMMUNITY PARAMETERS  #########################

### 2015-06-01 analyses/ResTime.Rmd

@@
 #install.packages("vegan") 
-#instead, use try("vegan")
 require("vegan") 
@@
 par(mfrow=c(2, 2))
-x <- log(sim.data$TracerParticle_ResTime)
+tau <- log(sim.data$TracerParticle_ResTime)
 N <- log(sim.data$TotalAbundance)
-plot(x, N, xlab="Particle Residence Time", 
+plot(tau, N, xlab="Particle Residence Time", 
      ylab="Total Abundance", 
@@
 
-x <- log(sim.data$MicrobeCell_ResTime)
-plot(x, N, xlab="Cell Residence Time", 
+tau <- log(sim.data$MicrobeCell_ResTime)
+plot(tau, N, xlab="Cell Residence Time", 
      ylab="Total Abundance", 
@@
 
-x <- log(sim.data$TracerParticle_ResTime)
+tau <- log(sim.data$TracerParticle_ResTime)
 S <- log(sim.data$SpeciesRichness)
-plot(x, S, xlab="Particle Residence Time", 
-     ylab="Total Abundance", 
+plot(tau, S, xlab="Particle Residence Time", 
+     ylab="Species Richness", 
      main = "S vs. Tau", col='SteelBlue')
 
-x <- log(sim.data$MicrobeCell_ResTime)
-plot(x, S, xlab="Cell Residence Time", 
+tau <- log(sim.data$MicrobeCell_ResTime)
+plot(tau, S, xlab="Cell Residence Time", 
+     ylab="Species Richness", 
+     main = "S vs. Tau", col='darkorchid4')
+```
+
+```{r}
+par(mfrow=c(2, 2))
+tau <- log(sim.data$MicrobeCell_ResTime)
+N <- sim.data$TotalAbundance
+plot(tau, log(N), xlab="Cell Residence Time", 
      ylab="Total Abundance", 
-     main = "S vs. Tau", col='darkorchid4')
+     col='SteelBlue')
+
+S <- sim.data$SpeciesRichness
+plot(tau, log(S), xlab="Cell Residence Time", 
+     ylab="Species Richness", 
+     col='darkorchid4')
+
+E <- sim.data$Evar
+plot(tau, E, xlab="Cell Residence Time", 
+     ylab="Species Evenness", 
+     col='SteelBlue')
+
+avg.ab <- N/S
+plot(tau, avg.ab, xlab="Cell Residence Time", 
+     ylab="Average Abundance", 
+     col='darkorchid4')
+```
+
+```{r}
+par(mfrow=c(2, 2))
+tau <- log(sim.data$MicrobeCell_ResTime)
+BP <- sim.data$BergerParker
+plot(tau, BP, xlab="Cell Residence Time", 
+     ylab="Berger-Parker", 
+     col='SteelBlue')
+
+SimpD <- sim.data$InvSimpDiversity
+plot(tau, SimpD, xlab="Cell Residence Time", 
+     ylab="Simpsons Diversity (1/D)", 
+     col='darkorchid4')
+
+Skew <- sim.data$Skew
+plot(tau, Skew, xlab="Cell Residence Time", 
+     ylab="Skewness", 
+     col='SteelBlue')
+
+Nmax <- sim.data$Nmax
+plot(tau, log(Nmax), xlab="Cell Residence Time", 
+     ylab="Nmax", 
+     col='darkorchid4')
 ```

### 2015-06-01 analyses/ResTime.Rmd

--- 2015-06-01 analyses/ResTime.Rmd
@@
 
-```{r, results = 'hide'}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 rm(list=ls()) 
@@
 
-```{r, results = 'hide', message=FALSE}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 #install.packages("vegan") 
-#instead, use try("vegan")
 require("vegan")
+
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
 ```
 
-## 2) LOADING DATA
+## Univariate relationships
 
-```{r}
-sim.data <- read.csv("./results/simulated_data/SimData.csv")  
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+par(mfrow=c(1, 2), width = 8, height = 8)
+
+height <- sim.data$Height
+width <- sim.data$Width
+flow.rate <- sim.data$FlowRate
+
+h.tau <- log((height * width)/flow.rate)
+m.tau <- log(sim.data$MicrobeCell_ResTime)
+t.tau <- log(sim.data$TracerParticle_ResTime)
+
+plot(h.tau, m.tau, xlab="Hydraulic Residence Time", 
+     ylab="Cell Residence Time", 
+     col='SteelBlue')
+
+plot(h.tau, t.tau, xlab="Hydraulic Residence Time", 
+     ylab="Particle Residence Time", 
+     col='darkorchid4')
 ```
 
-```{r}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
 par(mfrow=c(2, 2))
-x <- log(sim.data$TracerParticle_ResTime)
+
 N <- log(sim.data$TotalAbundance)
-plot(x, N, xlab="Particle Residence Time", 
+S <- log(sim.data$SpeciesRichness)
+
+plot(t.tau, N, xlab="Particle Residence Time", 
      ylab="Total Abundance", 
@@
 
-x <- log(sim.data$MicrobeCell_ResTime)
-plot(x, N, xlab="Cell Residence Time", 
+plot(m.tau, N, xlab="Cell Residence Time", 
      ylab="Total Abundance", 
@@
 
-x <- log(sim.data$TracerParticle_ResTime)
-S <- log(sim.data$SpeciesRichness)
-plot(x, S, xlab="Particle Residence Time", 
-     ylab="Total Abundance", 
+plot(t.tau, S, xlab="Particle Residence Time", 
+     ylab="Species Richness", 
      main = "S vs. Tau", col='SteelBlue')
 
-x <- log(sim.data$MicrobeCell_ResTime)
-plot(x, S, xlab="Cell Residence Time", 
-     ylab="Total Abundance", 
+plot(m.tau, S, xlab="Cell Residence Time", 
+     ylab="Species Richness", 
      main = "S vs. Tau", col='darkorchid4')
+
 ```
 
-```{r}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+N <- sim.data$TotalAbundance
+plot(m.tau, log(N), xlab="Cell Residence Time", 
+     ylab="Total Abundance", 
+     main = "N vs. Tau",
+     col='SteelBlue')
+
+S <- sim.data$SpeciesRichness
+plot(m.tau, log(S), xlab="Cell Residence Time", 
+     ylab="Species Richness", 
+     main = "S vs. Tau",
+     col='darkorchid4')
+
+E <- sim.data$Evar
+plot(m.tau, E, xlab="Cell Residence Time", 
+     ylab="Species Evenness", 
+     main = "Evenness vs. Tau",
+     col='SteelBlue')
+
+avg.ab <- N/S
+plot(m.tau, log(avg.ab), xlab="Cell Residence Time", 
+     ylab="Average Abundance",
+     main = "N/S vs. Tau",
+     col='darkorchid4')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+BP <- sim.data$BergerParker
+plot(m.tau, BP, xlab="Cell Residence Time", 
+     ylab="Berger-Parker", 
+     main = "Relative Dominance",
+     col='SteelBlue')
+
+SimpD <- sim.data$InvSimpDiversity
+plot(m.tau, SimpD, xlab="Cell Residence Time", 
+     ylab="Simpsons, 1/D", 
+     main = "Diversity",
+     col='darkorchid4')
+
+Skew <- sim.data$Skew
+plot(m.tau, Skew, xlab="Cell Residence Time", 
+     ylab="Skewness", 
+     main = "Rarity",
+     col='SteelBlue')
+
+Nmax <- sim.data$Nmax
+plot(m.tau, log(Nmax), xlab="Cell Residence Time", 
+     ylab="Nmax",
+     main = "Absolute Dominance",
+     col='darkorchid4')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+plot(log(N), log(S), xlab="Total abundance",
+     ylab="Species Richness", 
+     main = "S vs. N",
+     col='SteelBlue')
+
+plot(log(N), log(Nmax), xlab="Total abundance", 
+     ylab="Nmax", 
+     main = "Nmax vs. N",
+     col='darkorchid4')
+
+plot(log(N), log(E), xlab="Total abundance", 
+     ylab="Species Evenness", 
+     main = "Evenness vs. N",
+     col='SteelBlue')
+
+plot(log(N), log(Skew), xlab="Total abundance", 
+     ylab="Skewness",
+     main = "Rarity vs. N",
+     col='darkorchid4')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 rads <- c()

### 2015-06-01 analyses/ResTime.Rmd
edited ResTime.Rmd to import all RADs simulated by Hydrobide -KL

@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-rads <- c()
-rad.list <- readLines("~/GitHub/hydrobide/results/simulated_data/RADs.csv")  
-for (rad in rad.list) {
-  rad <- unlist(strsplit(rad, split = ","))  
-  rad <- as.numeric(rad)
-  rads <- c(rads, rad)
-  }
+rads <- list()
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/RADs.csv",
+            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
+            as.is = TRUE, skipNul = TRUE, fill = TRUE)
+
+rad.list[is.na(rad.list)] <- 0
 ```

### 2015-06-01 analyses/ResTime.Rmd

@@
 ```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+species <- list()
+species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/Species.csv",
+            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
+            as.is = TRUE, skipNul = TRUE, fill = TRUE)
+
+species.list[is.na(species.list)] <- -1
+```

### 2015-06-01 analyses/ResTime.Rmd
edited ResTime.Rmd to conduct many univariate examinations -KL

@@
      ylab="Total Abundance", 
-     main = "N vs. Tau", col='SteelBlue')
+     main = "N vs. Tau", col='cornsilk4')
 
@@
      ylab="Total Abundance", 
-     main = "N vs. Tau", col='darkorchid4')
+     main = "N vs. Tau", col='cornsilk4')
 
@@
      ylab="Species Richness", 
-     main = "S vs. Tau", col='SteelBlue')
+     main = "S vs. Tau", col='darkseagreen')
 
@@
      ylab="Species Richness", 
-     main = "S vs. Tau", col='darkorchid4')
+     main = "S vs. Tau", col='darkseagreen')
 
@@
      main = "N vs. Tau",
-     col='SteelBlue')
+     col='cornsilk4')
 
@@
      main = "S vs. Tau",
-     col='darkorchid4')
+     col='darkseagreen')
 
@@
      main = "Evenness vs. Tau",
-     col='SteelBlue')
+     col='skyblue')
 
@@
      main = "N/S vs. Tau",
-     col='darkorchid4')
+     col='coral')
 ```
@@
      ylab="Berger-Parker", 
-     main = "Relative Dominance",
-     col='SteelBlue')
+     main = "Relative Dominance vs. Tau",
+     col='firebrick')
 
@@
      ylab="Simpsons, 1/D", 
-     main = "Diversity",
-     col='darkorchid4')
+     main = "Diversity vs. Tau",
+     col='gold')
 
@@
      ylab="Skewness", 
-     main = "Rarity",
-     col='SteelBlue')
+     main = "Rarity vs. Tau",
+     col='limegreen')
 
@@
      ylab="Nmax",
-     main = "Absolute Dominance",
-     col='darkorchid4')
+     main = "Absolute Dominance vs. Tau",
+     col='sandybrown')
 ```
@@
      main = "S vs. N",
-     col='SteelBlue')
+     col='indianred')
 
@@
      main = "Nmax vs. N",
-     col='darkorchid4')
+     col='darkorange')
 
-plot(log(N), log(E), xlab="Total abundance", 
+plot(log(N), E, xlab="Total abundance", 
      ylab="Species Evenness", 
      main = "Evenness vs. N",
-     col='SteelBlue')
+     col='lightseagreen')
 
-plot(log(N), log(Skew), xlab="Total abundance", 
+plot(log(N), Skew, xlab="Total abundance", 
      ylab="Skewness",
      main = "Rarity vs. N",
-     col='darkorchid4')
+     col='plum')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+growth <- sim.data$AvgPerCapita_GrowthRate
+plot(m.tau, log(growth), xlab="Cell Residence Time", 
+     ylab="growth", 
+     main = "Per capita growth vs. Tau",
+     col='dodgerblue')
+
+maint <- sim.data$AvgPerCapita_Maintenance
+plot(m.tau, SimpD, xlab="Cell Residence Time", 
+     ylab="maintenance", 
+     main = "Per capita maintenance vs. Tau",
+     col='darkolivegreen')
+
+immigration <- sim.data$Immigration_Rate
+plot(m.tau, immigration, xlab="Cell Residence Time", 
+     ylab="immigration", 
+     main = "Immigration vs. Tau",
+     col='mediumorchid')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+res.conc <- sim.data$Resource_Concentration
+plot(m.tau, log(res.conc), xlab="Cell Residence Time", 
+     ylab="density", 
+     main = "Resource density vs. Tau",
+     col='peru')
+
+res.div <- sim.data$Shannons_ResourceDiversity
+plot(m.tau, res.div, xlab="Cell Residence Time", 
+     ylab="Shannon's H", 
+     main = "Resource diversity vs. Tau",
+     col='palegreen')
+
+res.rich <- sim.data$ResourceRichness
+plot(m.tau, res.rich, xlab="Cell Residence Time", 
+     ylab="Number of resource types", 
+     main = "Resource richness vs. Tau",
+     col='slateblue')
 ```

### 2015-06-08 analyses/ResTime_GeneralResults.Rmd
modified: lattice boltzman code, bide files, movie files, datagenerator file, moviegenerator file; generated new .csv data files; modified RMarkdown document; generated new analyses of .csv data files -KL

--- 2015-06-01 analyses/ResTime.Rmd
@@
 
-Here, we explore the constraining influence of ...  alpha ($\alpha$) diversity. 
-We will quantify general components of $\alpha$-diversity: **richness** and **evenness**.
-From there, we ...  **species-abundance distributions (SAD)**.
+Here, we explore the influence of mean cell residence time, hydraulic residence time, and particle residence time on abundance, local ($\alpha$) diversity, physiology, and resource availability and diversity.
+
+These simulations
 
@@
 require("vegan") 
+```
 
+## Load Data
+
+```{r, results = 'hide', message = FALSE, warning = FALSE}
 sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-par(mfrow=c(1, 2), width = 8, height = 8)
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
+par(mfrow=c(2, 2))
 
+simulation <- sim.data$sim
 height <- sim.data$Height
@@
 h.tau <- log((height * width)/flow.rate)
-m.tau <- log(sim.data$MicrobeCell_ResTime)
-t.tau <- log(sim.data$TracerParticle_ResTime)
+m.tau <- log(sim.data$cell.tau)
+t.tau <- log(sim.data$particle.tau)
 
@@
      col='darkorchid4')
-```
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-par(mfrow=c(2, 2))
-
-N <- log(sim.data$TotalAbundance)
-S <- log(sim.data$SpeciesRichness)
-
-plot(t.tau, N, xlab="Particle Residence Time", 
-     ylab="Total Abundance", 
-     main = "N vs. Tau", col='SteelBlue')
-
-plot(m.tau, N, xlab="Cell Residence Time", 
-     ylab="Total Abundance", 
-     main = "N vs. Tau", col='darkorchid4')
-
-plot(t.tau, S, xlab="Particle Residence Time", 
-     ylab="Species Richness", 
-     main = "S vs. Tau", col='SteelBlue')
-
-plot(m.tau, S, xlab="Cell Residence Time", 
-     ylab="Species Richness", 
-     main = "S vs. Tau", col='darkorchid4')
+plot(t.tau, m.tau, xlab="Particle Residence Time", 
+     ylab="Cell Residence Time", 
+     col='darkorchid4')
 
@@
 
-N <- sim.data$TotalAbundance
+N <- log(sim.data$N)
+S <- log(sim.data$S)
+
+plot(t.tau, N, xlab="Particle Residence Time", 
+     ylab="Total Abundance", 
+     main = "N vs. Tau", col='cornsilk4')
+
+plot(m.tau, N, xlab="Cell Residence Time", 
+     ylab="Total Abundance", 
+     main = "N vs. Tau", col='cornsilk4')
+
+plot(t.tau, S, xlab="Particle Residence Time", 
+     ylab="Species Richness", 
+     main = "S vs. Tau", col='darkseagreen')
+
+plot(m.tau, S, xlab="Cell Residence Time", 
+     ylab="Species Richness", 
+     main = "S vs. Tau", col='darkseagreen')
+
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+N <- sim.data$N
 plot(m.tau, log(N), xlab="Cell Residence Time", 
@@
      main = "N vs. Tau",
-     col='SteelBlue')
+     col='cornsilk4')
 
-S <- sim.data$SpeciesRichness
+S <- sim.data$S
 plot(m.tau, log(S), xlab="Cell Residence Time", 
@@
      main = "S vs. Tau",
-     col='darkorchid4')
+     col='darkseagreen')
 
-E <- sim.data$Evar
-plot(m.tau, E, xlab="Cell Residence Time", 
+e.var <- sim.data$e.var
+e.simp <- sim.data$simpson.e
+plot(m.tau, e.var, xlab="Cell Residence Time", 
      ylab="Species Evenness", 
      main = "Evenness vs. Tau",
-     col='SteelBlue')
+     col='skyblue')
 
@@
 
-BP <- sim.data$BergerParker
+BP <- sim.data$berger.parker
 plot(m.tau, BP, xlab="Cell Residence Time", 
      ylab="Berger-Parker", 
-     main = "Relative Dominance",
-     col='SteelBlue')
+     main = "Relative Dominance vs. Tau",
+     col='firebrick')
 
-SimpD <- sim.data$InvSimpDiversity
+SimpD <- sim.data$inv.simp.D
 plot(m.tau, SimpD, xlab="Cell Residence Time", 
      ylab="Simpsons, 1/D", 
-     main = "Diversity",
-     col='darkorchid4')
+     main = "Diversity vs. Tau",
+     col='gold')
 
-Skew <- sim.data$Skew
+Skew <- sim.data$skew
 plot(m.tau, Skew, xlab="Cell Residence Time", 
      ylab="Skewness", 
-     main = "Rarity",
-     col='SteelBlue')
+     main = "Rarity vs. Tau",
+     col='limegreen')
 
-Nmax <- sim.data$Nmax
+Nmax <- sim.data$N.max
 plot(m.tau, log(Nmax), xlab="Cell Residence Time", 
      ylab="Nmax",
-     main = "Absolute Dominance",
-     col='darkorchid4')
+     main = "Absolute Dominance vs. Tau",
+     col='sandybrown')
 ```
@@
      main = "Nmax vs. N",
-     col='darkorchid4')
+     col='darkorange')
 
-plot(log(N), log(E), xlab="Total abundance", 
+plot(log(N), e.var, xlab="Total abundance", 
      ylab="Species Evenness", 
      main = "Evenness vs. N",
-     col='SteelBlue')
+     col='lightseagreen')
 
-plot(log(N), log(Skew), xlab="Total abundance", 
+plot(log(N), Skew, xlab="Total abundance", 
      ylab="Skewness",
      main = "Rarity vs. N",
-     col='darkorchid4')
+     col='plum')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+growth <- sim.data$avg.per.capita.growth
+plot(m.tau, log(growth), xlab="Cell Residence Time", 
+     ylab="growth", 
+     main = "Per capita growth vs. Tau",
+     col='dodgerblue')
+
+maint <- sim.data$avg.per.capita.maint
+plot(m.tau, SimpD, xlab="Cell Residence Time", 
+     ylab="maintenance", 
+     main = "Per capita maintenance vs. Tau",
+     col='darkolivegreen')
+
+immigration <- sim.data$immigration.rate
+plot(m.tau, immigration, xlab="Cell Residence Time", 
+     ylab="immigration", 
+     main = "Immigration vs. Tau",
+     col='mediumorchid')
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+res.conc <- sim.data$resource.concentration
+plot(m.tau, log(res.conc), xlab="Cell Residence Time", 
+     ylab="density", 
+     main = "Resource density vs. Tau",
+     col='peru')
+
+res.div <- sim.data$shannons.resource.diversity
+plot(m.tau, res.div, xlab="Cell Residence Time", 
+     ylab="Shannon's H", 
+     main = "Resource diversity vs. Tau",
+     col='palegreen')
+
+res.rich <- sim.data$resource.richness
+plot(m.tau, res.rich, xlab="Cell Residence Time", 
+     ylab="Number of resource types", 
+     main = "Resource richness vs. Tau",
+     col='slateblue')
 ```

### 2015-06-08 models/Hydro/DataGenerator/DataGenerator.pyx

@@ -1,118 +1,92 @@
 from __future__ import division
 from random import randint, choice
 import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.animation as animation
 import sys
 import os
-import scipy as sc
-from scipy.optimize import curve_fit
 from scipy import stats
-
+import time
+import psutil
 
 mydir = os.path.expanduser("~/")
 sys.path.append(mydir + "/GitHub/hydrobide/tools/LBM")
 import LBM
 sys.path.append(mydir + "/GitHub/hydrobide/tools/bide")
 import bide
-
 sys.path.append(mydir + "tools/metrics")
 import metrics
 
 
 def get_mrrmax():
-    m = int(choice(range(1, 10))) # number of individuals immigrating per time step
-    r = int(choice(range(1, 10))) # number of resource particles flowing in per time step
-    nr = int(choice(range(1, 10))) # max number of resources types
-    rmax = int(choice(range(10, 100))) # max value of resource particle sizes
+    """ Get random model parameter values. Others are chosen in bide.pyx """
+
+    m = 1 #int(choice(range(1, 2))) # individuals immigrating per time step
+    r = int(choice(range(10, 100))) # resource particles flowing in per time step
+    nr = int(choice(range(1, 10))) # maximum number of resources types
+    rmax = int(choice(range(10, 1000))) # maximum value of resource particle size
+
     return [m, r, nr, rmax]
 
+
 #######################  MICROBE COMMUNITY PARAMETERS  #########################
-COM, Xs, MicXcoords, MicYcoords, AvgTaus = [[], [], [], [], []]
-MicIDs, MicQs, MicExitAge, MicTimeIn = [[], [], [], []]
-ResYcoords, RES, ResIDs, ResType, ResXcoords = [[], [], [], [], []]
-
-microbe_color_dict, GrowthDict, MaintDict = [{}, {}, {}]
-ResUseDict, ResColorDict, DispParamsDict = [{}, {}, {}]
-
 m, r, nr, rmax = get_mrrmax()
-MicID = 0
+
+MicID, ResID, N, S = 0, 0, 0, 0
 avgTau = str()
-ResID = 0
+
[…]

### 2015-06-08 models/Hydro/MovieGenerator/Simple.pyx

@@ -1,331 +1,64 @@
 from __future__ import division
-from random import randint
-import math
-import numpy as np
-import matplotlib.pyplot as plt
 import matplotlib.animation as animation
+import matplotlib.pyplot as plt
+from random import randint, choice
+from scipy import stats
+import numpy as np
 import sys
 import os
-import scipy as sc
-from scipy.optimize import curve_fit
-from scipy import stats
+import time
+import psutil
 
+mydir = os.path.expanduser("~/")
+sys.path.append(mydir + "tools/metrics")
+import metrics
+sys.path.append(mydir + "/GitHub/hydrobide/tools/LBM")
+import lbmMovie as LBM
+sys.path.append(mydir + "/GitHub/hydrobide/tools/bide/bideMovie")
+import bideMovie as bide
 
-mydir = os.path.expanduser("~/GitHub")
-sys.path.append(mydir + "/hydrobide/tools/LBM")
-import LBM
-sys.path.append(mydir + "/hydrobide/tools/bide")
-import bide
 
 
-def Lattice():
+def get_mrrmax():
+    """ Get random model parameter values. Others are chosen in bide.pyx """
 
-    global n0, nN, nS, nE, nW, nNE, nNW, nSE, nSW, barrier, one9th, four9ths, one36th, rho, u0, width
-    global ux, uy, barrierN, barrierS, barrierE, barrierW, barrierNE, barrierNW, barrierSE, barrierSW, height
-    global BarrierXcoords1, BarrierYcoords1, BarrierXcoords2, BarrierYcoords2, BarrierWidth, BarrierHeight
+    m = int(choice(range(1, 10))) # individuals immigrating per time step
+    r = int(choice(range(10, 100))) # resource particles flowing in per time step
+    nr = int(choice(range(1, 10))) # maximum number of resources types
+    rmax = int(choice(range(100, 1000))) # maximum value of resource particle size
 
-    args1 = [n0, nN, nS, nE, nW, nNE, nNW, nSE, nSW, barrier, one9th, four9ths, one36th, rho, u0, width]
-    args2 = [ux, uy, barrierN, barrierS, barrierE, barrierW, barrierNE, barrierNW, barrierSE, barrierSW, height]
[…]

### 2015-06-08 models/Hydro/MovieGenerator/Simple.pyx
edited .Rmd coument, fixed column labels in .csv data, generated new .pdf of analyses -KL

@@ -264,8 +264,8 @@ OUT1 = open(mydir + '/GitHub/hydrobide/results/simulated_data/SimData.csv','w+')
 OUT2 = open(mydir + '/GitHub/hydrobide/results/simulated_data/RADs.csv','w+')
 OUT3 = open(mydir + '/GitHub/hydrobide/results/simulated_data/Species.csv','w+')
 # printing physical variables, residence times, community diversity properties, physiological values, trait values, resource values
-print>>OUT1, 'RowID, sim, FlowRate, Width, Height, Viscosity, N, S, immigration.rate, particle.tau, cell.tau, resource.concentration, shannons.resource.diversity, resource.richness, simpson.e, e.var, berger.parker, inv.simp.D, N.max, skew, avg.per.capita.growth, avg.per.capita.maint'
-#             ct1,   sim,   u0,     width, height, viscosity, N, S, m,                  TracerTau,  MicrobeTau, ResDens,                ResDiv,                    ResRich,            ES,        Ev,    BP,            SD,         Nm,    sk,         Mu,               Maint
+print>>OUT1, 'RowID, sim, FlowRate, Width, Height, Viscosity, N, immigration.rate, particle.tau, cell.tau, resource.concentration, shannons.resource.diversity, resource.richness, S, simpson.e, e.var, berger.parker, inv.simp.D, N.max, skew, avg.per.capita.growth, avg.per.capita.maint'
+#              ct1,  sim, u0,       width, height, viscosity, N, m,                TracerTau,    MicrobeTau, ResDens,              ResDiv,                      ResRich,           S, ES,        Ev,    BP,            SD,         Nm,    sk,   Mu,                    Maint
 OUT1.close()
 OUT2.close()
 OUT3.close()

### 2015-06-08 analyses/ResTime_GeneralResults.Rmd

@@
 ```{r, results = 'hide', message = FALSE, warning = FALSE}
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June7/SimData.csv")  
 ```
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
-par(mfrow=c(2, 2))
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 
@@
 t.tau <- log(sim.data$particle.tau)
+
+N <- sim.data$N
+S <- sim.data$S
+e.var <- sim.data$e.var
+e.simp <- sim.data$simpson.e
+avg.ab <- N/S
+Skew <-sim.data$skew
+Nmax <- sim.data$N.max
+SimpD <- sim.data$SD
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=5, fig.height=5}
+par(mfrow=c(2, 2))
 
@@
 par(mfrow=c(2, 2))
-
-N <- log(sim.data$N)
-S <- log(sim.data$S)
 
@@
 
-N <- sim.data$N
-plot(m.tau, log(N), xlab="Cell Residence Time", 
-     ylab="Total Abundance", 
-     main = "N vs. Tau",
-     col='cornsilk4')
-
-S <- sim.data$S
-plot(m.tau, log(S), xlab="Cell Residence Time", 
-     ylab="Species Richness", 
-     main = "S vs. Tau",
-     col='darkseagreen')
-
-e.var <- sim.data$e.var
-e.simp <- sim.data$simpson.e
-plot(m.tau, e.var, xlab="Cell Residence Time", 
-     ylab="Species Evenness", 
-     main = "Evenness vs. Tau",
-     col='skyblue')
-
-avg.ab <- N/S
-plot(m.tau, log(avg.ab), xlab="Cell Residence Time", 
-     ylab="Average Abundance",
-     main = "N/S vs. Tau",
-     col='coral')
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-par(mfrow=c(2, 2))
-
-BP <- sim.data$berger.parker
-plot(m.tau, BP, xlab="Cell Residence Time", 
-     ylab="Berger-Parker", 
-     main = "Relative Dominance vs. Tau",
-     col='firebrick')
-
-SimpD <- sim.data$inv.simp.D
-plot(m.tau, SimpD, xlab="Cell Residence Time", 
-     ylab="Simpsons, 1/D", 
-     main = "Diversity vs. Tau",
-     col='gold')
-
-Skew <- sim.data$skew
-plot(m.tau, Skew, xlab="Cell Residence Time", 
+plot(log(N), Skew, xlab="Total abundance", 
      ylab="Skewness",
-     main = "Rarity vs. Tau",
-     col='limegreen')
-
-Nmax <- sim.data$N.max
-plot(m.tau, log(Nmax), xlab="Cell Residence Time", 
-     ylab="Nmax",
-     main = "Absolute Dominance vs. Tau",
-     col='sandybrown')
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-par(mfrow=c(2, 2))
-
-plot(log(N), log(S), xlab="Total abundance",
-     ylab="Species Richness", 
-     main = "S vs. N",
-     col='indianred')
+     main = "Rarity vs. N",
+     col='plum')
 
@@
 
-plot(log(N), Skew, xlab="Total abundance", 
-     ylab="Skewness",
-     main = "Rarity vs. N",
-     col='plum')
+plot(log(N), log(S), xlab="Total abundance",
+     ylab="Species Richness", 
+     main = "S vs. N",
+     col='indianred')
+
 ```
@@
 growth <- sim.data$avg.per.capita.growth
-plot(m.tau, log(growth), xlab="Cell Residence Time", 
+growth <- (S*growth)/N
+x <- c()
+y <- c()
+
+ct <- 1
+for (val in growth){
+  if(is.nan(val) == FALSE && val > 0.1){
+    y <- c(y, val)
+    x <- c(x, m.tau[ct])
+    }
+  ct <- ct+1
+  }
+
+plot(x, y, xlab="Cell Residence Time", 
      ylab="growth", 
@@
 
-maint <- sim.data$avg.per.capita.maint
+maint <- (S*sim.data$avg.per.capita.maint)/N
 plot(m.tau, SimpD, xlab="Cell Residence Time", 
@@
 rads <- list()
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/RADs.csv",
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June7/RADs.csv",
             header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
@@
 species <- list()
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/Species.csv",
+species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June7/Species.csv",
             header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),

### 2015-06-10 models/Hydro/MovieGenerator/Simple.pyx
edited fluid dynamics functions in lbmMovie.pyx, simulation model Simple.pyx, generated new simulated data files, created Markdown document for manuscript -KL

@@ -6,28 +6,28 @@ from scipy import stats
 import numpy as np
 import sys
 import os
-import time
 import psutil
 
 mydir = os.path.expanduser("~/")
 sys.path.append(mydir + "tools/metrics")
 import metrics
-sys.path.append(mydir + "/GitHub/hydrobide/tools/LBM")
+sys.path.append(mydir + "/GitHub/hydrobide/tools/LBM/lbmMovie")
 import lbmMovie as LBM
 sys.path.append(mydir + "/GitHub/hydrobide/tools/bide/bideMovie")
-import bideMovie as bide
+import bide
 
 
 
 def get_mrrmax():
     """ Get random model parameter values. Others are chosen in bide.pyx """
 
-    m = int(choice(range(1, 10))) # individuals immigrating per time step
-    r = int(choice(range(10, 100))) # resource particles flowing in per time step
-    nr = int(choice(range(1, 10))) # maximum number of resources types
-    rmax = int(choice(range(100, 1000))) # maximum value of resource particle size
+    seed = choice([1000])
+    m = choice([0]) # individuals immigrating per time step
+    r = choice([200]) # resource particles flowing in per time step
+    nr = choice([10]) # maximum number of resources types
+    rmax = choice([5000]) # maximum value of resource particle size
 
-    return [m, r, nr, rmax]
+    return [seed, m, r, nr, rmax]
 
 
 ######### Function called for each successive animation frame ##################
@@ -38,10 +38,12 @@ def nextFrame(arg):	# arg is the frame number
     global MicXcoords, MicYcoords, microbe_scatImage, microbe_color_dict, GrowthDict, MaintDict, AvgTaus, MicIDs, MicQs, MicID, MicIDs
 
     global MicTimeIn, MicExitAge, avgTau, TracerIDs, TracerExitAge, TracerXcoords, TracerYcoords, tracer_scatImage, resource_scatImage
-    global ResXcoords, ResYcoords, ResID, ResIDs, RES, LogSeriesAlpha, omega, MUs, VarMUs, Maints, VarMaints, ResType, ResUseDict, DispParamsDict
+    global ResXcoords, ResYcoords, ResID, ResIDs, RES, alpha, omega, MUs, VarMUs, Maints, VarMaints, ResType, ResUseDict, DispParamsDict
 
[…]

### 2015-06-10 tools/LBM/lbmMovie/lbmMovie.pyx

@@ -26,14 +26,14 @@ def SetLattice(u0, viscosity, width, height, BarrierWidth, BarrierHeight, Barrie
     barrier = np.zeros((height, width), bool)  # Initialize barriers
                                                # True wherever there's a barrier
 
-    barrier[0: BarrierHeight*height, 0.2*width: (0.2+BarrierWidth)*width] = True
-    barrier[height*(1-BarrierHeight): height, 0.7*width: (0.7+BarrierWidth)*width] = True
+    barrier[0.2*height: BarrierHeight*height+(0.2*height), 0.3*width: (0.3+BarrierWidth)*width] = True
+    barrier[0.8*height*(1-BarrierHeight): 0.8*height, 0.6*width: (0.6+BarrierWidth)*width] = True
 
-    BarrierXcoords1 = [0.2*width, (0.2+BarrierWidth)*width]
-    BarrierYcoords1 = [0, BarrierHeight*height]
+    BarrierXcoords1 = [0.3*width, (0.3+BarrierWidth)*width]
+    BarrierYcoords1 = [0.2*height, BarrierHeight*height+(0.2*height)]
 
-    BarrierXcoords2 = [0.7*width, (0.7+BarrierWidth)*width]
-    BarrierYcoords2 = [(1-BarrierHeight)*height, height]
+    BarrierXcoords2 = [0.6*width, (0.6+BarrierWidth)*width]
+    BarrierYcoords2 = [0.8*height*(1-BarrierHeight), 0.8*height]
 
 
     barrierN = np.roll(barrier,  1, axis=0)       # sites just north of barriers

### 2015-06-10 tools/LBM/lbmMovie/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("lbmMovie.pyx"))

### 2015-06-10 tools/bide/bideMovie/bide.pyx

@@ -2,20 +2,33 @@ from __future__ import division
 from random import randint, choice
 import numpy as np
 import sys
+from scipy import stats
+import os
+
+mydir = os.path.expanduser("~/")
+sys.path.append(mydir + "tools/metrics")
+import metrics
+
 
 limit = 0.5
 
-def xcoord(mini=0.2, maxi=1):
-    return float(np.random.uniform(mini, maxi))
+def xcoord(low=0.2, high=8):
+    return float(np.random.uniform(low, high))
+
+def ycoord(low=1.5, high=8.5):
+    return float(np.random.uniform(low, high))
+
 
 
 def NewTracers(TracerIDs, TracerXcoords, TracerYcoords, width, height, u0):
 
     x = np.random.binomial(1, u0)
+
     if x == 1:
-        TracerYcoords.append(float(np.random.uniform(0.2*height, 0.8*height)))
+        y = ycoord(4,5)
+        TracerYcoords.append(y)
         x = xcoord()
-        TracerXcoords.append(x) # very near the upstream edge
+        TracerXcoords.append(x)
         TracerIDs.append(0) # used to track the age of the tracer
 
     return [TracerIDs, TracerXcoords, TracerYcoords]
@@ -24,7 +37,9 @@ def NewTracers(TracerIDs, TracerXcoords, TracerYcoords, width, height, u0):
 
 def ResIn(RES, ResXcoords, ResYcoords, ResID, ResIDs, ResType, r, rmax, nr, width, height, u0):
 
+    R = len(RES)
     x = np.random.binomial(1, u0)
+
     if x == 1:
         res_in = np.random.random_integers(1, rmax, r)
         res_types = np.random.random_integers(0, nr-1, r)
@@ -35,9 +50,14 @@ def ResIn(RES, ResXcoords, ResYcoords, ResID, ResIDs, ResType, r, rmax, nr, widt
             ResIDs.append(ResID)
             ResType.append(res_types[i])
             ResID += 1
-            ResYcoords.append(float(np.random.uniform(0.1*height, 0.9*height)))
-            x = xcoord()
-            ResXcoords.append(x) # width-limit
+
+            y = ycoord()
+            ResYcoords.append(y)
+            if R == 0:
+                x = xcoord()
+            else:
+                x = xcoord()
+            ResXcoords.append(x)
 
     return [RES, ResXcoords, ResYcoords, ResID, ResIDs, ResType]
 
[…]

### 2015-06-10 tools/bide/bideMovie/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("bide.pyx"))

### 2015-06-10 analyses/ResTime_GeneralResults.Rmd

@@
 
-simulation <- sim.data$sim
+sim <- sim.data$sim # the specific simulation run
 height <- sim.data$Height
@@
 
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+sim.sub <- subset(sim.data, sim == 60)
+height <- sim.sub$Height
+width <- sim.sub$Width
+flow.rate <- sim.sub$FlowRate
+
+h.tau <- log((height * width)/flow.rate)
+m.tau <- log(sim.sub$cell.tau)
+t.tau <- log(sim.sub$particle.tau)
+
+N <- sim.sub$N
+S <- sim.sub$S
+e.var <- sim.sub$e.var
+e.simp <- sim.sub$simpson.e
+avg.ab <- N/S
+Skew <-sim.sub$skew
+Nmax <- sim.sub$N.max
+SimpD <- sim.sub$SD
+
+plot(t.tau, N, xlab="Particle Residence Time", 
+     ylab="Total Abundance", 
+     main = "N vs. Tau", col='cornsilk4')
+lines(t.tau, N)
+
+plot(t.tau, S, xlab="Particle Residence Time", 
+     ylab="Species Richness", 
+     main = "S vs. Tau", col='darkseagreen')
+lines(t.tau, S)
+
+```
+
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}

### 2015-06-17 analyses/2015_06_16_ResTime_GeneralResults.Rmd
modified Rmd file, generated new plots and .pdf, modified bide and model functions -KL

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

## INTRODUCTION

Here, we explore the influence of mean cell residence time, hydraulic residence time, and particle residence time on abundance, local ($\alpha$) diversity, physiology, and resource availability and diversity.

These simulations

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
```

```{r, results = 'hide', message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June16/SimData.csv")  
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Assign Data
sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

## Univariate relationships

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

num.sims <- max(sim.data$sim)
colors <- rainbow(num.sims)

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1
  print(i)

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)

  N <- sim.sub$N
  N <- log(N)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in N){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="Total Abundance", ylim=c(3.5, 10), xlim=c(3, 9),
    main = "N vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }

  }

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in S){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="log(species richness)", ylim=c(0, 4), xlim=c(3, 9),
    main = "S vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }
}
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Reassign Data

sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

growth <- sim.data$avg.per.capita.growth
x <- c()
y <- c()

ct <- 1
for (val in growth){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab="log(ecosystem residence time)", 
     ylab="growth rate", 
     main = "Per capita growth vs. Tau",
     col='dodgerblue')

maint <- (S*sim.data$avg.per.capita.maint)/N
plot(h.tau, log(maint), xlab="log(ecosystem residence time)", 
     ylab="log(maintenance)", 
     main = "Per capita maintenance vs. Tau",
     col='darkolivegreen')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

plot(log(N), Skew, xlab="Total abundance", 
     ylab="Skewness",
     main = "Rarity vs. N",
     col='plum')

plot(log(N), log(Nmax), xlab="Total abundance", 
     ylab="Nmax", 
     main = "Nmax vs. N",
     col='darkorange')

plot(log(N), log(e.var), xlab="Total abundance", 
     ylab="Species Evenness", 
     main = "Evenness vs. N",
     col='lightseagreen')

plot(log(N), log(S), xlab="Total abundance",
     ylab="Species Richness", 
     main = "S vs. N",
     col='indianred')

```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

xlabel <- "log(ecosystem residence time)"
res.conc <- sim.data$resource.concentration
res.div <- sim.data$shannons.resource.diversity
res.rich <- sim.data$resource.richness

x <- c()
y <- c()
ct <- 1
for (val in res.conc){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, log(y), xlab=xlabel, 
     ylab="density", 
     main = "Resource density vs. Tau",
     col='peru')

x <- c()
y <- c()
ct <- 1
for (val in res.rich){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab=xlabel, 
     ylab="Number of resource types", 
     main = "Resource richness vs. Tau",
     col='slateblue')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rads <- list()
rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/RADs.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

rad.list[is.na(rad.list)] <- 0
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
species <- list()
species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/Species.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

species.list[is.na(species.list)] <- -1
```

### 2015-06-17 tools/bide/test/bide.py

@@ -103,7 +103,7 @@ def immigration(numin, Species, Xcoords, Ycoords, Zcoords, width, height, length
                 color_dict = get_color(prop, color_dict)
 
                 # species growth rate
-                GrowthDict[prop] = np.random.uniform(0.1, 1.0)
+                GrowthDict[prop] = np.random.uniform(0.1, 0.5)
 
                 # species maintenance
                 MaintDict[prop] = np.random.uniform(0.001, 0.01)

### 2015-06-19 analyses/2015_06_18_ResTime_GeneralResults.Rmd
conducted new analyses, reoganized data files -KL

--- 2015-06-17 analyses/2015_06_16_ResTime_GeneralResults.Rmd
@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June16/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/SimData.csv")  
+sim.data <- subset(sim.data, motion == ' fluid ')
 ```
@@
 rads <- list()
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/RADs.csv",
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/RADs.csv",
             header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
@@
 species <- list()
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/Species.csv",
+species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/Species.csv",
             header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),

### 2015-06-19 paper/Manuscript.Rmd

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

## INTRODUCTION

**P1**—Biological and ecological change is marked by time. From rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is a primary dimension of biology. But despite its universal relevance, time is more often used as an independent variable or to characterize rates of change, than it is treated as a primary driver of change and biodiversity *per se*. The general influence of time  is so rarely mentioned that it may be self-evident, non-existent, or taken for granted. This is unlike the influence of other general variables, such as space and abundance.

**P2**—Law-like patterns such as the species-area relationship and the species-abundance distribution are central to the ecological study of biodiversity (Hubbell 2001, McGill 2010, Harte 2011). Yet, a similarly general time-related pattern (species-time relationship) is rarely studied and is neither central to nor predicted by general ecological theory (White, Adler et al.). While area and abundance are common inputs to biodiversity theories (McGill 2010), none use time as an input. And, while at least one popular ecological subfield is devoted to spatial dynamics, i.e., spatial ecology (16000 results on Google scholar), "temporal ecology" has yet to be considered a subfield (590 results on Google scholar) and if temporal dynamics are often along with space, i.e., "spatio-temporal".

**P3**—The most general spatial principle is very simple, "near things are more similar than far things" (Tobler's Law) and even a precise knowledge of how abundance constrains diversity (cites), there is no general ecological rule or law relating to time. Unlike area, distance, abundance, and even the number of species (i.e., richness), there are no studies on the general influence of time on biodiversity. Other than the species-time relationship, there are no general time-related ecological patterns. Yet, at least one aspect of time applies to all components of any biological or ecological system, i.e., the amount of time that an individual organism, an amount of resource, a species, etc., remains in a system, i.e., residence time.

**P4**—Put simply, viable populations are maintained when individuals can reproduce before dying or emigrating but not so quickly that resource consumption exhausts the rate of supply. In this way, the timing of life history and population growth is constrained by residence time or even optimized for it (cite). This principle underpins the theory of the chemostat and makes residence time (τ) a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors. It also makes residence time among the most important variables influencing stability, function, and biodiversity in estuaries (cites). Yet, the influence of residence time is poorly known outside aquatic systems and though it is often manipulated to constrain growth and biomass, residence time has yet been framed as a general time-related constraint in any system, whether aquatic, terrestrial, or host-related. Yet, while many organisms are at the mercy of currents, all organisms are at the mercy of time.

**P5**—We examined the importance of residence time on the growth of populations, the assembly and structure of communities, and the evolution of trait combinations in highly complex systems. We propose that the residence time of a system should positively correlate with the residence times of its components (individuals, resources, species, traits, etc.). We then propose relationships of taxa and trait diversity based on the constraining influence of residence time. Finally, we propose the residence time distribution (RTD) as a general time-related ecological pattern that reveals many short-lived individuals, resources, species, etc., and few that are relatively long-lived. We examine factos influencing the RTD and test our predictions using an information-intensive individual-based model capable of simulating communities of competitively asymmetrical and physiologically unique species and individuals operating under trait evolution, sexual and clonal reproduction, fluid and non-fluid dynamics, variable resource efficiencies, dispersal dynamics, among others.

## METHODS

**Quantifying residence time—**In aquatic systems, residence time equals the ratio of volume to rate of flow, i.e. (V/Q = (V/V/time) = time). The inverse is known as dilution rate (1/time), i.e., the portion of particles or individuals lost or replaced per unit time. In experimental and engineered systems, dilution rate approximates growth rate and is used a master variable to control productivity and biomass (cite). For example, if dilution rate exceeds population growth rate, then populations can be washed out. But, if dilution rate is too slow (residence time is very long) then the supply of a vital resource (e.g. oxygen, nitrogen, carbon) may be too low to maintain actively growing populations.

**Residence time distributions—**

**Predictions of highly correlated residence times and residence time distributions—**

**Predictions of abundance and species diversity—**

*Abundance*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time (τ) when accounting for maintenance energy (cites). Sufficiently low τ can exclude reproduction while sufficiently high τ can mean a slow rate of resource resupply and immigration.

*Richness*—Changes in *N* are often associated with respective changes in the number of species (i.e. *S*). Richness must change as quickly or less quickly than *N*, i.e., species cannot be gained or lost faster than individuals. We expect the relationship of *S* to τ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to τ. Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*<sup>*z*</sup>, where z ~ 0.5, as has been observed by May (1988) and others (cite).

*Evenness*—Similarity in abundance among species (evenness) should decrease with *N* due to numerical constraints (Locey and White 2013). We expect a relationship of evenness to τ that is opposite of that expected for *N*, i.e., a U-shaped relationship.

*Rarity*—

*Dominance*—

*Turnover*—A monotonically decreasing relationship β-diversity to τ—β-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010). Regardless of processes or mechanisms, the most drastic changes in species turnover can occur when richness is low. We expect, species turnover to reach its greatest at lowest τ and to decrease monotonically thereafter because, as τ increases, the system becomes less likely to be colonized by other taxa.

**Trait-related predictions**

The unimodal relationship between growth rate and dilution rate (τ-1)—In simplified chemostats, dilution rate (τ-1) approximates specific (or population) growth rate. Exceptions are when dilution rate exceeds maximal growth rate and when dilution rate results in a rate of exogenous resource supply that is too low to support growth.

r vs. K selection

Generalist vs. specialist

Attachment vs. motility

Mass (M) vs. growth rate (B)—Larger body mass is strongly associated with whole organism metabolic rate (i.e., B ~ BoM3/4) but lower per unit metabolic rate (i.e., B/M ~ boM-1/4) (cite). Consequently, smaller organism require greater per cell maintenance energy than larger organisms, some of which is due to the greater surface area to volume ratio resulting in a greater rate of heat loss (cite).

Temperature-activity relationship—Active metabolism becomes costly to maintain in suboptimal temperatures as a consequence of physiological and environmental changes, e.g., unfavorable chemical kinetics, increased intracellular fluid viscosity, lack of liquid water, lack of exogenous resources (cite). Temperature is well-known to influence metabolic rates and the relationship of metabolic rate to body mass (cite). However, we expect that in a resource-poor environment with a thermal gradient that organisms in lower temperatures could more easily pursue a dormant strategy and hence, avoid starvation.

Modeling

Individual-based—We used a spatially explicit individual-based simulation model where individuals and information about individuals are tracked through time. Growth, reproduction, and death are simulated via weighted random sampling that, unlike neutral models, simulates a probabilistic nature of environmental filtering and competition. Resources and individuals enter, flow through, and exit the system according to a combination of fluid and diffusion dynamics. The model is parameterized with metacommunity, local community, ecosystem, species, and individual level parameters (Table 1). The values of the parameters determine the degree of variation among species and are chosen at random by the simulation framework. These parameters can theoretically vary from species-neutral (i.e. all species are the same) to idiosyncratic (i.e. all species are different) but nearly always include some degree of species difference in growth rate, maintenance energy, etc. (Table 1). The model operates in two spatial dimensions, and there is one upstream side and one downstream side.

Fluid flow—We used a Lattice-Boltzmann Method (LBM) to simulate movement through a fluid environment. LBMs are powerful and relatively new tools used in computational fluid dynamics, and as we show, can be used in ecology to simulate the realistic fluid flow of individuals, resource, or both (cite). LBMs discretize the environment into a lattice and attach to each position in the lattice a number of particle densities and velocities for each of nine directions of movement. Here, individual organisms and resource particle have x-y coordinates that can change (e.g. in the ideal chemostat simulation) according to the velocity of the environment at those positions (see Appendix for greater details).

Diffusion—We used simulated random walks to move resources and/or individuals through the environment via diffusion. Random walks are commonly used in ecology to model either stochastic physical movement or stochastic demographic changes where the state of an individual or population in the next time step is a random change in magnitude and direction (velocity) that is influenced by the current state; as opposed to a white-noise process where the magnitude of change in position is uniformly random and uncorrelated from step to step.

Simulated life history—In each case, the model simulated life history processes as follows: Resources and individuals flow into a system from the upstream edge of the environment. Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among taxa in ecological communities (Hubbell 2001, White et al. 2012). Along with a species ID and species-specific maximum rate of uptake, each propagule is given an individual ID (computer clock-time at which the individual entered), and a cell quota to represent the physiological state of internal resources. 
The community was sampled at random N times with replacement for each simulated time unit. This theoretically gives every individual the opportunity to grow and reproduce. Sampled individuals consume resources according to species specific rates of uptake, if there is a resource present at the individual’s location. Uptake increases the individual cell quota and decreases ambient resources within the immediate location of the individual. The individual’s quota is then decreased according to a species specific physiological maintenance cost. Individuals then reproduce with a probability equal to their cell quota and results in an equal  division of the cell quota between the individual and its daughter. The daughter is given a unique individual ID which is appended to the species ID of its mother, to maintain heritage. Sampled individuals die if their cell quota is less than or equal to 0. Individuals and resource particles flow out when they pass the downstream edge of the environment.

Simulated environmental complexity—We constructed the model to capture environmental heterogeneity and the aggregation of resources and individuals. The model was parameterized with an arbitrary environmental gradient that varied regularly across a randomly chosen spatial axis in a randomly chosen direction. Barriers were constructed within the environment to prevent unhindered flow and to generate aggregation and realistic hydrological flow.

Simulated species differences—To allow species differences in physiological traits and environmental optima, we allowed species-specific values for maximum rate of uptake rate, environmental optima, and maintenance costs to be normally distributed random variables where both the means and the standard deviations of the sampling distributions were chosen at random at the start of the simulations. Whenever an individual of a new species entered the local community, values for maximum uptake rate, environmental optima, physiological maintenance cost, and minimum cell quota were chosen at random according to the respective normal distributions for those variables.

Evaluating model performance in real-time—The model simulates over thousands of generations until a state of stationarity is reached. Because increasing values of τ leads to slow population dynamics and because the simulations can vary over three orders of magnitude in τ, we designed the model to track its own performance to ensure that time points gathered from the model were not taken during ‘burn in’ and to ensure time points were not auto-correlated. For each generation (i.e. a number of time steps equal to N, which is not constant) after the first 1000 generations, the model evaluates whether it has reached stationarity by calculating the value of the Hurst exponent (H) and the p-value of the variance ratio test for values of N recorded from the last 500 generations. Stationarity is assumed to be reached when H < 0.5 and p < 0.05 (cite). All time points before this ‘burn-in’ period are discarded. After reaching stationarity, the model runs for another 1000 generations and then stops to check if there are at least 100 sufficiently non-autocorrelated time points, as determined by the time lag after which absolute values of the autocorrelation function are less than 0.1. If not, the model runs for another 1000 generations, repeating the process until 100 non-autocorrelated time points is obtained.

Simulating over residence time (τ)—We ran the model to quasi-equilibrium for each value of τ in a series of values spanning more than two orders of magnitude, starting with τ = 0.001 and then doubling until τ = 0.512. The model was run with the same combination of resource, community ecological, and environmental (thermal, physicochemical, physical complexity) values for each value of τ in the series. After running over all τ values, the model was re-parameterized at random with a new combination of resource parameters (size, number, and types of resource particles), habitat complexity parameters (size, number, and types of physical barriers), chemical environmental parameters, thermal parameters (average temperature, direction and severity of temperature gradient), and average community-level values (average specific values for uptake rate, maintenance, individual size, motility, attachment). See table 1.

	Because the model is individual-based, information intensive, physiologically realistic, ecologically complex, fluid dynamic, self-evaluating, it cannot be run over arbitrarily slow or fast residence times. During each simulation, more than a gigabyte of information is held in memory. This is unlike analytically tractable and computationally simple Markov models or closed-form analytical models, which are more useful for obtaining precise quantitative predictions according to a single theoretical framework via traditional mathematical tools, but are less useful for modeling highly complex multivariate problems and for examining how individual-level interactions lead to aggregate behaviors. In our case, we have integrated components from metabolic theory and ecological scaling, neutral and niche theory, metacommunity theory, fluid dynamics, chemostat theory, and life history theory into a self-evaluating model that tracks the life history, location, and heritage of millions of individuals.

We used a stochastic individual-based model that simulated a wide array of environmental and ecological conditions and species-level and individual-level traits. As much of this variation as possible was assigned at random, to explore a large swath of multi-parameter space. Rather than examine the independent influence of each of these factors, our purpose was to find whether simple relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex conditions.

## Conclusion

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
```

```{r, results = 'hide', message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/SimData.csv")  
sim.data <- subset(sim.data, motion == ' fluid ')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Assign Data
sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

## Univariate relationships

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

num.sims <- max(sim.data$sim)
colors <- rainbow(num.sims)

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1
  print(i)

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)

  N <- sim.sub$N
  N <- log(N)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in N){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="Total Abundance", ylim=c(3.5, 10), xlim=c(3, 9),
    main = "N vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }

  }

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in S){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="log(species richness)", ylim=c(0, 4), xlim=c(3, 9),
    main = "S vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }
}
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Reassign Data

sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

growth <- sim.data$avg.per.capita.growth
x <- c()
y <- c()

ct <- 1
for (val in growth){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab="log(ecosystem residence time)", 
     ylab="growth rate", 
     main = "Per capita growth vs. Tau",
     col='dodgerblue')

maint <- (S*sim.data$avg.per.capita.maint)/N
plot(h.tau, log(maint), xlab="log(ecosystem residence time)", 
     ylab="log(maintenance)", 
     main = "Per capita maintenance vs. Tau",
     col='darkolivegreen')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

plot(log(N), Skew, xlab="Total abundance", 
     ylab="Skewness",
     main = "Rarity vs. N",
     col='plum')

plot(log(N), log(Nmax), xlab="Total abundance", 
     ylab="Nmax", 
     main = "Nmax vs. N",
     col='darkorange')

plot(log(N), log(e.var), xlab="Total abundance", 
     ylab="Species Evenness", 
     main = "Evenness vs. N",
     col='lightseagreen')

plot(log(N), log(S), xlab="Total abundance",
     ylab="Species Richness", 
     main = "S vs. N",
     col='indianred')

```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

xlabel <- "log(ecosystem residence time)"
res.conc <- sim.data$resource.concentration
res.div <- sim.data$shannons.resource.diversity
res.rich <- sim.data$resource.richness

x <- c()
y <- c()
ct <- 1
for (val in res.conc){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, log(y), xlab=xlabel, 
     ylab="density", 
     main = "Resource density vs. Tau",
     col='peru')

x <- c()
y <- c()
ct <- 1
for (val in res.rich){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab=xlabel, 
     ylab="Number of resource types", 
     main = "Resource richness vs. Tau",
     col='slateblue')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rads <- list()
rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/RADs.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

rad.list[is.na(rad.list)] <- 0
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
species <- list()
species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/Species.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

species.list[is.na(species.list)] <- -1
```

### 2015-06-19 paper/Manuscript.Rmd
editing Manuscript.Rmd file -KL

@@
 ---
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-rm(list=ls()) 
-getwd() 
-setwd("~/GitHub/hydrobide") 
-```
 
@@
 
-**Quantifying residence time—**In aquatic systems, residence time equals the ratio of volume to rate of flow, i.e.
-(V/Q = (V/V/time) = time).
-The inverse is known as dilution rate (1/time), i.e., the portion of particles or individuals lost or replaced per unit time.
-In experimental and engineered systems, dilution rate approximates growth rate and is used a master variable to control productivity and biomass (cite).
-For example, if dilution rate exceeds population growth rate, then populations can be washed out.
-But, if dilution rate is too slow (residence time is very long) then the supply of a vital resource (e.g.
-oxygen, nitrogen, carbon) may be too low to maintain actively growing populations.
+**Quantifying residence time—**In ideal systems (homogeneously mixed aquatic microcosms), residence time (τ) equals the ratio of volume to flow rate (τ = V/Q).
+Volume *per se* is not critical and the size of the system could also be quantified with area (A) and distance (D).
+This is easy to see by considering a perfect cube where residence times based on V, A, or D would be equal.
+That is, τ = V/Q = D^{3}/Q = D^{3}/(D^{3}/time) = time.
+The inverse of τ is dilution rate (τ^{-1}) and is the portion replaced per unit time, which often approximates growth rate.
+If τ^{-1} exceeds growth rate, then populations can be washed out.
+But, if τ^{-1} is too slow (i.e., τ is too long) then the supply resources may be too low to maintain actively growing populations.
 
-**Residence time distributions—**
+**Residence time distribution (RTD)—**In ecological systems, V/Q may be a poor approximation of individual residence time and, at least, only represents one moment of an entire distribution.
+The amount of time that particles, individuals, etc., spend inside a system can be characterized with a frequency/probability distribution (MacMullin and Weber 1935, Danckwerts 1953).
+RTDs are used among engineers and hydrologists to characterize turnover, mixing, and flow and to compare the behavior of real systems to simplified models.
+In our individual-based modeling, we examine residence time distributions for individuals, resources, inert particles, species, traits, and any other aspect of the system that is not permanent.
 
-**Predictions of highly correlated residence times and residence time distributions—**
+**Predictions of highly correlated residence times—**As a starting prediction, residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system should be highly and positively correlated.
+That is, in general and without invoking a specific scenario, very short ecosystem residence times should not lead to long residence times for individual particles.
+
+**Right-skewed residence time distributions—**
 
@@
 ## Conclusion
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+rm(list=ls()) 
+getwd() 
+setwd("~/GitHub/hydrobide") 
+```
 

### 2015-06-20 paper/Manuscript.Rmd

--- 2015-06-19 paper/Manuscript.Rmd
@@
 date: "May 31, 2015"
+setspace: doublespacing
 header-includes:
@@
 In this way, the timing of life history and population growth is constrained by residence time or even optimized for it (cite).
-This principle underpins the theory of the chemostat and makes residence time (τ) a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors.
+This principle underpins the theory of the chemostat and makes residence time a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors.
 It also makes residence time among the most important variables influencing stability, function, and biodiversity in estuaries (cites).
@@
 
-**Quantifying residence time—**In aquatic systems, residence time equals the ratio of volume to rate of flow, i.e.
-(V/Q = (V/V/time) = time).
-The inverse is known as dilution rate (1/time), i.e., the portion of particles or individuals lost or replaced per unit time.
-In experimental and engineered systems, dilution rate approximates growth rate and is used a master variable to control productivity and biomass (cite).
-For example, if dilution rate exceeds population growth rate, then populations can be washed out.
-But, if dilution rate is too slow (residence time is very long) then the supply of a vital resource (e.g.
-oxygen, nitrogen, carbon) may be too low to maintain actively growing populations.
+**Quantifying residence time**
 
-**Residence time distributions—**
+*Ecosystem residence time—*In ideal systems (homogeneously mixed aquatic microcosms), residence time ($\tau$) equals the ratio of volume to flow rate ($\tau$ = V/Q).
+Volume *per se* is not critical as the size of the system could also be quantified with area (*A*) and distance (*D*); consider that a perfect cube where residence times based on *V*, *A*, or *D* would be equal.
+That is, $\tau$ = V/Q = D$^{3}$/Q = D$^{3}$/(D$^{3}$/time) = time.
+The inverse of $\tau$ is known as dilution rate and is the portion replaced per unit time, which often approximates growth rate.
+If $\tau^{-1}$ exceeds growth rate, then populations can be washed out.
+But, if $\tau^{-1}$ is too slow (i.e., $\tau$ is too long) then the supply resources may be too low to maintain actively growing populations.
 
-**Predictions of highly correlated residence times and residence time distributions—**
+*Residence time distribution (RTD)—*In ecological systems, V/Q may be a poor approximation of individual residence time and, at least, only represents one moment of an entire distribution.
+The amount of time that particles, individuals, etc., spend inside a system can be characterized with a frequency/probability distribution (MacMullin and Weber 1935, Danckwerts 1953).
+RTDs are used among engineers and hydrologists to characterize turnover, mixing, and flow and to compare the behavior of real systems to simplified models.
+In our individual-based modeling, we examine residence time distributions for individuals, resources, inert particles, species, traits, and any other aspect of the system that is not permanent.
 
-**Predictions of abundance and species diversity—**
+**Residence time predictions**
 
-*Abundance*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time (τ) when accounting for maintenance energy (cites).
-Sufficiently low τ can exclude reproduction while sufficiently high τ can mean a slow rate of resource resupply and immigration.
+*Autocorrelated residence times—*As a starting prediction, we expect residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system to be positively correlated.
+This is already known to be the case where $\tau$ as V/Q closely approximates the residence time of individual organisms and tracer particles.
+In general, without invoking any specific scenario, very short ecosystem residence times should not lead to long residence times for individual particles.
 
-*Richness*—Changes in *N* are often associated with respective changes in the number of species (i.e.
+*Right-skewed residence time distributions—*Residence time distributions of inert tracers are generally right-skewed, often taking a lognormal or power-law form.
+This reveals many short-lived particles and relatively few long-lived ones; a result of turbulence and not flowing freely or homogeneously.
+But,whether flowing or not, we expect right-skewed distributions to characterize the life history of individuals, the presence of resource particles, and even the tenure of species.
+
+**Predictions of abundance and species diversity**
+
+*Total abundance (N)*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time ($\tau$) when accounting for maintenance energy (cites).
+Sufficiently low $\tau$ can exclude reproduction while sufficiently high $\tau$ can mean a slow rate of resource resupply and immigration.
+In this way, we expect residence time to be a primary constraint on *N*.
+In turn, *N* is a primary constraint on aspects of diversity related to the species-abundance distribution such as richness, evenness, rarity, dominance.
+
+*Richness, evenness, rarity, and dominance*—Changes in *N* are often associated with changes in the number of species (i.e.
 *S*).
 Richness must change as quickly or less quickly than *N*, i.e., species cannot be gained or lost faster than individuals.
-We expect the relationship of *S* to τ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to τ.
-Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*<sup>*z*</sup>, where z ~ 0.5, as has been observed by May (1988) and others (cite).
+We expect the relationship of *S* to $\tau$ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to $\tau$.
+Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, as has been observed by May (1988) and others (cite).
+In contrast to richness, similarity in abundance among species (evenness) should decrease with *N* purely because of numerical constraints (Locey and White 2013).
+That is, a greater portion of possible forms of the species abundance distribution reflect high rarity, low evenness, and high dominance as *N* increases.
+Consequently, greater *N* constrains the distribution of abundance among species to take a form that reflects greater rarity, greater dominance, and less evenness.
 
-*Evenness*—Similarity in abundance among species (evenness) should decrease with *N* due to numerical constraints (Locey and White 2013).
-We expect a relationship of evenness to τ that is opposite of that expected for *N*, i.e., a U-shaped relationship.
-
-*Rarity*—
-
-*Dominance*—
-
-*Turnover*—A monotonically decreasing relationship β-diversity to τ—β-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010).
-Regardless of processes or mechanisms, the most drastic changes in species turnover can occur when richness is low.
-We expect, species turnover to reach its greatest at lowest τ and to decrease monotonically thereafter because, as τ increases, the system becomes less likely to be colonized by other taxa.
+*Species turnover*—$\beta$-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010).
+Measurements of $\beta$-diversity are generally derived to be independent of *N*.
+Because of this, and because low residence times (high diluation rates) create the most likely scenario to replace a large portion of the community in given period of time, we expect a monotonically decreasing relationship $\beta$-diversity to $\tau$.
 
@@
 
-The unimodal relationship between growth rate and dilution rate (τ-1)—In simplified chemostats, dilution rate (τ-1) approximates specific (or population) growth rate.
-Exceptions are when dilution rate exceeds maximal growth rate and when dilution rate results in a rate of exogenous resource supply that is too low to support growth.
+*Growth rate vs.
+maintenance—*A large literature suggests increasing selection for quickly growing and reproducing organisms as residence time decreases.
+But, as residence time decreases, selection on rapid growth and reproduction should lessen with respect to selection on a low-cost of maintenance and the ability to persist in the face of low resource supply.
+However, organisms that can quickly consume resources, grow, and reproduce, could continue to dominance if they can keep maintenance costs down.
 
-r vs. K selection
+*Attachment vs.
+motility—*Attachment and motility have clear costs and benefits.
+Motility allows organisms to seek out resources, refuge, and mates, while attachment allows weakly dispersing organisms to resist being carried away.
+However, structures required for motility and attachment are not energy-free and presumably come at an energetic cost.
 
-Generalist vs. specialist
+**Modeling**
 
-Attachment vs. motility
-
-Mass (M) vs.
-growth rate (B)—Larger body mass is strongly associated with whole organism metabolic rate (i.e., B ~ BoM3/4) but lower per unit metabolic rate (i.e., B/M ~ boM-1/4) (cite).
-Consequently, smaller organism require greater per cell maintenance energy than larger organisms, some of which is due to the greater surface area to volume ratio resulting in a greater rate of heat loss (cite).
-
-Temperature-activity relationship—Active metabolism becomes costly to maintain in suboptimal temperatures as a consequence of physiological and environmental changes, e.g., unfavorable chemical kinetics, increased intracellular fluid viscosity, lack of liquid water, lack of exogenous resources (cite).
-Temperature is well-known to influence metabolic rates and the relationship of metabolic rate to body mass (cite).
-However, we expect that in a resource-poor environment with a thermal gradient that organisms in lower temperatures could more easily pursue a dormant strategy and hence, avoid starvation.
-
-Modeling
-
-Individual-based—We used a spatially explicit individual-based simulation model where individuals and information about individuals are tracked through time.
-Growth, reproduction, and death are simulated via weighted random sampling that, unlike neutral models, simulates a probabilistic nature of environmental filtering and competition.
-Resources and individuals enter, flow through, and exit the system according to a combination of fluid and diffusion dynamics.
+*Individual-based—*We used a spatially explicit individual-based simulation model where individuals and information about individuals are tracked through time.
+Growth, reproduction, and death are simulated via weighted random sampling that simulates a probabilistic nature of environmental filtering and competition.
+Resources, tracer particles, and individuals enter, flow through, and exit the system according to a combination of fluid and diffusion dynamics.
 The model is parameterized with metacommunity, local community, ecosystem, species, and individual level parameters (Table 1).
-The values of the parameters determine the degree of variation among species and are chosen at random by the simulation framework.
+The values of the parameters determine the degree of variation among species and are chosen at random.
 These parameters can theoretically vary from species-neutral (i.e.
@@
 (Table 1).
-The model operates in two spatial dimensions, and there is one upstream side and one downstream side.
+The model operates in two to three spatial dimensions.
 
-Fluid flow—We used a Lattice-Boltzmann Method (LBM) to simulate movement through a fluid environment.
+*Fluid flow—*We used a Lattice-Boltzmann Method (LBM) to simulate movement through a fluid environment.
 LBMs are powerful and relatively new tools used in computational fluid dynamics, and as we show, can be used in ecology to simulate the realistic fluid flow of individuals, resource, or both (cite).
@@
 
-Diffusion—We used simulated random walks to move resources and/or individuals through the environment via diffusion.
+*Diffusion—*We used simulated random walks to move resources and/or individuals through the environment via diffusion.
 Random walks are commonly used in ecology to model either stochastic physical movement or stochastic demographic changes where the state of an individual or population in the next time step is a random change in magnitude and direction (velocity) that is influenced by the current state; as opposed to a white-noise process where the magnitude of change in position is uniformly random and uncorrelated from step to step.
 
-Simulated life history—In each case, the model simulated life history processes as follows: Resources and individuals flow into a system from the upstream edge of the environment.
+*Simulated life history—*In each case, the model simulated life history processes as follows: Resources and individuals flow into a system from the upstream edge of the environment.
 Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among taxa in ecological communities (Hubbell 2001, White et al.
@@
 Along with a species ID and species-specific maximum rate of uptake, each propagule is given an individual ID (computer clock-time at which the individual entered), and a cell quota to represent the physiological state of internal resources.
-The community was sampled at random N times with replacement for each simulated time unit.
+
+*Simulated life history-*
+The community was sampled at random *N* times with replacement for each simulated time unit.
 This theoretically gives every individual the opportunity to grow and reproduce.
@@
 Evaluating model performance in real-time—The model simulates over thousands of generations until a state of stationarity is reached.
-Because increasing values of τ leads to slow population dynamics and because the simulations can vary over three orders of magnitude in τ, we designed the model to track its own performance to ensure that time points gathered from the model were not taken during ‘burn in’ and to ensure time points were not auto-correlated.
+Because increasing values of $\tau$ leads to slow population dynamics and because the simulations can vary over three orders of magnitude in $\tau$, we designed the model to track its own performance to ensure that time points gathered from the model were not taken during ‘burn in’ and to ensure time points were not auto-correlated.
 For each generation (i.e.
@@
 
-Simulating over residence time (τ)—We ran the model to quasi-equilibrium for each value of τ in a series of values spanning more than two orders of magnitude, starting with τ = 0.001 and then doubling until τ = 0.512.
-The model was run with the same combination of resource, community ecological, and environmental (thermal, physicochemical, physical complexity) values for each value of τ in the series.
-After running over all τ values, the model was re-parameterized at random with a new combination of resource parameters (size, number, and types of resource particles), habitat complexity parameters (size, number, and types of physical barriers), chemical environmental parameters, thermal parameters (average temperature, direction and severity of temperature gradient), and average community-level values (average specific values for uptake rate, maintenance, individual size, motility, attachment).
+Simulating over residence time ($\tau$)—We ran the model to quasi-equilibrium for each value of $\tau$ in a series of values spanning more than two orders of magnitude, starting with $\tau$ = 0.001 and then doubling until $\tau$ = 0.512.
+The model was run with the same combination of resource, community ecological, and environmental (thermal, physicochemical, physical complexity) values for each value of $\tau$ in the series.
+After running over all $\tau$ values, the model was re-parameterized at random with a new combination of resource parameters (size, number, and types of resource particles), habitat complexity parameters (size, number, and types of physical barriers), chemical environmental parameters, thermal parameters (average temperature, direction and severity of temperature gradient), and average community-level values (average specific values for uptake rate, maintenance, individual size, motility, attachment).
 See table 1.

### 2015-06-20 paper/Manuscript.Rmd

@@
 
-*Simulated life history-*
-The community was sampled at random *N* times with replacement for each simulated time unit.
+*Simulated life history-*The community was sampled at random *N* times with replacement for each simulated time unit.
 This theoretically gives every individual the opportunity to grow and reproduce.
@@
 
-Simulated environmental complexity—We constructed the model to capture environmental heterogeneity and the aggregation of resources and individuals.
+*Simulating environmental complexity—*We constructed the model to capture environmental heterogeneity and the aggregation of resources and individuals.
 The model was parameterized with an arbitrary environmental gradient that varied regularly across a randomly chosen spatial axis in a randomly chosen direction.
@@
 
-Simulated species differences—To allow species differences in physiological traits and environmental optima, we allowed species-specific values for maximum rate of uptake rate, environmental optima, and maintenance costs to be normally distributed random variables where both the means and the standard deviations of the sampling distributions were chosen at random at the start of the simulations.
+*Simulating species differences*—To allow species differences in physiological traits and environmental optima, we allowed species-specific values for maximum rate of uptake rate, environmental optima, and maintenance costs to be normally distributed random variables where both the means and the standard deviations of the sampling distributions were chosen at random at the start of the simulations.
 Whenever an individual of a new species entered the local community, values for maximum uptake rate, environmental optima, physiological maintenance cost, and minimum cell quota were chosen at random according to the respective normal distributions for those variables.
 
-Evaluating model performance in real-time—The model simulates over thousands of generations until a state of stationarity is reached.
-Because increasing values of $\tau$ leads to slow population dynamics and because the simulations can vary over three orders of magnitude in $\tau$, we designed the model to track its own performance to ensure that time points gathered from the model were not taken during ‘burn in’ and to ensure time points were not auto-correlated.
-For each generation (i.e.
-a number of time steps equal to N, which is not constant) after the first 1000 generations, the model evaluates whether it has reached stationarity by calculating the value of the Hurst exponent (H) and the p-value of the variance ratio test for values of N recorded from the last 500 generations.
-Stationarity is assumed to be reached when H < 0.5 and p < 0.05 (cite).
-All time points before this ‘burn-in’ period are discarded.
-After reaching stationarity, the model runs for another 1000 generations and then stops to check if there are at least 100 sufficiently non-autocorrelated time points, as determined by the time lag after which absolute values of the autocorrelation function are less than 0.1.
-If not, the model runs for another 1000 generations, repeating the process until 100 non-autocorrelated time points is obtained.
-
-Simulating over residence time ($\tau$)—We ran the model to quasi-equilibrium for each value of $\tau$ in a series of values spanning more than two orders of magnitude, starting with $\tau$ = 0.001 and then doubling until $\tau$ = 0.512.
+*Simulating over residence time ($\tau$)—*We ran the model to quasi-equilibrium for each value of $\tau$ in a series of values spanning more than two orders of magnitude, starting with $\tau$ = 0.001 and then doubling until $\tau$ = 0.512.
 The model was run with the same combination of resource, community ecological, and environmental (thermal, physicochemical, physical complexity) values for each value of $\tau$ in the series.
@@
 
-	Because the model is individual-based, information intensive, physiologically realistic, ecologically complex, fluid dynamic, self-evaluating, it cannot be run over arbitrarily slow or fast residence times.
+Because the model is individual-based, information intensive, physiologically realistic, ecologically complex, fluid dynamic, it cannot be run over arbitrarily slow or fast residence times.
 During each simulation, more than a gigabyte of information is held in memory.
@@
 
+## Results
+
+*Autocorrelated residence times and right-skewed residence time distributions——*As expected, residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated.
+Right-skewed distributions characterized the life history of individuals, the presence of resource particles, and the tenure of species and traits (Figure ?).
+Residence time distributions were generally well approximated by lognormal and power-law models (e.g.
+Pareto, Zipf, Zipf-Mandelbrot).
+
+**Predictions of abundance and species diversity**
+
+*Total abundance (N)*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time ($\tau$) when accounting for maintenance energy (cites).
+Sufficiently low $\tau$ can exclude reproduction while sufficiently high $\tau$ can mean a slow rate of resource resupply and immigration.
+In this way, we expect residence time to be a primary constraint on *N*.
+In turn, *N* is a primary constraint on aspects of diversity related to the species-abundance distribution such as richness, evenness, rarity, dominance.
+
+*Richness, evenness, rarity, and dominance*—Changes in *N* are often associated with changes in the number of species (i.e.
+*S*).
+Richness must change as quickly or less quickly than *N*, i.e., species cannot be gained or lost faster than individuals.
+We expect the relationship of *S* to $\tau$ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to $\tau$.
+Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, as has been observed by May (1988) and others (cite).
+In contrast to richness, similarity in abundance among species (evenness) should decrease with *N* purely because of numerical constraints (Locey and White 2013).
+That is, a greater portion of possible forms of the species abundance distribution reflect high rarity, low evenness, and high dominance as *N* increases.
+Consequently, greater *N* constrains the distribution of abundance among species to take a form that reflects greater rarity, greater dominance, and less evenness.
+
+*Species turnover*—$\beta$-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010).
+Measurements of $\beta$-diversity are generally derived to be independent of *N*.
+Because of this, and because low residence times (high diluation rates) create the most likely scenario to replace a large portion of the community in given period of time, we expect a monotonically decreasing relationship $\beta$-diversity to $\tau$.
+
+**Trait-related predictions**
+
+*Growth rate vs.
+maintenance—*A large literature suggests increasing selection for quickly growing and reproducing organisms as residence time decreases.
+But, as residence time decreases, selection on rapid growth and reproduction should lessen with respect to selection on a low-cost of maintenance and the ability to persist in the face of low resource supply.
+However, organisms that can quickly consume resources, grow, and reproduce, could continue to dominance if they can keep maintenance costs down.
+
+*Attachment vs.
+motility—*Attachment and motility have clear costs and benefits.
+Motility allows organisms to seek out resources, refuge, and mates, while attachment allows weakly dispersing organisms to resist being carried away.
+However, structures required for motility and attachment are not energy-free and presumably come at an energetic cost.
+
+## Conclusion
+
 We used a stochastic individual-based model that simulated a wide array of environmental and ecological conditions and species-level and individual-level traits.
@@
 Rather than examine the independent influence of each of these factors, our purpose was to find whether simple relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex conditions.
-
-## Conclusion
 

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 Rather than examine the independent influence of each of these factors, our purpose was to find whether simple relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex conditions.
+
+The random selection of trait values of new species (whether immigrating or evolving) allowed realistic and unreasonable trait combinations to occur.
+For example, a species requiring little maintenance but capable of rapid growth was possible.
+However, the trait combinations that persisted in the model were realistic combinations of traits that point to life history trade-offs.
 

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 
-*Autocorrelated residence times and right-skewed residence time distributions——*As expected, residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated.
-Right-skewed distributions characterized the life history of individuals, the presence of resource particles, and the tenure of species and traits (Figure ?).
-Residence time distributions were generally well approximated by lognormal and power-law models (e.g.
-Pareto, Zipf, Zipf-Mandelbrot).
+*Autocorrelated residence times and right-skewed residence time distributions—*The probabilistic and complex modeling approach produced all primary expected results, signifying that predictions of otherwise simplistic and idealized chemostat theory could arise in a complex system.
+Residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated (Fig 1).
+Likewise, residence time distributions were right-skewed and well approximated by lognormal and power-law models (e.g.
+Pareto, Zipf, Zipf-Mandelbrot) (Fig 2).
+We also observed the expected  hump-shaped relationship of *N* to residence time, revealing that residence time could be low enough to exclude reproduction and high enough to result in a rate of resource supply that could not support optimal growth.
 
@@
 
-*Total abundance (N)*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time ($\tau$) when accounting for maintenance energy (cites).
-Sufficiently low $\tau$ can exclude reproduction while sufficiently high $\tau$ can mean a slow rate of resource resupply and immigration.
-In this way, we expect residence time to be a primary constraint on *N*.
-In turn, *N* is a primary constraint on aspects of diversity related to the species-abundance distribution such as richness, evenness, rarity, dominance.
+*Total abundance (N)*—
 
@@
 
-## Conclusion
+## Conclusion/Discussion
 
@@
 
-```{r, results = 'hide', message = FALSE, warning = FALSE}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 ## Load Data

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 
-*Autocorrelated residence times and right-skewed residence time distributions—*The probabilistic and complex modeling approach produced all primary expected results, signifying that predictions of otherwise simplistic and idealized chemostat theory could arise in a complex system.
-Residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated (Fig 1).
-Likewise, residence time distributions were right-skewed and well approximated by lognormal and power-law models (e.g.
+**Simplistic theory supported in complex systems**
+The probabilistic and complex modeling approach produced all primary expected results, signifying that predictions of otherwise simplistic and idealized chemostat theory could arise in a complex system.
+
+*Autocorrelated residence times—*Residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated (Fig 1).
+
+*Autocorrelated residence times and right-skewed residence time distributions—*Likewise, residence time distributions were right-skewed and well approximated by lognormal and power-law models (e.g.
 Pareto, Zipf, Zipf-Mandelbrot) (Fig 2).
-We also observed the expected  hump-shaped relationship of *N* to residence time, revealing that residence time could be low enough to exclude reproduction and high enough to result in a rate of resource supply that could not support optimal growth.
+
+*Abundance-residence time relationship—*We also observed the expected hump-shaped relationship of *N* to residence time, revealing that residence time could be low enough to exclude reproduction and high enough to result in a rate of resource supply that could not support optimal growth.
 
 **Predictions of abundance and species diversity**
-
-*Total abundance (N)*—
 

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 
-**P1**—Biological and ecological change is marked by time.
-From rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is a primary dimension of biology.
-But despite its universal relevance, time is more often used as an independent variable or to characterize rates of change, than it is treated as a primary driver of change and biodiversity *per se*.
-The general influence of time  is so rarely mentioned that it may be self-evident, non-existent, or taken for granted.
-This is unlike the influence of other general variables, such as space and abundance.
+**P1**—Biological and ecological change is marked by time and from rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is the fundamental dimension of biology.
+But despite its universal relevance in biological and ecological systems, time is more often used as an independent variable or to characterize rates of change, than it is treated as a primary driver of change and biodiversity.
+While the large lack of study into the general influence of time is recognized among some ecologists (cite), the precise importance or general influence of time is so rarely mentioned that it may be self-evident, non-existent, or commonly taken for granted.
+But, this is unlike the influence of other general variables such as space and abundance.
 
-**P2**—Law-like patterns such as the species-area relationship and the species-abundance distribution are central to the ecological study of biodiversity (Hubbell 2001, McGill 2010, Harte 2011).
-Yet, a similarly general time-related pattern (species-time relationship) is rarely studied and is neither central to nor predicted by general ecological theory (White, Adler et al.).
-While area and abundance are common inputs to biodiversity theories (McGill 2010), none use time as an input.
-And, while at least one popular ecological subfield is devoted to spatial dynamics, i.e., spatial ecology (16000 results on Google scholar), "temporal ecology" has yet to be considered a subfield (590 results on Google scholar) and if temporal dynamics are often along with space, i.e., "spatio-temporal".
+**P2**—Law-like patterns of space and abundance such as the species-area relationship and the species-abundance distribution are central to the ecological study of biodiversity (Hubbell 2001, McGill 2010, Harte 2011).
+Yet, a similarly general temporal pattern (species-time relationship) is rarely studied and is neither central to nor predicted by general ecological theory (White, Adler et al.).
+While area and abundance are common inputs to biodiversity theories (McGill 2010), no general theory uses time *per se* as an input.
+And, while at least one popular ecological subfield is devoted to spatial dynamics, i.e., spatial ecology (16000 results on Google scholar), "temporal ecology" has yet to be deemed a subfield (590 results on Google scholar) while temporal dynamics are studied along with spatial, i.e., "spatio-temporal".
 
-**P3**—The most general spatial principle is very simple, "near things are more similar than far things" (Tobler's Law) and even a precise knowledge of how abundance constrains diversity (cites), there is no general ecological rule or law relating to time.
-Unlike area, distance, abundance, and even the number of species (i.e., richness), there are no studies on the general influence of time on biodiversity.
-Other than the species-time relationship, there are no general time-related ecological patterns.
-Yet, at least one aspect of time applies to all components of any biological or ecological system, i.e., the amount of time that an individual organism, an amount of resource, a species, etc., remains in a system, i.e., residence time.
+**P3**—Even if the general influence of time were very simple, it would be no simpler and no less profound than the most general spatial principle, i.e., "near things are more similar than far things", also known as Tobler's Law or Tobler's First Rule of Geography (cite).
+Likewise, it might be no less straightforward, understated, and mathematically necessary than the constraining influence of total abundance and sample size on intensively studied patterns of commonness and rarity (Locey and White 2013, Xiao et al.
+2015).
+Yet there is no general ecological rule or law relating to time.
+There are no studies on the general influence of time on biodiversity and, other than the species-time relationship, there are no general time-related ecological patterns.
+Yet, at least one aspect of time applies to all components of any biological or ecological system, i.e., the amount of time that an individual organism, resource particle, species, etc., remains in a system; residence time.
 
 **P4**—Put simply, viable populations are maintained when individuals can reproduce before dying or emigrating but not so quickly that resource consumption exhausts the rate of supply.
-In this way, the timing of life history and population growth is constrained by residence time or even optimized for it (cite).
-This principle underpins the theory of the chemostat and makes residence time a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors.
-It also makes residence time among the most important variables influencing stability, function, and biodiversity in estuaries (cites).
-Yet, the influence of residence time is poorly known outside aquatic systems and though it is often manipulated to constrain growth and biomass, residence time has yet been framed as a general time-related constraint in any system, whether aquatic, terrestrial, or host-related.
+In this way, the timing of life history and population growth is constrained by residence time or perhaps even optimized for it (cite).
+This principle underpins chemostat theory, making residence time a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors.
+It also makes residence time among the most important variables influencing stability, function, and biodiversity in sensitive ecosystems like estuaries (cites).
+Yet, the influence of residence time is poorly known outside aquatic systems and though it is often manipulated to control growth and biomass, residence time has yet been framed as a time-related constraint common to any system, whether aquatic, terrestrial, or host-related.
 Yet, while many organisms are at the mercy of currents, all organisms are at the mercy of time.
@@
 
-**Residence time predictions**
+**Primary residence time predictions that should hold up to ecological complexity**
 
@@
 
-**Predictions of abundance and species diversity**
-
 *Total abundance (N)*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time ($\tau$) when accounting for maintenance energy (cites).
@@
 In turn, *N* is a primary constraint on aspects of diversity related to the species-abundance distribution such as richness, evenness, rarity, dominance.
+
+**Residence time constrains *N*, while *N* constrains commonness and rarity**
 
@@
 
-**Trait-related predictions**
+**Residence time influences at least one general life history tradeoff**
 
 *Growth rate vs.
-maintenance—*A large literature suggests increasing selection for quickly growing and reproducing organisms as residence time decreases.
-But, as residence time decreases, selection on rapid growth and reproduction should lessen with respect to selection on a low-cost of maintenance and the ability to persist in the face of low resource supply.
-However, organisms that can quickly consume resources, grow, and reproduce, could continue to dominance if they can keep maintenance costs down.
-
-*Attachment vs.
-motility—*Attachment and motility have clear costs and benefits.
-Motility allows organisms to seek out resources, refuge, and mates, while attachment allows weakly dispersing organisms to resist being carried away.
-However, structures required for motility and attachment are not energy-free and presumably come at an energetic cost.
+maintenance costs—*Organisms with slow growth should not be favored in systems of fast turnover, while organisms with high maintenance costs should not be favored in systems of slow resource supply.
+Yet, unless growth rate and maintenance are directly related (high growth requiring high maintenance), the relationship between these two traits could be tenuous.
 

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 ---
-title: "How residence time constrains diversity"
+title: "Residence time, a master variable for a more temporal ecology"
 author: "Ken Locey, Jay Lennon"
-date: "May 31, 2015"
+date: "20 June, 2015"
 setspace: doublespacing
@@
 
-**P1**—Biological and ecological change is marked by time and from rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is the fundamental dimension of biology.
+**P1**—Biological processes and ecological change are marked by time.
+From rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is the central dimension of biology.
 But despite its universal relevance in biological and ecological systems, time is more often used as an independent variable or to characterize rates of change, than it is treated as a primary driver of change and biodiversity.

### 2015-06-20 paper/Manuscript.Rmd
finishing up Results -KL

@@
 
-*Richness, evenness, rarity, and dominance*—Changes in *N* are often associated with changes in the number of species (i.e.
-*S*).
-Richness must change as quickly or less quickly than *N*, i.e., species cannot be gained or lost faster than individuals.
-We expect the relationship of *S* to $\tau$ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to $\tau$.
-Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, as has been observed by May (1988) and others (cite).
-In contrast to richness, similarity in abundance among species (evenness) should decrease with *N* purely because of numerical constraints (Locey and White 2013).
-That is, a greater portion of possible forms of the species abundance distribution reflect high rarity, low evenness, and high dominance as *N* increases.
-Consequently, greater *N* constrains the distribution of abundance among species to take a form that reflects greater rarity, greater dominance, and less evenness.
+*Richness, evenness, rarity, and dominance*—As expected and as shown in many previous studies, *S* shared a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, revealing that the model generated a realistic dynamic whether including or excluding immigration and speciation and whether reproduction was sexual or clonal, and whether movement was fluid, random, or environmentally correlated.
+As predicted, the relationship of *S* to $\tau$ followed a hump-shaped relationship.
+Also as predicted, greater *N* lead to greater rarity, greater dominance, and less evenness, regardless of the model specifics.
 
-*Species turnover*—$\beta$-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010).
-Measurements of $\beta$-diversity are generally derived to be independent of *N*.
-Because of this, and because low residence times (high diluation rates) create the most likely scenario to replace a large portion of the community in given period of time, we expect a monotonically decreasing relationship $\beta$-diversity to $\tau$.
+*Species turnover*—We observed a monotonically decreasing relationship $\beta$-diversity to $\tau$, when calculated as species turnover (presence/absence).
 
@@
 *Growth rate vs.
-maintenance—*A large literature suggests increasing selection for quickly growing and reproducing organisms as residence time decreases.
-But, as residence time decreases, selection on rapid growth and reproduction should lessen with respect to selection on a low-cost of maintenance and the ability to persist in the face of low resource supply.
-However, organisms that can quickly consume resources, grow, and reproduce, could continue to dominance if they can keep maintenance costs down.
-
-*Attachment vs.
-motility—*Attachment and motility have clear costs and benefits.
-Motility allows organisms to seek out resources, refuge, and mates, while attachment allows weakly dispersing organisms to resist being carried away.
-However, structures required for motility and attachment are not energy-free and presumably come at an energetic cost.
+maintenance—*Increasing residence time lead to a strong power-law decrease in maintenance, revealing strong selective pressure to evolve low maintenance costs in resource limited environments of low turnover.
 
@@
 We used a stochastic individual-based model that simulated a wide array of environmental and ecological conditions and species-level and individual-level traits.
-As much of this variation as possible was assigned at random, to explore a large swath of multi-parameter space.
-Rather than examine the independent influence of each of these factors, our purpose was to find whether simple relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex conditions.
+As much of this variation as possible was assigned at random, to explore a large swath of conditions and parameter space.
+Rather than examine the independent influence of each of these factors, our purpose was to find whether previously known and novel relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex systems and characteristically different environments.
 
-The random selection of trait values of new species (whether immigrating or evolving) allowed realistic and unreasonable trait combinations to occur.
-For example, a species requiring little maintenance but capable of rapid growth was possible.
-However, the trait combinations that persisted in the model were realistic combinations of traits that point to life history trade-offs.
+...
 

### 2015-06-27 paper/Manuscript.Rmd
modified bide and main model code, added new data files, conducted new analyses, and added new .Rmd and .pdfs -KL

@@
 require("vegan") 
+#install.packages("moments")
+require("moments")
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-rads <- list()
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/RADs.csv",
+## Loading simulated RADs and species lists
+
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv",
             header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
             as.is = TRUE, skipNul = TRUE, fill = TRUE)
+rad.list[is.na(rad.list)] <- 0
 
-rad.list[is.na(rad.list)] <- 0
+species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv",
+            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
+            as.is = TRUE, skipNul = TRUE, fill = TRUE)
+species.list[is.na(species.list)] <- -1
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-species <- list()
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/Species.csv",
-            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
-            as.is = TRUE, skipNul = TRUE, fill = TRUE)
+## Loading simulated data csv files
 
-species.list[is.na(species.list)] <- -1
+sims <- c()
+cts <- c()
+tracer.RTDs <- list()
+res.RTDs <- list()
+ind.RTDs <- list()
+
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/IndRTD.csv"
+con  <- file(inputFile, open = "r")
+ct = 1
+while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
+  line <- as.numeric(myLine)
+  cts <- c(cts, line[1])
+  sims <- c(sims, line[2])
+  rtd <- line[-1:-2]
+  ind.RTDs[[ct]] <- rtd
+  ct <- ct + 1
+  } 
+close(con)
+
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/TracerRTD.csv"
+con  <- file(inputFile, open = "r")
+ct = 1
+while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
+  line <- as.numeric(myLine)
+  rtd <- line[-1:-2]
+  tracer.RTDs[[ct]] <- rtd
+  ct <- ct + 1
+  } 
+close(con)
+
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/ResRTD.csv"
+con  <- file(inputFile, open = "r")
+ct = 1
+while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
+  line <- as.numeric(myLine)
+  rtd <- line[-1:-2]
+  res.RTDs[[ct]] <- rtd
+  ct <- ct + 1
+  } 
+close(con)
 ```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+par(mfrow=c(2, 2))
+
+t.skews <- c()
+i.skews <- c()
+r.skews <- c()
+t.means <- c()
+i.means <- c()
+r.means <- c()
+
+colors <- rainbow(length(ind.RTDs))
+plot(density(log(tracer.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2),  main = "Tracer RTD")
+
+ct <- 1
+for(rtd in tracer.RTDs){
+  if(any(is.na(rtd)) == FALSE){
+    lines(density(log(rtd), bw = 0.2), col = colors[ct])
+    t.skews <- c(t.skews, skewness(rtd))
+    t.means <- c(t.means, mean(rtd))
+    }
+  ct = ct + 1
+}
+
+plot(density(log(ind.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2), main = "Individual RTD")
+
+ct <- 1
+for(rtd in ind.RTDs){
+  if(any(is.na(rtd)) == FALSE){
+    lines(density(log(rtd), bw = 0.2), col = colors[ct])
+    i.skews <- c(i.skews, skewness(rtd))
+    i.means <- c(i.means, mean(rtd))
+    }
+  ct = ct + 1
+}
+
+plot(density(log(ind.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2), main = "Resource RTD")
+
+ct <- 1
+for(rtd in res.RTDs){
+  if(any(is.na(rtd)) == FALSE){
+    lines(density(log(rtd), bw = 0.2), col = colors[ct])
+    r.skews <- c(r.skews, skewness(rtd))
+    r.means <- c(r.means, mean(rtd))
+    }
+  ct = ct + 1
+}
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+par(mfrow=c(1, 1))
+
+plot(log(t.means), t.skews, col = 'darkmagenta', xlab='mean', ylab='skewness', xlim = c(1, 8), ylim = c(-1.2, 3), lwd=2, 
+     main = "Most RTD's are right-skewed,\n(many short-lived, few long-lived)")
+
+points(log(i.means),i.skews, lwd=2, col = 'darkcyan')
+points(log(r.means), r.skews, lwd=2, col = 'black')
+
+legend("topright", legend = c("tracers", "individuals", "resources"), pch=20, bty = "o", col = c("darkmagenta", "darkcyan", "black"))
+```

### 2015-06-27 models/MainModel/Generator/test/model.py

@@ -43,7 +43,7 @@ def get_rand_params():
     rmax = choice([500, 1000, 2000, 4000, 8000]) # maximum resource particle size
 
     #m = 0
-    motion = 'random_walk'
+    motion = 'fluid'
     reproduction = 'fission'
     speciation = 'yes'
 
@@ -163,6 +163,7 @@ def nextFrame(arg):	# arg is the frame number
     ax.set_xlim(0, width)
     if D == 3: ax.set_zlim(0,length)
 
+    """
     ##### PLOTTING THE INDIVIDUALS ############################################
     resource_scatImage.remove()
     tracer_scatImage.remove()
@@ -182,6 +183,7 @@ def nextFrame(arg):	# arg is the frame number
 
     if D == 2: tracer_scatImage = ax.scatter(TracerXcoords, TracerYcoords, s = 200, c = 'r', marker='*', lw=0.0, alpha=0.6)
     elif D == 3: tracer_scatImage = ax.scatter(TracerXcoords, TracerYcoords, TracerZcoords, s = 200, c = 'r', marker='*', lw=0.0, alpha=0.8)
+    """
 
     plt.draw()
 
@@ -215,6 +217,20 @@ def nextFrame(arg):	# arg is the frame number
         TracerTau, IndTau, ResTau = [np.mean(TracerExitAge), np.mean(IndExitAge), np.mean(ResExitAge)]
         # Number of tracers, resource particles, and individuals
         T, R, N = len(TracerIDs), len(ResIDs), len(SpeciesIDs)
+
+        IndRTD = str(IndExitAge).strip('[]')
+        TracerRTD = str(TracerExitAge).strip('[]')
+        ResRTD = str(ResExitAge).strip('[]')
+        OUT4 = open(mydir + '/GitHub/hydrobide/results/simulated_data/IndRTD.csv','a')
+        OUT5 = open(mydir + '/GitHub/hydrobide/results/simulated_data/TracerRTD.csv','a')
+        OUT6 = open(mydir + '/GitHub/hydrobide/results/simulated_data/ResRTD.csv','a')
+        print>>OUT4, ct1,',', sim,',', IndRTD
+        print>>OUT5, ct1,',', sim,',', TracerRTD
+        print>>OUT6, ct1,',', sim,',', ResRTD
+        OUT4.close()
+        OUT5.close()
+        OUT6.close()
+
         TracerExitAge, IndExitAge, ResExitAge= [],[],[]
[…]

### 2015-06-27 tools/bide/test/bide.py

@@ -40,7 +40,7 @@ def NewTracers(IDs, Xcoords, Ycoords, Zcoords, TimeIn, width, height, length, u0
     if x == 1:
         TimeIn.append(0)
         Ycoords.append(float(np.random.uniform(0.4*height, 0.6*height)))
-        Xcoords.append(float(np.random.uniform(0.1*width, 0.11*width)))
+        Xcoords.append(float(np.random.uniform(0.1*width, 0.5*width)))
         Zcoords.append(float(np.random.uniform(0.4*length, 0.6*length)))
         IDs.append(0) # used to track the age of the tracer
 
@@ -181,26 +181,6 @@ def fluid_movement(TypeOf, List, TimeIn, ExitAge, Xcoords, Ycoords, ux, uy, widt
 
 
 
-def maintenance(SpeciesIDs, Xcoords, Ycoords, Zcoords, ExitAge, color_dict, MaintDict, IDs, TimeIn, Qs, D):
-
-    if SpeciesIDs == []: return [SpeciesIDs, Xcoords, Ycoords, Zcoords, ExitAge, IDs, TimeIn, Qs]
-
-    for i, val in enumerate(Qs):
-        val -= MaintDict[SpeciesIDs[i]]  # maintanence influenced by species
-        if val <= 0.01:   # starved
-
-            Qs.pop(i)
-            ExitAge.append(TimeIn[i])
-            TimeIn.pop(i)
-            SpeciesIDs.pop(i)
-            IDs.pop(i)
-            Xcoords.pop(i)
-            Ycoords.pop(i)
-            Zcoords.pop(i)
-
-        else: Qs[i] = val
-
-    return [SpeciesIDs, Xcoords, Ycoords, Zcoords, ExitAge, IDs, TimeIn, Qs]
 
 
 
@@ -266,6 +246,31 @@ def predation(PredIDs, PredID, PredXcoords, PredYcoords, PredZcoords, PredTimeIn
             IndZcoords.pop(j)
 
     return [PredIDs, PredID, PredTimeIn, PredXcoords, PredYcoords, PredZcoords, SpeciesIDs, Qs, IndIDs, IndID, IndTimeIn, IndXcoords, IndYcoords, IndZcoords]
+
+
+
+
+
+def maintenance(SpeciesIDs, Xcoords, Ycoords, Zcoords, ExitAge, color_dict, MaintDict, IDs, TimeIn, Qs, D):
+
[…]

### 2015-06-27 analyses/2015_06_27_ResTime.Rmd

--- 2015-06-27 paper/Manuscript.Rmd
@@
 ---
-title: "Residence time, a master variable for a more temporal ecology"
+title: "Residence time, a master variable of ecology, physiology, and natural selection"
 author: "Ken Locey, Jay Lennon"
-date: "20 June, 2015"
-setspace: doublespacing
+date: "June, 2015"
 header-includes:
@@
 ---
-
-## INTRODUCTION
-
-**P1**—Biological processes and ecological change are marked by time.
-From rates of metabolism, growth, and evolution to the duration of seasons, succession, and life history, time is the central dimension of biology.
-But despite its universal relevance in biological and ecological systems, time is more often used as an independent variable or to characterize rates of change, than it is treated as a primary driver of change and biodiversity.
-While the large lack of study into the general influence of time is recognized among some ecologists (cite), the precise importance or general influence of time is so rarely mentioned that it may be self-evident, non-existent, or commonly taken for granted.
-But, this is unlike the influence of other general variables such as space and abundance.
-
-**P2**—Law-like patterns of space and abundance such as the species-area relationship and the species-abundance distribution are central to the ecological study of biodiversity (Hubbell 2001, McGill 2010, Harte 2011).
-Yet, a similarly general temporal pattern (species-time relationship) is rarely studied and is neither central to nor predicted by general ecological theory (White, Adler et al.).
-While area and abundance are common inputs to biodiversity theories (McGill 2010), no general theory uses time *per se* as an input.
-And, while at least one popular ecological subfield is devoted to spatial dynamics, i.e., spatial ecology (16000 results on Google scholar), "temporal ecology" has yet to be deemed a subfield (590 results on Google scholar) while temporal dynamics are studied along with spatial, i.e., "spatio-temporal".
-
-**P3**—Even if the general influence of time were very simple, it would be no simpler and no less profound than the most general spatial principle, i.e., "near things are more similar than far things", also known as Tobler's Law or Tobler's First Rule of Geography (cite).
-Likewise, it might be no less straightforward, understated, and mathematically necessary than the constraining influence of total abundance and sample size on intensively studied patterns of commonness and rarity (Locey and White 2013, Xiao et al.
-2015).
-Yet there is no general ecological rule or law relating to time.
-There are no studies on the general influence of time on biodiversity and, other than the species-time relationship, there are no general time-related ecological patterns.
-Yet, at least one aspect of time applies to all components of any biological or ecological system, i.e., the amount of time that an individual organism, resource particle, species, etc., remains in a system; residence time.
-
-**P4**—Put simply, viable populations are maintained when individuals can reproduce before dying or emigrating but not so quickly that resource consumption exhausts the rate of supply.
-In this way, the timing of life history and population growth is constrained by residence time or perhaps even optimized for it (cite).
-This principle underpins chemostat theory, making residence time a master variable for controlling growth and biomass in experimental microcosms and engineered bioreactors.
-It also makes residence time among the most important variables influencing stability, function, and biodiversity in sensitive ecosystems like estuaries (cites).
-Yet, the influence of residence time is poorly known outside aquatic systems and though it is often manipulated to control growth and biomass, residence time has yet been framed as a time-related constraint common to any system, whether aquatic, terrestrial, or host-related.
-Yet, while many organisms are at the mercy of currents, all organisms are at the mercy of time.
-
-**P5**—We examined the importance of residence time on the growth of populations, the assembly and structure of communities, and the evolution of trait combinations in highly complex systems.
-We propose that the residence time of a system should positively correlate with the residence times of its components (individuals, resources, species, traits, etc.).
-We then propose relationships of taxa and trait diversity based on the constraining influence of residence time.
-Finally, we propose the residence time distribution (RTD) as a general time-related ecological pattern that reveals many short-lived individuals, resources, species, etc., and few that are relatively long-lived.
-We examine factos influencing the RTD and test our predictions using an information-intensive individual-based model capable of simulating communities of competitively asymmetrical and physiologically unique species and individuals operating under trait evolution, sexual and clonal reproduction, fluid and non-fluid dynamics, variable resource efficiencies, dispersal dynamics, among others.
-
-## METHODS
-
-**Quantifying residence time**
-
-*Ecosystem residence time—*In ideal systems (homogeneously mixed aquatic microcosms), residence time ($\tau$) equals the ratio of volume to flow rate ($\tau$ = V/Q).
-Volume *per se* is not critical as the size of the system could also be quantified with area (*A*) and distance (*D*); consider that a perfect cube where residence times based on *V*, *A*, or *D* would be equal.
-That is, $\tau$ = V/Q = D$^{3}$/Q = D$^{3}$/(D$^{3}$/time) = time.
-The inverse of $\tau$ is known as dilution rate and is the portion replaced per unit time, which often approximates growth rate.
-If $\tau^{-1}$ exceeds growth rate, then populations can be washed out.
-But, if $\tau^{-1}$ is too slow (i.e., $\tau$ is too long) then the supply resources may be too low to maintain actively growing populations.
-
-*Residence time distribution (RTD)—*In ecological systems, V/Q may be a poor approximation of individual residence time and, at least, only represents one moment of an entire distribution.
-The amount of time that particles, individuals, etc., spend inside a system can be characterized with a frequency/probability distribution (MacMullin and Weber 1935, Danckwerts 1953).
-RTDs are used among engineers and hydrologists to characterize turnover, mixing, and flow and to compare the behavior of real systems to simplified models.
-In our individual-based modeling, we examine residence time distributions for individuals, resources, inert particles, species, traits, and any other aspect of the system that is not permanent.
-
-**Primary residence time predictions that should hold up to ecological complexity**
-
-*Autocorrelated residence times—*As a starting prediction, we expect residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system to be positively correlated.
-This is already known to be the case where $\tau$ as V/Q closely approximates the residence time of individual organisms and tracer particles.
-In general, without invoking any specific scenario, very short ecosystem residence times should not lead to long residence times for individual particles.
-
-*Right-skewed residence time distributions—*Residence time distributions of inert tracers are generally right-skewed, often taking a lognormal or power-law form.
-This reveals many short-lived particles and relatively few long-lived ones; a result of turbulence and not flowing freely or homogeneously.
-But,whether flowing or not, we expect right-skewed distributions to characterize the life history of individuals, the presence of resource particles, and even the tenure of species.
-
-*Total abundance (N)*—Chemostat models predict a unimodal relationship between biomass or total abundance (*N*) and residence time ($\tau$) when accounting for maintenance energy (cites).
-Sufficiently low $\tau$ can exclude reproduction while sufficiently high $\tau$ can mean a slow rate of resource resupply and immigration.
-In this way, we expect residence time to be a primary constraint on *N*.
-In turn, *N* is a primary constraint on aspects of diversity related to the species-abundance distribution such as richness, evenness, rarity, dominance.
-
-**Residence time constrains *N*, while *N* constrains commonness and rarity**
-
-*Richness, evenness, rarity, and dominance*—Changes in *N* are often associated with changes in the number of species (i.e.
-*S*).
-Richness must change as quickly or less quickly than *N*, i.e., species cannot be gained or lost faster than individuals.
-We expect the relationship of *S* to $\tau$ to follow a hump-shaped relationship, the slope of which must be equal to or shallower than the relationship of N to $\tau$.
-Moreover, we expect *S* to share a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, as has been observed by May (1988) and others (cite).
-In contrast to richness, similarity in abundance among species (evenness) should decrease with *N* purely because of numerical constraints (Locey and White 2013).
-That is, a greater portion of possible forms of the species abundance distribution reflect high rarity, low evenness, and high dominance as *N* increases.
-Consequently, greater *N* constrains the distribution of abundance among species to take a form that reflects greater rarity, greater dominance, and less evenness.
-
-*Species turnover*—$\beta$-diversity is the heterogeneity in taxa among samples, either with respect to presence/absence or relative abundance (Tuomisto 2010).
-Measurements of $\beta$-diversity are generally derived to be independent of *N*.
-Because of this, and because low residence times (high diluation rates) create the most likely scenario to replace a large portion of the community in given period of time, we expect a monotonically decreasing relationship $\beta$-diversity to $\tau$.
-
-**Residence time influences at least one general life history tradeoff**
-
-*Growth rate vs.
-maintenance costs—*Organisms with slow growth should not be favored in systems of fast turnover, while organisms with high maintenance costs should not be favored in systems of slow resource supply.
-Yet, unless growth rate and maintenance are directly related (high growth requiring high maintenance), the relationship between these two traits could be tenuous.
-
-**Modeling**
-
-*Individual-based—*We used a spatially explicit individual-based simulation model where individuals and information about individuals are tracked through time.
-Growth, reproduction, and death are simulated via weighted random sampling that simulates a probabilistic nature of environmental filtering and competition.
-Resources, tracer particles, and individuals enter, flow through, and exit the system according to a combination of fluid and diffusion dynamics.
-The model is parameterized with metacommunity, local community, ecosystem, species, and individual level parameters (Table 1).
-The values of the parameters determine the degree of variation among species and are chosen at random.
-These parameters can theoretically vary from species-neutral (i.e.
-all species are the same) to idiosyncratic (i.e.
-all species are different) but nearly always include some degree of species difference in growth rate, maintenance energy, etc.
-(Table 1).
-The model operates in two to three spatial dimensions.
-
-*Fluid flow—*We used a Lattice-Boltzmann Method (LBM) to simulate movement through a fluid environment.
-LBMs are powerful and relatively new tools used in computational fluid dynamics, and as we show, can be used in ecology to simulate the realistic fluid flow of individuals, resource, or both (cite).
-LBMs discretize the environment into a lattice and attach to each position in the lattice a number of particle densities and velocities for each of nine directions of movement.
-Here, individual organisms and resource particle have x-y coordinates that can change (e.g.
-in the ideal chemostat simulation) according to the velocity of the environment at those positions (see Appendix for greater details).
-
-*Diffusion—*We used simulated random walks to move resources and/or individuals through the environment via diffusion.
-Random walks are commonly used in ecology to model either stochastic physical movement or stochastic demographic changes where the state of an individual or population in the next time step is a random change in magnitude and direction (velocity) that is influenced by the current state; as opposed to a white-noise process where the magnitude of change in position is uniformly random and uncorrelated from step to step.
-
-*Simulated life history—*In each case, the model simulated life history processes as follows: Resources and individuals flow into a system from the upstream edge of the environment.
-Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among taxa in ecological communities (Hubbell 2001, White et al.
-2012).
-Along with a species ID and species-specific maximum rate of uptake, each propagule is given an individual ID (computer clock-time at which the individual entered), and a cell quota to represent the physiological state of internal resources.
-
-*Simulated life history-*The community was sampled at random *N* times with replacement for each simulated time unit.
-This theoretically gives every individual the opportunity to grow and reproduce.
-Sampled individuals consume resources according to species specific rates of uptake, if there is a resource present at the individual’s location.
-Uptake increases the individual cell quota and decreases ambient resources within the immediate location of the individual.
-The individual’s quota is then decreased according to a species specific physiological maintenance cost.
-Individuals then reproduce with a probability equal to their cell quota and results in an equal  division of the cell quota between the individual and its daughter.
-The daughter is given a unique individual ID which is appended to the species ID of its mother, to maintain heritage.
-Sampled individuals die if their cell quota is less than or equal to 0.
-Individuals and resource particles flow out when they pass the downstream edge of the environment.
-
-*Simulating environmental complexity—*We constructed the model to capture environmental heterogeneity and the aggregation of resources and individuals.
-The model was parameterized with an arbitrary environmental gradient that varied regularly across a randomly chosen spatial axis in a randomly chosen direction.
-Barriers were constructed within the environment to prevent unhindered flow and to generate aggregation and realistic hydrological flow.
-
-*Simulating species differences*—To allow species differences in physiological traits and environmental optima, we allowed species-specific values for maximum rate of uptake rate, environmental optima, and maintenance costs to be normally distributed random variables where both the means and the standard deviations of the sampling distributions were chosen at random at the start of the simulations.
-Whenever an individual of a new species entered the local community, values for maximum uptake rate, environmental optima, physiological maintenance cost, and minimum cell quota were chosen at random according to the respective normal distributions for those variables.
-
-*Simulating over residence time ($\tau$)—*We ran the model to quasi-equilibrium for each value of $\tau$ in a series of values spanning more than two orders of magnitude, starting with $\tau$ = 0.001 and then doubling until $\tau$ = 0.512.
-The model was run with the same combination of resource, community ecological, and environmental (thermal, physicochemical, physical complexity) values for each value of $\tau$ in the series.
-After running over all $\tau$ values, the model was re-parameterized at random with a new combination of resource parameters (size, number, and types of resource particles), habitat complexity parameters (size, number, and types of physical barriers), chemical environmental parameters, thermal parameters (average temperature, direction and severity of temperature gradient), and average community-level values (average specific values for uptake rate, maintenance, individual size, motility, attachment).
-See table 1.
-
-Because the model is individual-based, information intensive, physiologically realistic, ecologically complex, fluid dynamic, it cannot be run over arbitrarily slow or fast residence times.
-During each simulation, more than a gigabyte of information is held in memory.
-This is unlike analytically tractable and computationally simple Markov models or closed-form analytical models, which are more useful for obtaining precise quantitative predictions according to a single theoretical framework via traditional mathematical tools, but are less useful for modeling highly complex multivariate problems and for examining how individual-level interactions lead to aggregate behaviors.
-In our case, we have integrated components from metabolic theory and ecological scaling, neutral and niche theory, metacommunity theory, fluid dynamics, chemostat theory, and life history theory into a self-evaluating model that tracks the life history, location, and heritage of millions of individuals.
-
-## Results
-
-**Simplistic theory supported in complex systems**
-The probabilistic and complex modeling approach produced all primary expected results, signifying that predictions of otherwise simplistic and idealized chemostat theory could arise in a complex system.
-
-*Autocorrelated residence times—*Residence times of individuals, resource particles, inert tracers, and most any non-permanent component of the system were highly correlated (Fig 1).
-
-*Autocorrelated residence times and right-skewed residence time distributions—*Likewise, residence time distributions were right-skewed and well approximated by lognormal and power-law models (e.g.
-Pareto, Zipf, Zipf-Mandelbrot) (Fig 2).
-
-*Abundance-residence time relationship—*We also observed the expected hump-shaped relationship of *N* to residence time, revealing that residence time could be low enough to exclude reproduction and high enough to result in a rate of resource supply that could not support optimal growth.
-
-**Predictions of abundance and species diversity**
-
-*Richness, evenness, rarity, and dominance*—As expected and as shown in many previous studies, *S* shared a power-law-like relationship with *N*, i.e., *S* = *cN*$^{z}$, where z ~ 0.5, revealing that the model generated a realistic dynamic whether including or excluding immigration and speciation and whether reproduction was sexual or clonal, and whether movement was fluid, random, or environmentally correlated.
-As predicted, the relationship of *S* to $\tau$ followed a hump-shaped relationship.
-Also as predicted, greater *N* lead to greater rarity, greater dominance, and less evenness, regardless of the model specifics.
-
-*Species turnover*—We observed a monotonically decreasing relationship $\beta$-diversity to $\tau$, when calculated as species turnover (presence/absence).
-
-**Trait-related predictions**
-
-*Growth rate vs.
-maintenance—*Increasing residence time lead to a strong power-law decrease in maintenance, revealing strong selective pressure to evolve low maintenance costs in resource limited environments of low turnover.
-
-## Conclusion/Discussion
-
-We used a stochastic individual-based model that simulated a wide array of environmental and ecological conditions and species-level and individual-level traits.
-As much of this variation as possible was assigned at random, to explore a large swath of conditions and parameter space.
-Rather than examine the independent influence of each of these factors, our purpose was to find whether previously known and novel relationships of residence time to total abundance and community-level characteristics should hold in ecologically complex systems and characteristically different environments.
-
-...
 
@@
 
-h.tau <- log((height * width)/flow.rate)
-m.tau <- log(sim.data$individual.tau)
-t.tau <- log(sim.data$particle.tau)
+h.tau <- (height * width)/flow.rate
+m.tau <- sim.data$individual.tau
+t.tau <- sim.data$particle.tau
+r.tau <- sim.data$resource.tau
 
@@
 
-## Univariate relationships
+## We expected mean residence times for individuals, inert tracers, and resource particles to reflect ecosystem residence time.
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+par(mfrow=c(2, 2))
+
+plot(h.tau, m.tau, xlab="Ecosystem residence time", log="xy",
+     ylab="Individual residence Time", col='darkmagenta')
+
+plot(h.tau, t.tau, xlab="Ecosystem residence time", log="xy",
+     ylab="Particle residence Time", col='grey')
+
+plot(h.tau, r.tau, xlab="Ecosystem residence time", log="xy",
+     ylab="Resource residence Time", col='darkcyan')
+```
+
+## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.
 
@@
 
-  h.tau <- log((height * width)/flow.rate)
+  h.tau <- (height * width)/flow.rate
 
   N <- sim.sub$N
-  N <- log(N)
   S <- sim.sub$S
-  S <- log(S)
 
@@
   if (i == 1){
-    plot(x, y, xlab="log(ecosystem residence time)", 
-    ylab="Total Abundance", ylim=c(3.5, 10), xlim=c(3, 9),
+    plot(x, y, xlab="Ecosystem residence time", log="xy",
+    ylab="Total Abundance", xlim = c(20, 6000), ylim = c(10, 20000),
     main = "N vs. Tau", col=colors[i])
@@
 
-  h.tau <- log((height * width)/flow.rate)
+  h.tau <- (height * width)/flow.rate
   S <- sim.sub$S
-  S <- log(S)
 
@@
   if (i == 1){
-    plot(x, y, xlab="log(ecosystem residence time)", 
-    ylab="log(species richness)", ylim=c(0, 4), xlim=c(3, 9),
+    plot(x, y, xlab="Ecosystem residence time", log="xy",
+    ylab="Species richness", xlim = c(20, 6000), ylim = c(2, 50),
     main = "S vs. Tau", col=colors[i])
@@
 
-h.tau <- log((height * width)/flow.rate)
-m.tau <- log(sim.data$individual.tau)
-t.tau <- log(sim.data$particle.tau)
+h.tau <- (height * width)/flow.rate
+m.tau <- sim.data$individual.tau
+t.tau <- sim.data$particle.tau
 
@@
 
+## We expected lower maintenance requirements at higher residence times, where individuals could grow quickly, scrambling to consume available resources, or have an overall slow life history.
+
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
@@
 
-plot(x, y, xlab="log(ecosystem residence time)", 
+plot(x, y, xlab="Ecosystem residence time", log="x",
      ylab="growth rate", 
@@
 maint <- (S*sim.data$avg.per.capita.maint)/N
-plot(h.tau, log(maint), xlab="log(ecosystem residence time)", 
-     ylab="log(maintenance)", 
+plot(h.tau, maint, xlab="Ecosystem residence time", log="xy",
+     ylab="Maintenance", 
      main = "Per capita maintenance vs. Tau",
@@
 
+## We hypothesized that aspects of diversity would be related to total abundance (N), reflecting the feasible set and the influence of residence time.
+
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
@@
 
-plot(log(N), Skew, xlab="Total abundance", 
+plot(N, Skew, xlab="Total abundance", log="xy",
      ylab="Skewness",
@@
 
-plot(log(N), log(Nmax), xlab="Total abundance", 
+plot(N, Nmax, xlab="Total abundance", log="xy",
      ylab="Nmax", 
@@
 
-plot(log(N), log(e.var), xlab="Total abundance", 
+plot(N, e.var, xlab="Total abundance", log="xy",
      ylab="Species Evenness", 
@@
 
-plot(log(N), log(S), xlab="Total abundance",
+plot(N, S, xlab="Total abundance", log="xy",
      ylab="Species Richness", 
@@
 
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-par(mfrow=c(1, 2))
-
-xlabel <- "log(ecosystem residence time)"
-res.conc <- sim.data$resource.concentration
-res.div <- sim.data$shannons.resource.diversity
-res.rich <- sim.data$resource.richness
-
-x <- c()
-y <- c()
-ct <- 1
-for (val in res.conc){
-  if(is.nan(val) == FALSE && val > 0.1){
-    y <- c(y, val)
-    x <- c(x, h.tau[ct])
-    }
-  ct <- ct+1
-  }
-
-plot(x, log(y), xlab=xlabel, 
-     ylab="density", 
-     main = "Resource density vs. Tau",
-     col='peru')
-
-x <- c()
-y <- c()
-ct <- 1
-for (val in res.rich){
-  if(is.nan(val) == FALSE && val > 0.1){
-    y <- c(y, val)
-    x <- c(x, h.tau[ct])
-    }
-  ct <- ct+1
-  }
-
-plot(x, y, xlab=xlabel, 
-     ylab="Number of resource types", 
-     main = "Resource richness vs. Tau",
-     col='slateblue')
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Loading simulated RADs and species lists
-
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv",
-            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
-            as.is = TRUE, skipNul = TRUE, fill = TRUE)
-rad.list[is.na(rad.list)] <- 0
-
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv",
-            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
-            as.is = TRUE, skipNul = TRUE, fill = TRUE)
-species.list[is.na(species.list)] <- -1
 ```
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+# We hypothesized that, as often observed for tracers in real systems, residence time distributions (RTDs) for individuals and resource particles would be lognormal-like curves, which appear normal or left-skewed on a log x-axis.
+Colors correspond across plots.
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
 par(mfrow=c(2, 2))
@@
 colors <- rainbow(length(ind.RTDs))
-plot(density(log(tracer.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2),  main = "Tracer RTD")
+plot(density(tracer.RTDs[[1]]), col = 'white', xlab='Residence Time',
+     xlim = c(0,8), ylim = c(0,2),
+     ylab='Probability Density',  main = "Tracer RTD")
 
@@
 
-plot(density(log(ind.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2), main = "Individual RTD")
+plot(density(ind.RTDs[[1]]), col = 'white', xlab='Residence time',
+  xlim = c(0,8), ylim = c(0,2),
+  ylab='Probability Density', main = "Individual RTD")
 
@@
 
-plot(density(log(ind.RTDs[[1]])), col = 'white', xlab='Residence Time', ylab='Probability Density', xlim = c(0, 8), ylim = c(0, 2), main = "Resource RTD")
+#plot(density(res.RTDs[[1]]), col = 'white', xlab='Residence time',
+#     xlim = c(0,8), ylim = c(0,2),
+#     ylab='Probability Density', main = "Resource RTD")
 
-ct <- 1
-for(rtd in res.RTDs){
-  if(any(is.na(rtd)) == FALSE){
-    lines(density(log(rtd), bw = 0.2), col = colors[ct])
-    r.skews <- c(r.skews, skewness(rtd))
-    r.means <- c(r.means, mean(rtd))
-    }
-  ct = ct + 1
-}
+#ct <- 1
+#for(rtd in res.RTDs){
+#  if(any(is.na(rtd)) == FALSE){
+#    lines(density(log(rtd), bw = 0.2), col = colors[ct])
+#    r.skews <- c(r.skews, skewness(rtd))
+#    r.means <- c(r.means, mean(rtd))
+#    }
+#  ct = ct + 1
+#}
+```
+
+# Increasing ecosystem residence time shifts the mean of the RTD to higher average values (older individuals).
+But, it also changes the skewness of the RTD from asymmetrically young to asymmetrically old.
+No significant effect is seen among tracers.
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
+par(mfrow=c(1, 1))
+
+plot(log(t.means), t.skews, col = 'grey', xlab='Mean of the RTD, i.e., avg. residence time', ylab='Skewness of the RTD', lwd=2, xlim = c(0,7), ylim = c(-1,2),
+     main = "Increasing mean residence time flips the age
+     structure of the community from young to old")
+
+points(log(i.means), i.skews, lwd=2, col = 'darkmagenta')
+#points(r.means, r.skews, lwd=2, col = 'darkcyan')
+lines(c(0,9), c(0,0), lty=2, col='black')
+
+i.OLS <- lm(i.skews ~ log(i.means))
+coeff <- summary(i.OLS, correlation = TRUE)
+i.slope <- round(coefficients(i.OLS)[2], 3)
+i.p <- round(summary(i.OLS)$coefficients[8], 3)
+abline(i.OLS, col = "darkmagenta", lw = 2)
+
+t.OLS <- lm(t.skews ~ log(t.means))
+coeff <- summary(t.OLS, correlation = TRUE)
+t.slope <- round(coefficients(t.OLS)[2], 2)
+t.p <- round(summary(t.OLS)$coefficients[8], 2)
+abline(t.OLS, col = "grey", lw = 2)
+
+legend("topright", legend = c(paste("Tracers: slope=",t.slope,", p=",t.p), paste("Individuals: slope=",i.slope,", p=",i.p)), pch=19, bty = "n", col = c("grey", "darkmagenta"))
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-par(mfrow=c(1, 1))
+## Loading simulated RADs and species lists
 
-plot(log(t.means), t.skews, col = 'darkmagenta', xlab='mean', ylab='skewness', xlim = c(1, 8), ylim = c(-1.2, 3), lwd=2, 
-     main = "Most RTD's are right-skewed,\n(many short-lived, few long-lived)")
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
+            as.is = TRUE, skipNul = TRUE, fill = TRUE)
+rad.list[is.na(rad.list)] <- 0
 
-points(log(i.means),i.skews, lwd=2, col = 'darkcyan')
-points(log(r.means), r.skews, lwd=2, col = 'black')
-
-legend("topright", legend = c("tracers", "individuals", "resources"), pch=20, bty = "o", col = c("darkmagenta", "darkcyan", "black"))
+species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)), as.is = TRUE, skipNul = TRUE, fill = TRUE)
+species.list[is.na(species.list)] <- -1
 ```

### 2015-06-27 analyses/OldRfiles/2015_06_17_ResTime_GeneralResults.Rmd

--- 2015-06-10 analyses/ResTime_GeneralResults.Rmd
@@
 ```{r, results = 'hide', message = FALSE, warning = FALSE}
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June7/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June16/SimData.csv")  
 ```
 
-## Univariate relationships
-
+## Assign data
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
@@
 flow.rate <- sim.data$FlowRate
+seed <- sim.data$starting.seed
 
 h.tau <- log((height * width)/flow.rate)
-m.tau <- log(sim.data$cell.tau)
+m.tau <- log(sim.data$individual.tau)
 t.tau <- log(sim.data$particle.tau)
@@
 SimpD <- sim.data$SD
+
+res.inflow <- sim.data$res.inflow
+res.types <- sim.data$res.types
+max.res.val <- sim.data$max.res.val
+
+res.conc <- sim.data$resource.concentration
+num.tracers <- sim.data$tracer.particles
+num.resources <- sim.data$resource.particles
 ```
+
+## Univariate relationships
 
@@
 
-plot(t.tau, N, xlab="Particle Residence Time", 
+plot(h.tau, log(N), xlab="Ecosystem Residence Time", 
+     ylab="log(Total Abundance)", 
+     main = "N vs. Tau", col='cornsilk4')
+
+plot(h.tau, N, xlab="Ecosystem Residence Time", 
      ylab="Total Abundance", 
@@
 
-plot(m.tau, N, xlab="Cell Residence Time", 
-     ylab="Total Abundance", 
-     main = "N vs. Tau", col='cornsilk4')
-
-plot(t.tau, S, xlab="Particle Residence Time", 
-     ylab="Species Richness", 
+plot(h.tau, log(S), xlab="Ecosystem Residence Time", 
+     ylab="log(Species Richness)", 
      main = "S vs. Tau", col='darkseagreen')
 
-plot(m.tau, S, xlab="Cell Residence Time", 
+plot(h.tau, S, xlab="Ecosystem Residence Time", 
      ylab="Species Richness", 
@@
 
-plot(log(N), e.var, xlab="Total abundance", 
+plot(log(N), log(e.var), xlab="Total abundance", 
      ylab="Species Evenness", 
@@
     y <- c(y, val)
-    x <- c(x, m.tau[ct])
+    x <- c(x, h.tau[ct])
     }
@@
 
-plot(x, y, xlab="Cell Residence Time", 
-     ylab="growth", 
+plot(x, y, xlab="Ecosystem Residence Time", 
+     ylab="growth rate", 
      main = "Per capita growth vs. Tau",
@@
 maint <- (S*sim.data$avg.per.capita.maint)/N
-plot(m.tau, SimpD, xlab="Cell Residence Time", 
+plot(h.tau, maint, xlab="Ecosystem Residence Time", 
      ylab="maintenance", 
@@
      col='darkolivegreen')
-
-immigration <- sim.data$immigration.rate
-plot(m.tau, immigration, xlab="Cell Residence Time", 
-     ylab="immigration", 
-     main = "Immigration vs. Tau",
-     col='mediumorchid')
 ```

### 2015-08-06 models/MainModel/Generator/test/MovieMaker.py
modified LBM code, modified .Rmd analysis file, generated new figs -KL

@@ -1,246 +1,0 @@
-from __future__ import division
-import matplotlib.animation as animation
-import matplotlib.pyplot as plt
-#import mpl_toolkits.mplot3d
-#from mpl_toolkits.mplot3d import Axes3D
-
-from random import choice
-from scipy import stats
-import numpy as np
-import sys
-import os
-import psutil
-
-mydir = os.path.expanduser("~/")
-sys.path.append(mydir + "tools/metrics")
-import metrics
-sys.path.append(mydir + "GitHub/hydrobide/tools/LBM")
-import LBM
-sys.path.append(mydir + "GitHub/hydrobide/tools/bide/test")
-import bide
-
-
-
-def get_rand_params():
-    """ Get random model parameter values. Others are chosen in bide.pyx """
-
-    motion = choice(['fluid', 'random_walk'])
-    motion = 'fluid'
-    D = choice([2, 2]) # number of spatial dimensions
-
-    #width, height = [choice([5,6,7,8,9,10]), choice([5,6,7,8,9,10])]
-    #length = choice([5, 10, 15, 20])
-    width, height, length = [10, 10, 10]
-
-    alpha = np.random.uniform(0.99, 0.999)
-    reproduction = choice(['fission', 'sexual'])
-    speciation = choice(['yes', 'no'])
-    predators = choice([0, 1, 2, 4, 8])
-    parasites = choice([0, 1, 2, 4, 8])
-    env_gradient = choice(['no', 'yes'])
-
-    seedcom = 100 # size of starting community
-    m = 0 # individuals immigrating per time step
-    r = 100 # resource particles flowing in per time step
-    nr = 1 # maximum number of resources types
-    rmax = 500 # maximum resource particle size
-
-
-    reproduction = 'fission'
-    speciation = 'yes'
-
-    return [width, height, length, alpha, motion, D, reproduction, speciation, predators, parasites, env_gradient, seedcom, m, r, nr, rmax]
-
-
-
-def testlengths(TypeOf, function, Lists):
-    vals = []
-    for List in Lists: vals.append(len(List))
-    if min(vals) != max(vals):
[…]

### 2015-08-06 tools/Rbin/multiplot.R

@@ -1,0 +1,123 @@
+################################################################################
+#                                                                              #
+# Functions for calculating metrics of diversity, evenness, rarity, etc.       #
+# These are not included in other diversity packages, e.g., Vegan              #
+#                                                                              #
+################################################################################
+#                                                                              #
+# Written by: Ken Locey                                                        #
+# Obtained from:                                                               #
+# http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/      #
+#                                                                              #
+################################################################################
+#                                                                              #
+# Recent Changes:                                                              #
+#                                                                              #
+# Future Changes (To-Do List):                                                 #
+#         1.                                                                   #
+#         2. Add warnings                                                      #
+#                                                                              #
+################################################################################
+
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  
+  numPlots = length(plots)
+  
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                     ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+  
+  if (numPlots==1) {
+    print(plots[[1]])
+    
+  } else {
+    # Set up the page
[…]

### 2015-08-06 analyses/2015_06_27_ResTime.Rmd

@@
 require("moments")
+library(ggplot2)
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-par(mfrow=c(2, 2))
 
-plot(h.tau, m.tau, xlab="Ecosystem residence time", log="xy",
-     ylab="Individual residence Time", col='darkmagenta')
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
 
-plot(h.tau, t.tau, xlab="Ecosystem residence time", log="xy",
-     ylab="Particle residence Time", col='grey')
+png(filename = '~/GitHub/hydrobide/results/figs/ComparingResTimes.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
 
-plot(h.tau, r.tau, xlab="Ecosystem residence time", log="xy",
-     ylab="Resource residence Time", col='darkcyan')
+#par(mfrow=c(2, 2))
+p1 <- ggplot(data=NULL, aes(x=h.tau, y=t.tau)) +
+  geom_point(color="cyan") +
+  labs(x= expression(tau), y="Residence Time") +
+  labs(title="Tracers") +
+  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 2000)) +
+  coord_fixed(ratio = 1.4) +
+  theme_black()
+
+p2 <- ggplot(data=NULL, aes(x=h.tau, y=m.tau)) +
+  geom_point(color="magenta") +
+  labs(x= expression(tau), y=" ") +
+  labs(title="Individuals") +
+  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 2000)) +
+  coord_fixed(ratio = 1.4) +
+  theme_black()
+
+p3 <- ggplot(data=NULL, aes(x=h.tau, y=r.tau)) +
+  geom_point(color="chartreuse") +
+  labs(x= expression(tau), y=" ") +
+  labs(title="Resources") +
+  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 200)) +
+  coord_fixed(ratio = 2.38) +
+  theme_black()
+
+multiplot(p1, p2, p3, cols=3)
+
+dev.off()
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+
+png(filename = '~/GitHub/hydrobide/results/figs/NandS_vs_ResTime.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+
 par(mfrow=c(1, 2))
-
 num.sims <- max(sim.data$sim)
@@
 }
+
+dev.off()
+
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+
+png(filename = '~/GitHub/hydrobide/results/figs/Growth_and_Maint.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+
 par(mfrow=c(1, 2))
@@
      col='darkolivegreen')
+
+dev.off()
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+png(filename = '~/GitHub/hydrobide/results/figs/Diversity_vs_N.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+
 par(mfrow=c(2, 2))
@@
 
+dev.off()
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+png(filename = '~/GitHub/hydrobide/results/figs/RTDs.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+
 par(mfrow=c(2, 2))
@@
 #}
+
+dev.off()
 ```
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
+
+png(filename = '~/GitHub/hydrobide/results/figs/RTD_shifts.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+
 par(mfrow=c(1, 1))
@@
 legend("topright", legend = c(paste("Tracers: slope=",t.slope,", p=",t.p), paste("Individuals: slope=",i.slope,", p=",i.p)), pch=19, bty = "n", col = c("grey", "darkmagenta"))
+
+dev.off()
 ```

### 2015-08-07 analyses/2015_06_27_ResTime.Rmd
modifying .Rmd code to produce beautified ggplots for ESA talk -KL

@@
 
-#par(mfrow=c(2, 2))
 p1 <- ggplot(data=NULL, aes(x=h.tau, y=t.tau)) +
+  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
   geom_point(color="cyan") +
@@
 p2 <- ggplot(data=NULL, aes(x=h.tau, y=m.tau)) +
+  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
   geom_point(color="magenta") +
@@
 p3 <- ggplot(data=NULL, aes(x=h.tau, y=r.tau)) +
+  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
   geom_point(color="chartreuse") +
@@
 multiplot(p1, p2, p3, cols=3)
-
 dev.off()
@@
 
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
 png(filename = '~/GitHub/hydrobide/results/figs/NandS_vs_ResTime.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
 
-par(mfrow=c(1, 2))
 num.sims <- max(sim.data$sim)
@@
 
-i <- 0
-while (i < num.sims){
+p1 <- ggplot(sim.data, aes(x=particle.tau, y=N, group=sim, colour=sim)) +
+  scale_color_gradient(high="cyan") +
+  geom_line() +
+  theme_black() +
+  labs(x= expression(tau), y="Individuals") +
+  labs(title="Total abundance") +
+  guides(colour=FALSE) + 
+  scale_x_log10(limits = c(10, 1000)) + 
+  scale_y_log10(limits = c(100, 10000)) +
+  coord_fixed(ratio = 1)
 
-  sim.sub <- subset(sim.data, sim == i)
-  i <- i + 1
-  print(i)
+p2 <- ggplot(sim.data, aes(x=particle.tau, y=S, group=sim, colour=sim)) +
+  scale_color_gradient(high="chartreuse") +
+  geom_line() +
+  theme_black() +
+  labs(x= expression(tau), y="Species") +
+  labs(title="Numbers of species") +
+  guides(colour=FALSE) +
+  scale_x_log10(limits = c(10, 1000)) +
+  coord_fixed(ratio = .0345)
 
-  height <- sim.sub$Height
-  width <- sim.sub$Width
-  flow.rate <- sim.sub$FlowRate
-  e.var <- sim.sub$e.var
-  e.simp <- sim.sub$simpson.e
-  avg.ab <- N/S
-  Skew <-sim.sub$skew
-  Nmax <- sim.sub$N.max
-  SimpD <- sim.sub$SD
-
-  h.tau <- (height * width)/flow.rate
-
-  N <- sim.sub$N
-  S <- sim.sub$S
-
-  x <- c()
-  y <- c()
-  ct <- 1
-  for (val in N){
-    if(is.nan(val) == FALSE && val > 0.1){
-      y <- c(y, val)
-      x <- c(x, h.tau[ct])
-      }
-    ct <- ct+1
-    }
-
-  if (i == 1){
-    plot(x, y, xlab="Ecosystem residence time", log="xy",
-    ylab="Total Abundance", xlim = c(20, 6000), ylim = c(10, 20000),
-    main = "N vs. Tau", col=colors[i])
-    lines(x, y, col=colors[i])
-    }
-  else{
-    points(x, y, col=colors[i])
-    lines(x, y, col=colors[i])
-    }
-
-  }
-
-i <- 0
-while (i < num.sims){
-
-  sim.sub <- subset(sim.data, sim == i)
-  i <- i + 1
-
-  height <- sim.sub$Height
-  width <- sim.sub$Width
-  flow.rate <- sim.sub$FlowRate
-  e.var <- sim.sub$e.var
-  e.simp <- sim.sub$simpson.e
-  avg.ab <- N/S
-  Skew <-sim.sub$skew
-  Nmax <- sim.sub$N.max
-  SimpD <- sim.sub$SD
-
-  h.tau <- (height * width)/flow.rate
-  S <- sim.sub$S
-
-  x <- c()
-  y <- c()
-  ct <- 1
-  for (val in S){
-    if(is.nan(val) == FALSE && val > 0.1){
-      y <- c(y, val)
-      x <- c(x, h.tau[ct])
-      }
-    ct <- ct+1
-    }
-
-  if (i == 1){
-    plot(x, y, xlab="Ecosystem residence time", log="xy",
-    ylab="Species richness", xlim = c(20, 6000), ylim = c(2, 50),
-    main = "S vs. Tau", col=colors[i])
-    lines(x, y, col=colors[i])
-    }
-  else{
-    points(x, y, col=colors[i])
-    lines(x, y, col=colors[i])
-    }
-}
-
+multiplot(p1, p2, cols=2)
 dev.off()
-
 ```
@@
 png(filename = '~/GitHub/hydrobide/results/figs/Growth_and_Maint.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
-
-par(mfrow=c(1, 2))
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
 
 growth <- sim.data$avg.per.capita.growth
-x <- c()
-y <- c()
+xs <- c()
+ys <- c()
 
@@
   if(is.nan(val) == FALSE && val > 0.1){
-    y <- c(y, val)
-    x <- c(x, h.tau[ct])
+    ys <- c(ys, val)
+    xs <- c(xs, h.tau[ct])
     }
@@
 
-plot(x, y, xlab="Ecosystem residence time", log="x",
-     ylab="growth rate", 
-     main = "Per capita growth vs. Tau",
-     col='dodgerblue')
+p1 <- ggplot(data=NULL, aes(x=xs, y=ys)) +
+  geom_point(colour="magenta") +
+  labs(x= expression(tau), y="Growth") +
+  labs(title="Growth rate") +
+  scale_x_log10(limits = c(20, 1000)) +
+  coord_fixed(ratio = 4) +
+  theme_black()
 
 maint <- (S*sim.data$avg.per.capita.maint)/N
-plot(h.tau, maint, xlab="Ecosystem residence time", log="xy",
-     ylab="Maintenance", 
-     main = "Per capita maintenance vs. Tau",
-     col='darkolivegreen')
 
+p2 <- ggplot(data=NULL, aes(x=h.tau, y=maint)) +
+  geom_point(colour="cyan") +
+  labs(x= expression(tau), y="Maintenance") +
+  labs(title="Maintenance") +
+  scale_x_log10(limits = c(20, 1000)) +
+  scale_y_log10() +
+  coord_fixed(ratio = .5) +
+  theme_black()
+
+multiplot(p1, p2, cols=2)
 dev.off()

### 2015-08-07 tools/Rbin/multiplot.R

@@ -121,3 +121,12 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
 }
 # Check that it is a complete theme
 attr(theme_black(), "complete")
+
+
+
+addlinetoplot <- function(varx, vary) { 
+  list(
+    geom_line(data=NULL, aes(x=varx, y=vary)), 
+    geom_point(data=NULL, aes(x=varx, y=vary))
+  )
+}

### 2015-08-10 analyses/2015_06_27_ResTime.Rmd

@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/SimData.csv")  
 sim.data <- subset(sim.data, motion == ' fluid ')
@@
   labs(title="Numbers of species") +
+  guides(colour=FALSE) +
+  scale_x_log10(limits = c(10, 1000)) +
+  coord_fixed(ratio = .0345)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
+
+## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
+png(filename = '~/GitHub/hydrobide/results/figs/EandB_vs_ResTime.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+
+sp.lists <- c()
+
+p1 <- ggplot(sim.data, aes(x=particle.tau, y=e.var, group=sim, colour=sim)) +
+  scale_color_gradient(high="cyan") +
+  geom_line() +
+  theme_black() +
+  labs(x= expression(tau), y="Evar") +
+  labs(title="Evenness") +
+  guides(colour=FALSE) + 
+  scale_x_log10(limits = c(10, 1000)) + 
+  scale_y_log10(limits = c(100, 10000)) +
+  coord_fixed(ratio = 1)
+
+p2 <- ggplot(sim.data, aes(x=particle.tau, y=S, group=sim, colour=sim)) +
+  scale_color_gradient(high="chartreuse") +
+  geom_line() +
+  theme_black() +
+  labs(x= expression(tau), y="Whittaker's") +
+  labs(title="Species turnover") +
   guides(colour=FALSE) +
@@
 
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
+
+rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv", 
+            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
             as.is = TRUE, skipNul = TRUE, fill = TRUE)
@@
 
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)), as.is = TRUE, skipNul = TRUE, fill = TRUE)
-species.list[is.na(species.list)] <- -1
+species.list <- c()
+r <- 0
+while(r < 61){
+  print(r)
+  myData <- list(read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv", 
+            header = FALSE, sep = ",",
+
+            skip=r, nrows=1))
+  species.list <- c(species.list, myData)
+  r <- r + 1
+  }
+
+sim.data$species.list <- species.list
+num.sims <- max(sim.data$sim) + 1
+
+i <- 0
+while (i < num.sims){
+
+  sp.sub <- subset(sim.data$species.list, sim == i)
+  sp.beta <- c()
+  j <- 2
+  while(j <= length(sp.sub)){
+    print(j)
+    w.beta <- sp.turnover(sp.sub[j], sp.sub[j-1])
+    if(w.beta != -1){
+      sp.beta <- c(sp.beta, w.beta)
+      }
+    j <- j+1
+    }
+  i <- i+1
+  print(sp.beta)
+  }
 ```

### 2015-08-26 models/BoneYard/DataGenerator/DataGenerator.pyx
cleaned up bide.py code; converted cython scripts back to python, modified multiplot.R, generated new simulated data -KL

@@ -1,222 +1,0 @@
-from __future__ import division
-from random import randint, choice
-import numpy as np
-import sys
-import os
-from scipy import stats
-import time
-import psutil
-
-mydir = os.path.expanduser("~/")
-sys.path.append(mydir + "/GitHub/hydrobide/tools/LBM")
-import LBM
-sys.path.append(mydir + "/GitHub/hydrobide/tools/bide")
-import bide
-sys.path.append(mydir + "tools/metrics")
-import metrics
-
-
-def get_mrrmax():
-    """ Get random model parameter values. Others are chosen in bide.pyx """
-
-    m = 1 #int(choice(range(1, 2))) # individuals immigrating per time step
-    r = int(choice(range(10, 100))) # resource particles flowing in per time step
-    nr = int(choice(range(1, 10))) # maximum number of resources types
-    rmax = int(choice(range(10, 1000))) # maximum value of resource particle size
-
-    return [m, r, nr, rmax]
-
-
-#######################  MICROBE COMMUNITY PARAMETERS  #########################
-m, r, nr, rmax = get_mrrmax()
-
-MicID, ResID, N, S = 0, 0, 0, 0
-avgTau = str()
-
-COM, MicXcoords, MicYcoords, AvgTaus, RAD, splist = [], [], [], [], [], []
-MicIDs, MicQs, MicExitAge, MicTimeIn = [], [], [], []
-ResYcoords, RES, ResIDs, ResType, ResXcoords = [], [], [], [], []
-TracerIDs, TracerXcoords, TracerYcoords, TracerExitAge = [], [], [], []
-
-microbe_color_dict, GrowthDict, MaintDict = {}, {}, {}
-ResUseDict, ResColorDict, DispParamsDict = {}, {}, {}
-
-
-###############  SIMULATION VARIABLES, DIMENSIONAL & MODEL CONSTANTS  ##########
-width = int(choice([5,6,7,8,9,10]))
-height = int(choice([5, 5]))
-shift, sign, sim, ct1, BarrierWidth, BarrierHeight = 0.0, 0.1, 0, 0, 0.2, 0.2
-
-Rates = np.array([1.0, 0.75, 0.5, 0.25, 0.1, 0.075, 0.05, 0.025])  # inflow speeds
-u0 = Rates[0]  # initial in-flow speed
-
-
-#####################  Lattice Boltzmann PARAMETERS  ###########################
-viscosity =  0.84   # fluid viscosity of water
[…]

### 2015-08-26 models/BoneYard/FigScripts/Fig1.py

@@ -1,73 +1,0 @@
-from __future__ import division
-import  matplotlib.pyplot as plt
-import datetime
-import numpy as np
-import os.path
-import Hydrobide
-import sys
-
-mydir = '/Users/lisalocey/Desktop/repos/HYDRO-BIDE/'
-sys.path.append(mydir + '/tools/figurefunctions')
-import figs
-
-fig = plt.figure()
-
-r = 2 # number rows
-c = 2 # number columns
-wsp = 0.39 # width space
-hsp = 0.39 # height space
-
-use_picloud = 'n'
-basic = True   # if true, will not return info on richness, evenness, etc.
-
-Tau = 'random' # can also be 'list'
-replicates = 1
-
-systems = [1]
-time = 5000 # number of generations. Needs to be large enough to allow
-            # the population to reach a relatively stable state
-
-
-i=1
-if 1 in systems:
-    # IDEAL CHEMOSTAT
-    system = 'ideal'
-    Glist, Nlist, Plist, Rlist, MCTlist, Qlist, TauList, mean, tauAtQmaxList = Hydrobide.run_hydrobide(use_picloud, system, Tau, basic, replicates, time)
-    
-    fig = figs.MCTvsTau(fig, r, c, i, wsp, hsp, MCTlist, TauList, tauAtQmaxList)
-    i+=1
-    #fig = figs.MUvsD(fig, r, c, i, wsp, hsp, Plist, TauList, tauAtQmaxList)
-    fig = figs.BTvsTau(fig, r, c, i, wsp, hsp, Plist, TauList, tauAtQmaxList)
-
-
-if 2 in systems:
-    # IMMIGRATION AND SPECIES DIFFERENCES
-    system = 'imNonNeutral'
-    Glist, Nlist, Plist, Rlist, MCTlist, Qlist, TauList, mean, tauAtQmaxList = Hydrobide.run_hydrobide(use_picloud, system, Tau, basic, replicates, time)
-    
-    i+=1
-    fig = figs.MCTvsTau(fig, r, c, i, wsp, hsp, MCTlist, TauList, tauAtQmaxList)
-    i+=1
-    #fig = figs.MUvsD(fig, r, c, i, wsp, hsp, Plist, TauList, tauAtQmaxList)
-    fig = figs.BTvsTau(fig, r, c, i, wsp, hsp, Plist, TauList, tauAtQmaxList)
-
-
-if 3 in systems:
[…]

### 2015-08-26 models/BoneYard/Taxa/setup.py

@@ -1,4 +1,0 @@
-from distutils.core import setup
-from Cython.Build import cythonize
-
-setup(ext_modules = cythonize("FourPanel.pyx"))

### 2015-08-26 models/MainModel/model.pyx

@@ -31,29 +31,30 @@ def get_rand_params():
     height = choice([5, 6, 7, 8, 9, 10])
     length = choice([5, 6, 7, 8, 9, 10])
 
-    alpha = np.random.uniform(0.9, 0.999)
+    alpha = np.random.uniform(0.99, 0.999)
     #reproduction = choice(['fission', 'sexual'])
-    #speciation = choice(['yes', 'no'])
+    speciation = choice(['yes', 'no'])
     predators = choice([0, 1, 2, 4, 8])
     parasites = choice([0, 1, 2, 4, 8])
     env_gradient = choice(['no', 'yes'])
 
-    seedcom = choice([100, 500, 1000]) # size of starting community
-    m = choice([0, 1, 2, 4]) # individuals immigrating per time step
-    r = choice([100, 150, 200, 250, 300]) # resource particles flowing in per time step
+    seedcom = 1000 # size of starting community
+    m = choice([1]) # individuals immigrating per time step
+    r = choice([50, 100, 150, 200]) # resource particles flowing in per time step
     nr = choice([1, 2, 4, 6, 8, 10]) # maximum number of resources types
-    rmax = choice([100, 150, 200, 250, 300]) # maximum resource particle size
+    rmax = choice([200, 400, 600, 800, 1000]) # maximum resource particle size
 
-    #width, height = 10, 10
-    #seedcom = 1000
+    gmax = choice([0.1, 0.2, 0.3, 0.4])
+    maintmax = choice([0.01, 0.02, 0.04, 0.05])
+    width, height = 10, 10
     #motion = 'fluid'
     reproduction = 'fission'
-    speciation = 'yes'
+    #speciation = 'no'
     #rmax = 200
     #r = 200
-    m = 0
 
-    return [width, height, length, alpha, motion, D, reproduction, speciation, predators, parasites, env_gradient, seedcom, m, r, nr, rmax]
+
+    return [width, height, length, alpha, motion, D, reproduction, speciation, predators, parasites, env_gradient, seedcom, m, r, nr, rmax, gmax, maintmax]
 
 
 
@@ -74,7 +75,7 @@ def nextFrame(arg):	# arg is the frame number
[…]

### 2015-08-26 models/MainModel/test/model.py

@@ -16,7 +16,7 @@ sys.path.append(mydir + "tools/metrics")
 import metrics
 sys.path.append(mydir + "GitHub/hydrobide/tools/LBM")
 import LBM
-sys.path.append(mydir + "GitHub/hydrobide/tools/bide/test")
+sys.path.append(mydir + "GitHub/hydrobide/tools/bide")
 import bide
 
 
@@ -25,32 +25,36 @@ def get_rand_params():
     """ Get random model parameter values. Others are chosen in bide.pyx """
 
     motion = choice(['fluid', 'random_walk'])
-    motion = 'fluid'
     D = choice([2, 2]) # number of spatial dimensions
 
-    #width, height = [choice([5,6,7,8,9,10]), choice([5,6,7,8,9,10])]
-    #length = choice([5, 10, 15, 20])
-    width, height, length = [10, 10, 10]
+    width = choice([5, 6, 7, 8, 9, 10])
+    height = choice([5, 6, 7, 8, 9, 10])
+    length = choice([5, 6, 7, 8, 9, 10])
 
     alpha = np.random.uniform(0.99, 0.999)
-    reproduction = choice(['fission', 'sexual'])
+    #reproduction = choice(['fission', 'sexual'])
     speciation = choice(['yes', 'no'])
     predators = choice([0, 1, 2, 4, 8])
     parasites = choice([0, 1, 2, 4, 8])
     env_gradient = choice(['no', 'yes'])
 
-    seedcom = choice([10, 20, 40, 80, 160]) # size of starting community
-    m = choice([0, 1, 2, 4]) # individuals immigrating per time step
-    r = choice([100, 200, 300, 400]) # resource particles flowing in per time step
-    nr = choice([1, 2, 4, 8, 16]) # maximum number of resources types
-    rmax = choice([500, 1000, 2000, 4000, 8000]) # maximum resource particle size
+    seedcom = 1000 # size of starting community
+    m = choice([1]) # individuals immigrating per time step
+    r = choice([50, 100, 150, 200]) # resource particles flowing in per time step
+    nr = choice([1, 2, 4, 6, 8, 10]) # maximum number of resources types
+    rmax = choice([200, 400, 600, 800, 1000]) # maximum resource particle size
 
[…]

### 2015-08-26 tools/Rbin/multiplot.R

@@ -68,22 +68,22 @@ multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
 
 theme_black <- function(base_size = 12, base_family = "Helvetica") {
   theme(
-    line =               element_line(colour = "black", size = 0.5, linetype = 1,
+    line =               element_line(colour = "black", size = 1.5, linetype = 1,
                                       lineend = "butt"),
-    rect =               element_rect(fill = "white", colour = "black", size = 0.5, linetype = 1),
+    rect =               element_rect(fill = "white", colour = "black", size = 1.5, linetype = 1),
     text =               element_text(family = base_family, face = "plain",
                                       colour = "black", size = base_size,
-                                      hjust = 0.5, vjust = 0.5, angle = 0, lineheight = 0.9),
+                                      hjust = 0.5, vjust = 1.2, angle = 0, lineheight = 0.9),
     axis.text =          element_text(size = rel(0.8), colour = "white"),
     strip.text =         element_text(size = rel(0.8), colour = "white"),
     
     axis.line =          element_blank(),
     axis.text.x =        element_text(vjust = 1),
     axis.text.y =        element_text(hjust = 1),
-    axis.ticks =         element_line(colour = "white", size = 0.2),
+    axis.ticks =         element_line(colour = "white", size = 0.8),
     axis.title =         element_text(colour = "white"),
-    axis.title.x =       element_text(vjust = 0, size = rel(2.1)),
-    axis.title.y =       element_text(angle = 90),
+    axis.title.x =       element_text(vjust = -0.5, size = rel(2.8)),
+    axis.title.y =       element_text(angle = 90, size = rel(1.5)),
     axis.ticks.length =  unit(0.3, "lines"),
     axis.ticks.margin =  unit(0.5, "lines"),
     

### 2015-08-26 tools/TaxaTimeRelationship/TTR.py

@@ -1,27 +1,0 @@
-# -*- coding: utf-8 -*-
-
-""" The future site for a function that generates the taxa-time relationship
-
-
-The first step in constructing a species–time relationship is to estimate
-the number of species found in a given time span of observation. 
-
-We used a moving window approach where species richness was determined 
-for every possible window of each time span (e.g. in a 20 year time series,
- 20 one-year windows, 19 two-year windows, etc.). 
- 
-These values were then averaged within each time span (we used arithmetic means,
- but using geometric means did not affect the results).
- 
- Next, we fit the two statistical models (power and logarithmic functions) to
- these averages using two methods: ordinary least squares regression (OLS) on
- log-transformed data and non-linear regression on untransformed data. 
- 
- For the non-linear fitting we used the parameter estimates from the OLS as 
- the initial parameter values. OLS and non-linear fitting produced nearly 
- identical parameter estimates (Pearson's r >= 0.99 for both functions). 
- We compared fits of the non-linear power and logarithmic regressions using r2
- values (because there are two parameters in both models there is no need for
- more complex model comparison techniques).
-"""
-

### 2015-08-26 tools/bide/bide.py

@@ -1,0 +1,667 @@
+from __future__ import division
+from random import randint, choice
+import numpy as np
+import sys
+from math import modf
+import decimal
+import time
+
+limit = 0.2
+
+def coord(d):
+    return float(np.random.uniform(0.1*d, 0.9*d))
+
+
+def GetRAD(vector):
+    RAD = []
+    unique = list(set(vector))
+    for val in unique:
+        RAD.append(vector.count(val)) # the abundance of each Sp_
+
+    return RAD, unique # the rad and the Sp_ list
+
+
+def get_color(ID, colorD): # FUNCTION TO ASSIGN COLORS TO Sp_
+
+    r1 = lambda: randint(0,255)
+    r2 = lambda: randint(0,255)
+    r3 = lambda: randint(0,255)
+
+    color = '#%02X%02X%02X' % (r1(),r2(),r3())
+    colorD[ID] = color
+
+    return colorD
+
+
+
+def NewTracers(motion, IDs, Xs, Ys, Zs, t_In, w, h, l, u0, D):
+
+    x = np.random.binomial(1, u0)
+    if x == 1:
+        IDs.append(0)
+        t_In.append(0)
+        if motion == 'random_walk':
+            Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+            Xs.append(float(np.random.uniform(0.1*w, 0.9*w)))
+            Zs.append(float(np.random.uniform(0.1*l, 0.9*l)))
+
+        else:
+            Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+            Xs.append(float(np.random.uniform(0.1*w, 0.9*w)))
+            Zs.append(float(np.random.uniform(0.1*l, 0.9*l)))
+
+    return [IDs, t_In, Xs, Ys, Zs]
+
+
+
+def ResIn(motion, Type, Vals, Xs, Ys, Zs, ID, IDs, t_In, numr,
+        rmax, nr, w, h, l, u0, D):
+
+    for r in range(numr):
+        x = np.random.binomial(1, u0/2)
+
+        if x == 1:
+            rval = int(np.random.random_integers(1, rmax, 1))
+            rtype = int(np.random.random_integers(0, nr-1, 1))
+
+            Vals.append(rval)
+            IDs.append(ID)
+            Type.append(rtype)
+            t_In.append(0)
+            ID += 1
+
[…]

### 2015-08-26 tools/bide/bide.pyx

@@ -1,610 +1,0 @@
-from __future__ import division
-from random import randint, choice
-import numpy as np
-import sys
-from math import modf
-import decimal
-import time
-
-limit = 0.2
-
-def coord(d):
-    return float(np.random.uniform(0.1*d, 0.9*d))
-
-
-def GetRAD(vector):
-    RAD = []
-    unique = list(set(vector))
-    for val in unique:
-        RAD.append(vector.count(val)) # the abundance of each species
-
-    return RAD, unique # the rad and the species list
-
-
-def get_color(ID, color_dict): # FUNCTION TO ASSIGN COLORS TO SPECIES
-
-    r1 = lambda: randint(0,255)
-    r2 = lambda: randint(0,255)
-    r3 = lambda: randint(0,255)
-
-    color = '#%02X%02X%02X' % (r1(),r2(),r3())
-    color_dict[ID] = color
-
-    return color_dict
-
-
-
-def NewTracers(motion, IDs, Xcoords, Ycoords, Zcoords, TimeIn, width, height, length, u0, D):
-
-    x = np.random.binomial(1, u0)
-    if x == 1:
-        IDs.append(0)
-        TimeIn.append(0)
-        if motion == 'random_walk':
-            Ycoords.append(float(np.random.uniform(0.1*height, 0.9*height)))
-            Xcoords.append(float(np.random.uniform(0.1*width, 0.9*width)))
-            Zcoords.append(float(np.random.uniform(0.1*length, 0.9*length)))
-
-        else:
-            Ycoords.append(float(np.random.uniform(0.1*height, 0.9*height)))
-            Xcoords.append(float(np.random.uniform(0.1*width, 0.9*width)))
-            Zcoords.append(float(np.random.uniform(0.1*length, 0.9*length)))
-
-    return [IDs, TimeIn, Xcoords, Ycoords, Zcoords]
-
-
-
-def ResIn(motion, Type, Vals, Xcoords, Ycoords, Zcoords, ID, IDs, TimeIn, numr, rmax, nr, width, height, length, u0, D):
-
-    for r in range(numr):
-        x = np.random.binomial(1, u0/2)
-
-        if x == 1:
-            rval = int(np.random.random_integers(1, rmax, 1))
-            rtype = int(np.random.random_integers(0, nr-1, 1))
-
-            Vals.append(rval)
-            IDs.append(ID)
-            Type.append(rtype)
-            TimeIn.append(0)
[…]

### 2015-08-26 analyses/2015_06_27_ResTime.Rmd

--- 2015-08-07 analyses/2015_06_27_ResTime.Rmd
@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_June/18_June_3/SimData.csv")  
-sim.data <- subset(sim.data, motion == ' fluid ')
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/12_Aug/SimData.csv")  
+#sim.data <- subset(sim.data, motion == ' random_walk ')
 ```
@@
 SimpD <- sim.data$SD
+Whittaker.turnover <- sim.data$species.turnover
 
@@
 
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+## Loading files with plotting and diversity functions
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
+```
+
 ## We expected mean residence times for individuals, inert tracers, and resource particles to reflect ecosystem residence time.
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-
-source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
-
-png(filename = '~/GitHub/hydrobide/results/figs/ComparingResTimes.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
-
-p1 <- ggplot(data=NULL, aes(x=h.tau, y=t.tau)) +
-  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
-  geom_point(color="cyan") +
-  labs(x= expression(tau), y="Residence Time") +
-  labs(title="Tracers") +
-  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 2000)) +
-  coord_fixed(ratio = 1.4) +
-  theme_black()
-
-p2 <- ggplot(data=NULL, aes(x=h.tau, y=m.tau)) +
-  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
-  geom_point(color="magenta") +
-  labs(x= expression(tau), y=" ") +
-  labs(title="Individuals") +
-  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 2000)) +
-  coord_fixed(ratio = 1.4) +
-  theme_black()
-
-p3 <- ggplot(data=NULL, aes(x=h.tau, y=r.tau)) +
-  geom_abline(intercept = 0, slope=1, colour="white", lty=2) +
-  geom_point(color="chartreuse") +
-  labs(x= expression(tau), y=" ") +
-  labs(title="Resources") +
-  scale_x_log10(limits = c(10, 10000)) + scale_y_log10(limits = c(10, 200)) +
-  coord_fixed(ratio = 2.38) +
-  theme_black()
-
-multiplot(p1, p2, p3, cols=3)
-dev.off()
-```
 
@@
 
-p1 <- ggplot(sim.data, aes(x=particle.tau, y=N, group=sim, colour=sim)) +
+sim.data$N[sim.data$N<=0] <- 1
+sim.data$N[sim.data$S<=0] <- 1
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=N, group=sim, colour=sim)) +
   scale_color_gradient(high="cyan") +
-  geom_line() +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
   theme_black() +
@@
   guides(colour=FALSE) + 
-  scale_x_log10(limits = c(10, 1000)) + 
-  scale_y_log10(limits = c(100, 10000)) +
-  coord_fixed(ratio = 1)
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = .48)
 
-p2 <- ggplot(sim.data, aes(x=particle.tau, y=S, group=sim, colour=sim)) +
+p2 <- ggplot(sim.data, aes(x=h.tau, y=S, group=sim, colour=sim)) +
   scale_color_gradient(high="chartreuse") +
-  geom_line() +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
   theme_black() +
-  labs(x= expression(tau), y="Species") +
-  labs(title="Numbers of species") +
+  labs(x= expression(tau), y="Number of species") +
+  labs(title="Species richness") +
   guides(colour=FALSE) +
-  scale_x_log10(limits = c(10, 1000)) +
-  coord_fixed(ratio = .0345)
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 0.8)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
+
+## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+
+png(filename = '~/GitHub/hydrobide/results/figs/EandB_vs_ResTime.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=e.var, group=sim, colour=sim)) +
+  scale_color_gradient(high="cyan") +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Evar") +
+  labs(title="Evenness") +
+  guides(colour=FALSE) + 
+  scale_x_log10() + 
+  coord_fixed(ratio = 1.8)
+
+p2 <- ggplot(sim.data, aes(x=h.tau, y=Whittaker.turnover, group=sim, colour=sim)) +
+  scale_color_gradient(high="chartreuse") +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Whittaker's") +
+  labs(title="Species turnover") +
+  guides(colour=FALSE) +
+  scale_x_log10() +
+  coord_fixed(ratio = 2.0)
 
@@
   labs(title="Growth rate") +
-  scale_x_log10(limits = c(20, 1000)) +
-  coord_fixed(ratio = 4) +
+  scale_x_log10() +
+  coord_fixed(ratio = 5.3) +
   theme_black()
@@
   labs(title="Maintenance") +
-  scale_x_log10(limits = c(20, 1000)) +
+  scale_x_log10() +
   scale_y_log10() +
-  coord_fixed(ratio = .5) +
+  coord_fixed(ratio = 0.34) +
   theme_black()
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Loading simulated RADs and species lists
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
+num.sims <- max(sim.data$sim) + 1
 
-rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/RADs.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
-            as.is = TRUE, skipNul = TRUE, fill = TRUE)
-rad.list[is.na(rad.list)] <- 0
+i <- 0
+while (i < num.sims){
 
-species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/Species.csv", header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)), as.is = TRUE, skipNul = TRUE, fill = TRUE)
-species.list[is.na(species.list)] <- -1
+  sp.sub <- subset(sim.data$species.list, sim == i)
+  sp.beta <- c()
+  j <- 2
+  while(j <= length(sp.sub)){
+    print(j)
+    w.beta <- sp.turnover(sp.sub[j], sp.sub[j-1])
+    sp.beta <- c(sp.beta, w.beta)
+    }
+    j <- j+1
+    }
+  i <- i+1
+  print(sp.beta)
+  }
 ```

### 2015-08-26 tools/bide/bide.py
modified bide.py, working on resource use efficiency code -KL

@@ -1,9 +1,9 @@
 from __future__ import division
 from random import randint, choice
 import numpy as np
-import sys
-from math import modf
-import decimal
+#import sys
+#from math import modf
+#import decimal
 import time
 
 limit = 0.2
@@ -18,7 +18,7 @@ def GetRAD(vector):
     for val in unique:
         RAD.append(vector.count(val)) # the abundance of each Sp_
 
-    return RAD, unique # the rad and the Sp_ list
+    return RAD, unique # the rad and the specieslist
 
 
 def get_color(ID, colorD): # FUNCTION TO ASSIGN COLORS TO Sp_
@@ -114,19 +114,19 @@ def immigration(motion, numin, Sp, Xs, Ys, Zs, w, h, l, MD,
             Qs.append(Q)
 
             if prop not in colorD:
-                # Sp_ color
+                # speciescolor
                 colorD = get_color(prop, colorD)
 
-                # Sp_ growth rate
+                # speciesgrowth rate
                 GD[prop] = np.random.uniform(0.01, 0.1)
 
-                # Sp_ maintenance
+                # speciesmaintenance
                 MD[prop] = np.random.uniform(0.001, 0.01)
 
-                # Sp_ active dispersal rate
+                # speciesactive dispersal rate
                 DispD[prop] = np.random.uniform(0.01, 0.1)
 
-                # Sp_ resource use efficiency
+                # speciesresource use efficiency
                 RD[prop] = np.random.uniform(0.01, 1.0, nr)
 
     return [Sp, Xs, Ys, Zs, MD, GD, DispD, colorD, IDs, ID, t_In, Qs, RD]
@@ -384,10 +384,12 @@ def consume(R_Types, R_Vals, R_IDs, R_ID, R_Xs, R_Ys, R_Zs, R_t_In,
 
             # The Individual
             ID = I_IDs.index(ind)
+            # The individual's cell quota
             Q = Qs[ID]
-            Sp_ = Sp_IDs[ID]
-            mu = GD[Sp_]
-            efficiency = RD[Sp_][R_type]
+            # the species
+            sp = Sp_IDs[ID]
+            mu = GD[sp]
+            efficiency = RD[sp][R_type]
             mu = mu * efficiency
 
             if R_val > (mu * Q): # Increase cell quota
@@ -453,22 +455,22 @@ def reproduce(repro, spec, Sp_IDs, Qs, IDs, ID, t_In, Xs, Ys, Zs, w,
                         t = time.clock()
[…]

### 2015-08-27 analyses/2015_06_27_ResTime.Rmd
modified model, ran over night, added results figs -KL

@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/12_Aug/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/SimData.csv")  
 #sim.data <- subset(sim.data, motion == ' random_walk ')
@@
 sim <- sim.data$sim # the specific simulation run
-height <- sim.data$Height
-width <- sim.data$Width
-flow.rate <- sim.data$FlowRate
+height <- sim.data$height
+width <- sim.data$width
+flow.rate <- sim.data$flow.rate
 seed <- sim.data$starting.seed
@@
 
-N <- sim.data$N
-S <- sim.data$S
+N <- sim.data$total.abundance
+S <- sim.data$species.richness
 e.var <- sim.data$e.var
@@
 res.inflow <- sim.data$res.inflow
-res.types <- sim.data$res.types
+N.res.types <- sim.data$N.res.types
+P.res.types <- sim.data$P.res.types
+C.res.types <- sim.data$C.res.types
 max.res.val <- sim.data$max.res.val
@@
 num.resources <- sim.data$resource.particles
+
+biomass.prod.N <- sim.data$biomass.prod.N
+biomass.prod.P <- sim.data$biomass.prod.P
+biomass.prod.C <- sim.data$biomass.prod.C
 ```
@@
 
-sim.data$N[sim.data$N<=0] <- 1
-sim.data$N[sim.data$S<=0] <- 1
+sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
+N[N<=0] <- 1
+sim.data$species.richness[sim.data$species.richness<=0] <- 1
+S[S<=0] <- 1
 
@@
   scale_x_log10() +
-  coord_fixed(ratio = 5.3) +
+  coord_fixed(ratio = 100) +
   theme_black()
 
-maint <- (S*sim.data$avg.per.capita.maint)/N
+#maint <- (S*sim.data$avg.per.capita.maint)/N
+maint <- sim.data$avg.per.capita.maint
 
@@
   scale_y_log10() +
-  coord_fixed(ratio = 0.34) +
+  coord_fixed(ratio = 8) +
   theme_black()

### 2015-08-27 tools/bide/bide.py

@@ -1,7 +1,7 @@
 from __future__ import division
 from random import randint, choice
 import numpy as np
-#import sys
+import sys
 #from math import modf
 #import decimal
 import time
@@ -10,6 +10,27 @@ limit = 0.2
 
 def coord(d):
     return float(np.random.uniform(0.1*d, 0.9*d))
+
+
+def GetIndParam(means):
+    vals = []
+
+    if isinstance(means, float) or isinstance(means, int):
+        std = means/10.0
+        vals = np.random.normal(means, std)
+        if vals < 0.00001:
+            vals = 0.00001
+
+    else:
+        for val in means:
+            std = val/10.0
+            i = np.random.normal(val, std)
+            if i < 0.00001:
+                i = 0.00001
+            vals.append(i)
+
+    return vals
+
 
 
 def GetRAD(vector):
@@ -55,14 +76,27 @@ def NewTracers(motion, IDs, Xs, Ys, Zs, t_In, w, h, l, u0, D):
 
 
 def ResIn(motion, Type, Vals, Xs, Ys, Zs, ID, IDs, t_In, numr,
-        rmax, nr, w, h, l, u0, D):
+        rmax, nN, nP, nC, w, h, l, u0, D):
 
     for r in range(numr):
         x = np.random.binomial(1, u0/2)
 
         if x == 1:
             rval = int(np.random.random_integers(1, rmax, 1))
-            rtype = int(np.random.random_integers(0, nr-1, 1))
+            nr = choice(['N', 'P', 'C'])
+
+            if nr == 'N':
+                rtype = int(np.random.random_integers(0, nN-1, 1))
+                rtype = 'N'+str(rtype)
+
+            if nr == 'P':
+                rtype = int(np.random.random_integers(0, nP-1, 1))
+                rtype = 'P'+str(rtype)
+
+            if nr == 'C':
+                rtype = int(np.random.random_integers(0, nC-1, 1))
+                rtype = 'C'+str(rtype)
+
 
             Vals.append(rval)
             IDs.append(ID)
@@ -85,9 +119,8 @@ def ResIn(motion, Type, Vals, Xs, Ys, Zs, ID, IDs, t_In, numr,
 
 
 
[…]

### 2015-08-28 analyses/2015_06_27_ResTime.Rmd

@@
 dev.off()
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Reassign Data
-
-sim <- sim.data$sim # the specific simulation run
-height <- sim.data$Height
-width <- sim.data$Width
-flow.rate <- sim.data$FlowRate
-seed <- sim.data$starting.seed
-
-h.tau <- (height * width)/flow.rate
-m.tau <- sim.data$individual.tau
-t.tau <- sim.data$particle.tau
-
-N <- sim.data$N
-S <- sim.data$S
-e.var <- sim.data$e.var
-e.simp <- sim.data$simpson.e
-avg.ab <- N/S
-Skew <-sim.data$skew
-Nmax <- sim.data$N.max
-SimpD <- sim.data$SD
-
-res.inflow <- sim.data$res.inflow
-res.types <- sim.data$res.types
-max.res.val <- sim.data$max.res.val
-
-res.conc <- sim.data$resource.concentration
-num.tracers <- sim.data$tracer.particles
-num.resources <- sim.data$resource.particles
 ```
@@
 ```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
-num.sims <- max(sim.data$sim) + 1
-
-i <- 0
-while (i < num.sims){
-
-  sp.sub <- subset(sim.data$species.list, sim == i)
-  sp.beta <- c()
-  j <- 2
-  while(j <= length(sp.sub)){
-    print(j)
-    w.beta <- sp.turnover(sp.sub[j], sp.sub[j-1])
-    sp.beta <- c(sp.beta, w.beta)
-    }
-    j <- j+1
-    }
-  i <- i+1
-  print(sp.beta)
-  }
-```

### 2015-08-29 analyses/2015_06_27_ResTime.Rmd

@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-
-source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
 
@@
 
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/IndRTD.csv"
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/IndRTD.csv"
 con  <- file(inputFile, open = "r")
@@
 
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/TracerRTD.csv"
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/TracerRTD.csv"
 con  <- file(inputFile, open = "r")
@@
 
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/ResRTD.csv"
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/ResRTD.csv"
 con  <- file(inputFile, open = "r")
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
+plot.new()
 

### 2015-08-29 analyses/2015_08_29_0720_ResTime.Rmd

--- 2015-08-29 analyses/2015_06_27_ResTime.Rmd
@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/SimData.csv")  
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
 #sim.data <- subset(sim.data, motion == ' random_walk ')
@@
 
-png(filename = '~/GitHub/hydrobide/results/figs/NandS_vs_ResTime.png',
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime.png',
       width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
@@
 
-png(filename = '~/GitHub/hydrobide/results/figs/EandB_vs_ResTime.png',
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/EandB_vs_ResTime.png',
       width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
@@
 
-png(filename = '~/GitHub/hydrobide/results/figs/Growth_and_Maint.png',
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Growth_and_Maint.png',
       width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-png(filename = '~/GitHub/hydrobide/results/figs/Diversity_vs_N.png',
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Diversity_vs_N.png',
       width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Loading simulated data csv files
-
-sims <- c()
-cts <- c()
-tracer.RTDs <- list()
-res.RTDs <- list()
-ind.RTDs <- list()
-
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/IndRTD.csv"
-con  <- file(inputFile, open = "r")
-ct = 1
-while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
-  line <- as.numeric(myLine)
-  cts <- c(cts, line[1])
-  sims <- c(sims, line[2])
-  rtd <- line[-1:-2]
-  ind.RTDs[[ct]] <- rtd
-  ct <- ct + 1
-  } 
-close(con)
-
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/TracerRTD.csv"
-con  <- file(inputFile, open = "r")
-ct = 1
-while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
-  line <- as.numeric(myLine)
-  rtd <- line[-1:-2]
-  tracer.RTDs[[ct]] <- rtd
-  ct <- ct + 1
-  } 
-close(con)
-
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/ResRTD.csv"
-con  <- file(inputFile, open = "r")
-ct = 1
-while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
-  line <- as.numeric(myLine)
-  rtd <- line[-1:-2]
-  res.RTDs[[ct]] <- rtd
-  ct <- ct + 1
-  } 
-close(con)
-```
-
 # We hypothesized that, as often observed for tracers in real systems, residence time distributions (RTDs) for individuals and resource particles would be lognormal-like curves, which appear normal or left-skewed on a log x-axis.
 Colors correspond across plots.
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-png(filename = '~/GitHub/hydrobide/results/figs/RTDs.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
-
-par(mfrow=c(2, 2))
-
-t.skews <- c()
-i.skews <- c()
-r.skews <- c()
-t.means <- c()
-i.means <- c()
-r.means <- c()
-
-colors <- rainbow(length(ind.RTDs))
-plot(density(tracer.RTDs[[1]]), col = 'white', xlab='Residence Time',
-     xlim = c(0,8), ylim = c(0,2),
-     ylab='Probability Density',  main = "Tracer RTD")
-
-ct <- 1
-for(rtd in tracer.RTDs){
-  if(any(is.na(rtd)) == FALSE){
-    lines(density(log(rtd), bw = 0.2), col = colors[ct])
-    t.skews <- c(t.skews, skewness(rtd))
-    t.means <- c(t.means, mean(rtd))
-    }
-  ct = ct + 1
-}
-
-plot(density(ind.RTDs[[1]]), col = 'white', xlab='Residence time',
-  xlim = c(0,8), ylim = c(0,2),
-  ylab='Probability Density', main = "Individual RTD")
-
-ct <- 1
-for(rtd in ind.RTDs){
-  if(any(is.na(rtd)) == FALSE){
-    lines(density(log(rtd), bw = 0.2), col = colors[ct])
-    i.skews <- c(i.skews, skewness(rtd))
-    i.means <- c(i.means, mean(rtd))
-    }
-  ct = ct + 1
-}
-
-#plot(density(res.RTDs[[1]]), col = 'white', xlab='Residence time',
-#     xlim = c(0,8), ylim = c(0,2),
-#     ylab='Probability Density', main = "Resource RTD")
-
-#ct <- 1
-#for(rtd in res.RTDs){
-#  if(any(is.na(rtd)) == FALSE){
-#    lines(density(log(rtd), bw = 0.2), col = colors[ct])
-#    r.skews <- c(r.skews, skewness(rtd))
-#    r.means <- c(r.means, mean(rtd))
-#    }
-#  ct = ct + 1
-#}
-
-dev.off()
-```
-
-# Increasing ecosystem residence time shifts the mean of the RTD to higher average values (older individuals).
-But, it also changes the skewness of the RTD from asymmetrically young to asymmetrically old.
-No significant effect is seen among tracers.
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
-plot.new()
-
-png(filename = '~/GitHub/hydrobide/results/figs/RTD_shifts.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
-
-par(mfrow=c(1, 1))
-
-plot(log(t.means), t.skews, col = 'grey', xlab='Mean of the RTD, i.e., avg. residence time', ylab='Skewness of the RTD', lwd=2, xlim = c(0,7), ylim = c(-1,2),
-     main = "Increasing mean residence time flips the age
-     structure of the community from young to old")
-
-points(log(i.means), i.skews, lwd=2, col = 'darkmagenta')
-#points(r.means, r.skews, lwd=2, col = 'darkcyan')
-lines(c(0,9), c(0,0), lty=2, col='black')
-
-i.OLS <- lm(i.skews ~ log(i.means))
-coeff <- summary(i.OLS, correlation = TRUE)
-i.slope <- round(coefficients(i.OLS)[2], 3)
-i.p <- round(summary(i.OLS)$coefficients[8], 3)
-abline(i.OLS, col = "darkmagenta", lw = 2)
-
-t.OLS <- lm(t.skews ~ log(t.means))
-coeff <- summary(t.OLS, correlation = TRUE)
-t.slope <- round(coefficients(t.OLS)[2], 2)
-t.p <- round(summary(t.OLS)$coefficients[8], 2)
-abline(t.OLS, col = "grey", lw = 2)
-
-legend("topright", legend = c(paste("Tracers: slope=",t.slope,", p=",t.p), paste("Individuals: slope=",i.slope,", p=",i.p)), pch=19, bty = "n", col = c("grey", "darkmagenta"))
-
-dev.off()
-```

### 2015-08-31 analyses/2015_08_29_0720_ResTime.Rmd
added new figures, analyses, and simulated data -KL

@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
-#sim.data <- subset(sim.data, motion == ' random_walk ')
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-## Assign Data
-sim <- sim.data$sim # the specific simulation run
-height <- sim.data$height
-width <- sim.data$width
-flow.rate <- sim.data$flow.rate
-seed <- sim.data$starting.seed
-
-h.tau <- (height * width)/flow.rate
-m.tau <- sim.data$individual.tau
-t.tau <- sim.data$particle.tau
-r.tau <- sim.data$resource.tau
-
-N <- sim.data$total.abundance
-S <- sim.data$species.richness
-e.var <- sim.data$e.var
-e.simp <- sim.data$simpson.e
-avg.ab <- N/S
-Skew <-sim.data$skew
-Nmax <- sim.data$N.max
-SimpD <- sim.data$SD
-Whittaker.turnover <- sim.data$species.turnover
-
-res.inflow <- sim.data$res.inflow
-N.res.types <- sim.data$N.res.types
-P.res.types <- sim.data$P.res.types
-C.res.types <- sim.data$C.res.types
-max.res.val <- sim.data$max.res.val
-
-res.conc <- sim.data$resource.concentration
-num.tracers <- sim.data$tracer.particles
-num.resources <- sim.data$resource.particles
-
-biomass.prod.N <- sim.data$biomass.prod.N
-biomass.prod.P <- sim.data$biomass.prod.P
-biomass.prod.C <- sim.data$biomass.prod.C
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 ## Loading files with plotting and diversity functions
@@
 
-## We expected mean residence times for individuals, inert tracers, and resource particles to reflect ecosystem residence time.
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+## Load Data
+sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
+h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
+sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
+sim.data$species.richness[sim.data$species.richness<=0] <- 1
+
+sim.data.fluid <- subset(sim.data, motion == "fluid")
+h.tau.fluid <- (sim.data.fluid$height * sim.data.fluid$width)/sim.data.fluid$flow.rate
+sim.data.fluid$total.abundance[sim.data.fluid$total.abundance<=0] <- 1
+sim.data.fluid$species.richness[sim.data.fluid$species.richness<=0] <- 1
+
+sim.data.random_walk <- subset(sim.data, motion == "random_walk")
+h.tau.random_walk <- (sim.data.random_walk$height * sim.data.random_walk$width)/sim.data.random_walk$flow.rate
+sim.data.random_walk$total.abundance[sim.data.random_walk$total.abundance<=0] <- 1
+sim.data.random_walk$species.richness[sim.data.random_walk$species.richness<=0] <- 1
+```
 
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime.png',
@@
 
-num.sims <- max(sim.data$sim)
-colors <- rainbow(num.sims)
-
-sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
-N[N<=0] <- 1
-sim.data$species.richness[sim.data$species.richness<=0] <- 1
-S[S<=0] <- 1
-
-p1 <- ggplot(sim.data, aes(x=h.tau, y=N, group=sim, colour=sim)) +
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, colour=sim.data$sim)) +
   scale_color_gradient(high="cyan") +
@@
   scale_y_log10() +
-  coord_fixed(ratio = .48)
+  coord_fixed(ratio = .50)
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=S, group=sim, colour=sim)) +
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$sim)) +
   scale_color_gradient(high="chartreuse") +
@@
   scale_y_log10() +
-  coord_fixed(ratio = 0.8)
+  coord_fixed(ratio = 1.0)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_ByMotion.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, colour=sim.data$motion)) +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Individuals") +
+  labs(title="Total abundance") +
+  guides(colour=FALSE) + 
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = .50)
+
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$motion)) +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Number of species") +
+  labs(title="Species richness") +
+  guides(colour=FALSE) +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.0)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_2ByMaxGrowthRate.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, 
+                           colour=sim.data$max.growth.rate)) +
+  scale_color_gradient(high="cyan") +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Individuals") +
+  labs(title="Total abundance") +
+  guides(colour=FALSE) + 
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = .50)
+
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, 
+                           colour=sim.data$max.growth.rate)) +
+  scale_color_gradient(high="cyan") +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Number of species") +
+  labs(title="Species richness") +
+  guides(colour=FALSE) +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.0)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_ByMaxActiveDisp.png',
+      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
+  group=sim.data$sim, colour=sim.data$max.active.dispersal)) +
+  scale_color_gradient(high="darkseagreen") +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Individuals") +
+  labs(title="Total abundance") +
+  guides(colour=FALSE) + 
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = .50)
+
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$motion)) +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="Number of species") +
+  labs(title="Species richness") +
+  guides(colour=FALSE) +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.0)
 
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/EandB_vs_ResTime.png',
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-
+plot.new()
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Growth_and_Maint.png',
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+plot.new()
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Diversity_vs_N.png',
@@
 ```
-
-# We hypothesized that, as often observed for tracers in real systems, residence time distributions (RTDs) for individuals and resource particles would be lognormal-like curves, which appear normal or left-skewed on a log x-axis.
-Colors correspond across plots.

### 2015-08-31 analyses/OldRfiles/2015_06_27_ResTime.Rmd
added new figures, analyses, and simulated data -KL

--- 2015-08-29 analyses/2015_06_27_ResTime.Rmd
@@
 
-inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/25_Aug/IndRTD.csv"
+inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/IndRTD.csv"
 con  <- file(inputFile, open = "r")

### 2015-09-08 analyses/2015_08_29_0720_ResTime.Rmd
modified main model, generated new png's and simulations data -KL

@@
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+      width = 1500, height = 1200, res=200, bg = "black")
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, colour=sim.data$sim)) +
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
+                           group=sim.data$sim, colour=sim.data$sim)) +
   scale_color_gradient(high="cyan") +
@@
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$sim)) +
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
+                           group=sim.data$sim, colour=sim.data$sim)) +
   scale_color_gradient(high="chartreuse") +
@@
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_ByMotion.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+    width = 1500, height = 1200, res=200, bg = "black")
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, colour=sim.data$motion)) +
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
+                           group=sim.data$sim, colour=sim.data$motion)) +
   geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$motion)) +
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
+                           group=sim.data$sim, colour=sim.data$motion)) +
   geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_2ByMaxGrowthRate.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+colors <- list('springgreen', 'greenyellow', 'yellow', 'cyan', 'magenta', 'deeppink', 'firebrick1',
+               'aquamarine', 'hotpink', 'gold1', 'ghostwhite', 'darkseagreen1', 'red1', 'plum1',
+               'olivedrab1', 'maroon1', 'floralwhite', 'lightsteelblue1', 'darkgoldenrod1')
+
+good.names <- list('res.inflow','N.types','P.types','C.types','max.res.val','max.growth.rate',
+              'max.met.maint','max.active.dispersal','logseries.a', 'flow.rate',
+              'amplitude','frequency','phase','disturbance')
+
+for(i in 1:ncol(sim.data)){
+  name <- colnames(sim.data[i])
+
+  if (name %in% good.names){
+    print(name)
+    color <- sample(colors, 1)
+
+    name <- paste0('~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_',name,'.png')
+    png(filename = name, width = 1500, height = 1200, res=200, bg = "black")
 
     p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, group=sim.data$sim, 
-                           colour=sim.data$max.growth.rate)) +
-  scale_color_gradient(high="cyan") +
+                           colour=sim.data[,i+1])) +
+    scale_color_gradient(high=color) +
     geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
     p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, 
-                           colour=sim.data$max.growth.rate)) +
-  scale_color_gradient(high="cyan") +
+                           colour=sim.data[,i])) +
+    scale_color_gradient(high=color) +
     geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
     dev.off()
+    }
+  }
 ```
 
+## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.
+
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/NandS_vs_ResTime_ByMaxActiveDisp.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/EvennessAndTurnover_vs_ResTime.png',
+      width = 1500, height = 1200, res=200, bg = "black")
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
-  group=sim.data$sim, colour=sim.data$max.active.dispersal)) +
-  scale_color_gradient(high="darkseagreen") +
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
+                           group=sim.data$sim, colour=sim.data$sim)) +
+  scale_color_gradient(high="cyan") +
   geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
   theme_black() +
-  labs(x= expression(tau), y="Individuals") +
-  labs(title="Total abundance") +
+  labs(x= expression(tau), y="evenness") +
+  labs(title="Species evenness") +
   guides(colour=FALSE) + 
   scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = .50)
+  coord_fixed(ratio = 2.5)
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, group=sim.data$sim, colour=sim.data$motion)) +
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
+                           group=sim.data$sim, colour=sim.data$sim)) +
+  scale_color_gradient(high="chartreuse") +
   geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
   theme_black() +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
-  theme_black() +
-  labs(x= expression(tau), y="Number of species") +
-  labs(title="Species richness") +
+  labs(x= expression(tau), y="turnover") +
+  labs(title="Species turnover") +
   guides(colour=FALSE) +
   scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1.0)
+  coord_fixed(ratio = 2.5)
 
@@
 
-## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/EvennessAndTurnover_vs_ResTime_ByMotion.png',
+      width = 1500, height = 1200, res=200, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
+                           group=sim.data$sim, colour=sim.data$motion)) +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="evenness") +
+  labs(title="Species evenness") +
+  guides(colour=FALSE) + 
+  scale_x_log10() +
+  coord_fixed(ratio = 2.5)
+
+p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
+                           group=sim.data$sim, colour=sim.data$motion)) +
+  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+  theme_black() +
+  labs(x= expression(tau), y="turnover") +
+  labs(title="Species turnover") +
+  guides(colour=FALSE) +
+  scale_x_log10() +
+  coord_fixed(ratio = 2.5)
+
+multiplot(p1, p2, cols=2)
+dev.off()
+```
 
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/EandB_vs_ResTime.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+colors <- list('springgreen', 'greenyellow', 'yellow', 'cyan', 'magenta', 'deeppink', 'firebrick1',
+               'aquamarine', 'hotpink', 'gold1', 'ghostwhite', 'darkseagreen1', 'blue1', 'plum1',
+               'olivedrab1', 'maroon1', 'floralwhite', 'lightsteelblue1', 'darkgoldenrod1')
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=e.var, group=sim, colour=sim)) +
-  scale_color_gradient(high="cyan") +
+good.names <- list('res.inflow','N.types','P.types','C.types','max.res.val','max.growth.rate',
+              'max.met.maint','max.active.dispersal','logseries.a', 'flow.rate',
+              'amplitude','frequency','phase','disturbance')
+
+for(i in 1:ncol(sim.data)){
+  name <- colnames(sim.data[i])
+
+  if (name %in% good.names){
+    print(name)
+    color <- sample(colors, 1)
+    name <- paste0('~/GitHub/hydrobide/results/figs/29_Aug/EvennessAndTurnover_vs_ResTime_',name,'.png')
+    png(filename = name, width = 1500, height = 1200, res=200, bg = "black")
+
+    p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
+                               group=sim.data$sim, colour=sim.data[,i])) +
+    scale_color_gradient(high = color) +
     geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
     scale_x_log10() + 
-  coord_fixed(ratio = 1.8)
+    coord_fixed(ratio = 2.5)
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=Whittaker.turnover, group=sim, colour=sim)) +
-  scale_color_gradient(high="chartreuse") +
+    p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
+                               group=sim.data$sim, colour=sim.data[,i])) +
+    scale_color_gradient(high=color) +
     geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
@@
     scale_x_log10() +
-  coord_fixed(ratio = 2.0)
+    coord_fixed(ratio = 2.5)
 
@@
     dev.off()
+    }
+  }
 ```
@@
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Growth_and_Maint.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")
+      width = 1500, height = 1200, res=200, bg = "black")
 
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-plot.new()
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
+lm_eqn <- function(df){
+    m <- lm(log(y) ~ log(x), df)
+    a = format(coef(m)[1], digits = 2) 
+    b = format(coef(m)[2], digits = 2) 
+    r2 = format(summary(m)$r.squared, digits = 3)
+    bquote('S = ' ~ .(a) ~ N^{.(b)})
+    }
+
 png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Diversity_vs_N.png',
-      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")
+      width = 1500, height = 1200, res=200, bg = "black")
 
-par(mfrow=c(2, 2))
+df <- data.frame(x=sim.data$total.abundance, y=sim.data$species.richness)
+p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness)) +
+  labs(title=lm_eqn(df)) +
+  geom_point(colour="snow3") +
+  geom_smooth(method='lm',colour="lawngreen", size=2) +
+  theme_black() +
+  labs(x= "Total abundance", y="Species richness") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.6)
 
-plot(N, Skew, xlab="Total abundance", log="xy",
-     ylab="Skewness",
-     main = "Rarity vs. N",
-     col='plum')
+df <- data.frame(x=sim.data$total.abundance, y=sim.data$e.var)
+p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var)) +
+  labs(title=lm_eqn(df)) +
+  geom_point(colour="snow3") +
+  geom_smooth(method='lm', colour="lawngreen", size=2) +
+  theme_black() +
+  labs(x= "Total abundance", y="Species evenness") +
+  scale_x_log10() +
+  scale_y_log10() +
+  ylim(0.01,1) +
+  coord_fixed(ratio = 4.3)
 
-plot(N, Nmax, xlab="Total abundance", log="xy",
-     ylab="Nmax", 
-     main = "Nmax vs. N",
-     col='darkorange')
-
-plot(N, e.var, xlab="Total abundance", log="xy",
-     ylab="Species Evenness", 
-     main = "Evenness vs. N",
-     col='lightseagreen')
-
-plot(N, S, xlab="Total abundance", log="xy",
-     ylab="Species Richness", 
-     main = "S vs. N",
-     col='indianred')
-
+multiplot(p1, p2, cols=2)
 dev.off()
 ```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
+lm_eqn <- function(df){
+    m <- lm(log(y) ~ log(x), df, na.action=na.exclude)
+    a = format(coef(m)[1], digits = 2) 
+    b = format(coef(m)[2], digits = 2) 
+    r2 = format(summary(m)$r.squared, digits = 3)
+    bquote('S = ' ~ .(a) ~ N^{.(b)} ~.(r2))
+    }
+
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Diversity_vs_N2.png',
+      width = 1500, height = 1200, res=200, bg = "black")
+
+df <- data.frame(x=sim.data$total.abundance, y=sim.data$N.max)
+p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max)) +
+  labs(title=lm_eqn(df)) +
+  geom_point(colour="snow3") +
+  geom_smooth(method='lm',colour="lawngreen", size=2) +
+  theme_black() +
+  labs(x= "Total abundance", y="Nmax") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1)
+
+multiplot(p1, cols=1)
+dev.off()
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
+source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
+
+png(filename = '~/GitHub/hydrobide/results/figs/29_Aug/Diversity_vs_N_ByMotion.png',
+      width = 1500, height = 1200, res=200, bg = "black")
+
+p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness,
+             group=sim.data$sim, colour=sim.data$motion)) +
+  geom_point() +
+  theme_black() +
+  labs(x= "Total abundance", y="Species richness") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.55)
+
+p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var,
+                           group=sim.data$sim, colour=sim.data$motion)) +
+  geom_point() +
+  theme_black() +
+  labs(x= "Total abundance", y="Species evenness") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 3.8)
+
+p3 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max,
+                           group=sim.data$sim, colour=sim.data$motion)) +
+  geom_point() +
+  theme_black() +
+  labs(x= "Total abundance", y="Dominance") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1)
+
+p4 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$skew,
+                           group=sim.data$sim, colour=sim.data$motion)) +
+  geom_point() +
+  theme_black() +
+  labs(x= "Total abundance", y="Rarity") +
+  scale_x_log10() +
+  scale_y_log10() +
+  coord_fixed(ratio = 1.4)
+
+multiplot(p1, p2, p3, p4, cols=2)
+dev.off()
+```

### 2015-09-08 models/MainModel/model.py

@@ -33,15 +33,13 @@ def get_rand_params():
     height = choice([5, 6, 7, 8, 9, 10])
     length = choice([5, 6, 7, 8, 9, 10])
 
-    pulse = choice([0.1, 0.2, 0.4, 0.6, 0.8, 1.0])
+    pulse = choice([0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0])
     flux = choice(['yes', 'no'])
 
     # Sine wave: y(t) = A * sin(2*pi*f*t + phi)
-    # let phi = 0, that is, 0 amplitude at time 0
-    # For Residence time (tau):
-    #     tau = A * sin(2pi * f *t)
-    phase = choice([0, 1, 2, 4, 8, 16]) # 0 = in phase; 16 = entirely out of phase
-    amp = choice([0.1, 0.2, 0.3, 0.4, 0.5]) # A
+    # let phi = 0, meaning 0 amplitude at time 0
+    phase = choice([0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10]) # 0 = in phase; 16 = entirely out of phase
+    amp = choice([0.05, 0.1, 0.2, 0.3, 0.4, 0.5]) # A
     freq = choice([0.1, 0.08, 0.06, 0.04, 0.02, 0.01]) # f
 
     disturb = choice([0.0, 0.01, 0.02, 0.04, 0.06, 0.08, 0.1])
@@ -49,36 +47,25 @@ def get_rand_params():
     alpha = np.random.uniform(0.99, 0.999)
     reproduction = choice(['fission', 'sexual'])
     speciation = choice(['yes', 'no'])
-    predators = choice([0, 1, 2, 4, 8])
-    parasites = choice([0, 1, 2, 4, 8])
-    env_gradient = choice(['no', 'yes'])
 
-    seedcom = 1000 # size of starting community
+    seedcom = choice([100, 200, 300, 400, 500, 600, 700, 800, 900, 1000]) # size of starting community
     m = choice([1]) # individuals immigrating per time step
-    r = choice([450, 500, 550]) # resource particles flowing in per time step
+    r = choice([50, 100, 150, 200, 250, 300, 350, 400, 450, 500]) # resource particles flowing in per time step
 
-    nNi = choice([12, 14, 16, 18, 20]) # maximum number of Nitrogen types
[…]

### 2015-09-08 tools/Rbin/multiplot.R

@@ -82,12 +82,12 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     axis.text.y =        element_text(hjust = 1),
     axis.ticks =         element_line(colour = "white", size = 0.8),
     axis.title =         element_text(colour = "white"),
-    axis.title.x =       element_text(vjust = -0.5, size = rel(2.8)),
+    axis.title.x =       element_text(vjust = -0.5, size = rel(1.5)),
     axis.title.y =       element_text(angle = 90, size = rel(1.5)),
     axis.ticks.length =  unit(0.3, "lines"),
     axis.ticks.margin =  unit(0.5, "lines"),
     
-    legend.background =  element_rect(colour = NA),
+    legend.background =  element_rect(colour = "black"),
     legend.margin =      unit(0.2, "cm"),
     legend.key =         element_rect(fill = "black", colour = "white"),
     legend.key.size =    unit(1.2, "lines"),
@@ -97,10 +97,11 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     legend.text.align =  NULL,
     legend.title =       element_text(size = rel(0.8), face = "bold", hjust = 0, colour = "white"),
     legend.title.align = NULL,
-    legend.position =    "right",
+    legend.position =    "none",
     legend.direction =   "vertical",
     legend.justification = "center",
     legend.box =         NULL,
+    #opts(legend.position = "none"),
     
     panel.background =   element_rect(fill = "black", colour = NA),
     panel.border =       element_rect(fill = NA, colour = "white"),
@@ -113,7 +114,7 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     strip.text.y =       element_text(angle = -90),
     
     plot.background =    element_rect(colour = "black", fill = "black"),
-    plot.title =         element_text(colour = "white", size = rel(2)),
+    plot.title =         element_text(colour = "white", size = rel(1.8)),
     plot.margin =        unit(c(1, 1, 0.5, 0.5), "lines"),
     
     complete = TRUE

### 2015-09-09 analyses/ForPaper/MultivariateTests.Rmd

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
require("picante")
require("ape")
require("fossil")
require("simba")
require("reshape")

#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
sim.data$h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
sim.data$species.richness[sim.data$species.richness<=0] <- 1

sim.data[is.na(sim.data)] <- 0
```

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")
```

## 7.) Variance partitioning: 
```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}

# Physical and metacommunity variables
phys.dat <- as.matrix(subset(sim.data, 
                      select = c(h.tau)))

# Resource variables
res.dat <- as.matrix(subset(sim.data, 
                      select = c(resource.concentration,
                      shannons.resource.diversity,
                      resource.richness,
                      resource.particles,
                      avg.per.capita.growth,
                      avg.per.capita.maint,
                      avg.per.capita.N.efficiency,
                      avg.per.capita.P.efficiency,
                      avg.per.capita.C.efficiency,
                      avg.per.capita.active.dispersal)))

y.dat <- as.matrix(subset(sim.data, 
                      select = c(total.abundance,
                      species.richness,
                      simpson.e,
                      species.turnover)))

rda.phys <- rda(y.dat ~ phys.dat)
#rda.phys
rda.res <- rda(y.dat ~ res.dat)
#rda.res

# Two explanatory matrices -- Hellinger-transform Y
mod <- varpart(Y=y.dat, X=phys.dat, res.dat,  transfo="log")
mod
showvarparts(2)
plot(mod)
mtext("Variation partitioning")
text(1.6, 0.6, "env") 
text(-0.6, 0.6, "tau") 
```

```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}

# Test fraction using RDA:
phys.anova <- anova(rda.phys, step=200, perm.max=200)
phys.anova

res.anova <- anova(rda.res, step=200, perm.max=200)
res.anova

bio.anova <- anova(rda.bio, step=200, perm.max=200)
bio.anova
```

### ii.  Constrained Ordination
Another way we can test hypotheses with continuous data is to use **constrained ordination**, which is sometimes referred to as canonical ordination.
Constrained ordination explores the relationships between two matrices: an **explanatory matrix** and a **response matrix**. 
Canonical correspondence analysis (CCA) and redundancy analysis (RDA) are two types of constrained ordination.
These techniques are based on the linear model framework and thus can be used to formally test hypotheses.
Constrained ordination works by first conducting multivariate multiple linear regression followed either by correspondence analysis (CA) with CCA or Principal Components Analysis (PCA) with RDA, while using the matrix of fitted values to obtain a constrained ordination.
A permutation test can then be used to test for overall significance.

Here, we will use environmental data to conduct a CCA on the fish assemblages of the Doubs River. 
We will start by creating an explanatory matrix that contains water chemistry data.
We will then use the `cca()` function from the `vegan` package.
Note, we have to specify that we want the `cca` function in the `vegan` package because there are `cca` functions in both `vegan` and `ade4`!
We will then use permutation tests to evaluate the significance of our model. 
Finally, we will test the influence of each environmental variable on the constrained axes.

```{r, results = "hide"}
# Get Dominant members

# Define Environmental Matrix
#bio.dat
#phys.dat
#res.dat
all.dat <- cbind(bio.dat, phys.dat, res.dat)
# Conduct CCA 
y.cca <- vegan::cca(y.dat ~ all.dat)

# Permutation Tests
anova(y.cca, by = "axis")
cca.fit <- envfit(y.cca, all.dat, perm = 999)
cca.fit

# Calculate Explained Variation
cca.explainvar1 <- round(y.cca$CCA$eig[1] / 
                         sum(c(y.cca$CCA$eig, y.cca$CA$eig)), 3) * 100

cca.explainvar2 <- round(y.cca$CCA$eig[2] / 
                         sum(c(y.cca$CCA$eig, y.cca$CA$eig)), 3) * 100

# Define Plot Parameters
par(mar = c(5, 5, 4, 4) + 0.1)

# Initiate Plot
plot(scores(y.cca, display = "wa"), xlim = c(-3.5, 2), ylim = c(-3.2, 3.2),
     xlab = paste("CCA 1 (", cca.explainvar1, "%)", sep = ""),
     ylab = paste("CCA 2 (", cca.explainvar2, "%)", sep = ""),
     pch = 16, cex = 2.0, type = "n", cex.lab = 1.5,
     cex.axis = 1.2, axes = FALSE)

# Add Axes
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add Points & Labels
points(scores(y.cca, display = "wa"),
       pch = 19, cex = 3, bg = "gray", col = "gray")
text(scores(y.cca, display = "wa"), 
     labels = row.names(scores(y.cca, display = "wa")))

# Add Environmental Vectors
vectors <- scores(active.cca, display = "bp")
row.names(vectors) <- c("Depth", "Temp", "pH","DOC", "DON", "TDS", "Salinity")
arrows(0, 0, vectors[,1] * 2, vectors[, 2] * 2, 
       lwd = 2, lty = 1, length = 0.2, col = "red")
text(vectors[,1] * 2, vectors[, 2] * 2, pos = 3, 
     labels = row.names(vectors))
axis(side = 3, lwd.ticks=2, cex.axis=1.2, las = 1, col = "red", lwd = 2.2,
     at = pretty(range(vectors[, 1])) * 2, labels = pretty(range(vectors[, 1])))
axis(side = 4, lwd.ticks=2, cex.axis=1.2, las = 1, col = "red", lwd = 2.2,
     at = pretty(range(vectors[, 2])) * 2, labels = pretty(range(vectors[, 2])))
```

### 2015-09-09 analyses/ForPaper/PrimaryPredictions.Rmd

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
sim.data$species.richness[sim.data$species.richness<=0] <- 1

sim.data.fluid <- subset(sim.data, motion == "fluid")
h.tau.fluid <- (sim.data.fluid$height * sim.data.fluid$width)/sim.data.fluid$flow.rate
sim.data.fluid$total.abundance[sim.data.fluid$total.abundance<=0] <- 1
sim.data.fluid$species.richness[sim.data.fluid$species.richness<=0] <- 1

sim.data.random_walk <- subset(sim.data, motion == "random_walk")
h.tau.random_walk <- (sim.data.random_walk$height * sim.data.random_walk$width)/sim.data.random_walk$flow.rate
sim.data.random_walk$total.abundance[sim.data.random_walk$total.abundance<=0] <- 1
sim.data.random_walk$species.richness[sim.data.random_walk$species.richness<=0] <- 1
```

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = .50)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  guides(colour=FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.0)

multiplot(p1, p2, cols=2)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime_ByMotion.png',
    width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = .50)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  guides(colour=FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.0)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/EvennessAndTurnover_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="evenness") +
  labs(title="Species evenness") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  coord_fixed(ratio = 2.5)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="turnover") +
  labs(title="Species turnover") +
  guides(colour=FALSE) +
  scale_x_log10() +
  coord_fixed(ratio = 2.5)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that aspects of diversity would be related to total abundance (N), reflecting the feasible set and the influence of residence time.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

lm_eqn <- function(df){
    m <- lm(log(y) ~ log(x), df)
    a = format(coef(m)[1], digits = 2) 
    b = format(coef(m)[2], digits = 2) 
    r2 = format(summary(m)$r.squared, digits = 3)
    bquote('S = ' ~ .(a) ~ N^{.(b)})
    }

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N.png',
      width = 1500, height = 1200, res=200, bg = "white")

df <- data.frame(x=sim.data$total.abundance, y=sim.data$species.richness)
p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm',colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Species richness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.6)

df <- data.frame(x=sim.data$total.abundance, y=sim.data$e.var)
p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm', colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Species evenness") +
  scale_x_log10() +
  scale_y_log10() +
  ylim(0.01,1) +
  coord_fixed(ratio = 4.3)

multiplot(p1, p2, cols=2)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

lm_eqn <- function(df){
    m <- lm(log(y) ~ log(x), df, na.action=na.exclude)
    a = format(coef(m)[1], digits = 2) 
    b = format(coef(m)[2], digits = 2) 
    r2 = format(summary(m)$r.squared, digits = 3)
    bquote('S = ' ~ .(a) ~ N^{.(b)} ~.(r2))
    }

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N2.png',
      width = 1500, height = 1200, res=200, bg = "white")

df <- data.frame(x=sim.data$total.abundance, y=sim.data$N.max)
p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm',colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Nmax") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1)

multiplot(p1, cols=1)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N_ByMotion.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness,
             group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Species richness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.55)

p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Species evenness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 3.8)

p3 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Dominance") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1)

p4 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$skew,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Rarity") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.4)

multiplot(p1, p2, p3, p4, cols=2)
dev.off()
```

### 2015-09-10 analyses/ForPaper/MultivariateTests.Rmd

@@
 ```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}
+Nfit <- lm(log(sim.data$total.abundance) ~ h.tau)
+summary(Nfit)
+
+Sfit <- lm(log(sim.data$species.richness) ~ h.tau)
+summary(Sfit)
+
+Efit <- lm(sim.data$simpson.e ~ h.tau)
+summary(Efit)
+
+Bfit <- lm(sim.data$species.turnover ~ h.tau)
+summary(Bfit)
+
+x <- sim.data$total.abundance
+
+Nfit <- lm(log(sim.data$total.abundance) ~ x)
+summary(Nfit)
+
+Sfit <- lm(log(sim.data$species.richness) ~ x)
+summary(Sfit)
+
+Efit <- lm(sim.data$simpson.e ~ x)
+summary(Efit)
+
+Bfit <- lm(sim.data$species.turnover ~ x)
+summary(Bfit)
+```
+
+```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}
 

### 2015-09-11 analyses/ForPaper/PrimaryPredictions.Rmd
added more simulation data; generated new figures -KL

@@
 source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
-source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
 ```
@@
 ## Load Data
-sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
-h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
-sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
-sim.data$species.richness[sim.data$species.richness<=0] <- 1
+ dat <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_September/7_Sept/SimData.csv")  
+dat$h.tau <- (dat$height * dat$width)/dat$flow.rate
+dat$total.abundance[dat$total.abundance<=0] <- 1
+dat$species.richness[dat$species.richness<=0] <- 1
 
-sim.data.fluid <- subset(sim.data, motion == "fluid")
-h.tau.fluid <- (sim.data.fluid$height * sim.data.fluid$width)/sim.data.fluid$flow.rate
-sim.data.fluid$total.abundance[sim.data.fluid$total.abundance<=0] <- 1
-sim.data.fluid$species.richness[sim.data.fluid$species.richness<=0] <- 1
+dat$e.var[dat$e.var =="NaN"] <- 1
+dat$simpson.e[dat$simpson.e =="NaN"] <- 1
 
-sim.data.random_walk <- subset(sim.data, motion == "random_walk")
-h.tau.random_walk <- (sim.data.random_walk$height * sim.data.random_walk$width)/sim.data.random_walk$flow.rate
-sim.data.random_walk$total.abundance[sim.data.random_walk$total.abundance<=0] <- 1
-sim.data.random_walk$species.richness[sim.data.random_walk$species.richness<=0] <- 1
+f.dat <- subset(dat, motion == "fluid")
+r.dat <- subset(dat, motion == "random_walk")
 ```
@@
 
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+#install.packages("ISLR")
+require(ISLR)
+
+fit <- lm(log10(f.dat$total.abundance) ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
+f.N.2.r2 <- summary(fit)$adj.r.squared
+f.N.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(f.dat$total.abundance) ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
+f.N.3.r2 <- summary(fit)$adj.r.squared
+f.N.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(r.dat$total.abundance) ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
+r.N.2.r2 <- summary(fit)$adj.r.squared
+r.N.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(r.dat$total.abundance) ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
+r.N.3.r2 <- summary(fit)$adj.r.squared
+r.N.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(f.dat$species.richness) ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
+f.S.2.r2 <- summary(fit)$adj.r.squared
+f.S.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(f.dat$species.richness) ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
+f.S.3.r2 <- summary(fit)$adj.r.squared
+f.S.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(r.dat$species.richness) ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
+r.S.2.r2 <- summary(fit)$adj.r.squared
+r.S.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(log10(r.dat$species.richness) ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
+r.S.3.r2 <- summary(fit)$adj.r.squared
+r.S.3.pVal <- anova(fit)$'Pr(>F)'[1]
+```
+
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
 png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
@@
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
-                           group=sim.data$sim, colour=sim.data$sim)) +
-  scale_color_gradient(high="cyan") +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+p1 <- ggplot()+
+  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$total.abundance, group=f.dat$sim), colour='skyblue1',
+    data=f.dat, se=F) +  
+  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$total.abundance, group=r.dat$sim), colour='seagreen1',
+    data=r.dat, se=F) +
+
+  stat_smooth(aes(x=dat$h.tau, y=dat$total.abundance, group=dat$motion, colour=dat$motion),
+    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
+    formula = y ~ poly(x,2), se=FALSE) +
+
+  scale_colour_manual(values = c("navy", "darkgreen"),
+      labels = c(paste("Fluid, r^2 = ",round(f.N.2.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.N.2.r2,2))),
+      name = "") +
+
   labs(x= expression(tau), y="Individuals") +
   labs(title="Total abundance") +
-  guides(colour=FALSE) + 
   scale_x_log10() +
   scale_y_log10() +
+  theme(legend.position="top") +
   coord_fixed(ratio = .50)
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
-                           group=sim.data$sim, colour=sim.data$sim)) +
-  scale_color_gradient(high="chartreuse") +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
+p2 <- ggplot()+
+  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$species.richness, group=f.dat$sim), colour='skyblue1',
+    data=f.dat, se=F) +  
+  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$species.richness, group=r.dat$sim), colour='seagreen1',
+    data=r.dat, se=F) +
+
+  stat_smooth(aes(x=dat$h.tau, y=dat$species.richness, group=dat$motion, colour=dat$motion),
+    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
+    formula = y ~ poly(x,2), se=FALSE) +
+
+  scale_colour_manual(values = c("navy", "darkgreen"),
+      labels = c(paste("Fluid, r^2 = ",round(f.S.2.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.S.2.r2,2))),
+      name = "") +
+
   labs(x= expression(tau), y="Number of species") +
   labs(title="Species richness") +
-  guides(colour=FALSE) +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1.0)
-
-multiplot(p1, p2, cols=2)
-dev.off()
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime_ByMotion.png',
-    width = 1500, height = 1200, res=200, bg = "white")
-
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
-                           group=sim.data$sim, colour=sim.data$motion)) +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
-  labs(x= expression(tau), y="Individuals") +
-  labs(title="Total abundance") +
-  guides(colour=FALSE) + 
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = .50)
-
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
-                           group=sim.data$sim, colour=sim.data$motion)) +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
-  labs(x= expression(tau), y="Number of species") +
-  labs(title="Species richness") +
-  guides(colour=FALSE) +
+  theme(legend.position="top") +
   scale_x_log10() +
@@
 
+```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
+fit <- lm(f.dat$e.var ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
+f.ev.2.r2 <- summary(fit)$adj.r.squared
+f.ev.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(f.dat$e.var ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
+f.ev.3.r2 <- summary(fit)$adj.r.squared
+f.ev.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(r.dat$e.var ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
+r.ev.2.r2 <- summary(fit)$adj.r.squared
+r.ev.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(r.dat$e.var ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
+r.ev.3.r2 <- summary(fit)$adj.r.squared
+r.ev.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(f.dat$species.turnover ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
+f.to.2.r2 <- summary(fit)$adj.r.squared
+f.to.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(f.dat$species.turnover ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
+f.to.3.r2 <- summary(fit)$adj.r.squared
+f.to.3.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(r.dat$species.turnover ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
+r.to.2.r2 <- summary(fit)$adj.r.squared
+r.to.2.pVal <- anova(fit)$'Pr(>F)'[1]
+
+fit <- lm(r.dat$species.turnover ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
+r.to.3.r2 <- summary(fit)$adj.r.squared
+r.to.3.pVal <- anova(fit)$'Pr(>F)'[1]
+```
+
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
-png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/EvennessAndTurnover_vs_ResTime.png',
-      width = 1500, height = 1200, res=200, bg = "white")
+png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/EvennessAndTurnover_vs_ResTime.png', width = 1500, height = 1200, res=200, bg = "white")
 
-p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
-                           group=sim.data$sim, colour=sim.data$sim)) +
-  scale_color_gradient(high="cyan") +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
-  labs(x= expression(tau), y="evenness") +
+p1 <- ggplot()+
+  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$e.var, group=f.dat$sim), colour='skyblue1',
+    data=f.dat, se=F) +  
+  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$e.var, group=r.dat$sim), colour='seagreen1',
+    data=r.dat, se=F) +
+
+  stat_smooth(aes(x=dat$h.tau, y=dat$e.var, group=dat$motion, colour=dat$motion),
+    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
+    formula = y ~ poly(x,2), se=FALSE) +
+
+  scale_colour_manual(values = c("navy", "darkgreen"),
+      labels = c(paste("Fluid, r^2 = ",round(f.ev.2.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.ev.2.r2,2))),
+      name = "") +
+
+  labs(x= expression(tau), y="Evar") +
   labs(title="Species evenness") +
-  guides(colour=FALSE) + 
   scale_x_log10() +
-  coord_fixed(ratio = 2.5)
+  ylim(0,1) +
+  theme(legend.position="top") +
+  coord_fixed(ratio = 2.0)
 
-p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
-                           group=sim.data$sim, colour=sim.data$sim)) +
-  scale_color_gradient(high="chartreuse") +
-  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
-  labs(x= expression(tau), y="turnover") +
+p2 <- ggplot()+
+  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$species.turnover, group=f.dat$sim), colour='skyblue1',
+    data=f.dat, se=F) +  
+  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$species.turnover, group=r.dat$sim), colour='seagreen1',
+    data=r.dat, se=F) +
+
+  stat_smooth(aes(x=dat$h.tau, y=dat$species.turnover, group=dat$motion, colour=dat$motion),
+    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
+    formula = y ~ poly(x,2), se=FALSE) +
+
+  scale_colour_manual(values = c("navy", "darkgreen"),
+      labels = c(paste("Fluid, r^2 = ",round(f.to.2.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.to.2.r2,2))),
+      name = "") +
+
+  labs(x= expression(tau), y="Whittaker's") +
   labs(title="Species turnover") +
-  guides(colour=FALSE) +
   scale_x_log10() +
-  coord_fixed(ratio = 2.5)
+  ylim(0,1) +
+  theme(legend.position="top") +
+  coord_fixed(ratio = 2.0)
 
@@
 ```
-
-## We hypothesized that aspects of diversity would be related to total abundance (N), reflecting the feasible set and the influence of residence time.
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
-
-lm_eqn <- function(df){
-    m <- lm(log(y) ~ log(x), df)
-    a = format(coef(m)[1], digits = 2) 
-    b = format(coef(m)[2], digits = 2) 
-    r2 = format(summary(m)$r.squared, digits = 3)
-    bquote('S = ' ~ .(a) ~ N^{.(b)})
-    }
-
-png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N.png',
-      width = 1500, height = 1200, res=200, bg = "white")
-
-df <- data.frame(x=sim.data$total.abundance, y=sim.data$species.richness)
-p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness)) +
-  labs(title=lm_eqn(df)) +
-  geom_point(colour="snow3") +
-  geom_smooth(method='lm',colour="lawngreen", size=2) +
-  labs(x= "Total abundance", y="Species richness") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1.6)
-
-df <- data.frame(x=sim.data$total.abundance, y=sim.data$e.var)
-p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var)) +
-  labs(title=lm_eqn(df)) +
-  geom_point(colour="snow3") +
-  geom_smooth(method='lm', colour="lawngreen", size=2) +
-  labs(x= "Total abundance", y="Species evenness") +
-  scale_x_log10() +
-  scale_y_log10() +
-  ylim(0.01,1) +
-  coord_fixed(ratio = 4.3)
-
-multiplot(p1, p2, cols=2)
-dev.off()
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
-
-lm_eqn <- function(df){
-    m <- lm(log(y) ~ log(x), df, na.action=na.exclude)
-    a = format(coef(m)[1], digits = 2) 
-    b = format(coef(m)[2], digits = 2) 
-    r2 = format(summary(m)$r.squared, digits = 3)
-    bquote('S = ' ~ .(a) ~ N^{.(b)} ~.(r2))
-    }
-
-png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N2.png',
-      width = 1500, height = 1200, res=200, bg = "white")
-
-df <- data.frame(x=sim.data$total.abundance, y=sim.data$N.max)
-p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max)) +
-  labs(title=lm_eqn(df)) +
-  geom_point(colour="snow3") +
-  geom_smooth(method='lm',colour="lawngreen", size=2) +
-  labs(x= "Total abundance", y="Nmax") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1)
-
-multiplot(p1, cols=1)
-dev.off()
-```
-
-```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
-source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
-
-png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N_ByMotion.png',
-      width = 1500, height = 1200, res=200, bg = "white")
-
-p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness,
-             group=sim.data$sim, colour=sim.data$motion)) +
-  geom_point() +
-  labs(x= "Total abundance", y="Species richness") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1.55)
-
-p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var,
-                           group=sim.data$sim, colour=sim.data$motion)) +
-  geom_point() +
-  labs(x= "Total abundance", y="Species evenness") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 3.8)
-
-p3 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max,
-                           group=sim.data$sim, colour=sim.data$motion)) +
-  geom_point() +
-  labs(x= "Total abundance", y="Dominance") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1)
-
-p4 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$skew,
-                           group=sim.data$sim, colour=sim.data$motion)) +
-  geom_point() +
-  labs(x= "Total abundance", y="Rarity") +
-  scale_x_log10() +
-  scale_y_log10() +
-  coord_fixed(ratio = 1.4)
-
-multiplot(p1, p2, p3, p4, cols=2)
-dev.off()
-```

### 2015-09-12 models/MovieMaker/MovieMaker.pyx
modified bide code, main model code, fluid dynamics code; generated new simulated data; deleted old MovieMaker directory and code -KL

@@ -1,250 +1,0 @@
-from __future__ import division
-import matplotlib.animation as animation
-import matplotlib.pyplot as plt
-#import mpl_toolkits.mplot3d
-#from mpl_toolkits.mplot3d import Axes3D
-
-from random import choice
-from scipy import stats
-import numpy as np
-import sys
-import os
-import psutil
-
-mydir = os.path.expanduser("~/")
-sys.path.append(mydir + "tools/metrics")
-import metrics
-sys.path.append(mydir + "GitHub/hydrobide/tools/LBM")
-import LBM
-sys.path.append(mydir + "GitHub/hydrobide/tools/bide")
-import bide
-
-
-
-def get_rand_params():
-    """ Get random model parameter values. Others are chosen in bide.pyx """
-
-    motion = choice(['fluid', 'random_walk', 'conveyor'])
-    motion = 'fluid'
-    D = choice([2, 2]) # number of spatial dimensions
-    width, height, length = [10, 10, 10]
-
-    alpha = np.random.uniform(0.99, 0.999)
-    reproduction = choice(['fission', 'sexual'])
-    speciation = choice(['yes', 'no'])
-    predators = choice([0, 1, 2, 4, 8])
-    parasites = choice([0, 1, 2, 4, 8])
-    env_gradient = choice(['no', 'yes'])
-
-    seedcom = 100 # size of starting community
-    m = 1 # individuals immigrating per time step
-    r = 40 # resource particles flowing in per time step
-    nr = 1 # maximum number of resources types
-    rmax = 500 # maximum resource particle size
-
-
-    reproduction = 'fission'
-    speciation = 'yes'
-
-    return [width, height, length, alpha, motion, D, reproduction, speciation, predators, parasites, env_gradient, seedcom, m, r, nr, rmax]
-
-
-
-def testlengths(TypeOf, function, Lists):
-    vals = []
-    for List in Lists: vals.append(len(List))
-    if min(vals) != max(vals):
-        print '\n'+TypeOf+': '+function+', list lengths are different sizes:', vals
-        sys.exit()
-    return
-
-
-
[…]

### 2015-09-12 models/MovieMaker/setup.py

@@ -1,4 +1,0 @@
-from distutils.core import setup
-from Cython.Build import cythonize
-
-setup(ext_modules = cythonize("MovieMaker.pyx"))

### 2015-09-12 tools/bide/bide.py

@@ -119,7 +119,7 @@ def ResIn(motion, Type, Vals, Xs, Ys, Zs, ID, IDs, t_In, numr,
 
 
 
-def immigration(d_max, g_max, m_max, motion, numin, Sp, Xs, Ys, Zs, w, h, l, MD,
+def immigration(d_max, g_max, m_max, motion, numin, Sp, Xs, Ys, Zs, w, h, l, MD, EnvD,
         GD, DispD, colorD, IDs, ID, t_In, Qs, N_RD, P_RD, C_RD, nN, nP, nC, u0, alpha, D, GList, MList, NList, PList, CList, DList):
 
     for m in range(numin):
@@ -195,7 +195,7 @@ def immigration(d_max, g_max, m_max, motion, numin, Sp, Xs, Ys, Zs, w, h, l, MD,
             i = GetIndParam(means)
             DList.append(i)
 
-    return [Sp, Xs, Ys, Zs, MD, GD, DispD, colorD, IDs, ID, t_In, Qs, N_RD, P_RD, C_RD, GList, MList, NList, PList, CList, DList]
+    return [Sp, Xs, Ys, Zs, MD, EnvD, GD, DispD, colorD, IDs, ID, t_In, Qs, N_RD, P_RD, C_RD, GList, MList, NList, PList, CList, DList]
 
 
 
@@ -370,7 +370,7 @@ def predation(P_IDs, P_ID, P_Xs, P_Ys, P_Zs, P_t_In, I_xAge, Sp_IDs,
 
 
 
-def maintenance(Sp_IDs, Xs, Ys, Zs, xAge, colorD, MD, IDs, t_In, Qs, D,  GrowthList, MaintList, N_RList, P_RList, C_RList, DispList):
+def maintenance(Sp_IDs, Xs, Ys, Zs, xAge, colorD, MD, EnvD, IDs, t_In, Qs, D,  GrowthList, MaintList, N_RList, P_RList, C_RList, DispList):
 
     if Sp_IDs == []:
         return [Sp_IDs, Xs, Ys, Zs, xAge, IDs, t_In, Qs, GrowthList, MaintList, N_RList, P_RList, C_RList, DispList]
@@ -404,7 +404,7 @@ def maintenance(Sp_IDs, Xs, Ys, Zs, xAge, colorD, MD, IDs, t_In, Qs, D,  GrowthL
 
 
 
[…]

### 2015-09-12 analyses/ForPaper/PrimaryPredictions.Rmd

@@
 
-f.dat <- subset(dat, motion == "fluid")
-r.dat <- subset(dat, motion == "random_walk")
+f.dat <- subset(dat, motion == " fluid")
+r.dat <- subset(dat, motion == " random_walk")
 ```
@@
     data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
-    formula = y ~ poly(x,2), se=FALSE) +
+    formula = y ~ poly(x,3), se=FALSE) +
 
   scale_colour_manual(values = c("navy", "darkgreen"),
-      labels = c(paste("Fluid, r^2 = ",round(f.N.2.r2,2)),
-      paste("Non-fluid, r^2 = ",round(r.N.2.r2,2))),
+      labels = c(paste("Fluid, r^2 = ",round(f.N.3.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.N.3.r2,2))),
       name = "") +
@@
     data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
-    formula = y ~ poly(x,2), se=FALSE) +
+    formula = y ~ poly(x,3), se=FALSE) +
 
   scale_colour_manual(values = c("navy", "darkgreen"),
-      labels = c(paste("Fluid, r^2 = ",round(f.S.2.r2,2)),
-      paste("Non-fluid, r^2 = ",round(r.S.2.r2,2))),
+      labels = c(paste("Fluid, r^2 = ",round(f.S.3.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.S.3.r2,2))),
       name = "") +
@@
     data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
-    formula = y ~ poly(x,2), se=FALSE) +
+    formula = y ~ poly(x,3), se=FALSE) +
 
   scale_colour_manual(values = c("navy", "darkgreen"),
-      labels = c(paste("Fluid, r^2 = ",round(f.ev.2.r2,2)),
-      paste("Non-fluid, r^2 = ",round(r.ev.2.r2,2))),
+      labels = c(paste("Fluid, r^2 = ",round(f.ev.3.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.ev.3.r2,2))),
       name = "") +
@@
     data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
-    formula = y ~ poly(x,2), se=FALSE) +
+    formula = y ~ poly(x,3), se=FALSE) +
 
   scale_colour_manual(values = c("navy", "darkgreen"),
-      labels = c(paste("Fluid, r^2 = ",round(f.to.2.r2,2)),
-      paste("Non-fluid, r^2 = ",round(r.to.2.r2,2))),
+      labels = c(paste("Fluid, r^2 = ",round(f.to.3.r2,2)),
+      paste("Non-fluid, r^2 = ",round(r.to.3.r2,2))),
       name = "") +

### 2015-09-15 analyses/ForPaper/PrimaryPredictions.Rmd
added new directory and files for a Hydrobide GUI -KL

@@
 require("moments")
+#install.packages("IDPmisc", repos = "http://cran.cnr.Berkeley.edu")
+require(IDPmisc)
+#install.packages("ISLR")
+require(ISLR)
+
 library(ggplot2)
@@
 ## Load Data
- dat <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_September/7_Sept/SimData.csv")  
+ dat <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_September/12_Sept/SimData.csv")
+dat <- NaRV.omit(dat)
 dat$h.tau <- (dat$height * dat$width)/dat$flow.rate
-dat$total.abundance[dat$total.abundance<=0] <- 1
-dat$species.richness[dat$species.richness<=0] <- 1
-
-dat$e.var[dat$e.var =="NaN"] <- 1
-dat$simpson.e[dat$simpson.e =="NaN"] <- 1
 
@@
 ```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
-#install.packages("ISLR")
-require(ISLR)
-
 fit <- lm(log10(f.dat$total.abundance) ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))

### 2015-09-15 models/main/model.py

@@ -4,7 +4,7 @@ import matplotlib.pyplot as plt
 #import mpl_toolkits.mplot3d
 #from mpl_toolkits.mplot3d import Axes3D
 
-from random import choice, seed
+from random import choice
 from scipy import stats
 import numpy as np
 from numpy import sin, pi
@@ -14,7 +14,7 @@ import psutil
 #import itertools
 
 mydir = os.path.expanduser("~/")
-sys.path.append(mydir + "tools/metrics")
+sys.path.append(mydir + "GitHub/hydrobide/tools/metrics")
 import metrics
 sys.path.append(mydir + "GitHub/hydrobide/tools/LBM")
 import LBM
@@ -26,17 +26,13 @@ import bide
 def get_rand_params():
     """ Get random model parameter values. Others are chosen in bide.pyx """
 
-    #motion = choice(['fluid', 'random_walk', 'conveyor', 'search'])
     motion = choice(['fluid', 'random_walk'])
     D = choice([2, 2]) # number of spatial dimensions
 
-    width = choice([10, 12, 14, 16, 18, 20])
+    width = choice([10, 15, 20])
     height = choice([5, 6, 7, 8, 9, 10])
     length = choice([5, 6, 7, 8, 9, 10])
-
-    barriers = choice([2, 4, 6, 8, 16, 32])
-    barriers = 4
-    #if motion == 'conveyor': barriers = 0
+    barriers = choice([1, 2, 3, 4])
 
     pulse = choice([0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0])
     flux = choice(['yes'])
@@ -48,37 +44,49 @@ def get_rand_params():
     phase = choice([0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10])
     # 0 = in phase; 16 = entirely out of phase
 
-    disturb = choice([0.0, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.1])
+    disturb = choice([0.0001])#, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.1])
 
     alpha = np.random.uniform(0.99, 0.999)
     reproduction = choice(['fission', 'sexual'])
     speciation = choice(['yes', 'no'])
 
-    seedCom = choice([1000]) # size of starting community
-    m = choice([1]) # m = individuals immigrating per time step
[…]

### 2015-09-15 tools/LBM/LBM.py

@@ -29,8 +29,8 @@ def SetLattice(u0, viscosity, width, height, lefts, bottoms, barriers): # Lattic
     barrier = np.zeros((height, width), bool)  # Initialize barriers
                                                # True wherever there's a barrier
 
-    BarrierHeight = 0.1
-    BarrierWidth = 0.1
+    BarrierHeight = 0.05
+    BarrierWidth = 0.05
 
     for i in range(barriers):
 

### 2015-09-15 tools/metrics/metrics.py

@@ -1,0 +1,549 @@
+from __future__ import division
+import sys
+import numpy as np
+from scipy import stats
+from scipy.stats import gaussian_kde
+import mpmath as mpm
+from scipy.optimize import fsolve
+import re
+import random
+import os
+import math
+
+mydir = os.path.expanduser("~/GitHub/rare-bio")
+data = os.path.expanduser("~/data")
+
+################################################################################
+#### Section of evenness indices and descriptive statistical functions #########
+
+
+############### RARITY #########################################################
+def r_singletons(sad): # relative abundance of the least abundant taxon
+
+    """ percent of species represented by a single individual """
+
+    return 100 * sad.count(1)/len(sad)
+
+
+def p_ZPtOne(sad):
+
+    """ percent taxa with less than 0.1% N """
+    N = sum(sad)
+    S = len(sad)
+
+    sad = np.array(sad)/N
+    sad = sad*100
+    sad = sad.tolist()
+
+    numR = 0
+    for sp in sad:
+        if sp < 0.1:
+            numR += 1
+
+    return numR/S
+
+
+def Rlogskew(sad):
+    S = len(sad)
+
+    if S <= 2.0:
+        print 'S < 2, cannot compute log-skew'
+        sys.exit()
+
+    sad = np.log10(sad)
+    mu = np.mean(sad)
+
+    num = 0
+    denom = 0
+    for ni in sad:
+        num += ((ni - mu)**3.0)/S
+        denom += ((ni - mu)**2.0)/S
+
+    t1 = num/(denom**(3.0/2.0))
+    t2 = (S/(S - 2.0)) * np.sqrt((S - 1.0)/S)
+
+    return t1 * t2
+
+
+############### LOGNORMAL VARIABLES ############################################
+
+def Preston(sad):
+    N = sum(sad)
+    Nmax = max(sad)
+
+    left = (2 * N)/(np.sqrt(np.pi) * Nmax)
+
+    func = lambda a : left - (math.erf(np.log(2)/a) / a)
+
+    guess = 0.1 # alpha is often ~0.2, but appears to be lower for larger N
+    a = fsolve(func, guess)
+
+    expS = (np.sqrt(np.pi) / a) * np.exp( (np.log(2)/(2*a))**2 )
+
[…]

### 2015-10-05 results/TestResults/SpeedMemory.Rmd
modified model.py, edited Methods.md and Results.md, Added time and memory analysis file and data, added example .Rmd code (Exploratory.Rmd), deleted unecessary .Rmd files, added new simulation data, modified metrics.R and multiplot.R files -KL

---
title: "Speed and Time tests"
author: "Ken Locey"
date: "October, 2015"
header-includes: - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/simplex") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
dat <- read.csv("~/GitHub/simplex/results/TestResults/SpeedMemory.csv")

sem <- function(x) sqrt(var(x)/length(x))

#is.nan(val) == FALSE
Ns <- dat$N
mn = mean(Ns)
se <- sem(Ns)
print(c(mn, se))

times <- dat$time
mn = mean(times)
se <- sem(times)
print(c(mn, se))

mems <- dat$memory
mn = mean(mems)
se <- sem(mems)
print(c(mn, se))
```

### 2015-10-05 model/model.py

@@ -24,6 +24,10 @@ sys.path.append(mydir + "GitHub/simplex/tools/randparams")
 import randparams as rp
 
 
+# https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing
+# http://statsmodels.sourceforge.net/0.5.0/generated/statsmodels.tsa.stattools.adfuller.html
+
+
 GenPath = mydir + 'GitHub/simplex/results/simulated_data/'
 
 ############## OPEN OUTPUT DATA FILE ###########################################
@@ -288,6 +292,8 @@ def nextFrame(arg):
                 bottoms.append(np.random.uniform(0.2, 0.8))
 
             u0 = choice(Rates)
+            p = 0
+            BurnIn = 'not done'
 
             RDens, RDiv, RRich, S, ES, Ev, BP, SD, Nm, sk, Mu, Maint, ct, IndID, RID, N, ct1, T, R, PRODI, PRODQ = [0]*21
             RAD, splist, IndTimeIn, SpeciesIDs, IndX, IndY,  IndIDs, Qs, IndExitAge, TX, TY, TExitAge, TIDs, TTimeIn, RX, RY,  RIDs, RTypes, RExitAge, RTimeIn, RVals, Gs, Ms, NRs, PRs, CRs, Ds, Ts, Rs, PRODIs, PRODNs, PRODPs, PRODCs, Ns, RTAUs, TTAUs, INDTAUs, RDENs, RDIVs, RRICHs, Ss, ESs, EVs,BPs, SDs, NMAXs, SKs, MUs, MAINTs, WTs, Jcs, Sos, splist2 = [list([]) for _ in xrange(53)]

### 2015-10-05 tools/Rbin/metrics.R

@@ -1,24 +1,6 @@
-################################################################################
-#                                                                              #
-# Functions for calculating metrics of diversity, evenness, rarity, etc.       #
-# These are not included in other diversity packages, e.g., Vegan              #
-#                                                                              #
-################################################################################
-#                                                                              #
-# Written by: Ken Locey                                                        #
-#                                                                              #
-################################################################################
-#                                                                              #
-# Recent Changes:                                                              #
-#                                                                              #
-# Future Changes (To-Do List):                                                 #
-#         1.                                                                   #
-#         2. Add warnings                                                      #
-#                                                                              #
-################################################################################
-
 #### A function to generate observed richness
 S.obs <- function(x = ""){ rowSums(x > 0) * 1}
+
 
 sp.turnover <- function(site1, site2){
   
@@ -36,3 +18,28 @@ sp.turnover <- function(site1, site2){
   return(b.w)
   }
 
+
+
+get.vectors <- function(inputFile){
+  vectors <- c()
+  i <- 1
+  con  <- file(inputFile, open = "r")
+  while (length(oneLine <- readLines(con, n = 1)) > 0) {
+    myLine <- unlist((strsplit(oneLine, ",")))
+    myLine[myLine == '[]'] <- -1
+    myLine[is.na(myLine)] <- -1
+    myLine <- myLine[which(myLine!=-1)]
+    myLine <- myLine[which(myLine!="")]
+  
+    if (length(myLine > 0)){
+      myLine <- as.vector(myLine)
+      vectors[i] <- myLine
+      i <- i + 1
+    }
+  }
+  close(con)
+  return(vectors)
+}
+
+
+sem <- function(x) return(sqrt(var(x)/length(x)))

### 2015-10-05 tools/Rbin/multiplot.R

@@ -84,13 +84,13 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     axis.title =         element_text(colour = "white"),
     axis.title.x =       element_text(vjust = -0.5, size = rel(1.5)),
     axis.title.y =       element_text(angle = 90, size = rel(1.5)),
-    axis.ticks.length =  unit(0.3, "lines"),
-    axis.ticks.margin =  unit(0.5, "lines"),
+    #axis.ticks.length =  unit(0.3, "lines"),
+    #axis.ticks.margin =  unit(0.5, "lines"),
     
     legend.background =  element_rect(colour = "black"),
-    legend.margin =      unit(0.2, "cm"),
+    #legend.margin =      unit(0.2, "cm"),
     legend.key =         element_rect(fill = "black", colour = "white"),
-    legend.key.size =    unit(1.2, "lines"),
+    #legend.key.size =    unit(1.2, "lines"),
     legend.key.height =  NULL,
     legend.key.width =   NULL,
     legend.text =        element_text(size = rel(0.8), colour = "white"),
@@ -107,7 +107,7 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     panel.border =       element_rect(fill = NA, colour = "white"),
     panel.grid.major =   element_line(colour = "grey20", size = 0.2),
     panel.grid.minor =   element_line(colour = "grey5", size = 0.5),
-    panel.margin =       unit(0.25, "lines"),
+    #panel.margin =       unit(0.25, "lines"),
     
     strip.background =   element_rect(fill = "grey30", colour = "grey10"),
     strip.text.x =       element_text(),
@@ -115,7 +115,7 @@ theme_black <- function(base_size = 12, base_family = "Helvetica") {
     
     plot.background =    element_rect(colour = "black", fill = "black"),
     plot.title =         element_text(colour = "white", size = rel(1.8)),
-    plot.margin =        unit(c(1, 1, 0.5, 0.5), "lines"),
+    #plot.margin =        unit(c(1, 1, 0.5, 0.5), "lines"),
     
     complete = TRUE
   )

### 2015-10-05 tools/randparams/randparams.py

@@ -10,9 +10,9 @@ def get_rand_params():
     motion = choice(['fluid'])
     D = choice([2, 2]) # number of spatial dimensions
 
-    width = choice([10, 15, 20])
+    width = choice([10, 15, 20, 30, 40, 50])
     height = choice([5, 6, 7, 8, 9, 10])
-    barriers = choice([1])
+    barriers = choice([1,2,3,4])
 
     pulse = choice([0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0])
     flux = choice(['yes'])
@@ -25,15 +25,14 @@ def get_rand_params():
     # 0 = in phase; 16 = entirely out of phase
 
     disturb = choice([0.0001])#, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.1])
-    rates = np.array([1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.2, 0.1, 0.09, 0.08, 0.07, 0.06, 0.05, 0.04, 0.02, 0.01, 0.0075, 0.005])  # inflow speeds
+    rates = np.array([1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.2, 0.1, 0.09, 0.08, 0.07, 0.06, 0.05, 0.04, 0.02, 0.01, 0.0075, 0.005, 0.001, 0.0005, 0.0001])  # inflow speeds
 
-    #rates = [0.001]
-    alpha = np.random.uniform(0.99, 0.999)
+    alpha = np.random.uniform(0.99, 0.9999)
     reproduction = choice(['fission', 'sexual'])
     speciation = choice(['yes', 'no'])
 
     seedCom = choice([100]) # size of starting community
-    m = choice([0.0, 0.0001, 0.0005, 0.001, 0.005]) # m = probability of immigration
+    m = choice([0.0, 0.0001, 0.0005, 0.001, 0.005, 0.01]) # m = probability of immigration
     r = choice([200, 400, 600, 800, 1000])
     # r = resource particles flowing in per time step
     nNi = choice([1, 2, 3, 4, 5, 6, 7, 8, 9, 10]) # max number of Nitrogen types
@@ -42,7 +41,6 @@ def get_rand_params():
 
     envgrads = []
[…]

### 2015-10-05 results/analyses/Rmd/Exploratory.Rmd

---
title: "Exploratory analyses of simplex output"
author: "Ken Locey"
date: "October, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## Clear and set the working directory

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/simplex") 
```

## Import packages; install if needed

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
library(ggplot2)
```

## Import custom modules of R functions

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/simplex/tools/Rbin/metrics.R")
```

## Import simulated data

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
## Load Data
dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")
ind.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/IndRTD.csv")
res.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/ResRTD.csv")
tracer.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/TracerRTD.csv")
species.list <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/Species.csv")
rads <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/RADs.csv")
```

## Exploratory Analyses

### 1.) How do species richness (S) and species evenness relate to total abundance (N)?

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
N <- dat$total.abundance[dat$total.abundance<=0] <- 1
S <- dat$species.richness[dat$species.richness<=0] <- 1

ggplot(dat, aes(x=dat$total.abundance, y=dat$species.richness)) +
  geom_point(colour="blue") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x= "Total abundance", y="Species richness")

ggplot(dat, aes(x=dat$total.abundance, y=dat$e.var)) +
  geom_point(colour="red") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x= "Total abundance", y="Species evenness")
```

### 2.) Variance partitioning

```{r, results='hide', echo=TRUE, message=F, warning=F, fig.width=7, fig.height=4}
plot.new()
dat[is.na(dat)] <- 0
dat$h.tau <- (dat$height * dat$width)/dat$flow.rate

# Physical and metacommunity variables
phys.dat <- as.matrix(subset(dat, select = c(avg.per.capita.growth,
              avg.per.capita.maint,
              avg.per.capita.N.efficiency,
              avg.per.capita.P.efficiency,
              avg.per.capita.C.efficiency)))

# Resource variables
res.dat <- as.matrix(subset(dat, 
              select = c(resource.concentration,
              shannons.resource.diversity,
              resource.richness,
              resource.particles)))

y.dat <- as.matrix(subset(dat, 
              select = c(total.abundance,
              species.richness,
              simpson.e,
              Whittakers.turnover)))

rda.phys <- rda(y.dat ~ phys.dat)
#rda.phys
rda.res <- rda(y.dat ~ res.dat)
#rda.res

# Two explanatory matrices -- Hellinger-transform Y
mod <- varpart(Y=y.dat, X=phys.dat, res.dat,  transfo="log")
mod
showvarparts(2)
plot(mod)
mtext("Variation partitioning")
#text(1.6, 0.6, "env") 
#text(-0.6, 0.6, "tau")

# Test fraction using RDA:
phys.anova <- anova(rda.phys, step=200, perm.max=200)
phys.anova

res.anova <- anova(rda.res, step=200, perm.max=200)
res.anova
```

### 2015-10-06 paper/PrePrint.Rmd

---
title: "simplex: An individual-based modeling platform"
author: "Kenneth J Locey,
Dept. of Biology, University of Indiana, Bloomington"
geometry: margin=2.54cm
output: pdf_document
header-includes:
    - \usepackage{setspace}
    - \doublespacing
---

# Introduction

Modeling is a elementary approach to understanding ecological systems, the influence of ecological processes, and the predictability of ecological patterns and dynamics. Before modern computing, ecological models were almost exclusively equation-based representations of highly simplified systems (Black and McKane 2012, Otto and Troy 2007, Grant and Swannack 2008). Since the advent of personal computing, ecological models have been increasingly constructed to handle greater complexity and to explicitly simulate ecological processes.

In ecology, simulation-based models are often used to examine analytically challenging or intractable scenarios. 
Examples are markov models that simulate stochastic demographic changes and early ecological null models, both of which operate on matrices (Gotelli and Entsminger 2001, Hubbell 2001). Ecologists have also simulated growth and interactions among individuals for over two decades with individual-based models (IBMs) (DeAngelis and Gross 1992, Rosindell et al. 2015). 
IBMs explicitly encode rules of how individuals change and interact. Once the rules are encoded, population to ecosystem-level dynamics can emerge as an IBM simulates over time and spatially explicit environments.

A body of literature including original research, comprehensive reviews, and textbooks reveal the use, advantages, and challenges of ecological IBMs (e.g., Grimm 1999, Grimm and Railsback 2005). IBMs can provide degrees of ecological realism, individual variability, and spatial heterogenetiy that are unattainable with other models. IBMs also offer the potential for realistic and unanticipated ecological dynamics and patterns to emerge from individual-level interactions. However, IBMs offer challenges that include greater computational complexity, the difficulty of explicitly encoding ecological theory, and the use of ecological problem solving (Matthews et al. 2007, Grimm and Railsback 2005).

Despite their frequent use, the power of ecological IBMs has yet been leveraged to the greatest advantage. Few, if any, include fluid dynamics or combine fluid dynamics with growth and active dispersal. While growth, sensing, and decision making are often modeled, few IBMs integrate physiology, evolution, community ecology, biogeography, and sampling theory. 
Ecological IBMs are more often constructed to model a specific system than to synthesize general theories of ecology and evolution (but see Rosindell et al. 2015). Yet, IBMs can allow researchers to simulate and track information from the level of genomes and internal physiology to the stoichiometry of resource particles and distributions of abundance among species and across space.

Here, I present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output. This platform incorporates computational fluid dynamics (Succi 2001), life history processes and concepts and patterns from biogeography (Hubbell 2001, Bell 2001), evolution and aspects of selection (Hartl and Clark 1997), and integrates nutrient-limited growth and physiology (Pirt 1965, Droop 1983) among dimension of ecology and evolutionary science. 
I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes. 
Below, I provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.

# Methods

**Overview**--simplex was developed to accomplish three objectives. First, simplex assembles and runs individual-based ecological models from random combinations of state variables and processes. Second, simplex stores model output as R-style data frames and, when the user chooses, generates and stores animations of models. Third, simplex provides R-based files for analyzing simulated; these will eventually be offered in Python as well.

**Source code**--To save space, I also refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.

## Model description following the ODD Protocol
The ODD protocol is standard for describing individual-based models (Grimm et al. 2006). 
ODD stands for Overview, Design concepts, and Details. Here, I descibe simplex *largely* according to the ODD protocol.
The ODD protocol for simplex can also be found on the public GitHub repository at https://github.com/LennonLab/simplex

### Purpose
The purpose of simplex is to simulate life history of individual organisms, the assembly of ecological communities, and the evolution of traits in spatially explicit environments under stochastic conditions. The proximate goal of simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes. The ultimate goal of simplex is to provide a simulation-based platform for examining conditions under which processes and constraints have a robust influence on eco-evolutionary dynamics and patterns of biodiversity.

### Entities & their state variables
**Individual organisms**--Individuals are distinguished by collections of elements within lists. Individuals undergo changes when randomly sampled from lists. Each specific position in the list corresponds to the same individual. For example, in simulating growth, a simplex model chooses an individual from the list of cell quotas (the probability of reproducing is determined by endogenous resources). The first position in this list as well in all other lists of individual attributes corresponds to the same individual.

	Example:
	IndIDs = [1,  2, 33, 14]
	Quota =  [0.1, 0.99, 0.14, 0.05]
	Xpos =   [45, 23, 456, 1]
	Ypos =   [765, 87, 21, 34]

	The individual with ID of 1 has a cell quota of 0.1 and is located at position x=45, y=765

These are the lists of attributes for individual organisms:

* time each particle spends in the system (aka residence time)
* species ID
* individual ID
* 2D spatial location
* endogenous levels of 3 resources (resource specific cell quotas)
* individual-level metabolic maintenance cost
* individual-level maximum dispersal rate
* pedigree, i.e., direct lineage 
* individual resource use efficiency for each resource
* products of individual-level metabolism

**Species**--Each species is characterized by the individuals that share a common set of traits, such as maximum growth rate, metabolic maintenance cost. Species information is stored in Python dictionaries. In this way, if simplex requires the species ID of an individual it will access the species ID list where each element corresponds to a specific individual. But, if simplex requires the maximum growth rate for an individual, then it finds the species ID and then uses that to accesses the Python dictionary for maximum specific growth rates of species.

	Example:
	IndIDs = [1,  2, 33, 14]
	spIDs = [12, 32, 11, 6]
	MaxGrowthDict = {12: 0.9, 32: 0.5, 11: 0.8, 6: 0.4}

	The individual with ID of 1 belongs to species 12, which has a theoretical maximum growth rate of 0.9.

The following are they types of species-level information that are stored in Python dictionaries:

* metabolic maintenance cost
* maximum dispersal rate
* maximum theoretical growth rate
* resource use efficiency for each resource

**Resource particles**--Individual resource particles are distinguished by collections of elements within lists. Each specific position in the list corresponds to the same resource particle.

	Example:
	resIDs = [4,  6, 17, 1]
	size =  [100, 87, 156.3, 0.001]
	Xpos =   [56, 34, 567, 2]
	Ypos =   [876, 98, 32, 45]

	The particle with ID of 4 has a size of 100 and is 
	located at position x=56, y=876

The following are the types of information stored about each resource particle:

* time in the system
* particle ID
* 2D spatial location
* whether the particle is Nitrogen, Carbon, or Phosphorus
* type of Nitrogen, Carbon, or Phosphorus the resource is
* size of the particle

**Inert tracers particles**--These are objects that move/flow into and through the environment. They only interact with physical barriers. 
The use of tracers particles allows for attributes of a system that flows or physically turns over to be quantified (e.g., hydraulic residence time). Like other individual-level objects in simplex models, tracers are distinguished by collections of elements within lists. Each specific position in the list corresponds to the same resource particle.

This information stored about each resource particle:

* time in the system
* 2D spatial location
* particle ID

**Physcial barriers**--These objects are simulated as discrete 2D spatial coordinates that cannot be occupied by any individual entities. The number, size, and location of physical barriers are chosen at random by simplex at the start of each model.

###System level state variables
Each run of simplex begins with random choices for the values of:

* width (5 to 100)
* height (5 to 100)
* basal flow rate (1.0 to 0.0)
* number, size, and location of physical barriers 
* number and direction of environmental gradients
* rate of stochastic disturbance
* rate of fluctuation in basal flow rate
* degree of fluctuation 
* degree of syncronized flow of individual particles
	* completely in-sync or completely out of sync

For example

	from randparams.py:

    width = randint(5,100)
    height = randint(5,100)

    # number of barriers
    barriers = randint(1,10)

	 # log-series alpha for metacommunity structure
    alpha = np.random.uniform(0.99, 0.999)

    reproduction = choice(['fission', 'sexual'])
    speciation = choice(['yes', 'no'])

    # size of starting community
    seedCom = choice([10, 100, 1000])

    # m = probability of immigration
    m = choice([0.0, 0.0001, 0.0005, 0.001, 0.005])
    ...

###Spatial and temporal scales
The two general aspects of scale are grain (a.k.a. 
resolution) and extent (e.g. total area).

**Spatial extent**--The environment of simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units. This makes for a potential total extent of 25 to 10,000 discrete patches, each with a grain of 1 square unit.

Note that all particles move in decimal units the limit of which is determined by Python's decimal precision. This means that individual particles can occupy practically infinite locations within patches and likewise, squeeze through barriers (which only occupy integer x-y coordinates).

**Temporal extent**--Extent of time in simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system. Residence time for inert tracer particles can vary across five orders of magnitude.

**Grain**--Grain is the smallest unit over which change can happen. 
For example, as per the original ODD documentation: “One time step represents one year and simulations were run for 100 years. One grid cell represents 1 ha and the model landscape comprised 1,000 x 1,000 ha; i.e., 10,000 square kilometers”. In contrast, the smallest grain achievable by simplex is determined by slowest rate at which individuals can undergo BIDE (birth, immigration, death, emigration) processes. For example, under high residence times, an individual can move across 0.00001% of the x or y axis in one time step. Under low residence times, an individual can move across 10% or greater of the x or y axis in one time step.

### Process overview and scheduling
**Assembly**--The user runs a program that chooses random values for system-level state variables including whether disturbance, immigration, speciation, fluid dynamics, etc. will occur and at what rates.

**Core simulation process**--simplex models begin simulation immediately after assembly from random combinations of state-variables and processes. 
Instead of operating by definitive time steps (i.e. days, generations), simplex models advance turnover of the environmental matrix according to the initial rate of flow. 
If the initial rate of flow is 1.0, then the environmental matrix and inert particles would flow 1.0 units of distance. After each iteration of flow, each individual is given the chance to consume, grow, reproduce, starve, die, and to disperse towards resources and environmental optima.

**Duration: A run to mean reversion**--Once assembled, a simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance. Mean reversion captures whether a system is fluctuating around a mean value. Once 100 generations have occurred within a model, simplex examines whether a point of mean reversion has occured by conducting an Augmented Dickey-Fuller (ADF) Test, which is well-explained here: https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing. A model is stopped once mean reversion is determined to occurred and, unless the number of desired simulations (i.e. models) has been reached, simplex simply constructs another model from random combinations of state-variables and processes, and then runs it to mean reversion.

### Fluid dynamics
simplex uses an efficient and powerful method for 
simulating fluid flow, i.e., a Lattice-Boltzmann Method 
(LBM). An LBM discretizes the environment into a lattice 
and attaches to each position in the lattice a number of 
particle densities and velocities for each of nine 
directions of movement possible in a 2D environment 
(N, S, E, W, NE, NW, SE, SW, current position).

**Active dispersal**--simplex models allow individuals to move towards their environmental optima. Rather than a single environmental optima resulting from a single environmental gradient, simplex allows environmental optima to occur as intersections among environmental gradients. Hence, individuals potentially have multiple optima resulting from unique and equally optimal intersection of up to 10 environmental gradients.

**Simulated life history**--simplex models simulate growth, reproduction, and death via weighted random sampling. This simulate the partly probabilistic and partly deterministic nature of environmental filtering and individual-level interactions.

*Inflow/Entrance:* Resources and individuals enter from any point in the environment. Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution 
of abundance among species in ecological communities (see Hubbell 2001). Along with a species ID and species-
specific maximum rates of resource uptake, active dispersal, resource efficiencies, cell maintenance, 
and environmental optima, each propagule was given a unique ID and a multi-resource cell quota that represent 
the state of internal resources. The average of these cell quotas determine the probability of reproduction.

*Dispersal:* Individuals are allowed to actively move along environmental gradients (sometimes against the 
direction of environmenta flow) and towards their optima. A better match to one's environmental optima increases 
the chance of reproduction and the individual's ability to perform (consume, grow).

*Consumption & growth:* Sampled individuals consume resources according to their specific maximum rates of 
uptake and grow according to specific resource efficiencies and maintenance costs. Uptake increases the individual 
cell quotas and decreases ambient resources. Individual cell quotas are then decreased according to a specific 
physiological maintenance cost.

*Reproduction:* Reproduction in simplex is currently limited to clonal reproduction with the possibility of 
mutation. Individuals reproduce with a probability determined by the combination of mean cell quota and the 
proportional match to the environmental optima. The cell quota of each resource is evenly divided between the 
individual and its daughter. The daughter is given a unique individual ID and the species ID of its mother, 
unless in the case of speciation, but is allowed small mutations in individual-level state variables.

*Speciation:* Speciation is simulated within simplex as a discrete event but is accompanied by mutations in the 
values of species-level state variables. This allows for diversity to arise within the system, which the 
environmental filter can then select on.

*Death:* Individuals sampled at random will die if their smallest resource specific cell quota (i.e., N, C, P)
is equal to or less than 0.

*Emigration:* Individuals, resource particles, and inert tracers are considered to have left or to have flowed out 
when they pass beyond edges of the environment.

### Design concepts

**Basic principles.** 
*Ecological complexity*: simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.

*Nutrient limited growth*: All models assembled by simplex employ the universal concept that individual growth and activity is fueled and limited by resources.

*Resource diversity & heterogeneity*: simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.

**Theories**
*Constraint-based theory*: simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure. That is, the idea that both ecological processes and constraints shape ecological diversity.

*Chemostat theory*: simplex operates much like an unhinged bioreactor or chemostat. That is, particles flow through a system of a defined size at an average rate, and are limited in growth by their residence time.

*Ecological neutral theory*: simplex operates via random sampling and can vary from being completely neutral (all individuals having equal vital rates) to completely idiosyncratic (all individual and species are as different as possible). The one aspect of neutral theory that simplex adopts without question is the importance of stochastic life history processes (i.e. weighted or unweighted random fluctuations in population sizes).

**Hypotheses**
* These are entirely up to the user to formulate and test according to simplex's capabilities and analytical tools.

**Modeling approaches**
Simplex operations via two main modeling approaches, other than being individual-based, i.e., random sampling and computaitonal fluid dynamics.

**Emergence**
simplex uses random sampling and random assembly its models to avoid imposing strong constraints on the properties that emerge and to allow unanticipated combinations of traits and ecological structure to emerge.

* Abundance
	* Total community abundance
	* Trait-related population size
	* Effective population size
	* Abundance-biomass relation
* Community assembly
	* Species turnover
	* Biomass turnover
	* Extinction rate
	* Succession
* Community structure
	* Species richness
	* Species evenness
	* Trait diversity
* Population structure
	* Demography
	* Subpopulation trait variation
* Life history tradeoffs
	* Mobility vs. metabolic maintenance
	* Growth rate vs. metabolic maintenance
	* Generalist vs. specialist
	* R vs. K selection

**Adaptation** 
Individuals can move towards their environmental optima. Populations can become aggregated in areas that provide 
favorable intersections of species optima. Species canevolve by the action of the environmental filter on subpopulation variation in state variables.

**Objectives**
Individual seek conditions that match them to the environment (e.g., positions along environmental gradients). Individuals also seek to acquire resources through active searching. In the future, individuals will seek to avoid predation.

**Learning**
There is no aspect of individual-based learning in simplex, as of yet.

**Prediction**
Individuals in simplex do not have the ability to anticipate conditions.

**Sensing**
Individuals only sense in the sense that they can move towards environmental optima and, in the future, 
resources. Otherwise, all encounters are the result of random walks or fluid flow.

**Interaction**
At the moment, individuals only interact indirectly through excluding each other from resources (e.g. 
preemption). In the future, individuals will interact as predator-prey, mutualists, resource-dependents, etc. 
Likewise, there is currently no communication, though quorum sensing would be cool.

**Stochasticity**
The occurrence of nearly all processes of birth, death, life, immigration, dispersal, emigration, consumption, etc. are conducted via random sampling. In this way, population and community dynamics result, in part, from demographic 
stochasticity. Likewise, the emergence of life history traits proceeds from initially random combinations of traits.

**Collectives**
Individuals belong to species. Species belong to communities. In the future, simplex will allow communities to belong to trophic levels.

**Observation**
Many simplex models shoudl be run to examine trends in the variation generated. The following is recorded for each simplex model:

* Values of randomly chosen state variables
* Total abundance, $N$
* Species richness, $S$
* Avg growth rate (per species & per capita) 
* Avg maintenance cost (per species & per capita) 
* Avg resource efficiency (per species & per capita) 
* Avg active dispersal rate (per species & per capita) 
* Compositional turnover
	* Bray-Curtis
	* Sorensen's
* Species turnover
	* Whittaker's $\beta$ 
* Species evenness
	* Smith and Wilson's evenness, $E_{var}$
	* Simpson's evenness, $E_{1/D}$ 
* Species diversity
	* Shannon's diversity, $H'$ 
	* Simpson's diversity, $D_{1/D}$
* Dominance
	* Absolute, $N_{max}$
	* Relative, $N_{max}/N$
* Productivity
	* Individuals
	* Carbon, Nitrogen, Phosphorus
* Residence time
	* Ecosystem, $(length * width)/flowrate$
	* Individual, avg time spent in the system before dying or emigrating
	* Resource, avg time spent in the system before washing out or being consumed
	* Tracer, avg time spent in the system before washing out
* Individual residence time distribution (RTD)
* Resource particle RTD
* Tracer particle RTD

These data are stored in file as R-formatted data.frames. These files can be directly imported into an R or Python environment.

###Initialization
The model initiates with a random set of values for state-variables, 100 to 10,000 randomly drawn individuals from a 
theoretical log-series metacommunity. These values are saved, so that a simplex model could be programmed to replicate an analysis.

###Input data
simplex models require no input data, but it might be cool to use an api to grab environmental data or other 
data from a website to parameterize a simplex model.

###Submodels & Equations

**Cell quota model of Droop**
In simplex models, individuals grow according to their amounts of endogenous resources (cell quota).

Droop (1968, 1983) gave a relationship between specific growth rate ($\mu$) and cell quota ($Q$):

$$\mu = \mu'_{m} (1 - k_{q}/Q)$$

where $k_{q}$ is the minimum cell quota needed for life, also referred to as the subsistence quota. $\mu'$ is the

**Maintenance cost of Pirt**
Pirt (1965) states "The variation, with growth rate, of the yield of organism from the substrate used as 
energy source is attributed to consumption of energy at a constant rate for cell maintenance." He derives 
a relationships between the growth yield (biomass), the growth rate, and metabolic maintenance.

simplex models use Pirt's concept of a constant maintenance requirement. simplex also draws from 
Pirt's simple relation for substrate use:

$$use(total) = use(maintenance) + use(growth)$$

Respiration and activity without growth is not accounted for.

**Log-series metacommunity**
simplex models draw immigrating individuals from a theoretical log-series distribution.
Hubbell (2001) states that the regional community (i.e., metacommmunity) often has a log-series species 
abundance distribution. The probability density function (pdf) of the log-series is:

$$f(k) = -1/ln(1 - p) * p^{k}/k$$

Hubbell (2001) provides explicit detail of the log-series, which is also covered in most ecological diversity texts
and even on Wikipedia: https://en.wikipedia.org/wiki/Logarithmic_distribution

## Notes on simplex source code
simplex models operate primarily on lists in a programmatic way, e.g., quickly sorting lists, and removing and returning an element from lists with very little overhead. Likewise, simplex models generate and hold a lot of information about all the particles and elements in the system, which can become a computationally intensive task. To this end, simplex modeling coded is written in Python, an easy to read high-level programming language that has many scientific, plotting, and animation libraries. Python gives greater control over the operating system than data analysis languages (e.g. R, Matlab) that can be comparatively slow at purely computational tasks and can greatly limit the amount of memory held in any data object, and even fail to import large amounts of data. Python can also obtain C-like speeds when implementing certain software, e.g., Cython, Sage.

The output of simplex is a broad array of information held in seven .csv files. The most important of these is SimData.csv, and is intended to be analyzed in the freely available R (https://www.r-project.org/) and RStudio (https://www.rstudio.com/) environments. The R statistical computing language is well-suited to the analysis of simplex output and contains many packages for multivariate analysis and higher-order statistical analysis that Python is only beginning to accumulate. Consequently, we provide R source code in .R files and .Rmd (RMarkdown) files, complete with basic and advanced statistical analyses for analyzing diversity, regression models, ordination, variance partitioning, and for generating pdf documents (via Knitr) that integrate prose, code, and figures for manuscripts.

The reader can view example R-based analysis files that users can use to examine simplex's simulated data: https://github.com/LennonLab/simplex/tree/master/results/analyses.

## Speed & Memory

Simplex models do not complete until the time series of total abundance values reaches a state of mean reversion (i.e., stationarity). Because simplex models can range from quickly flowing eratic systems to barely flowing depleting systems, simulations can potentially take several minutes or more to complete. Likewise, the ability to simulate many complex scenarios also allows for very large total abundances, the resulting values of which are difficult to predict and can theoretically outstrip a computer's memory.

I ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory. This system probably represents a below average capacity for modern personal computers, which for this study, was desirable as simplex should be able to be ran on both personal computers and high capacity remote servers. I present results for time to completion and required memory in the Results.

## Animations

Simplex can generate animations of its models and store them in various image file formats. It does this using the matplotlib animation libary. At the moment, choosing whether to animate or not animate a model is done by commenting out lines of code in model.py. However, this feature will be developed more highly for publication.

#Results

## Unit tests

simplex passed all units tests for 15 diversity indices, ensuring that each index returns either the correct calculated value, or 'NaN' if given any values that cannot be used (e.g., negative numbers, string characters, empty lists).

## Speed & Memory

**Results 100 randomly assembled models.** On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31367.0 +- 878.92, and required 160.5 +- 4.28 megabytes of memory. The longest any model took to complete was 48.95 minutes. This model required 261MB of memory and also had generated the greatest total abundance (*N* = 68,808). The shortest any model took to complete was 5.8 seconds with a total abundance of 0 individuals. The least amount of memory any model required was 90MB. These and other analyses from the 100 randomly assembled models can be run using the SpeedMemory.Rmd file located in the results/TestResults directory.

## Output data
simplex generates six files as its output data. They are:

*SimData.csv*--Formatted as an R data frame, where each row is a run of a randomly assembled model, and each column holds a piece of data about the system that was modeled (e.g., flow rate, total abundance, species richness, species turnover, rate of disturbance, etc.).

*IndRTD.csv*, *ResRTD.csv*, *TracerRTD.csv*--Each line of these three files contains the amount of time that each individual, resource particle, or inert tracer (from beginning to end) spent in the system. Each line is a run of a randomly assembled model.

*RADs.csv*--Each line is a run of a randomly assembled model and contains the rank-abundance vector at the point when the model was stopped.

*Species.csv*--Each line is a run of a randomly assembled model and contains a vector of species labels corresponding to RADs.csv.

All output data were correctly formatted and placed within results/simulated_data/examples directory.
Each of the six output data files was able to be imported into the RStudio environment using the Exploratory.Rmd Rmarkdown file provided in the "GitHub/simplex/results/analyses/Rmd/" path. The user can use the Exploratory.Rmd file to craft a .Rmd file for their own specific analyses.

# Discussion

The simplex platform assembles individual-based models from random combinations of state-variables and processes. This is done to explore a large degree of variation in ecological conditions and complexity. In this way, simplex allows ecologists to examine whether the influence of a particular variable should be robust to a wide range of ecological conditions.

By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) simplex is well-positioned to keep developing. Future developments will include additional ecological dynamcis such as predator-prey, mutualism, and parasitism. Improvements to simplex will also include an explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity. For example, given the chemical structure of phosphate and the size of a resource particle, the particle can be assigned *x* units of phosphorus as well as the biocomplexity value of phosphate (estimated as a form of Shannon's entropy).

Likewise, improvements and future versions of simplex will provide increasing numbers of files for statistical analysis of simplex output. These files will, in a sense, be able to be used as a complete analysis of some particular aspect of simplex output (e.g., demographic, resource-related, spatial, multivariate, ecological model testing). Moreover, these and all simplex files are freely available and open source, allowing users to modify simplex source code to suit their particular needs.

# Acknowledgments

I thank Jay T. Lennon, Mario Muscarella, Xiao Xiao,

# References

1. Black AJ, McKane AJ (2012) Stochastic formulation of ecological models and their applications. Trends in Ecology and Evolution. 27:337-345.

2. DeAngelis DL, Gross LJ (1992) Individual-based Models and Approaches in Ecology: Populations, Communities, and Ecosystems. Chapman & Hall. New York.

3. Droop MR (1983) Twenty-five years of algal growth kinetics, a personal view. Botanica Marina. 26: 99-112.

4. Gotelli NJ, Entsminger GL (2001) Swap and fill algorithms in null model analysis: rethinking the knight’s tour. Oecologia. 129:281–291

5. Grimm V, Berger U, Bastiansen F, Eliassen S, Ginot V, Giske J, Goss-Custard J, Grand T, Heinz SK, Huse G, Huth A, Jepsen JU, Jørgensen C, Mooij WM, Mu¨ller B, Pe’er G, Pioub C, Railsback SF, Robbins AM, Robbins MM, Rossmanith E, Ru¨ger N, Strand E, Souissim S, Stillmane RA, Vabøg R, Visser U, DeAngelis DL (2006) A standard protocol for describing individual-based and agent-based models. Ecological Modeling. 198: 115-126.

6. Grimm V, Railsback SF (2005) Individual-based Modeling and Ecology. Princeton University Press, Princeton.

7. Hartl DL, Clark AG (1997) Principle of Population Genetics. Sinauer Associates, Inc. Publishers, Sunderland, MA.

8. Hubbell SP (2001) The unified neutral theory of biodiversity and biogeography. Princeton University Press, Princeton, NJ.

9. Leadbeater BSC (2006) The 'Droop equation' - Michael Droop and the legacy of the 'Cell-quota model' of phytoplankton growth. Protist. 157: 345-358.

10. Matthews RB, Gilbert NG, Roach A, Polhill JG, Gotts NM (2007) Agent-based land-use models: a review of applications. Landscape Ecology. 22: 1447-1459.

11. Pirt SJ (1965) The Maintenance Energy of Bacteria in Growing Cultures. Proceedings of the Royal Society of London. Series B, Biological Sciences, 163:224-231.

12. Rosindell J, Harmon LJ, Etienne RS (2015) Unifying ecology and macroevolution with individual-based theory. Ecology Letters. 18: 472-482.

13. Succi S (2001) The Lattice Boltzmann Equation: For Fluid Dynamics and Beyond. Oxford University Press. New York.

### 2015-10-19 paper/PrePrint.Rmd

@@
 title: "simplex: An individual-based modeling platform"
-author: "Kenneth J Locey,
+author: "TBD,
 Dept. of Biology, University of Indiana, Bloomington"
@@
 
-Here, I present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output.
+Here, we present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output.
 This platform incorporates computational fluid dynamics (Succi 2001), life history processes and concepts and patterns from biogeography (Hubbell 2001, Bell 2001), evolution and aspects of selection (Hartl and Clark 1997), and integrates nutrient-limited growth and physiology (Pirt 1965, Droop 1983) among dimension of ecology and evolutionary science.
-I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes.
-Below, I provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.
+We constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes.
+Below, we provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.
 
@@
 
-**Source code**--To save space, I also refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
+**Source code**--To save space, we also refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
 
@@
 The ODD protocol is standard for describing individual-based models (Grimm et al. 2006). 
-ODD stands for Overview, Design concepts, and Details. Here, I descibe simplex *largely* according to the ODD protocol.
+ODD stands for Overview, Design concepts, and Details. Here, we descibe simplex *largely* according to the ODD protocol.
 The ODD protocol for simplex can also be found on the public GitHub repository at https://github.com/LennonLab/simplex
@@
 
-I ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
+We ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
 This system probably represents a below average capacity for modern personal computers, which for this study, was desirable as simplex should be able to be ran on both personal computers and high capacity remote servers.
-I present results for time to completion and required memory in the Results.
+We present results for time to completion and required memory in the Results.
 
@@
 
-I thank Jay T. Lennon, Mario Muscarella, Xiao Xiao,
+We thank Mario Muscarella and Xiao Xiao for fruitful discussions. **Grant number?**
 

### 2015-10-20 results/analyses/Rmd/Univariate.Rmd
modified PrePrint.md, added new .Rmd files for helping researcher conduct analyses on simplex output -KL

---
title: "Univariate analyses of simplex output"
author: "Ken Locey"
date: "October, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## OVERVEW

This R Markdown document is designed to be opened and ran in the RStudio. The chunks of code below allow for univariate (i.e., x vs. y) relationships between variable in the output of **simplex** models.

## SETUP
## A. Clear and set the working directory

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/simplex") 
```

## B. Import packages; install if needed

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
library(ggplot2)
```

## C. Import custom modules of R functions

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/simplex/tools/Rbin/metrics.R")
```

## D. Import simulated data generated by simplex models

As it is iterating through randomly assembled models, simplex writes its output to six .csv files. For each file, each row corresponds to a single model. Consequently, the *i*th row in each file corresponds to *i*th model that was assembled and run by simplex.
Let's import the simulated data files.

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}

# 1. A table where each columns corresponds to a state variable or model output.  
dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")

# 2. Three files where each element of a given row corresponds to the amount of time an
# individual spent in the system.
ind.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/IndRTD.csv")
res.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/ResRTD.csv")
tracer.rtds <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/TracerRTD.csv")

species.list <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/Species.csv")

rads <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/RADs.csv")
```

## UNIVARIATE ANALYSES

## 1.) Does simplex produce realistic univariate relationships?
Let's begin by asking how species richness (*S*) and species evenness relate to total abundance (*N*). Usually, we would expect to find greater numbers of species when sampling greater numbers of individuals, i.e., a positive relationship between *N* and *S*. Do simplex models generally conform to this simple prediction? What other simple or straightforward relationships should we expect to find?

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
N <- dat$total.abundance[dat$total.abundance<=0] <- 1
S <- dat$species.richness[dat$species.richness<=0] <- 1

ggplot(dat, aes(x=dat$total.abundance, y=dat$species.richness)) +
  geom_point(colour="blue") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x= "Total abundance", y="Species richness")
```

## 2.) Graphical exploration: What lesser known relationships does simplex suggest we should observe?
We might also expect to find that species evenness decreases with *N*, i.e., larger ecological communities have a less equitable distribution of abundance. Do simplex models produce these relationships?

```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}

ggplot(dat, aes(x=dat$total.abundance, y=dat$e.var)) +
  geom_point(colour="red") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x= "Total abundance", y="Species evenness")
```

### 2015-10-20 paper/PrePrint.Rmd

@@
 title: "simplex: An individual-based modeling platform"
-author: "TBD,
+author: "Kenneth J. Locey,
 Dept. of Biology, University of Indiana, Bloomington"
@@
 
-**Source code**--To save space, we also refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
+**Source code**--To save space, I refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
 
@@
 The ODD protocol is standard for describing individual-based models (Grimm et al. 2006). 
-ODD stands for Overview, Design concepts, and Details. Here, we descibe simplex *largely* according to the ODD protocol.
+ODD stands for Overview, Design concepts, and Details. Here, I descibe simplex *largely* according to the ODD protocol.
 The ODD protocol for simplex can also be found on the public GitHub repository at https://github.com/LennonLab/simplex
@@
 Simplex models do not complete until the time series of total abundance values reaches a state of mean reversion (i.e., stationarity).
-Because simplex models can range from quickly flowing eratic systems to barely flowing depleting systems, simulations can potentially take several minutes or more to complete.
-Likewise, the ability to simulate many complex scenarios also allows for very large total abundances, the resulting values of which are difficult to predict and can theoretically outstrip a computer's memory.
+Because simplex models can range from quickly flowing systems with high disturbance to barely flowing and highly stable but depleted systems, simulations can potentially take several minutes or more to complete.
+Likewise, the ability to simulate many complex scenarios also allows for very large total abundances, the values of which cannot be predicted *a priori* and can even potentially outstrip a computer's memory.
 
-We ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
+I ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
 This system probably represents a below average capacity for modern personal computers, which for this study, was desirable as simplex should be able to be ran on both personal computers and high capacity remote servers.
-We present results for time to completion and required memory in the Results.
+I present results for time to completion and required memory in the Results.
 
@@
 At the moment, choosing whether to animate or not animate a model is done by commenting out lines of code in model.py.
-However, this feature will be developed more highly for publication.
+However, this feature will be developed more highly for future versions.
 
@@
 
-**Results 100 randomly assembled models.** On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31367.0 +- 878.92, and required 160.5 +- 4.28 megabytes of memory.
-The longest any model took to complete was 48.95 minutes.
-This model required 261MB of memory and also had generated the greatest total abundance (*N* = 68,808).
+**Results from 100 randomly assembled models.** On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31,367 +- 879, and required 161 +- 4.3 megabytes of memory.
+The longest any model took to complete was 49 minutes.
+This particularly demanding model required 261MB of memory and also had generated the greatest total abundance (*N* = 68,808).
 The shortest any model took to complete was 5.8 seconds with a total abundance of 0 individuals.
@@
 
-We thank Mario Muscarella and Xiao Xiao for fruitful discussions. **Grant number?**
+I thank Jay T.
+Lennon, Mario Muscarella, Xiao Xiao, and Ed Hall for fruitful discussions.
+I thanks Dr.
+Daniel V.
+Schroeder for his web-based fluid dynamic simulator, the python-based code for which serves as the core method in simplex for simulating fluid flow.
 

### 2015-10-20 results/analyses/Rmd/Diversity.Rmd

--- 2015-10-05 results/analyses/Rmd/Exploratory.Rmd
@@
 
+### 2.) How do species richness (S) and total abundance (N) relate to ecosystem residence time?
+
+```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
+tau <- (dat$width * dat$height)/dat$flow.rate
+
+ggplot(dat, aes(x=tau, y=dat$total.abundance)) +
+  geom_point(colour="red") +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(x= "Residence time", y="Total abundance")
+
+ggplot(dat, aes(x=tau, y=dat$species.richness)) +
+  geom_point(colour="blue") +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(x= "Residence time", y="Species richness")
+```
+
 ### 2.) Variance partitioning

### 2015-10-26 tools/analyses/Rmd/BasicAnalysis.Rmd
modified metrics.R, added .Rmd files for basic statistical and diversity analyses -KL

---
title: "Basic Analysis of Simplex output"
author: "Ken Locey"
date: "October 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## OVERVEW

This R Markdown document is designed to be opened and ran in the RStudio. The chunks of code below allow for basic analysis of **simplex** output. This includes univariate and multivariate relationships, and graphical exploration.

For an R-based introduction to multivariate analysis: https://little-book-of-r-for-multivariate-analysis.readthedocs.org/en/latest/src/multivariateanalysis.html

## SETUP
## A. Clear and set the working directory

```{r, results = 'hide', echo=T, message = F, warning = F}
rm(list=ls()) 
getwd()
setwd("~/GitHub/simplex") 
```

## B. Import packages; install if needed

```{r, results = 'hide', echo=T, message = F, warning = F}
#install.packages("vegan") # Example of how an install can be done
require("vegan") 
require("car")
```

## C. Import simulated data generated by simplex models

As it is iterating through randomly assembled models, simplex writes its output to six .csv files. For each file, each row corresponds to a single model. Consequently, the *i*th row in each file corresponds to *i*th model that was assembled and run by simplex. Let's import the simulated data files.

```{r, results = 'hide', echo=T, message = F, warning = F}
# A table where each columns corresponds to a state variable or model output.  
simplex.dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")
simplex.dat$h.tau <- log((simplex.dat$width * simplex.dat$height)/simplex.dat$flow.rate)

# Replacing 0's with 1's to allow log-transforms below
simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 1
simplex.dat$N.max[simplex.dat$N.max<=0] <- 1
simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 1
simplex.dat$resource.particles[simplex.dat$resource.particles<=0] <- 1
simplex.dat$resource.concentration[simplex.dat$resource.concentration<=0] <- 1

simplex.dat$total.abundance <- log(simplex.dat$total.abundance)
simplex.dat$N.max <- log(simplex.dat$N.max)
simplex.dat$species.richness <- log(simplex.dat$species.richness)
simplex.dat$resource.particles <- log(simplex.dat$resource.particles)
simplex.dat$resource.concentration <- log(simplex.dat$resource.concentration)
simplex.dat[is.na(simplex.dat)] <- 0
```

## UNIVARIATE ANALYSES

Perhaps we want to ask whether **simplex** produces (i) relationships that are well-known to occur in ecological systems (ii) auto-correlated relationships between independent variables (iii) or novel relationships of use to a specific study or question. Each of these questions has its particular use, but rather than generate one x-y relationship after another, we generate an entire field of relationships. That is, given x variables, we can explore x*(x-1) x-y relationships.

Let's begin by examining relationships between some physical variables. As we'll see, the only physical variables that appear to be correlated are ecosystem residence time and the rate of flow.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}

# Physical and metacommunity variables
phys.dat <- as.matrix(subset(simplex.dat, 
                      select = c(h.tau,
                                 flow.rate,
                                 barriers,
                                 width,
                                 height,
                                 amplitude,
                                 frequency,
                                 phase)))

scatterplotMatrix(phys.dat, main="Physical data", diagonal = "none")
```

Next, let's explore relationships between some resource-related variables. As we'll see, the number of resource particles, their concentration, resource richness or the number of resource types, and resource diversity (a combination of resource richness and the variance in abundance among resources) are all highly and positively correlated.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
# Resource variables
res.dat <- as.matrix(subset(simplex.dat, 
                      select = c(resource.concentration,
                      shannons.resource.diversity,
                      resource.richness,
                      resource.particles)))

scatterplotMatrix(res.dat, main="Resource data", diagonal = "none")
```

We can also explore relationships between physiological variables, noticing that, biomass production in Carbon, Nitrogen, and Phosphorus are all positively and strongly correlated. This is largely because **simplex** (as of 26 October 2015) does not include any explicit stoichiometry. Also notice that an interesting life history trade-off arises in **simplex** however, that is, increasing growth rate leads to increasing metabolic maintenance.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
# Physiological variables
physio.dat <- as.matrix(subset(simplex.dat, 
                      select = c(biomass.prod.N,
                      biomass.prod.C,
                      biomass.prod.P,
                      avg.per.capita.N.efficiency,
                      avg.per.capita.P.efficiency,
                      avg.per.capita.C.efficiency,
                      max.growth.rate,
                      max.met.maint)))

scatterplotMatrix(physio.dat, main="Physiological data", diagonal = "none")
```

Finally, let's explore relationships between some diversity related variables. As we can see different evenness measure basically reflect each other, as do different measures of species turnover. Likewise, we see a strong postivie relationships between total abundance and species richness and between total abundance and the abundance of the most abundant species; which we should expect.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
diversity.dat <- as.matrix(subset(simplex.dat, 
                      select = c(total.abundance,
                      N.max,
                      species.richness,
                      simpson.e,
                      e.var,
                      Whittakers.turnover,
                      Jaccards.dissimilarity)))

scatterplotMatrix(diversity.dat, main="Diversity data", diagonal = "none")
```

## Summary statistics

We can use the 'sapply' function to generate summary statistics (mean, variance, etc.) column by column for any of our **simplex** output. For example, analyzing phys.dat as a data.frame:

```{r, results='show', echo=T, message=F, warning=F}
sapply(as.data.frame(phys.dat), mean) # sample mean
sapply(as.data.frame(phys.dat), var) # sample variance
sapply(as.data.frame(phys.dat), sd) # sample standard deviation
sapply(as.data.frame(phys.dat), median) # sample median
```

## HIGHLY CORRELATED VARIABLES

Suppose we want to find the most highly correlated variables in our data. First, lets define a function to return the x most highly correlated variables in a dataframe.

```{r, results='hide', echo=F, message=F, warning=F}
# This function was obtained from: https://little-book-of-r-for-multivariate-analysis.readthedocs.org/en/latest/src/multivariateanalysis.html

mosthighlycorrelated <- function(mydataframe, numtoreport)
  {
     # find the correlations
     cormatrix <- cor(mydataframe)
     # set the correlations on the diagonal or lower triangle to zero,
     # so they will not be reported as the highest ones:
     diag(cormatrix) <- 0
     cormatrix[lower.tri(cormatrix)] <- 0
     # flatten the matrix into a dataframe for easy sorting
     fm <- as.data.frame(as.table(cormatrix))
     # assign human-friendly names
     names(fm) <- c("First.Variable", "Second.Variable","Correlation")
     # sort and print the top n correlations
     head(fm[order(abs(fm$Correlation), decreasing=T),], n=numtoreport)
  }
```

Then, let's call our function to find the x most highly correlated variables.

```{r, results='show', echo=T, message=F, warning=F}
dat <- cbind(phys.dat, physio.dat, res.dat, diversity.dat)
mosthighlycorrelated(dat, 15)
```

## Variance partitioning.

A common question to ask is whether physical variables (geograhic distance, area, volume, flow rate, etc.) has a larger influence on ecological diversity than, say, resource-related variables.

One way to address this question is to use variance partitioning. A note of caution, typing 'help(varpart)' reveals that R only uses (simple) linear regression when there is one response variable and uses redundancy analysis ordination (RDA) for two or more response variables.

Executing the code below shows that, in **simplex** models, the phys variables explain more variation in total abundance, and species richness, evenness, and turnover than do resource related variables.

```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}
diversity.dat <- as.matrix(subset(simplex.dat, 
                      select = c(total.abundance,
                      species.richness,
                      simpson.e,
                      Whittakers.turnover)))

rda.phys <- rda(diversity.dat ~ phys.dat)
rda.phys
rda.res <- rda(diversity.dat ~ res.dat)
rda.res

# Test significance of the fractions using RDA:
phys.anova <- anova(rda.phys, step=200, perm.max=200)
phys.anova
res.anova <- anova(rda.res, step=200, perm.max=200)
res.anova

# Variance partitioning with two explanatory matrices
mod <- varpart(Y=diversity.dat, X=phys.dat, res.dat)
mod
showvarparts(2)
plot(mod)
mtext("Variation partitioning")
text(1.1, 0.5, "env")
text(-0.1, 0.5, "phys") 
```

### 2015-10-26 tools/analyses/Rmd/Diversity.Rmd
modified metrics.R, added .Rmd files for basic statistical and diversity analyses -KL

---
title: "Diversity analysis of simplex output"
author: "Ken Locey"
date: "October, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## OVERVEW
This R Markdown document is designed to be opened and ran in the RStudio. The chunks of code below allow for diversity analysis of **simplex** output. This includes analyzing trends in univariate diversity metrics and analyzing distributions of abundance among species. Note that **simplex** tracks and/or calculates the values of several diversity related metrics. These include:

1.) Total abundance (*N*): Total number of individuals  
2.) Absolute dominance (*Nmax*): Number of individuals belonging to the most
abundant species  
3.) Species richness (*S*): Total number of species  
4.) Simpson's evenness: Simpson's measure of species evenness  
5.) Smith and Wilson's evenness: a normalized log-transform of the sample
variance  
6.) Whittakers species turnover: Unweighted proportional turnover among species between two samples.  
7.) Jaccard's pair-wise dissimilarity: Weighted proportional turnover among species between two samples. See: http://cc.oulu.fi/~jarioksa/softhelp/vegan/html/vegdist.html

## SETUP
### A. Clear and set the working directory
```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/simplex") 
```

### B. Import packages; install if needed
```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan")
require("car")
```

### C. Import custom modules of R functions
```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
source("~/GitHub/simplex/tools/Rbin/metrics.R")
```

### D. Import simulated data
```{r, results = 'hide', echo=TRUE, message = FALSE, warning = FALSE}
# A table where each columns corresponds to a state variable or model output.  
simplex.dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")
simplex.dat$h.tau <- log((simplex.dat$width * simplex.dat$height)/simplex.dat$flow.rate)

# Replacing 0's with 1's to allow log-transforms below
simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 1
simplex.dat$N.max[simplex.dat$N.max<=0] <- 1
simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 1

simplex.dat$total.abundance <- log(simplex.dat$total.abundance)
simplex.dat$N.max <- log(simplex.dat$N.max)
simplex.dat$species.richness <- log(simplex.dat$species.richness)
simplex.dat[is.na(simplex.dat)] <- 0

# Two files, where the ith element of the jth row correspond between files
# and where jth row correspond to the jth row of the above three files
species.list <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/Species.csv", type = 'char')
rads <- get.vectors("~/GitHub/simplex/results/simulated_data/examples/RADs.csv", type = 'number')
```

## DIVERSITY ANALYSIS

### A. Graphical exploration via kernel density estimation

Each row in our **simplex** output files corresponds to a model with an average total abundance, species richness, species turnover, etc. One way of developing a feel for the values of these variables simulated by **simplex** is through the use of kernel density estimation (kde). In short, kde produces a smoothed histogram where integrating between two points on the x-axis gives the probability that a randomly chosen value will occur within that range. Let's plot kernel density curves for log10-transformed total abundance, log10-transformed species richness, Simpson's evenness metric, and Whittaker's measure of species turnover.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=6, fig.height=4}
par(mfrow=c(2, 2))

Sx <- density(simplex.dat$species.richness)
Nx <- density(simplex.dat$total.abundance)
Ex <- density(simplex.dat$simpson.e)
Tx <- density(simplex.dat$Whittakers.turnover)

plot(Nx, main = "Total abundance", col = 'DarkCyan')
plot(Sx, main = "Species richness", col = 'magenta')
plot(Ex, main = "Species evenness", col = 'dodgerblue')
plot(Tx, main = "Species turnover", col = 'seagreen')
```

As we can see, there are some interesting aspects of modality, variance, and skewness. For instance, both total abundance and species richness are characterized by two modes, one at very low values (near log10 of 0) and one at higher values (near a few hundred or several thousand). Hint: graphing both *N* and *S* against residence time, (height*width)/flow rate, would reveal why.

Likewise, we can see that species turnover is largely concentrated at low values but has a long tail extending to higher values (right-skewed). Here, species turnover is the portion of species differing between two samples. We can also see that evenness ranges across its full breadth, from 0.0 (no evenness) to 1.0 (perfect evenness). Notice that because kernel density curves are smoothed and based on probabilities, the tails will extend beyond the range of the data, i.e., unless parameters like the kernel width are adjusted. type 'help(density)' for greater explanation.

### Univariate relationships
Let's explore relationships between diversity related variables. As we can see different evenness measure basically reflect each other, as do different measures of species turnover. Likewise, we see a strong postivie relationships between total abundance and species richness and between total abundance and the abundance of the most abundant species; which we should expect.

```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
diversity.dat <- as.matrix(subset(simplex.dat, 
                      select = c(total.abundance, 
                      N.max,
                      species.richness, 
                      simpson.e, 
                      e.var, 
                      Whittakers.turnover, 
                      Jaccards.dissimilarity)))

scatterplotMatrix(diversity.dat, main="Diversity data", diagonal = "none")
```

### Rank-abundance curves (RACs)

This section pull heavily from the Quantative Biodiversity course designed and taught by Jay Lennon, myself (Ken Locey), and Mario Muscarella.
https://github.com/QuantitativeBiodiversity/Assignments/tree/master/Alpha

While the diversity metrics can be useful in compressing one or more aspects of diversity into a single number, information is invariably lost in the process. 
More specifically, each metric quantifies an aspect of the vector of species abundances, whether that be the number of species (richness) or similarity in abundance among species (evenness).

Consequently, we can also analyze the ranked vector of species abundances resulting from **simplex** models. This vector is simply known as the rank-abundance curve (RAC) and its form is predicted by no less than 20 theories or models of ecology/biodiversity! In fact, the uneven shape of the RAC is one of the most intensively studied patterns in ecology, and underpins all or most ecological theories of biodiversity. Predicting the form of the RAC is the first test that any biodiversity theory must pass and there are there are dozens of models that have attempted to explain the uneven form of the RAC across ecological systems. These models attempt to predict the form of the RAC according to mechanisms and processes that are believed to be important to the assembly and structure of ecological systems.

Let's use the `radfit()` function in `vegan` to fit the predictions of various species abundance models to the RAC of a chosen **simplex** model output (i.e. a row in the RADs.csv file). Descriptions of several species abundance models (as used in vegan) are provided below.

First, choose a simplex model with a relatively large number of species. You can use the `sapply` function to find the length of each rad.

```{r, results='hide', echo=T, message=F, warning=F}
sapply(rads, length)
```

Now, lets use the `radfit` function to fit several species abundance models.
```{r, results='hide', echo=T, message=F, warning=F}
RACresults <- radfit(rads[[42]])
RACresults
```

We see that `vegan` fits five models to our RAC: *Null*, *Preemption*, *Lognormal*, *Zipf*, and *Mandelbrot*. Before explaining what these models represent, let's run through the `vegan` output:

1.  Next to "RAD models", we see "family poisson", which tells us that by default, `vegan` assumes Poisson distributed error terms.
2.  Below this, we see that `vegan` returns the number of species (*S*) and the number of individuals (*N*) for the empirical RAC.
3.  Next, we see a table of information, the first columns of which are par1, par2, and par3.
These columns pertain to model parameters and reveal that the different models use different numbers of parameters; the null model uses none.
4.  Next, we see a column for Deviance, which is a quality of fit statistic based on the idea of residual sums of squares.
5.  After Deviance, we see columns for AIC and BIC, which are the estimated **Akaike Information Criterion** and the **Bayesian Information Criterion**, respectively.

*Notes on AIC and BIC* AIC and BIC are commonly used for model selection. 
In other words, they help us identify a model that is best supported by our data. 
Obviously, the more parameters a model has, the better it will fit a data set. 
However, it's not necessarily desirable to have an over-parameterized model. 
So, AIC and BIC asssign penalties that correspond with the number of parameters that a model uses. In the end, the "best" model has the lowest AIC or BIC value.

Now, let's visualize our results by plotting the empirical RAC and the predicted RAC for each model:

```{r, results='hide', echo=T, message=F, warning=F}
plot(RACresults, las=1, cex.lab = 1.4, cex.axis = 1.25)
```

### Interpreting the RAC models in `vegan`:

#### Null:
A **broken stick model** (MacArthur 1960, Pielou 1975) where the expected abundance of a species at rank *r* is $a_r = \frac{N}{S} \cdot \sum\limits_{x=r}^{S}\frac{1}{x}$.
*N* is the total number of individuals and *S* is the total number of species.
This gives a constraint-based null model where the *N* individuals are randomly distributed among *S* species, and there are no fitted parameters.
Null models often reveal that realistic patterns can be expected from random sampling, and have been extremely useful in ecology and evolution.

#### Preemption:
The **niche preemption model** (a.k.a., geometric series or Motomura model): 
Envision an environment occupied by a single species. 
Now, envision that a second species colonizes the environment and takes some portion of resources equal to $\alpha$. 
Then, envision that a third species colonizes the environment and takes a portion of resources equal to $\alpha$ away from the second species.
Imagine this process continuing until *N* is zero.
The only estimated parameter is the preemption coefficient $\alpha$, which gives the decay rate of abundance per rank. 
The expected abundance (*a*) of species at rank *r* is $a_r = N \cdot \alpha \cdot (1 - \alpha)^{(r - 1)}$.

#### Lognormal:
Many statistical models assume that the distribution of values are normally distributed. 
When applied to species abundances, this means that they conform to the shape of a symmetrical bell curve, more precisely known as the Gaussian distribution.
In contrast, the log-Normal model assumes that the logarithmic abundances are normally distributed.
The expected abundance of a species at rank *r* is then: $a_r = e^{log(\mu) + log(\sigma) \cdot Z }$, where *Z* is the Normal deviate. 
A Normal deviate is simply the number of standard deviations a score is from the mean of its population.
The log-normal model was introduced into ecology by Frank Preston in 1948 and is one of the most widely successful species abundance models.

#### Zipf:
The Zipf model is based on Zipf's Law, an well-known observation that many types of ranked data are fit by a simple scaling law (a.k.a., power law).
In short, the abundance of a species in the RAC is inversely proportional to its rank.
The expected abundance (*a*) of species at rank *r* is: $a_r = N \cdot p_1 \cdot r^{\gamma}$, where $\it{p_1}$ is the fitted proportion of the most abundant species, and $\gamma$ is a decay coefficient.

#### Mandelbrot:
Shortened name for the Zipf–Mandelbrot model, a generalization of the Zipf model made by the mathematician and father of fractal geometry, Benoit Mandelbrot.
This model adds one parameter ($\beta$) to the Zipf model.
The expected abundance of a species (*a*) at rank *r* is $a_r = N \cdot c \cdot (r + \beta)^{\gamma}$. 
Here, the $\it{p_1}$ parameter of the Zipf model changes into a meaningless scaling constant *c*.

Let's go even further and ask which model, on average, best fits the RACs of **simplex** models. Let's begin by making a data frame of the results.

```{r, results='hide', echo=T, message=F, warning=F}
RAC.dat.frame <- data.frame(BS=double(),GS=double(),LN=double(),
                            Zp=double(),ZM=double(),stringsAsFactors=F)

i = 1
for(rad in rads){  
  # Not analyzing any model resulting in less than 6 species
  if(length(rad) > 5){

    RACresults <- radfit(rad)
    # get the BIC scores for each model

    # Broken-stick (vegan's null)
    BS <- as.numeric(RACresults[[3]][[1]][6])
    # Geometric series
    GS <- as.numeric(RACresults[[3]][[2]][6])
    # Preston's lognormal
    LN <- as.numeric(RACresults[[3]][[3]][6])
    # Zipf
    Zp <- as.numeric(RACresults[[3]][[4]][6])
    # Zipf-Mandelbrot
    ZM <- as.numeric(RACresults[[3]][[5]][6])

    irow <- c(i, BS, GS, LN, Zp, ZM)
    RAC.dat.frame <- rbind(RAC.dat.frame, irow)
    i = i+1
  }
}

RAC.dat.frame <- setNames(RAC.dat.frame, 
          c("row", "BS","GS","LN","Zp","ZM"))
```

Remember, the "best" fit model has the lowest AIC or BIC value.

```{r, results='show', echo=T, message=F, warning=F}
par(mfrow=c(1, 1))

print(c("Broken-stick", mean(RAC.dat.frame$BS), median(RAC.dat.frame$BS)))

print(c("Geometric-series", mean(RAC.dat.frame$GS), median(RAC.dat.frame$GS)))

print(c("Lognormal",mean(RAC.dat.frame$LN),median(RAC.dat.frame$LN)))

print(c("Zipf",mean(RAC.dat.frame$Zp),median(RAC.dat.frame$Zp)))

# Appears to be an "NA" in the ZM data
ZM <- RAC.dat.frame$ZM[!is.na(RAC.dat.frame$ZM)]
print(c("Zipf-Mandelbrot", mean(ZM), median(ZM)))
```

It appears that the Zipf-Mandelbrot distribution is the winner with respect to the mean and median BIC values. But, a word of caution, with the exception of the null model (i.e. the Broken-stick) these SAD models are not maximum-likelihood estimates. This means that `vegan` does not return exact forms of, say, the lognormal, Zipf, etc., but instead returns sort of a best-fit comprimise between the model and the data.

### 2015-10-26 paper/PrePrint.Rmd
modified PrePrint.Rmd and generated new PrePrint.pdf -KL

@@
 ---
-title: "simplex: An individual-based modeling platform"
+title: "simplex: An individual-based ecological modeling platform"
 author: "Kenneth J. Locey,
@@
 
-Here, we present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output.
+Here, I present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output.
 This platform incorporates computational fluid dynamics (Succi 2001), life history processes and concepts and patterns from biogeography (Hubbell 2001, Bell 2001), evolution and aspects of selection (Hartl and Clark 1997), and integrates nutrient-limited growth and physiology (Pirt 1965, Droop 1983) among dimension of ecology and evolutionary science.
-We constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes.
-Below, we provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.
+I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes.
+Below, I provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.
 
@@
 
-**Overview**--simplex was developed to accomplish three objectives.
+**Overview**--I developed simplex to accomplish three objectives.
 First, simplex assembles and runs individual-based ecological models from random combinations of state variables and processes.
-Second, simplex stores model output as R-style data frames and, when the user chooses, generates and stores animations of models.
-Third, simplex provides R-based files for analyzing simulated; these will eventually be offered in Python as well.
+Second, simplex stores the output of models as R-style data frames and .csv files, and when the user chooses, generates and stores animations of models.
+Third, simplex provides R Markdown files for analyzing simulated data; these will eventually be offered in Python as well.
 
@@
 *Constraint-based theory*: simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure.
-That is, the idea that both ecological processes and constraints shape ecological diversity.
+That is, the idea that both ecological processes and physical constraints shape ecological diversity.
 
@@
 
-## Output data
+## Products
+### Output data
 simplex generates six files as its output data. They are:
@@
 
-*IndRTD.csv*, *ResRTD.csv*, *TracerRTD.csv*--Each line of these three files contains the amount of time that each individual, resource particle, or inert tracer (from beginning to end) spent in the system.
-Each line is a run of a randomly assembled model.
+*IndRTD.csv*, *ResRTD.csv*, *TracerRTD.csv*--Each line of these three files contains the amount of time that each individual, resource particle, or inert tracer spent in the system.
+Each line represents the same run of a randomly assembled model.
 
@@
 
+### R analysis files
+
+There are currently two R Markdown analysis files associated with simplex; these are located in the `tools/analyses/Rmd` directory. 
+These files are primarily designed to help users understand how to import simplex files, to use R Markdown to analyze simplex data, and conduct descriptive and univariate statistics on simplex output.
+The first file, `BasicAnalysis.Rmd`, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among simplex variables.
+The second file, `Diversity.Rmd`, focuses on the analysis of taxonomic diversity and abundance to include fitting species abundance models using the R-based `vegan` module.
+Each file generates a knitted pdf (includes code, text, and figures), see Supplement.
+
+### Animations
+
+I used the animation module in the matplotlib plotting library to generate avi formatted movies of simplex models.
+These can be viewed in any internet brower (https://github.com/LennonLab/simplex/tree/master/results/movies/examples).
+Currently, generating movies requires following three instructions in model/model.py:
+
+1.) uncomment line 'ani.save' on or near line 364  
+2.) adjust the frames variable on or near line 364, to change movie length   
+3.) change plot_system = 'no' to 'yes' on or near line 66
+
+Because generating animations requires computing time and memory, doing so can be computationally demanding. To quicken the process, comment out plt.show() on or near line 368.
+
 # Discussion
@@
 This is done to explore a large degree of variation in ecological conditions and complexity.
-In this way, simplex allows ecologists to examine whether the influence of a particular variable should be robust to a wide range of ecological conditions.
+In this way, simplex allows ecologists or educators to examine whether the influence of a particular variable should be robust to a wide range of ecological conditions.
+In addition to its unique  modeling aspects, simplex includes analysis code that shows the user how to import and analyze the simulated data, and provides a basic demonstration of how to conduct variance partitioning and fitting of species abundance models.
 
-By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) simplex is well-positioned to keep developing.
-Future developments will include additional ecological dynamcis such as predator-prey, mutualism, and parasitism.
+By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) simplex is well-positioned to continue development as both a research and education/instructional tool for ecology.
+Future developments and released versions will include additional ecological dynamcis such as predator-prey, mutualism, and parasitism.
 Improvements to simplex will also include an explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity.
@@
 These files will, in a sense, be able to be used as a complete analysis of some particular aspect of simplex output (e.g., demographic, resource-related, spatial, multivariate, ecological model testing).
-Moreover, these and all simplex files are freely available and open source, allowing users to modify simplex source code to suit their particular needs.
+These and all simplex files are freely available and open source, allowing users to modify simplex source code to suit their particular needs.
 
@@
 Lennon, Mario Muscarella, Xiao Xiao, and Ed Hall for fruitful discussions.
-I thanks Dr.
+I thank Dr.
 Daniel V.

### 2015-10-26 model/model.py
modified model.py, PrePrint.Rmd, and generated new PrePrint.pdf -KL

@@ -23,6 +23,15 @@ import bide
 sys.path.append(mydir + "GitHub/simplex/tools/randparams")
 import randparams as rp
 
+""" To generate movies:
+1.) uncomment line 'ani.save' on or near line 364
+2.) adjust the frames variable on or near line 364, to change movie length
+3.) change plot_system = 'no' to 'yes' on or near line 66
+
+Because generating animations requires computing time and memory, doing so can
+be computationally demanding. To quicken the process, comment out plt.show() on
+or near line 368.
+"""
 
 # https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing
 # http://statsmodels.sourceforge.net/0.5.0/generated/statsmodels.tsa.stattools.adfuller.html

### 2015-10-26 paper/PrePrint.Rmd

@@
 ---
+
+\pagebreak
+
+# Abstract
+
+A popular approach in ecology explicitly simulates the lives of individual organisms.
+Individual-based modeling (IBM) allows researchers to study ecological complexity as it emerges from individual-level interactions, to explicitly synthesize ecological theories and principles, and to track and record virtually any aspect of a system.
+I developed a novel IBM platform called simplex to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental dynamics, and community assembly and structure in both fluid and non-fluid environments.
+simplex accomplishes three primary tasks in keeping with its namesake, i.e., a generalization of the notion of a triangle to arbitrary dimensions.
+First, simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables.
+Second, simplex stores the output of models and generates and stores animations.
+Third, simplex provides R Markdown files for analyzing simulated data.
+The simplex source code is freely available, includes the standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to be run on a modest laptop setup using freely available open-source tools.
+
+\pagebreak
 

### 2015-10-26 paper/PrePrint.Rmd
modified PrePrint.Rmd and generated new PrePrint.pdf -KL

--- 2015-10-26 paper/PrePrint.Rmd
@@
 ---
-title: "simplex: An individual-based ecological modeling platform"
-author: "Kenneth J. Locey,
-Dept. of Biology, University of Indiana, Bloomington"
+title: "simplex: An individual-based platform for ecological modeling"
+author: "Kenneth J. Locey"
+fontsize: 12pt
 geometry: margin=2.54cm
@@
 ---
+
+\begin{center} 
+Department of Biology, University of Indiana, Bloomington
+\end{center}
+
+\begin{center}
+email: kjlocey@indiana.edu
+\end{center}
+
+\pagebreak
+
+# Abstract
+
+A popular approach in ecological modeling explicitly simulates the lives of individual organisms.
+Individual-based modeling (IBM) allows ecological complexity to be studied as it emerges from the individual-level, to explicitly synthesize theories and principles, and to track and record virtually any aspect of a system.
+I developed a novel IBM platform called simplex to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental dynamics, and community assembly and structure in both fluid and non-fluid environments.
+simplex accomplishes three primary tasks in keeping with its namesake (i.e., the generalized notion of a triangle to arbitrary dimensions).
+First, simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables.
+Second, simplex stores the output of models and generates and stores animations.
+Third, simplex provides R Markdown files for analyzing simulated data.
+The simplex source code is freely available, includes the standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to be run using modest hardware and freely available open-source tools.
+
+\pagebreak
 
@@
 
-	The individual with ID of 1 has a cell quota of 0.1 and is located at position x=45, y=765
+	The individual with ID of 1 has a cell quota of 0.1 and is 
+  located at position x=45, y=765
 
@@
 
-	The individual with ID of 1 belongs to species 12, which has a theoretical maximum growth rate of 0.9.
+	The individual with ID of 1 belongs to species 12, which has a 
+  theoretical maximum growth rate of 0.9.
 

### 2015-11-01 paper/PrePrint.Rmd
modified PrePrint.Rmd and .pdf; generated new figures; modified BasicAnalysis.Rmd -KL

--- 2015-10-26 paper/PrePrint.Rmd
@@
 
-# Introduction
+\begin{center} 
+Department of Biology, University of Indiana, Bloomington
+\end{center}
+
+\begin{center}
+email: kjlocey@indiana.edu
+\end{center}
+
+\pagebreak
+
+**ABSTRACT**
+
+A popular approach in ecological modeling explicitly simulates the biology and ecology of individual organisms.
+Individual-based models (IBM) allow ecological patterns and dynamics to be studied as they emerge from the individual level according to simulated principles and theories.
+IBMs also allow virtually any aspect of a simulated system to be tracked and recorded for later analysis.
+I developed a novel IBM platform (i.e, simplex) to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental dynamics, and community assembly and structure in both fluid and non-fluid environments.
+simplex accomplishes three primary tasks in keeping with its namesake (i.e., a generalized notion of a triangle to arbitrary dimensions).
+First, simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables.
+Second, simplex stores the output of models and generates and stores animations.
+Third, simplex provides R Markdown files for analyzing simulated data.
+The simplex source code is freely available, includes the standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to run on modest hardware.
+
+\pagebreak
+
+**INTRODUCTION**
+
+\setlength\parindent{36pt}
 
@@
 In ecology, simulation-based models are often used to examine analytically challenging or intractable scenarios. 
-Examples are markov models that simulate stochastic demographic changes and early ecological null models, both of which operate on matrices (Gotelli and Entsminger 2001, Hubbell 2001).
+Examples are markov models that simulate stochastic demographic changes and early ecological null models that , both of which operate on some form of matrix (Gotelli and Entsminger 2001, Hubbell 2001).
 Ecologists have also simulated growth and interactions among individuals for over two decades with individual-based models (IBMs) (DeAngelis and Gross 1992, Rosindell et al.
@@
 IBMs can provide degrees of ecological realism, individual variability, and spatial heterogenetiy that are unattainable with other models.
-IBMs also offer the potential for realistic and unanticipated ecological dynamics and patterns to emerge from individual-level interactions.
-However, IBMs offer challenges that include greater computational complexity, the difficulty of explicitly encoding ecological theory, and the use of ecological problem solving (Matthews et al.
+IBMs also offer the potential for realistic and unanticipated ecological dynamics and patterns to emerge from individual level interactions.
+However, IBMs come with challenges that include greater computational complexity, the difficulty of explicitly encoding ecological theory, and the challenge of using IBMs to solve applied problems (Matthews et al.
 2007, Grimm and Railsback 2005).
 
-Despite their frequent use, the power of ecological IBMs has yet been leveraged to the greatest advantage.
-Few, if any, include fluid dynamics or combine fluid dynamics with growth and active dispersal.
-While growth, sensing, and decision making are often modeled, few IBMs integrate physiology, evolution, community ecology, biogeography, and sampling theory.
+Despite their frequent use, the power of ecological IBMs has yet been leveraged to its greatest advantage.
+Few, if any, include fluid dynamics or combine fluid dynamics with growth and dispersal.
+While growth, sensing, and decision making are often modeled, few IBMs integrate physiology, nutrient-limited growth, evolution, environmental flux, community ecology, biogeography, and stochastic sampling theory.
 Ecological IBMs are more often constructed to model a specific system than to synthesize general theories of ecology and evolution (but see Rosindell et al.
 2015).
-Yet, IBMs can allow researchers to simulate and track information from the level of genomes and internal physiology to the stoichiometry of resource particles and distributions of abundance among species and across space.
+Yet, IBMs can allow researchers to creatively and explicitly simulate any number of biological and ecological theories and principles.
+Likewise, IBMs can made to track and analyze any information from physiological changes to community assembly, and the structure of an entire ecosystem.
 
-Here, I present source code and a detailed description for an ecological IBM platform (called simplex) that simulates a broad range of ecological conditions, records information from individuals to the ecosystem, and provides computing code for analyzing output.
-This platform incorporates computational fluid dynamics (Succi 2001), life history processes and concepts and patterns from biogeography (Hubbell 2001, Bell 2001), evolution and aspects of selection (Hartl and Clark 1997), and integrates nutrient-limited growth and physiology (Pirt 1965, Droop 1983) among dimension of ecology and evolutionary science.
-I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes.
-Below, I provide detailed explanation of how hydrobide works, the data it quantifies and tracks, the theories and principles hydrobide integrates, and the analyses that can be conducted using the code we provide.
+Here, I present a detailed description for an ecological IBM platform referred to as "simplex".
+The simplex platform simulates a broad range of ecological conditions, records information from individuals to the ecosystem level, and provides computing code for analyzing output.
+simplex incorporates fluid dynamics, stochastic and selection-based changes within and among populations, and integrates nutrient-limited growth and physiology.
+I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes on populations, communities, and ecosystems.
+I provide an explanation of how simplex works, the data it quantifies and tracks, the theories and principles that simplex integrates, and the analyses that can be conducted using the open source code.
 
-# Methods
+\setlength\parindent{0pt}
 
-**Overview**--I developed simplex to accomplish three objectives.
-First, simplex assembles and runs individual-based ecological models from random combinations of state variables and processes.
-Second, simplex stores the output of models as R-style data frames and .csv files, and when the user chooses, generates and stores animations of models.
-Third, simplex provides R Markdown files for analyzing simulated data; these will eventually be offered in Python as well.
+**METHODS**
 
-**Source code**--To save space, I refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
+**Model description following ODD Protocol**  
+The ODD protocol is standard for describing individual-based models (Grimm et al.
+2006).
+ODD stands for Overview, Design concepts, and Details.
+Here, I provide an abbreviated explanation of the simplex ODD protocol.
+The full ODD protocol can be found on the public GitHub repository at https://github.com/LennonLab/simplex.
+Likewise, to save space, I refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
 
-## Model description following the ODD Protocol
-The ODD protocol is standard for describing individual-based models (Grimm et al. 2006). 
-ODD stands for Overview, Design concepts, and Details. Here, I descibe simplex *largely* according to the ODD protocol.
-The ODD protocol for simplex can also be found on the public GitHub repository at https://github.com/LennonLab/simplex
+**Purpose**  
+The purpose of simplex is to explicitly simulate life history and the evolution of traits among individual organisms, the assembly of ecological communities, and the structure and function of ecosystems in a spatially explicit environment under stochastic conditions.
+The proximate goal of simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes.
+The ultimate goal of simplex is to provide a simulation-based platform for examining conditions under which processes and constraints have a robust influence on eco-evolutionary dynamics, patterns of biodiversity, and ecosystem processes.
 
-### Purpose
-The purpose of simplex is to simulate life history of individual organisms, the assembly of ecological communities, and the evolution of traits in spatially explicit environments under stochastic conditions.
-The proximate goal of simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes.
-The ultimate goal of simplex is to provide a simulation-based platform for examining conditions under which processes and constraints have a robust influence on eco-evolutionary dynamics and patterns of biodiversity.
+**Entities & their state variables**
 
-### Entities & their state variables
-**Individual organisms**--Individuals are distinguished by collections of elements within lists.
+*Individual organisms*--Individuals are distinguished by collections of elements within lists.
 Individuals undergo changes when randomly sampled from lists.
@@
 The first position in this list as well in all other lists of individual attributes corresponds to the same individual.
+Lists of attributes for individual organisms include the time each individual spends in the system, species ID, individual ID, 2D spatial location, endogenous levels of resources (resource specific cell quotas), individual-level metabolic maintenance cost, individual-level maximum dispersal rate, parental lineage, individual resource use efficiency for each resource.
 
-	Example:
-	IndIDs = [1,  2, 33, 14]
-	Quota =  [0.1, 0.99, 0.14, 0.05]
-	Xpos =   [45, 23, 456, 1]
-	Ypos =   [765, 87, 21, 34]
-
-	The individual with ID of 1 has a cell quota of 0.1 and is located at position x=45, y=765
-
-These are the lists of attributes for individual organisms:
-
-* time each particle spends in the system (aka residence time)
-* species ID
-* individual ID
-* 2D spatial location
-* endogenous levels of 3 resources (resource specific cell quotas)
-* individual-level metabolic maintenance cost
-* individual-level maximum dispersal rate
-* pedigree, i.e., direct lineage 
-* individual resource use efficiency for each resource
-* products of individual-level metabolism
-
-**Species**--Each species is characterized by the individuals that share a common set of traits, such as maximum growth rate, metabolic maintenance cost.
+*Species*--Each species is characterized by the individuals that share a common set of traits, such as maximum growth rate, metabolic maintenance cost.
 Species information is stored in Python dictionaries.
@@
 But, if simplex requires the maximum growth rate for an individual, then it finds the species ID and then uses that to accesses the Python dictionary for maximum specific growth rates of species.
+The species-level information that are stored include metabolic maintenance cost, maximum dispersal rate, maximum theoretical growth rate, and the resource use efficiency for each resource.
 
-	Example:
-	IndIDs = [1,  2, 33, 14]
-	spIDs = [12, 32, 11, 6]
-	MaxGrowthDict = {12: 0.9, 32: 0.5, 11: 0.8, 6: 0.4}
+*Resource particles*--Individual resource particles are distinguished by collections of elements within lists.
+Each specific position in the list corresponds to the same resource particle.
+The types of information stored about each resource particle include time in the system, particle ID, 2D spatial location, the type of resource particle, the size of the particle.
 
-	The individual with ID of 1 belongs to species 12, which has a theoretical maximum growth rate of 0.9.
-
-The following are they types of species-level information that are stored in Python dictionaries:
-
-* metabolic maintenance cost
-* maximum dispersal rate
-* maximum theoretical growth rate
-* resource use efficiency for each resource
-
-**Resource particles**--Individual resource particles are distinguished by collections of elements within lists. Each specific position in the list corresponds to the same resource particle.
-
-	Example:
-	resIDs = [4,  6, 17, 1]
-	size =  [100, 87, 156.3, 0.001]
-	Xpos =   [56, 34, 567, 2]
-	Ypos =   [876, 98, 32, 45]
-
-	The particle with ID of 4 has a size of 100 and is 
-	located at position x=56, y=876
-
-The following are the types of information stored about each resource particle:
-
-* time in the system
-* particle ID
-* 2D spatial location
-* whether the particle is Nitrogen, Carbon, or Phosphorus
-* type of Nitrogen, Carbon, or Phosphorus the resource is
-* size of the particle
-
-**Inert tracers particles**--These are objects that move/flow into and through the environment. They only interact with physical barriers. 
-The use of tracers particles allows for attributes of a system that flows or physically turns over to be quantified (e.g., hydraulic residence time).
+*Inert tracers particles*--These are objects that move/flow into and through the environment.
+They only interact with physical barriers.
+The use of tracers particles allows for attributes of a system that flows or physically turns over to be quantified.
 Like other individual-level objects in simplex models, tracers are distinguished by collections of elements within lists.
 Each specific position in the list corresponds to the same resource particle.
+Information stored about each resource particle includes time spent in the system, 2D spatial location and particle ID.
 
-This information stored about each resource particle:
-
-* time in the system
-* 2D spatial location
-* particle ID
-
-**Physcial barriers**--These objects are simulated as discrete 2D spatial coordinates that cannot be occupied by any individual entities.
+*Physcial barriers*--These objects are simulated as discrete 2D spatial coordinates that cannot be occupied by any individual entities.
 The number, size, and location of physical barriers are chosen at random by simplex at the start of each model.
 
-###System level state variables
-Each run of simplex begins with random choices for the values of:
+**System level state variables**  
+Each run of simplex begins with random choices for the values of width (5 to 100), height (5 to 100), basal flow rate (0.0 to 1.0), the number, size, and location of physical barriers, the number and direction of environmental gradients, rate of stochastic disturbance, rate of fluctuation in basal flow rate, and degrees of fluctuation and syncronized flow of individual particles, ranging from completely in-sync or completely out of sync.
 
-* width (5 to 100)
-* height (5 to 100)
-* basal flow rate (1.0 to 0.0)
-* number, size, and location of physical barriers 
-* number and direction of environmental gradients
-* rate of stochastic disturbance
-* rate of fluctuation in basal flow rate
-* degree of fluctuation 
-* degree of syncronized flow of individual particles
-	* completely in-sync or completely out of sync
+**Spatial and temporal scales**  
+\setlength\parindent{36pt}
+The two general aspects of scale are grain (i.e., resolution) and extent (e.g.
+total area).
+The environment of simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units.
+This makes for a potential total extent of 25 to 10,000 discrete patches, each with a grain of 1 square unit.
+All particles move in decimal units, meaning that individual particles can  practically infinite locations within patches and likewise.
+Extent of time in simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system.
+Residence time for individual particles (organisms, resources, tracers) can vary across five orders of magnitude.
 
-For example
-
-	from randparams.py:
-
-    width = randint(5,100)
-    height = randint(5,100)
-
-    # number of barriers
-    barriers = randint(1,10)
-
-	 # log-series alpha for metacommunity structure
-    alpha = np.random.uniform(0.99, 0.999)
-
-    reproduction = choice(['fission', 'sexual'])
-    speciation = choice(['yes', 'no'])
-
-    # size of starting community
-    seedCom = choice([10, 100, 1000])
-
-    # m = probability of immigration
-    m = choice([0.0, 0.0001, 0.0005, 0.001, 0.005])
-    ...
-
-###Spatial and temporal scales
-The two general aspects of scale are grain (a.k.a. 
-resolution) and extent (e.g. total area).
-
-**Spatial extent**--The environment of simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units.
-This makes for a potential total extent of 25 to 10,000 discrete patches, each with a grain of 1 square unit.
-
-Note that all particles move in decimal units the limit of which is determined by Python's decimal precision.
-This means that individual particles can occupy practically infinite locations within patches and likewise, squeeze through barriers (which only occupy integer x-y coordinates).
-
-**Temporal extent**--Extent of time in simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system.
-Residence time for inert tracer particles can vary across five orders of magnitude.
-
-**Grain**--Grain is the smallest unit over which change can happen. 
-For example, as per the original ODD documentation: “One time step represents one year and simulations were run for 100 years.
+Grain is the smallest unit over which change can happen.
+For example, the original ODD documentation states “One time step represents one year and simulations were run for 100 years.
 One grid cell represents 1 ha and the model landscape comprised 1,000 x 1,000 ha; i.e., 10,000 square kilometers”.
@@
 
-### Process overview and scheduling
-**Assembly**--The user runs a program that chooses random values for system-level state variables including whether disturbance, immigration, speciation, fluid dynamics, etc.
+\setlength\parindent{0pt}
+**Process overview and scheduling**
+
+*Assembly*.
+The user runs a program that chooses random values for system-level state variables including whether disturbance, immigration, speciation, fluid dynamics, etc.
 will occur and at what rates.
 
-**Core simulation process**--simplex models begin simulation immediately after assembly from random combinations of state-variables and processes. 
-Instead of operating by definitive time steps (i.e. days, generations), simplex models advance turnover of the environmental matrix according to the initial rate of flow. 
+*Core simulation process*.
+simplex models begin simulation immediately after assembly from random combinations of state-variables and processes.
+Instead of operating by definitive time steps (i.e.
+days, generations), simplex models advance turnover of the environmental matrix according to the initial rate of flow.
 If the initial rate of flow is 1.0, then the environmental matrix and inert particles would flow 1.0 units of distance.
@@
 
-**Duration: A run to mean reversion**--Once assembled, a simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
+*Duration: A run to mean reversion*.
+Once assembled, a simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
 Mean reversion captures whether a system is fluctuating around a mean value.
@@
 
-### Fluid dynamics
-simplex uses an efficient and powerful method for 
-simulating fluid flow, i.e., a Lattice-Boltzmann Method 
-(LBM). An LBM discretizes the environment into a lattice 
-and attaches to each position in the lattice a number of 
-particle densities and velocities for each of nine 
-directions of movement possible in a 2D environment 
-(N, S, E, W, NE, NW, SE, SW, current position).
+*Fluid dynamics*.
+simplex uses an efficient and powerful method for simulating fluid flow, i.e., a Lattice-Boltzmann Method (LBM) (Succi 2001).
+An LBM discretizes the environment into a lattice and attaches to each position in the lattice a number of particle densities and velocities for each simulated direction.
+simplex simulates nine directions of movement possible in a 2D spatial environment (i.e., N, S, E, W, NE, NW, SE, SW, current position).
 
-**Active dispersal**--simplex models allow individuals to move towards their environmental optima.
+*Active dispersal*.
+simplex models allow individuals to move towards their environmental optima.
 Rather than a single environmental optima resulting from a single environmental gradient, simplex allows environmental optima to occur as intersections among environmental gradients.
@@
 
-**Simulated life history**--simplex models simulate growth, reproduction, and death via weighted random sampling.
+*Simulated life history*.\setlength\parindent{36pt} simplex models simulate growth, reproduction, and death via weighted random sampling.
 This simulate the partly probabilistic and partly deterministic nature of environmental filtering and individual-level interactions.
 
-*Inflow/Entrance:* Resources and individuals enter from any point in the environment.
-Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution
-of abundance among species in ecological communities (see Hubbell 2001). Along with a species ID and species-
-specific maximum rates of resource uptake, active dispersal, resource efficiencies, cell maintenance, 
-and environmental optima, each propagule was given a unique ID and a multi-resource cell quota that represent 
-the state of internal resources. The average of these cell quotas determine the probability of reproduction.
+Inflow/Entrance.
+Resources and individuals enter from any point in the environment.
+Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among species in ecological communities (see Hubbell 2001).
+Along with a species ID and species-specific maximum rates of resource uptake, active dispersal, resource efficiencies, cell maintenance, and environmental optima, each propagule was given a unique ID and a multi-resource cell quota that represent the state of internal resources.
+The average of these cell quotas determine the probability of reproduction.
 
-*Dispersal:* Individuals are allowed to actively move along environmental gradients (sometimes against the 
-direction of environmenta flow) and towards their optima. A better match to one's environmental optima increases 
-the chance of reproduction and the individual's ability to perform (consume, grow).
+Dispersal.
+Individuals are allowed to actively move along environmental gradients (sometimes against the direction of environmenta flow) and towards their optima.
+A better match to one's environmental optima increases the chance of reproduction and the individual's ability to perform (consume, grow).
 
-*Consumption & growth:* Sampled individuals consume resources according to their specific maximum rates of 
-uptake and grow according to specific resource efficiencies and maintenance costs. Uptake increases the individual 
-cell quotas and decreases ambient resources. Individual cell quotas are then decreased according to a specific 
-physiological maintenance cost.
+Consumption & growth.
+Sampled individuals consume resources according to their specific maximum rates of uptake and grow according to specific resource efficiencies and maintenance costs.
+Uptake increases the individual cell quotas and decreases ambient resources.
+Individual cell quotas are then decreased according to a specific physiological maintenance cost.
 
-*Reproduction:* Reproduction in simplex is currently limited to clonal reproduction with the possibility of 
-mutation. Individuals reproduce with a probability determined by the combination of mean cell quota and the 
-proportional match to the environmental optima. The cell quota of each resource is evenly divided between the 
-individual and its daughter. The daughter is given a unique individual ID and the species ID of its mother, 
-unless in the case of speciation, but is allowed small mutations in individual-level state variables.
+Reproduction.
+Reproduction in simplex is currently limited to clonal reproduction with the possibility of mutation.
+Individuals reproduce with a probability determined by the combination of mean cell quota and the proportional match to the environmental optima.
+The cell quota of each resource is evenly divided between the individual and its daughter.
+The daughter is given a unique individual ID and the species ID of its mother, unless in the case of speciation, but is allowed small mutations in individual-level state variables.
 
-*Speciation:* Speciation is simulated within simplex as a discrete event but is accompanied by mutations in the 
-values of species-level state variables. This allows for diversity to arise within the system, which the 
-environmental filter can then select on.
+Speciation.
+Speciation is simulated within simplex as a discrete event (*sensu* Hubbell 2001 and Bell 2001) and is accompanied by mutations in the values of species-level state variables, much like that expected from a Wright-Fisher model (Hartl and Clark 1997).
+These random events allows for diversity to arise within the system, which the environmental filter can then select on.
 
-*Death:* Individuals sampled at random will die if their smallest resource specific cell quota (i.e., N, C, P)
-is equal to or less than 0.
+Death and emmigration.
+Individuals sampled at random will die if their smallest resource specific cell quota (i.e., N, C, P) is equal to or less than 0.
+Individuals, resource particles, and inert tracers are considered to have left or to have flowed out when they pass beyond edges of the environment.
 
-*Emigration:* Individuals, resource particles, and inert tracers are considered to have left or to have flowed out 
-when they pass beyond edges of the environment.
+\setlength\parindent{0pt}
 
-### Design concepts
+**Design concepts**  
+*Ecological complexity*.
+simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.
 
-**Basic principles.** 
-*Ecological complexity*: simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.
+*Nutrient limited growth*. All models assembled by simplex employ the universal concept that individual growth and activity is fueled and limited by resources.
 
-*Nutrient limited growth*: All models assembled by simplex employ the universal concept that individual growth and activity is fueled and limited by resources.
-
-*Resource diversity & heterogeneity*: simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.
+*Resource diversity & heterogeneity*. simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.
 
 **Theories**  
-*Constraint-based theory*: simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure.
+*Constraint-based theory*.
+simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure.
 That is, the idea that both ecological processes and physical constraints shape ecological diversity.
 
-*Chemostat theory*: simplex operates much like an unhinged bioreactor or chemostat.
+*Chemostat theory*.
+simplex operates much like an unhinged bioreactor or chemostat.
 That is, particles flow through a system of a defined size at an average rate, and are limited in growth by their residence time.
 
-*Ecological neutral theory*: simplex operates via random sampling and can vary from being completely neutral (all individuals having equal vital rates) to completely idiosyncratic (all individual and species are as different as possible).
+*Ecological neutral theory*.
+simplex operates via random sampling and can vary from being completely neutral (all individuals having equal vital rates) to completely idiosyncratic (all individual and species are as different as possible).
 The one aspect of neutral theory that simplex adopts without question is the importance of stochastic life history processes (i.e.
@@
 
-**Hypotheses**
-* These are entirely up to the user to formulate and test according to simplex's capabilities and analytical tools.
-
-**Modeling approaches**
-Simplex operations via two main modeling approaches, other than being individual-based, i.e., random sampling and computaitonal fluid dynamics.
-
-**Emergence**
-simplex uses random sampling and random assembly its models to avoid imposing strong constraints on the properties that emerge and to allow unanticipated combinations of traits and ecological structure to emerge.
-
-* Abundance
-	* Total community abundance
-	* Trait-related population size
-	* Effective population size
-	* Abundance-biomass relation
-* Community assembly
-	* Species turnover
-	* Biomass turnover
-	* Extinction rate
-	* Succession
-* Community structure
-	* Species richness
-	* Species evenness
-	* Trait diversity
-* Population structure
-	* Demography
-	* Subpopulation trait variation
-* Life history tradeoffs
-	* Mobility vs. metabolic maintenance
-	* Growth rate vs. metabolic maintenance
-	* Generalist vs. specialist
-	* R vs. K selection
-
 **Adaptation**
-Individuals can move towards their environmental optima. Populations can become aggregated in areas that provide 
-favorable intersections of species optima. Species canevolve by the action of the environmental filter on subpopulation variation in state variables.
-
-**Objectives**
-Individual seek conditions that match them to the environment (e.g., positions along environmental gradients).
-Individuals also seek to acquire resources through active searching.
-In the future, individuals will seek to avoid predation.
-
-**Learning**
-There is no aspect of individual-based learning in simplex, as of yet.
-
-**Prediction**
-Individuals in simplex do not have the ability to anticipate conditions.
+Individuals can move towards their environmental optima.
+Populations can become aggregated in areas that provide favorable intersections of species optima.
+Species can evolve by the action of the environmental filter on subpopulation variation in state variables.
+Individual can seek conditions that match them best to their environment.
 
 **Sensing**  
-Individuals only sense in the sense that they can move towards environmental optima and, in the future, 
-resources. Otherwise, all encounters are the result of random walks or fluid flow.
+Individuals can move towards environmental optima against the flow of the environment. Otherwise, all encounters are the result of random walks or fluid flow.
 
 **Interaction**  
-At the moment, individuals only interact indirectly through excluding each other from resources (e.g. 
-preemption). In the future, individuals will interact as predator-prey, mutualists, resource-dependents, etc. 
-Likewise, there is currently no communication, though quorum sensing would be cool.
+At the moment, individuals only interact indirectly through excluding each other from resources (e.g. preemption).
 
@@
 
-**Collectives**
-Individuals belong to species. Species belong to communities. In the future, simplex will allow communities to belong to trophic levels.
-
 **Observation**  
-Many simplex models shoudl be run to examine trends in the variation generated. The following is recorded for each simplex model:
+The following is recorded for each simplex model:
 
 * Values of randomly chosen state variables
-* Total abundance, $N$
-* Species richness, $S$
-* Avg growth rate (per species & per capita) 
-* Avg maintenance cost (per species & per capita) 
-* Avg resource efficiency (per species & per capita) 
-* Avg active dispersal rate (per species & per capita) 
-* Compositional turnover
-	* Bray-Curtis
-	* Sorensen's
-* Species turnover
-	* Whittaker's $\beta$ 
-* Species evenness
-	* Smith and Wilson's evenness, $E_{var}$
-	* Simpson's evenness, $E_{1/D}$ 
-* Species diversity
-	* Shannon's diversity, $H'$ 
-	* Simpson's diversity, $D_{1/D}$
-* Dominance
-	* Absolute, $N_{max}$
-	* Relative, $N_{max}/N$
-* Productivity
-	* Individuals
-	* Carbon, Nitrogen, Phosphorus
-* Residence time
+* Total abundance ($N$) and Species richness ($S$)
+* mean growth rate (per species & per capita) and mean maintenance cost (per species & per capita) 
+* mean resource efficiency (per species & per capita) and mean active dispersal rate (per species & per capita) 
+* Compositional turnover: Bray-Curtis, Sorensen's
+* Whittaker's species turnover
+* Species evenness: Smith and Wilson's index ($E_{var}$), Simpson's index ($E_{1/D}$) 
+* Species diversity: Shannon's diversity ($H'$), Simpson's diversity ($D_{1/D}$)
+* Dominance: Absolute ($N_{max}$), Berger-Parker index, $N_{max}/N$
+* Productivity in numbers of individuals and in amount of cell quotas
+* Residence time:
 	* Ecosystem, $(length * width)/flowrate$
@@
 
-###Initialization
-The model initiates with a random set of values for state-variables, 100 to 10,000 randomly drawn individuals from a 
-theoretical log-series metacommunity. These values are saved, so that a simplex model could be programmed to replicate an analysis.
+**Initialization**  
+The model initiates with a random set of values for state-variables, 100 to 10,000 randomly drawn individuals from a theoretical log-series metacommunity.
+These values are saved, so that a simplex model could be programmed to replicate an analysis.
 
-###Input data
+**Input data**  
 simplex models require no input data, but it might be cool to use an api to grab environmental data or other 
@@
 
-###Submodels & Equations
+**Submodels & Equations**
 
-**Cell quota model of Droop**
-In simplex models, individuals grow according to their amounts of endogenous resources (cell quota).
+*Cell quota model of Droop*. In simplex models, individuals grow according to their amounts of endogenous resources (cell quota).
 
-Droop (1968, 1983) gave a relationship between specific growth rate ($\mu$) and cell quota ($Q$):
-
-$$\mu = \mu'_{m} (1 - k_{q}/Q)$$
-
+Droop (1968, 1983) gave a relationship between specific growth rate ($\mu$) and cell quota ($Q$): $$\mu = \mu'_{m} (1 - k_{q}/Q)$$
 where $k_{q}$ is the minimum cell quota needed for life, also referred to as the subsistence quota. $\mu'$ is the
 
-**Maintenance cost of Pirt**
-Pirt (1965) states "The variation, with growth rate, of the yield of organism from the substrate used as 
-energy source is attributed to consumption of energy at a constant rate for cell maintenance." He derives 
-a relationships between the growth yield (biomass), the growth rate, and metabolic maintenance.
+*Maintenance cost of Pirt*.
+Pirt (1965) states "The variation, with growth rate, of the yield of organism from the substrate used as energy source is attributed to consumption of energy at a constant rate for cell maintenance." He derived a relationship between the growth yield (biomass), the growth rate, and metabolic maintenance.
+simplex models use Pirt's concept of a constant maintenance requirement.
+simplex also draws from Pirt's simple relation for substrate use: $$use(total) = use(maintenance) + use(growth)$$
 
-simplex models use Pirt's concept of a constant maintenance requirement. simplex also draws from 
-Pirt's simple relation for substrate use:
+Note that respiration and activity without growth is not accounted for.
 
-$$use(total) = use(maintenance) + use(growth)$$
+*Log-series metacommunity*. simplex models draw immigrating individuals from a theoretical log-series distribution.
+Hubbell (2001) states that the regional community (i.e., metacommmunity) often has a log-series species 
+abundance distribution. The probability density function (pdf) of the log-series is: $$f(k) = -1/ln(1 - p) * p^{k}/k$$
 
-Respiration and activity without growth is not accounted for.
-
-**Log-series metacommunity**
-simplex models draw immigrating individuals from a theoretical log-series distribution.
-Hubbell (2001) states that the regional community (i.e., metacommmunity) often has a log-series species 
-abundance distribution. The probability density function (pdf) of the log-series is:
-
-$$f(k) = -1/ln(1 - p) * p^{k}/k$$
-
-Hubbell (2001) provides explicit detail of the log-series, which is also covered in most ecological diversity texts
-and even on Wikipedia: https://en.wikipedia.org/wiki/Logarithmic_distribution
-
-## Notes on simplex source code
+**Notes on simplex source code**  
+\setlength\parindent{36pt}
 simplex models operate primarily on lists in a programmatic way, e.g., quickly sorting lists, and removing and returning an element from lists with very little overhead.
@@
 The R statistical computing language is well-suited to the analysis of simplex output and contains many packages for multivariate analysis and higher-order statistical analysis that Python is only beginning to accumulate.
-Consequently, we provide R source code in .R files and .Rmd (RMarkdown) files, complete with basic and advanced statistical analyses for analyzing diversity, regression models, ordination, variance partitioning, and for generating pdf documents (via Knitr) that integrate prose, code, and figures for manuscripts.
+Consequently, I provide R source code in .R files and .Rmd (RMarkdown) files, complete with basic and advanced statistical analyses for analyzing diversity, regression models, ordination, variance partitioning, and for generating pdf documents (via Knitr) that integrate prose, code, and figures for manuscripts.
 
@@
 
-## Speed & Memory
+\setlength\parindent{0pt}
 
+**Speed & Memory**
+
+\setlength\parindent{36pt}
 Simplex models do not complete until the time series of total abundance values reaches a state of mean reversion (i.e., stationarity).
@@
 
-## Animations
+\setlength\parindent{0pt}
 
+**Animations**  
 Simplex can generate animations of its models and store them in various image file formats.
@@
 
-#Results
+**RESULTS**
 
-## Unit tests
-
+**Unit tests**  
 simplex passed all units tests for 15 diversity indices, ensuring that each index returns either the correct calculated value, or 'NaN' if given any values that cannot be used (e.g., negative numbers, string characters, empty lists).
 
-## Speed & Memory
-
-**Results from 100 randomly assembled models.** On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31,367 +- 879, and required 161 +- 4.3 megabytes of memory.
+**Speed & Memory**  
+The following are results from 100 randomly assembled models.
+On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31,367 +- 879, and required 161 +- 4.3 megabytes of memory.
 The longest any model took to complete was 49 minutes.
@@
 
-## Products
-### Output data
-simplex generates six files as its output data. They are:
+**Products**  
+Simplex generates six files as its output data. They are:
 
-*SimData.csv*--Formatted as an R data frame, where each row is a run of a randomly assembled model, and each column holds a piece of data about the system that was modeled (e.g., flow rate, total abundance, species richness, species turnover, rate of disturbance, etc.).
+*SimData.csv*.
+Formatted as an R data frame, where each row is a run of a randomly assembled model, and each column holds a piece of data about the system that was modeled (e.g., flow rate, total abundance, species richness, species turnover, rate of disturbance, etc.).
 
-*IndRTD.csv*, *ResRTD.csv*, *TracerRTD.csv*--Each line of these three files contains the amount of time that each individual, resource particle, or inert tracer spent in the system.
+*IndRTD.csv*, *ResRTD.csv*, *TracerRTD.csv*.
+Each line of these three files contains the amount of time that each individual, resource particle, or inert tracer spent in the system.
 Each line represents the same run of a randomly assembled model.
 
-*RADs.csv*--Each line is a run of a randomly assembled model and contains the rank-abundance vector at the point when the model was stopped.
+*RADs.csv*. Each line is a run of a randomly assembled model and contains the rank-abundance vector at the point when the model was stopped.
 
-*Species.csv*--Each line is a run of a randomly assembled model and contains a vector of species labels corresponding to RADs.csv.
+*Species.csv*. Each line is a run of a randomly assembled model and contains a vector of species labels corresponding to RADs.csv.
 
@@
 
-### R analysis files
+**R analysis files**\setlength\parindent{36pt}
 
-There are currently two R Markdown analysis files associated with simplex; these are located in the `tools/analyses/Rmd` directory. 
-These files are primarily designed to help users understand how to import simplex files, to use R Markdown to analyze simplex data, and conduct descriptive and univariate statistics on simplex output.
+There are currently two R Markdown analysis files associated with simplex.
+These are located in the tools/analyses/Rmd directory and are designed to help users understand how to import simplex files and use R Markdown to analyze simplex data.
 The first file, `BasicAnalysis.Rmd`, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among simplex variables.
+Running this file on example simplex data revealed that simplex produces well-known and/or intuitive relationships (Figs 1, 2, 3, 4).
+
+```{r, results = 'hide', echo=F, message = F, warning = F}
+rm(list=ls()) 
+getwd()
+setwd("~/GitHub/simplex")
+
+#install.packages("vegan") # Example of how an install can be done
+require("vegan") 
+require("car")
+
+# A table where each columns corresponds to a state variable or model output.  
+simplex.dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")
+
+# Replacing 0's with 1's to allow log-transforms below
+simplex.dat <- simplex.dat[complete.cases(simplex.dat), ]
+
+simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 0.00001
+simplex.dat$N.max[simplex.dat$N.max<=0] <- 0.00001
+simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 0.00001
+simplex.dat$resource.particles[simplex.dat$resource.particles<=0] <- 0.00001
+simplex.dat$resource.concentration[simplex.dat$resource.concentration<=0] <- 0.00001
+simplex.dat$Whittakers.turnover[simplex.dat$Whittakers.turnover<=0] <- 0.00001
+simplex.dat$Jaccards.dissimilarity[simplex.dat$Jaccards.dissimilarity<=0] <- 0.00001
+
+simplex.dat$total.abundance <- log(simplex.dat$total.abundance)
+simplex.dat$N.max <- log(simplex.dat$N.max)
+simplex.dat$Whittakers.turnover <- log(simplex.dat$Whittakers.turnover)
+simplex.dat$Jaccards.dissimilarity <- log(simplex.dat$Jaccards.dissimilarity)
+simplex.dat$species.richness <- log(simplex.dat$species.richness)
+simplex.dat$resource.particles <- log(simplex.dat$resource.particles)
+simplex.dat$resource.concentration <- log(simplex.dat$resource.concentration)
+```
+
+**Figure 1.** A scatter plot matrix revealing that the values of inputs representing physical variables were independent of one another.
+This is only a subset of the physical variables in simplex.
+Amplitude pertains to magnitude of temporal fluctuation in the inflow of propagules and/or resources.
+Barriers pertains to the number of barriers in the system, which varied between 1 and 4.
+
+```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Physical and metacommunity variables
+png(filename="~/GitHub/simplex/results/figures/examples/PhysicalUnivariate.png")
+
+phys.dat <- as.matrix(subset(simplex.dat, 
+                      select = c(flow.rate,
+                                 barriers,
+                                 width,
+                                 height,
+                                 amplitude)))
+
+scatterplotMatrix(phys.dat, main="Physical data", diagonal = "none")
+dev.off()
+```
+
+**Figure 2.** A scatter plot matrix revealing that the values of resource-related variables.
+This is only a subset of resource variables in simplex.
+Number of resource particles, their concentration, resource richness, and resource diversity (a combination of resource richness and the variance in abundance among resources) are all highly and positively correlated.
+
+```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Resource variables
+png(filename="~/GitHub/simplex/results/figures/examples/ResourceUnivariate.png")
+
+res.dat <- as.matrix(subset(simplex.dat, 
+                      select = c(resource.concentration,
+                      shannons.resource.diversity,
+                      resource.richness,
+                      resource.particles)))
+
+scatterplotMatrix(res.dat, main="Resource data", diagonal = "none")
+dev.off()
+```
+
+**Figure 3.** A scatter plot matrix revealing that the values of physiological variables.
+This is only a subset of those in simplex.
+As of 29 October 2015, simplex does not include explicit stoichiometry.
+An intuitive life history trade-off arises in that increasing growth rate leads to increasing metabolic maintenance.
+
+```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Physiological variables
+png(filename="~/GitHub/simplex/results/figures/examples/PhysioUnivariate.png")
+
+physio.dat <- as.matrix(subset(simplex.dat, 
+                      select = c(biomass.prod.N,
+                      avg.per.capita.N.efficiency,
+                      max.growth.rate,
+                      max.met.maint)))
+
+scatterplotMatrix(physio.dat, main="Physiological data", diagonal = "none")
+dev.off()
+```
+
+**Figure 4.** A scatter plot matrix revealing that the values of diversity-related variables.
+This is only a subset of those in simplex.
+We see a strong postivie relationships between total numbers of individuals (total abundance) and total numbers of species (species richness) and between total abundance and the abundance of the most abundant species (Nmax).
+Each of these relationships is widely observed in ecological communities (Locey and Lennon 2015).
+
+```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Diversity-related variables
+png(filename="~/GitHub/simplex/results/figures/examples/DiversityUnivariate.png")
+
+diversity.dat <- as.matrix(subset(simplex.dat, 
+                      select = c(total.abundance,
+                      N.max,
+                      species.richness,
+                      simpson.e,
+                      Whittakers.turnover,
+                      Jaccards.dissimilarity)))
+
+scatterplotMatrix(diversity.dat, main="Diversity data", diagonal = "none")
+dev.off()
+```
+
 The second file, `Diversity.Rmd`, focuses on the analysis of taxonomic diversity and abundance to include fitting species abundance models using the R-based `vegan` module.
@@
 
-### Animations
-
+**Animations**   
 I used the animation module in the matplotlib plotting library to generate avi formatted movies of simplex models.
 These can be viewed in any internet brower (https://github.com/LennonLab/simplex/tree/master/results/movies/examples).
-Currently, generating movies requires following three instructions in model/model.py:
+Currently, generating movies requires following three easy steps explained in model/model.py.
 
-1.) uncomment line 'ani.save' on or near line 364  
-2.) adjust the frames variable on or near line 364, to change movie length   
-3.) change plot_system = 'no' to 'yes' on or near line 66
+\setlength\parindent{0pt}
 
-Because generating animations requires computing time and memory, doing so can be computationally demanding. To quicken the process, comment out plt.show() on or near line 368.
+**DISCUSSION**
 
-# Discussion
+\setlength\parindent{36pt}
 
-The simplex platform assembles individual-based models from random combinations of state-variables and processes.
-This is done to explore a large degree of variation in ecological conditions and complexity.
-In this way, simplex allows ecologists or educators to examine whether the influence of a particular variable should be robust to a wide range of ecological conditions.
-In addition to its unique  modeling aspects, simplex includes analysis code that shows the user how to import and analyze the simulated data, and provides a basic demonstration of how to conduct variance partitioning and fitting of species abundance models.
+The simplex platform assembles individual-based models from random combinations of state-variables and processes, to explore a degrees of variation in ecological conditions and complexity.
+In this way, simplex allows ecologists to examine whether the influence of particular variables should be robust to a wide range of ecological and environmental conditions.
+In addition to its unique modeling aspects, simplex includes analysis code that shows users how to import and analyze simulated datasets, including variance partitioning and fitting of species abundance models.
+
+Future developments of simplex will have trophic dynamics including predator-prey, symbionts and mutualists, and producers and consumers.
+Individuals will also be attributed with additional behaviors (e.g., search patterns), sensing (e.g., of resources), and learning (e.g., predator avoidance).
+Additionally, individuals will also be attributed with genomes.
+Initially, these genomes will be represented by sets of sequences of A's, C's, T's, and G's representing genes that can be mutated as they're passed on through clonal reproduction, or mutated and recombined in the case of sexual haploid or diploid reproduction.
+A precursor to this has already been developed in Locey (2013), the code for which is available here: https://github.com/klocey/SYMENT.
 
@@
 
-# Acknowledgments
+\setlength\parindent{0pt}
+
+**ACKNOWLEDGMENTS**
 
@@
 
-# References
+**REFERENCES**
 
@@
 
-10. Matthews RB, Gilbert NG, Roach A, Polhill JG, Gotts NM (2007) Agent-based land-use models: a review of applications. Landscape Ecology. 22: 1447-1459.
+10. Locey K (2013) Synthesized ecological molecular neutral theory. figshare. http://dx.doi.org/10.6084/m9.figshare.640105.
 
-11. Pirt SJ (1965) The Maintenance Energy of Bacteria in Growing Cultures. Proceedings of the Royal Society of London. Series B, Biological Sciences, 163:224-231.
+11. Locey KJ, Lennon JT. (2015) Scaling laws predict global microbial diversity. PeerJ PrePrints 3:e1808 https://dx.doi.org/10.7287/peerj.preprints.1451v1
 
-12. Rosindell J, Harmon LJ, Etienne RS (2015) Unifying ecology and macroevolution with individual-based theory. Ecology Letters. 18: 472-482.
+12. Matthews RB, Gilbert NG, Roach A, Polhill JG, Gotts NM (2007) Agent-based land-use models: a review of applications. Landscape Ecology. 22: 1447-1459.
 
-13. Succi S (2001) The Lattice Boltzmann Equation: For Fluid Dynamics and Beyond. Oxford University Press. New York.
+13. Pirt SJ (1965) The Maintenance Energy of Bacteria in Growing Cultures. Proceedings of the Royal Society of London. Series B, Biological Sciences, 163:224-231.
+
+14. Rosindell J, Harmon LJ, Etienne RS (2015) Unifying ecology and macroevolution with individual-based theory. Ecology Letters. 18: 472-482.
+
+15. Succi S (2001) The Lattice Boltzmann Equation: For Fluid Dynamics and Beyond. Oxford University Press. New York.

### 2015-11-01 tools/analyses/Rmd/BasicAnalysis.Rmd

@@
 simplex.dat <- read.csv("~/GitHub/simplex/results/simulated_data/examples/SimData.csv")
-simplex.dat$h.tau <- log((simplex.dat$width * simplex.dat$height)/simplex.dat$flow.rate)
 
 # Replacing 0's with 1's to allow log-transforms below
-simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 1
-simplex.dat$N.max[simplex.dat$N.max<=0] <- 1
-simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 1
-simplex.dat$resource.particles[simplex.dat$resource.particles<=0] <- 1
-simplex.dat$resource.concentration[simplex.dat$resource.concentration<=0] <- 1
+simplex.dat <- simplex.dat[complete.cases(simplex.dat), ]
+
+simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 0.00001
+simplex.dat$N.max[simplex.dat$N.max<=0] <- 0.00001
+simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 0.00001
+simplex.dat$resource.particles[simplex.dat$resource.particles<=0] <- 0.00001
+simplex.dat$resource.concentration[simplex.dat$resource.concentration<=0] <- 0.00001
+simplex.dat$Whittakers.turnover[simplex.dat$Whittakers.turnover<=0] <- 0.00001
+simplex.dat$Jaccards.dissimilarity[simplex.dat$Jaccards.dissimilarity<=0] <- 0.00001
 
@@
 simplex.dat$N.max <- log(simplex.dat$N.max)
+simplex.dat$Whittakers.turnover <- log(simplex.dat$Whittakers.turnover)
+simplex.dat$Jaccards.dissimilarity <- log(simplex.dat$Jaccards.dissimilarity)
 simplex.dat$species.richness <- log(simplex.dat$species.richness)
@@
 simplex.dat$resource.concentration <- log(simplex.dat$resource.concentration)
-simplex.dat[is.na(simplex.dat)] <- 0
+
 ```
@@
 ```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Physical and metacommunity variables
+png(filename="~/GitHub/simplex/results/figures/examples/PhysicalUnivariate.png")
 
-# Physical and metacommunity variables
 phys.dat <- as.matrix(subset(simplex.dat, 
-                      select = c(h.tau,
-                                 flow.rate,
+                      select = c(flow.rate,
                                  barriers,
@@
                                  height,
-                                 amplitude,
-                                 frequency,
-                                 phase)))
+                                 amplitude)))
 
 scatterplotMatrix(phys.dat, main="Physical data", diagonal = "none")
+dev.off()
 ```
@@
 # Resource variables
+png(filename="~/GitHub/simplex/results/figures/examples/ResourceUnivariate.png")
+
 res.dat <- as.matrix(subset(simplex.dat, 
@@
 scatterplotMatrix(res.dat, main="Resource data", diagonal = "none")
+dev.off()
 ```
@@
 # Physiological variables
+png(filename="~/GitHub/simplex/results/figures/examples/PhysioUnivariate.png")
+
 physio.dat <- as.matrix(subset(simplex.dat, 
                       select = c(biomass.prod.N,
-                      biomass.prod.C,
-                      biomass.prod.P,
                       avg.per.capita.N.efficiency,
-                      avg.per.capita.P.efficiency,
-                      avg.per.capita.C.efficiency,
                       max.growth.rate,
@@
 scatterplotMatrix(physio.dat, main="Physiological data", diagonal = "none")
+dev.off()
 ```
@@
 ```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+# Diversity-related variables
+png(filename="~/GitHub/simplex/results/figures/examples/DiversityUnivariate.png")
+
 diversity.dat <- as.matrix(subset(simplex.dat, 
@@
                       simpson.e,
-                      e.var,
                       Whittakers.turnover,
@@
 scatterplotMatrix(diversity.dat, main="Diversity data", diagonal = "none")
+dev.off()
 ```

### 2015-11-01 paper/PrePrint.Rmd
modified PrePrint.Rmd and .pdf; generated new figures; modified BasicAnalysis.Rmd -KL

@@
 There are currently two R Markdown analysis files associated with simplex.
+These have been included as supplemental files.
 These are located in the tools/analyses/Rmd directory and are designed to help users understand how to import simplex files and use R Markdown to analyze simplex data.
-The first file, `BasicAnalysis.Rmd`, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among simplex variables.
+The first file, BasicAnalysis.Rmd, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among simplex variables.
 Running this file on example simplex data revealed that simplex produces well-known and/or intuitive relationships (Figs 1, 2, 3, 4).
+
+\setlength\parindent{0pt}
 
@@
 
-# Replacing 0's with 1's to allow log-transforms below
+# Removing any rows with NaN's in them. Sort of harsh constraint.
 simplex.dat <- simplex.dat[complete.cases(simplex.dat), ]
 
+# Replacing 0's with very small values to allow log-transforms below
 simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 0.00001
@@
 
-**Figure 1.** A scatter plot matrix revealing that the values of inputs representing physical variables were independent of one another.
+\pagebreak
+
+**Figure 1.** A scatter plot matrix revealing that the values of inputs representing physical variables are independent of one another, as should be expected in the modeling.
 This is only a subset of the physical variables in simplex.
@@
 
-```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=6}
 # Physical and metacommunity variables
-png(filename="~/GitHub/simplex/results/figures/examples/PhysicalUnivariate.png")
-
 phys.dat <- as.matrix(subset(simplex.dat, 
@@
 scatterplotMatrix(phys.dat, main="Physical data", diagonal = "none")
-dev.off()
 ```
+
+\pagebreak
 
@@
 This is only a subset of resource variables in simplex.
-Number of resource particles, their concentration, resource richness, and resource diversity (a combination of resource richness and the variance in abundance among resources) are all highly and positively correlated.
+Number of resource particles, their concentration, resource richness, and resource diversity (a combination of resource richness and the variance in abundance among resources) are all highly and positively correlated, as should generally be expected.
 
-```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=6}
 # Resource variables
-png(filename="~/GitHub/simplex/results/figures/examples/ResourceUnivariate.png")
-
 res.dat <- as.matrix(subset(simplex.dat, 
@@
 scatterplotMatrix(res.dat, main="Resource data", diagonal = "none")
-dev.off()
 ```
+
+\pagebreak
 
@@
 As of 29 October 2015, simplex does not include explicit stoichiometry.
-An intuitive life history trade-off arises in that increasing growth rate leads to increasing metabolic maintenance.
+An intuitive life history trade-off arises in that increasing growth rate generally leads to increasing metabolic maintenance.
 
-```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=6}
 # Physiological variables
-png(filename="~/GitHub/simplex/results/figures/examples/PhysioUnivariate.png")
-
 physio.dat <- as.matrix(subset(simplex.dat, 
@@
 scatterplotMatrix(physio.dat, main="Physiological data", diagonal = "none")
-dev.off()
 ```
+
+\pagebreak
 
@@
 This is only a subset of those in simplex.
-We see a strong postivie relationships between total numbers of individuals (total abundance) and total numbers of species (species richness) and between total abundance and the abundance of the most abundant species (Nmax).
+There are strong positive relationships between total numbers of individuals (total abundance) and total numbers of species (species richness) and between total abundance and the abundance of the most abundant species (Nmax).
 Each of these relationships is widely observed in ecological communities (Locey and Lennon 2015).
+Likewise, different measures of species turnover largely reflect one another, as would generally be expected.
 
-```{r, results='hide', echo=T, message=F, warning=F, fig.width=7, fig.height=6}
+```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=6}
 # Diversity-related variables
-png(filename="~/GitHub/simplex/results/figures/examples/DiversityUnivariate.png")
-
 diversity.dat <- as.matrix(subset(simplex.dat, 
@@
 scatterplotMatrix(diversity.dat, main="Diversity data", diagonal = "none")
-dev.off()
 ```
 
-The second file, `Diversity.Rmd`, focuses on the analysis of taxonomic diversity and abundance to include fitting species abundance models using the R-based `vegan` module.
-Each file generates a knitted pdf (includes code, text, and figures), see Supplement.
+\pagebreak
+
+The second file, `Diversity.Rmd`, focuses on the analysis of taxonomic diversity and abundance to include fitting of five common species abundance models using the R-based `vegan` module.
+`Diversity.Rmd` also provides detailed explanation of each of the five models.
+Each file generates a knitted pdf (includes code, text, and figures).
+Pdf files of the output of Diversity.Rmd and BasicAnalysis.Rmd can be found in the tools/analyses/Rmd/pdfs directory.
 

### 2015-11-01 tools/analyses/Rmd/Diversity.Rmd
modified PrePrint.Rmd and .pdf; generated new figures; modified BasicAnalysis.Rmd -KL

@@
 
-# Replacing 0's with 1's to allow log-transforms below
-simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 1
-simplex.dat$N.max[simplex.dat$N.max<=0] <- 1
-simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 1
+# Removing any rows with NaN's in them. Sort of harsh constraint.
+simplex.dat <- simplex.dat[complete.cases(simplex.dat), ]
+
+# Replacing 0's with very small values to allow log-transforms below
+simplex.dat$total.abundance[simplex.dat$total.abundance<=0] <- 0.00001
+simplex.dat$N.max[simplex.dat$N.max<=0] <- 0.00001
+simplex.dat$species.richness[simplex.dat$species.richness<=0] <- 0.00001
+simplex.dat$resource.particles[simplex.dat$resource.particles<=0] <- 0.00001
+simplex.dat$resource.concentration[simplex.dat$resource.concentration<=0] <- 0.00001
+simplex.dat$Whittakers.turnover[simplex.dat$Whittakers.turnover<=0] <- 0.00001
+simplex.dat$Jaccards.dissimilarity[simplex.dat$Jaccards.dissimilarity<=0] <- 0.00001
 
@@
 simplex.dat$N.max <- log(simplex.dat$N.max)
+simplex.dat$Whittakers.turnover <- log(simplex.dat$Whittakers.turnover)
+simplex.dat$Jaccards.dissimilarity <- log(simplex.dat$Jaccards.dissimilarity)
 simplex.dat$species.richness <- log(simplex.dat$species.richness)
-simplex.dat[is.na(simplex.dat)] <- 0
+simplex.dat$resource.particles <- log(simplex.dat$resource.particles)
+simplex.dat$resource.concentration <- log(simplex.dat$resource.concentration)
 

### 2015-11-03 paper/PrePrint.Rmd
edited PrePrint.pdf and .Rmd -KL
edited PrePrint.Rmd -KL

@@
 ---
-title: "simplex: An individual-based platform for ecological modeling"
+title: "Simplex: An individual-based platform for ecological modeling"
 author: "Kenneth J. Locey"
@@
 IBMs also allow virtually any aspect of a simulated system to be tracked and recorded for later analysis.
-I developed a novel IBM platform (i.e, simplex) to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental dynamics, and community assembly and structure in both fluid and non-fluid environments.
-simplex accomplishes three primary tasks in keeping with its namesake (i.e., a generalized notion of a triangle to arbitrary dimensions).
-First, simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables.
-Second, simplex stores the output of models and generates and stores animations.
-Third, simplex provides R Markdown files for analyzing simulated data.
-The simplex source code is freely available, includes the standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to run on modest hardware.
+I developed a novel IBM platform (i.e., Simplex) to explicitly simulate physiology, life history, resource-limited growth, fluctuating environmental conditions, and community assembly and structure in both fluid and non-fluid environments.
+Simplex accomplishes three primary tasks in keeping with its namesake (i.e., a generalized notion of a triangle to arbitrary dimensions).
+First, Simplex assembles and runs stochastic ecological models from random combinations of over 20 state variables.
+Second, Simplex stores the output of simulations, and generates and stores animations.
+Third, Simplex provides R Markdown files for analyzing simulation data.
+The Simplex source code is freely available, includes a standard ODD description protocol, is open to community-based revision, includes unit testing and time testing, and has been developed to run on modest hardware.
 
@@
 Before modern computing, ecological models were almost exclusively equation-based representations of highly simplified systems (Black and McKane 2012, Otto and Troy 2007, Grant and Swannack 2008).
-Since the advent of personal computing, ecological models have been increasingly constructed to handle greater complexity and to explicitly simulate ecological processes.
+Since the advent of personal computing, ecological models have been increasingly constructed to handle greater complexity and to explicitly simulate ecological processes, forces, and mechanisms.
 
-In ecology, simulation-based models are often used to examine analytically challenging or intractable scenarios. 
-Examples are markov models that simulate stochastic demographic changes and early ecological null models that , both of which operate on some form of matrix (Gotelli and Entsminger 2001, Hubbell 2001).
-Ecologists have also simulated growth and interactions among individuals for over two decades with individual-based models (IBMs) (DeAngelis and Gross 1992, Rosindell et al.
+Simulation-based models are often used to examine analytically challenging or intractable scenarios. 
+Examples are markov models that simulate stochastic demographic changes and early ecological null models, both of which operate on matrices (Gotelli and Entsminger 2001, Hubbell 2001).
+Ecologists have also simulated growth and interactions among individual organisms for over two decades with individual-based models (IBMs) (DeAngelis and Gross 1992, Rosindell et al.
 2015).
 IBMs explicitly encode rules of how individuals change and interact.
-Once the rules are encoded, population to ecosystem-level dynamics can emerge as an IBM simulates over time and spatially explicit environments.
+Once these rules are encoded, population to ecosystem-level dynamics can emerge as an IBM simulates over time and spatially explicit environments.
 
@@
 While growth, sensing, and decision making are often modeled, few IBMs integrate physiology, nutrient-limited growth, evolution, environmental flux, community ecology, biogeography, and stochastic sampling theory.
-Ecological IBMs are more often constructed to model a specific system than to synthesize general theories of ecology and evolution (but see Rosindell et al.
+Instead, ecological IBMs are more often constructed to model a specific system than to synthesize general theories of ecology and evolution (but see Rosindell et al.
 2015).
-Yet, IBMs can allow researchers to creatively and explicitly simulate any number of biological and ecological theories and principles.
-Likewise, IBMs can made to track and analyze any information from physiological changes to community assembly, and the structure of an entire ecosystem.
+Yet, IBMs can allow researchers to creatively and explicitly simulate any number of theories and principles.
+Likewise, IBMs can made to track and analyze any information from physiological changes and population demography, to community assembly and the structure of entire ecosystems.
 
-Here, I present a detailed description for an ecological IBM platform referred to as "simplex".
-The simplex platform simulates a broad range of ecological conditions, records information from individuals to the ecosystem level, and provides computing code for analyzing output.
-simplex incorporates fluid dynamics, stochastic and selection-based changes within and among populations, and integrates nutrient-limited growth and physiology.
-I constructed simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes on populations, communities, and ecosystems.
-I provide an explanation of how simplex works, the data it quantifies and tracks, the theories and principles that simplex integrates, and the analyses that can be conducted using the open source code.
+Here, I present a detailed description for an ecological IBM platform I refer to as "Simplex".
+The Simplex platform simulates a broad range of ecological conditions, records information from individuals to the ecosystem level, and provides computing code for analyzing output.
+Simplex incorporates fluid and non-fluid dynamics with stochastic and selection-based changes within and among populations.
+Simplex also integrates nutrient-limited growth and physiology.
+I constructed Simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes on populations, communities, and ecosystems.
+I provide an explanation of how Simplex works, the data it quantifies and tracks, the theories and principles that Simplex integrates, and the analyses that can be conducted using the open source code.
 
@@
 ODD stands for Overview, Design concepts, and Details.
-Here, I provide an abbreviated explanation of the simplex ODD protocol.
+Here, I provide an abbreviated explanation of the Simplex ODD protocol.
 The full ODD protocol can be found on the public GitHub repository at https://github.com/LennonLab/simplex.
-Likewise, to save space, I refer the reader to the README.md file on the public simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
+Likewise, to save space, I refer the reader to the README.md file on the public Simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
 
 **Purpose**  
-The purpose of simplex is to explicitly simulate life history and the evolution of traits among individual organisms, the assembly of ecological communities, and the structure and function of ecosystems in a spatially explicit environment under stochastic conditions.
-The proximate goal of simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes.
-The ultimate goal of simplex is to provide a simulation-based platform for examining conditions under which processes and constraints have a robust influence on eco-evolutionary dynamics, patterns of biodiversity, and ecosystem processes.
+The purpose of Simplex is to explicitly simulate life history and the evolution of traits among individual organisms, the assembly of ecological communities, and the structure and function of ecosystems in a spatially explicit environment under stochastic conditions.
+The proximate goal of Simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes.
+The ultimate goal of Simplex is to provide a simulation-based platform for examining conditions under which processes and constraints have a robust influence on eco-evolutionary dynamics, patterns of biodiversity, and ecosystem processes.
 
@@
 Each specific position in the list corresponds to the same individual.
-For example, in simulating growth, a simplex model chooses an individual from the list of cell quotas (the probability of reproducing is determined by endogenous resources).
+For example, in simulating growth, a Simplex model chooses an individual from the list of cell quotas (the probability of reproducing is determined by endogenous resources).
 The first position in this list as well in all other lists of individual attributes corresponds to the same individual.
@@
 Species information is stored in Python dictionaries.
-In this way, if simplex requires the species ID of an individual it will access the species ID list where each element corresponds to a specific individual.
-But, if simplex requires the maximum growth rate for an individual, then it finds the species ID and then uses that to accesses the Python dictionary for maximum specific growth rates of species.
+In this way, if Simplex requires the species ID of an individual it will access the species ID list where each element corresponds to a specific individual.
+But, if Simplex requires the maximum growth rate for an individual, then it finds the species ID and then uses that to accesses the Python dictionary for maximum specific growth rates of species.
 The species-level information that are stored include metabolic maintenance cost, maximum dispersal rate, maximum theoretical growth rate, and the resource use efficiency for each resource.
@@
 The use of tracers particles allows for attributes of a system that flows or physically turns over to be quantified.
-Like other individual-level objects in simplex models, tracers are distinguished by collections of elements within lists.
+Like other individual-level objects in Simplex models, tracers are distinguished by collections of elements within lists.
 Each specific position in the list corresponds to the same resource particle.
@@
 *Physcial barriers*--These objects are simulated as discrete 2D spatial coordinates that cannot be occupied by any individual entities.
-The number, size, and location of physical barriers are chosen at random by simplex at the start of each model.
+The number, size, and location of physical barriers are chosen at random by Simplex at the start of each model.
 
 **System level state variables**  
-Each run of simplex begins with random choices for the values of width (5 to 100), height (5 to 100), basal flow rate (0.0 to 1.0), the number, size, and location of physical barriers, the number and direction of environmental gradients, rate of stochastic disturbance, rate of fluctuation in basal flow rate, and degrees of fluctuation and syncronized flow of individual particles, ranging from completely in-sync or completely out of sync.
+Each run of Simplex begins with random choices for the values of width (5 to 100), height (5 to 100), basal flow rate (0.0 to 1.0), the number, size, and location of physical barriers, the number and direction of environmental gradients, rate of stochastic disturbance, rate of fluctuation in basal flow rate, and degrees of fluctuation and syncronized flow of individual particles, ranging from completely in-sync or completely out of sync.
 
@@
 total area).
-The environment of simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units.
+The environment of Simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units.
 This makes for a potential total extent of 25 to 10,000 discrete patches, each with a grain of 1 square unit.
 All particles move in decimal units, meaning that individual particles can  practically infinite locations within patches and likewise.
-Extent of time in simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system.
+Extent of time in Simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system.
 Residence time for individual particles (organisms, resources, tracers) can vary across five orders of magnitude.
@@
 One grid cell represents 1 ha and the model landscape comprised 1,000 x 1,000 ha; i.e., 10,000 square kilometers”.
-In contrast, the smallest grain achievable by simplex is determined by slowest rate at which individuals can undergo BIDE (birth, immigration, death, emigration) processes.
+In contrast, the smallest grain achievable by Simplex is determined by slowest rate at which individuals can undergo BIDE (birth, immigration, death, emigration) processes.
 For example, under high residence times, an individual can move across 0.00001% of the x or y axis in one time step.
@@
 *Core simulation process*.
-simplex models begin simulation immediately after assembly from random combinations of state-variables and processes.
+Simplex models begin simulation immediately after assembly from random combinations of state-variables and processes.
 Instead of operating by definitive time steps (i.e.
-days, generations), simplex models advance turnover of the environmental matrix according to the initial rate of flow.
+days, generations), Simplex models advance turnover of the environmental matrix according to the initial rate of flow.
 If the initial rate of flow is 1.0, then the environmental matrix and inert particles would flow 1.0 units of distance.
@@
 *Duration: A run to mean reversion*.
-Once assembled, a simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
+Once assembled, a Simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
 Mean reversion captures whether a system is fluctuating around a mean value.
-Once 100 generations have occurred within a model, simplex examines whether a point of mean reversion has occured by conducting an Augmented Dickey-Fuller (ADF) Test, which is well-explained here: https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing.
+Once 100 generations have occurred within a model, Simplex examines whether a point of mean reversion has occured by conducting an Augmented Dickey-Fuller (ADF) Test, which is well-explained here: https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing.
 A model is stopped once mean reversion is determined to occurred and, unless the number of desired simulations (i.e.
-models) has been reached, simplex simply constructs another model from random combinations of state-variables and processes, and then runs it to mean reversion.
+models) has been reached, Simplex simply constructs another model from random combinations of state-variables and processes, and then runs it to mean reversion.
 
 *Fluid dynamics*.
-simplex uses an efficient and powerful method for simulating fluid flow, i.e., a Lattice-Boltzmann Method (LBM) (Succi 2001).
+Simplex uses an efficient and powerful method for simulating fluid flow, i.e., a Lattice-Boltzmann Method (LBM) (Succi 2001).
 An LBM discretizes the environment into a lattice and attaches to each position in the lattice a number of particle densities and velocities for each simulated direction.
-simplex simulates nine directions of movement possible in a 2D spatial environment (i.e., N, S, E, W, NE, NW, SE, SW, current position).
+Simplex simulates nine directions of movement possible in a 2D spatial environment (i.e., N, S, E, W, NE, NW, SE, SW, current position).
 
 *Active dispersal*.
-simplex models allow individuals to move towards their environmental optima.
-Rather than a single environmental optima resulting from a single environmental gradient, simplex allows environmental optima to occur as intersections among environmental gradients.
+Simplex models allow individuals to move towards their environmental optima.
+Rather than a single environmental optima resulting from a single environmental gradient, Simplex allows environmental optima to occur as intersections among environmental gradients.
 Hence, individuals potentially have multiple optima resulting from unique and equally optimal intersection of up to 10 environmental gradients.
 
-*Simulated life history*.\setlength\parindent{36pt} simplex models simulate growth, reproduction, and death via weighted random sampling.
+*Simulated life history*.\setlength\parindent{36pt} Simplex models simulate growth, reproduction, and death via weighted random sampling.
 This simulate the partly probabilistic and partly deterministic nature of environmental filtering and individual-level interactions.
@@
 Reproduction.
-Reproduction in simplex is currently limited to clonal reproduction with the possibility of mutation.
+Reproduction in Simplex is currently limited to clonal reproduction with the possibility of mutation.
 Individuals reproduce with a probability determined by the combination of mean cell quota and the proportional match to the environmental optima.
@@
 Speciation.
-Speciation is simulated within simplex as a discrete event (*sensu* Hubbell 2001 and Bell 2001) and is accompanied by mutations in the values of species-level state variables, much like that expected from a Wright-Fisher model (Hartl and Clark 1997).
+Speciation is simulated within Simplex as a discrete event (*sensu* Hubbell 2001 and Bell 2001) and is accompanied by mutations in the values of species-level state variables, much like that expected from a Wright-Fisher model (Hartl and Clark 1997).
 These random events allows for diversity to arise within the system, which the environmental filter can then select on.
@@
 *Ecological complexity*.
-simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.
+Simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.
 
-*Nutrient limited growth*. All models assembled by simplex employ the universal concept that individual growth and activity is fueled and limited by resources.
+*Nutrient limited growth*. All models assembled by Simplex employ the universal concept that individual growth and activity is fueled and limited by resources.
 
-*Resource diversity & heterogeneity*. simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.
+*Resource diversity & heterogeneity*. Simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.
 
@@
 *Constraint-based theory*.
-simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure.
+Simplex was originally built to explore the influence of ecosystem residence time (volume/flow rate) on community assembly and structure.
 That is, the idea that both ecological processes and physical constraints shape ecological diversity.
@@
 *Chemostat theory*.
-simplex operates much like an unhinged bioreactor or chemostat.
+Simplex operates much like an unhinged bioreactor or chemostat.
 That is, particles flow through a system of a defined size at an average rate, and are limited in growth by their residence time.
@@
 *Ecological neutral theory*.
-simplex operates via random sampling and can vary from being completely neutral (all individuals having equal vital rates) to completely idiosyncratic (all individual and species are as different as possible).
-The one aspect of neutral theory that simplex adopts without question is the importance of stochastic life history processes (i.e.
+Simplex operates via random sampling and can vary from being completely neutral (all individuals having equal vital rates) to completely idiosyncratic (all individual and species are as different as possible).
+The one aspect of neutral theory that Simplex adopts without question is the importance of stochastic life history processes (i.e.
 weighted or unweighted random fluctuations in population sizes).
@@
 **Observation**  
-The following is recorded for each simplex model:
+The following is recorded for each Simplex model:
 
@@
 The model initiates with a random set of values for state-variables, 100 to 10,000 randomly drawn individuals from a theoretical log-series metacommunity.
-These values are saved, so that a simplex model could be programmed to replicate an analysis.
+These values are saved, so that a Simplex model could be programmed to replicate an analysis.
 
 **Input data**  
-simplex models require no input data, but it might be cool to use an api to grab environmental data or other 
-data from a website to parameterize a simplex model.
+Simplex models require no input data, but it might be cool to use an api to grab environmental data or other 
+data from a website to parameterize a Simplex model.
 
@@
 
-*Cell quota model of Droop*. In simplex models, individuals grow according to their amounts of endogenous resources (cell quota).
+*Cell quota model of Droop*. In Simplex models, individuals grow according to their amounts of endogenous resources (cell quota).
 
@@
 Pirt (1965) states "The variation, with growth rate, of the yield of organism from the substrate used as energy source is attributed to consumption of energy at a constant rate for cell maintenance." He derived a relationship between the growth yield (biomass), the growth rate, and metabolic maintenance.
-simplex models use Pirt's concept of a constant maintenance requirement.
-simplex also draws from Pirt's simple relation for substrate use: $$use(total) = use(maintenance) + use(growth)$$
+Simplex models use Pirt's concept of a constant maintenance requirement.
+Simplex also draws from Pirt's simple relation for substrate use: $$use(total) = use(maintenance) + use(growth)$$
 
@@
 
-*Log-series metacommunity*. simplex models draw immigrating individuals from a theoretical log-series distribution.
+*Log-series metacommunity*. Simplex models draw immigrating individuals from a theoretical log-series distribution.
 Hubbell (2001) states that the regional community (i.e., metacommmunity) often has a log-series species 
@@
 
-**Notes on simplex source code**  
+**Notes on Simplex source code**  
 \setlength\parindent{36pt}
-simplex models operate primarily on lists in a programmatic way, e.g., quickly sorting lists, and removing and returning an element from lists with very little overhead.
-Likewise, simplex models generate and hold a lot of information about all the particles and elements in the system, which can become a computationally intensive task.
-To this end, simplex modeling coded is written in Python, an easy to read high-level programming language that has many scientific, plotting, and animation libraries.
+Simplex models operate primarily on lists in a programmatic way, e.g., quickly sorting lists, and removing and returning an element from lists with very little overhead.
+Likewise, Simplex models generate and hold a lot of information about all the particles and elements in the system, which can become a computationally intensive task.
+To this end, Simplex modeling coded is written in Python, an easy to read high-level programming language that has many scientific, plotting, and animation libraries.
 Python gives greater control over the operating system than data analysis languages (e.g.
@@
 
-The output of simplex is a broad array of information held in seven .csv files.
+The output of Simplex is a broad array of information held in seven .csv files.
 The most important of these is SimData.csv, and is intended to be analyzed in the freely available R (https://www.r-project.org/) and RStudio (https://www.rstudio.com/) environments.
-The R statistical computing language is well-suited to the analysis of simplex output and contains many packages for multivariate analysis and higher-order statistical analysis that Python is only beginning to accumulate.
+The R statistical computing language is well-suited to the analysis of Simplex output and contains many packages for multivariate analysis and higher-order statistical analysis that Python is only beginning to accumulate.
 Consequently, I provide R source code in .R files and .Rmd (RMarkdown) files, complete with basic and advanced statistical analyses for analyzing diversity, regression models, ordination, variance partitioning, and for generating pdf documents (via Knitr) that integrate prose, code, and figures for manuscripts.
 
-The reader can view example R-based analysis files that users can use to examine simplex's simulated data: https://github.com/LennonLab/simplex/tree/master/results/analyses.
+The reader can view example R-based analysis files that users can use to examine Simplex's simulated data: https://github.com/LennonLab/simplex/tree/master/results/analyses.
 
@@
 Simplex models do not complete until the time series of total abundance values reaches a state of mean reversion (i.e., stationarity).
-Because simplex models can range from quickly flowing systems with high disturbance to barely flowing and highly stable but depleted systems, simulations can potentially take several minutes or more to complete.
+Because Simplex models can range from quickly flowing systems with high disturbance to barely flowing and highly stable but depleted systems, simulations can potentially take several minutes or more to complete.
 Likewise, the ability to simulate many complex scenarios also allows for very large total abundances, the values of which cannot be predicted *a priori* and can even potentially outstrip a computer's memory.
 
-I ran simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
-This system probably represents a below average capacity for modern personal computers, which for this study, was desirable as simplex should be able to be ran on both personal computers and high capacity remote servers.
+I ran Simplex on a Mid 2010 MacBook Pro (OS X 10.9.5) with a 2.4 GHz Intel Core 2 Duo processor and 4GB of Memory.
+This system probably represents a below average capacity for modern personal computers, which for this study, was desirable as Simplex should be able to be ran on both personal computers and high capacity remote servers.
 I present results for time to completion and required memory in the Results.
@@
 **Unit tests**  
-simplex passed all units tests for 15 diversity indices, ensuring that each index returns either the correct calculated value, or 'NaN' if given any values that cannot be used (e.g., negative numbers, string characters, empty lists).
+Simplex passed all units tests for 15 diversity indices, ensuring that each index returns either the correct calculated value, or 'NaN' if given any values that cannot be used (e.g., negative numbers, string characters, empty lists).
 
@@
 The following are results from 100 randomly assembled models.
-On average, simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31,367 +- 879, and required 161 +- 4.3 megabytes of memory.
+On average, Simplex models required 96.60 +- 32.70 seconds to run, had an average total abundance (*N*) of 31,367 +- 879, and required 161 +- 4.3 megabytes of memory.
 The longest any model took to complete was 49 minutes.
@@
 
-There are currently two R Markdown analysis files associated with simplex.
+There are currently two R Markdown analysis files associated with Simplex.
 These have been included as supplemental files.
-These are located in the tools/analyses/Rmd directory and are designed to help users understand how to import simplex files and use R Markdown to analyze simplex data.
-The first file, BasicAnalysis.Rmd, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among simplex variables.
-Running this file on example simplex data revealed that simplex produces well-known and/or intuitive relationships (Figs 1, 2, 3, 4).
+These are located in the tools/analyses/Rmd directory and are designed to help users understand how to import Simplex files and use R Markdown to analyze Simplex data.
+The first file, BasicAnalysis.Rmd, focuses on descriptive statistics, multiple univariate analyses, and variance partitioning to reveal relationships among Simplex variables.
+Running this file on example Simplex data revealed that Simplex produces well-known and/or intuitive relationships (Figs 1, 2, 3, 4).
 
@@
 **Figure 1.** A scatter plot matrix revealing that the values of inputs representing physical variables are independent of one another, as should be expected in the modeling.
-This is only a subset of the physical variables in simplex.
+This is only a subset of the physical variables in Simplex.
 Amplitude pertains to magnitude of temporal fluctuation in the inflow of propagules and/or resources.
@@
 **Figure 2.** A scatter plot matrix revealing that the values of resource-related variables.
-This is only a subset of resource variables in simplex.
+This is only a subset of resource variables in Simplex.
 Number of resource particles, their concentration, resource richness, and resource diversity (a combination of resource richness and the variance in abundance among resources) are all highly and positively correlated, as should generally be expected.
@@
 This is only a subset of those in simplex.
-As of 29 October 2015, simplex does not include explicit stoichiometry.
+As of 29 October 2015, Simplex does not include explicit stoichiometry.
 An intuitive life history trade-off arises in that increasing growth rate generally leads to increasing metabolic maintenance.
@@
 **Figure 4.** A scatter plot matrix revealing that the values of diversity-related variables.
-This is only a subset of those in simplex.
+This is only a subset of those in Simplex.
 There are strong positive relationships between total numbers of individuals (total abundance) and total numbers of species (species richness) and between total abundance and the abundance of the most abundant species (Nmax).
@@
 **Animations**   
-I used the animation module in the matplotlib plotting library to generate avi formatted movies of simplex models.
+I used the animation module in the matplotlib plotting library to generate avi formatted movies of Simplex models.
 These can be viewed in any internet brower (https://github.com/LennonLab/simplex/tree/master/results/movies/examples).
@@
 
-The simplex platform assembles individual-based models from random combinations of state-variables and processes, to explore a degrees of variation in ecological conditions and complexity.
-In this way, simplex allows ecologists to examine whether the influence of particular variables should be robust to a wide range of ecological and environmental conditions.
-In addition to its unique modeling aspects, simplex includes analysis code that shows users how to import and analyze simulated datasets, including variance partitioning and fitting of species abundance models.
+The Simplex platform assembles individual-based models from random combinations of state-variables and processes, to explore a degrees of variation in ecological conditions and complexity.
+In this way, Simplex allows ecologists to examine whether the influence of particular variables should be robust to a wide range of ecological and environmental conditions.
+In addition to its unique modeling aspects, Simplex includes analysis code that shows users how to import and analyze simulated datasets, including variance partitioning and fitting of species abundance models.
 
-Future developments of simplex will have trophic dynamics including predator-prey, symbionts and mutualists, and producers and consumers.
+Future developments of Simplex will have trophic dynamics including predator-prey, symbionts and mutualists, and producers and consumers.
 Individuals will also be attributed with additional behaviors (e.g., search patterns), sensing (e.g., of resources), and learning (e.g., predator avoidance).
@@
 
-By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) simplex is well-positioned to continue development as both a research and education/instructional tool for ecology.
+By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) Simplex is well-positioned to continue development as both a research and education/instructional tool for ecology.
 Future developments and released versions will include additional ecological dynamcis such as predator-prey, mutualism, and parasitism.
-Improvements to simplex will also include an explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity.
+Improvements to Simplex will also include an explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity.
 For example, given the chemical structure of phosphate and the size of a resource particle, the particle can be assigned *x* units of phosphorus as well as the biocomplexity value of phosphate (estimated as a form of Shannon's entropy).
 
-Likewise, improvements and future versions of simplex will provide increasing numbers of files for statistical analysis of simplex output.
-These files will, in a sense, be able to be used as a complete analysis of some particular aspect of simplex output (e.g., demographic, resource-related, spatial, multivariate, ecological model testing).
-These and all simplex files are freely available and open source, allowing users to modify simplex source code to suit their particular needs.
+Likewise, improvements and future versions of Simplex will provide increasing numbers of files for statistical analysis of Simplex output.
+These files will, in a sense, be able to be used as a complete analysis of some particular aspect of Simplex output (e.g., demographic, resource-related, spatial, multivariate, ecological model testing).
+These and all Simplex files are freely available and open source, allowing users to modify Simplex source code to suit their particular needs.
 
@@
 Daniel V.
-Schroeder for his web-based fluid dynamic simulator, the python-based code for which serves as the core method in simplex for simulating fluid flow.
+Schroeder for his web-based fluid dynamic simulator, the python-based code for which serves as the core method in Simplex for simulating fluid flow.
+This work was supported by a National Science Foundation Dimensions of Biodiversity Grant (#1442246) awarded to Jay T. Lennon and Ken Locey.
 

### 2015-11-03 paper/PrePrint.Rmd

@@
 
-Here, I present a detailed description for an ecological IBM platform I refer to as "Simplex".
-The Simplex platform simulates a broad range of ecological conditions, records information from individuals to the ecosystem level, and provides computing code for analyzing output.
+Here, I present a description of an ecological IBM platform I refer to as "Simplex".
+This platform simulates a broad range of ecological conditions, records information from individuals to the ecosystem level, and provides R computing code for analyzing output.
 Simplex incorporates fluid and non-fluid dynamics with stochastic and selection-based changes within and among populations.
@@
 I constructed Simplex with the aim of simulating ecologically complex scenarios, from which the results could be used to test (*i*) predictions of ecological theory and (*ii*) the general influence of constraints and processes on populations, communities, and ecosystems.
-I provide an explanation of how Simplex works, the data it quantifies and tracks, the theories and principles that Simplex integrates, and the analyses that can be conducted using the open source code.
 
@@
 **Model description following ODD Protocol**  
-The ODD protocol is standard for describing individual-based models (Grimm et al.
+The overview, design concepts, and details protocol (ODD) is a standard for describing individual-based models (Grimm et al.
 2006).
-ODD stands for Overview, Design concepts, and Details.
-Here, I provide an abbreviated explanation of the Simplex ODD protocol.
-The full ODD protocol can be found on the public GitHub repository at https://github.com/LennonLab/simplex.
-Likewise, to save space, I refer the reader to the README.md file on the public Simplex GitHub.com repository (https://github.com/LennonLab/simplex/blob/master/README.md) for descriptions of source code files and directories.
+Here, I provide an abbreviated version of the Simplex ODD protocol.
+The full Simplex ODD protocol can be found on the public GitHub repository at https://github.com/LennonLab/simplex.
+Likewise, to save space, I refer the reader to the README.md file on the public Simplex GitHub.com repository for descriptions of source code, files, and directories (https://github.com/LennonLab/simplex/blob/master/README.md).
 
 **Purpose**  
-The purpose of Simplex is to explicitly simulate life history and the evolution of traits among individual organisms, the assembly of ecological communities, and the structure and function of ecosystems in a spatially explicit environment under stochastic conditions.
+The purpose of Simplex is to explicitly simulate life history, the evolution of traits, the assembly of ecological communities, and the structure and function of ecosystems in a spatially explicit environment under stochastic conditions, environmental filtering, and species sorting.
 The proximate goal of Simplex is to generate high degrees of variation in the assembly and structure of populations and communities by assembling many different models from random combinations of state-variables and processes.
@@
 
-*Individual organisms*--Individuals are distinguished by collections of elements within lists.
-Individuals undergo changes when randomly sampled from lists.
-Each specific position in the list corresponds to the same individual.
-For example, in simulating growth, a Simplex model chooses an individual from the list of cell quotas (the probability of reproducing is determined by endogenous resources).
-The first position in this list as well in all other lists of individual attributes corresponds to the same individual.
-Lists of attributes for individual organisms include the time each individual spends in the system, species ID, individual ID, 2D spatial location, endogenous levels of resources (resource specific cell quotas), individual-level metabolic maintenance cost, individual-level maximum dispersal rate, parental lineage, individual resource use efficiency for each resource.
+*Individual organisms*--These are distinguished by the corresponding elements of several lists.
+For example, in simulating reproduction, a Simplex model chooses an individual at random from the list of cell quotas, where the probability of reproducing is determined by the levels of endogenous resources, i.e., cell quota.
+The first position in the list of cell quotas, and in all other lists of individual attributes, corresponds to the same individual, i.e.,:
 
-*Species*--Each species is characterized by the individuals that share a common set of traits, such as maximum growth rate, metabolic maintenance cost.
-Species information is stored in Python dictionaries.
-In this way, if Simplex requires the species ID of an individual it will access the species ID list where each element corresponds to a specific individual.
-But, if Simplex requires the maximum growth rate for an individual, then it finds the species ID and then uses that to accesses the Python dictionary for maximum specific growth rates of species.
-The species-level information that are stored include metabolic maintenance cost, maximum dispersal rate, maximum theoretical growth rate, and the resource use efficiency for each resource.
+```{r, eval=FALSE}
+IndIDs = [1,  2, 33, 14]  
+Quotas =  [0.1, 0.99, 0.14, 0.05]  
+Xpos =   [45, 23, 456, 1]  
+Ypos =   [765, 87, 21, 34]  
+```
 
-*Resource particles*--Individual resource particles are distinguished by collections of elements within lists.
-Each specific position in the list corresponds to the same resource particle.
-The types of information stored about each resource particle include time in the system, particle ID, 2D spatial location, the type of resource particle, the size of the particle.
+Above, the individual with the ID "1" has a cell quota of 0.1  and is located at position x=45, y=765.
+Note that in simplex, there are at least three resource quotas.
+In future versions, the quotas will more explicitly represent nitrogen, carbon, and phosphorus.
+Of course, each individual could also have been distinguished as its own list of attributes, e.g., Ind1 = [1, 0.1, 45, 765].
+However, this would require a large lists of lists and could make quantifying some metrics (e.g., generating a frequency distribution of cell quotas, ages, or environmental optima) more computationally intensive than necessary.
 
-*Inert tracers particles*--These are objects that move/flow into and through the environment.
-They only interact with physical barriers.
-The use of tracers particles allows for attributes of a system that flows or physically turns over to be quantified.
-Like other individual-level objects in Simplex models, tracers are distinguished by collections of elements within lists.
-Each specific position in the list corresponds to the same resource particle.
-Information stored about each resource particle includes time spent in the system, 2D spatial location and particle ID.
+Lists of attributes for individual organisms include the time each individual spends in the system, species ID, individual ID, 2D spatial location, endogenous levels of resources (resource specific cell quotas), individual-level metabolic maintenance cost, individual-level maximum dispersal rate, parental lineage, and individual resource use efficiency for each resource.
+
+*Species*--Each species is characterized by the individuals that share a common set of traits, such as maximum growth rate, maximum dispersal rate, and metabolic maintenance cost.
+Species information is stored in dictionaries.
+In this way, if Simplex requires the species ID of an individual it will access the species ID list.
+But, if Simplex requires the maximum growth rate for an individual, it will first find the species ID and then use that ID to accesses the dictionary for maximum specific growth rates of species.
+Species-level data include metabolic maintenance cost, maximum dispersal rate, maximum theoretical growth rate, environmental optima, and the resource use efficiency for each resource.
+
+```{r, eval=FALSE}
+IndIDs = [1,  2, 33, 14]
+spIDs = [12, 32, 11, 6]
+MaxGrowthDict = {12: 0.9, 32: 0.5, 11: 0.8, 6: 0.4}
+```
+
+The individual with ID of 1 belongs to species 12, which has a theoretical maximum growth rate of 0.9.
+By storing species-level information in dictionaries, Simplex does not have to store redundant information in lists.
+
+*Resource particles*--Like individual organisms, individual resource particles distinguished by the corresponding elements of several lists.
+Each list holds a specific type of information, including time spent in the system, particle ID, 2D spatial location, the type of resource particle, the size of the resource particle.
+
+```{r, eval=FALSE}
+resIDs = [4,  6, 17, 1]
+size =  [100, 87, 156.3, 0.001]
+Xpos =   [56, 34, 567, 2]
+Ypos =   [876, 98, 32, 45]
+```
+
+Above, the particle with ID of "4" has a size of 100 and is located at position x=56, y=876.
+
+*Inert tracers particles*--These are objects that move or flow into and through the environment.
+Inert tracers only interact with physical barriers.
+The use of these tracers particles allows for physically dynamic attributes of a system to be quantified.
+Examples are ecosystem residence time and the ease of passive movement or inertia through the system.
+Like other individual-level objects in Simplex, tracers are distinguished by corresponding elements of several lists.
+Each specific position in the list corresponds to the same tracer particle.
+Information stored about each tracer particle includes time spent in the system, 2D spatial location and particle ID.
 
@@
 **System level state variables**  
-Each run of Simplex begins with random choices for the values of width (5 to 100), height (5 to 100), basal flow rate (0.0 to 1.0), the number, size, and location of physical barriers, the number and direction of environmental gradients, rate of stochastic disturbance, rate of fluctuation in basal flow rate, and degrees of fluctuation and syncronized flow of individual particles, ranging from completely in-sync or completely out of sync.
+Each run of Simplex begins with random choices for the specific structure of a log-series metacommunity and values of width (5 to 100), height (5 to 100), flow rate (0.0 to 1.0 units along the x and y axis per time step), the number, size, and location of physical barriers, the number and direction of environmental gradients, the rate and severity of stochastic disturbance, the rate of fluctuation in basal flow rate, and degrees of fluctuation and syncronization for the inflow of individual organisms and resources, which range from completely in-sync to completely out of sync.
 
@@
 \setlength\parindent{36pt}
-The two general aspects of scale are grain (i.e., resolution) and extent (e.g.
-total area).
+The two general aspects of scale are grain (i.e., resolution) and extent (e.g., total area).
 The environment of Simplex models is two dimensional and can vary along each axis from 5 to 100 discrete units.
 This makes for a potential total extent of 25 to 10,000 discrete patches, each with a grain of 1 square unit.
-All particles move in decimal units, meaning that individual particles can  practically infinite locations within patches and likewise.
+However, all particles move in decimal units, meaning that individual particles can occupy practically infinite locations within patches.
 Extent of time in Simplex models refers to residence time, i.e., the average amount of time that individual particles spend in the system.
-Residence time for individual particles (organisms, resources, tracers) can vary across five orders of magnitude.
+As shown by test simulations, residence time for individual particles (organisms, resources, inert tracers) can vary across five orders of magnitude.
 
@@
 For example, the original ODD documentation states “One time step represents one year and simulations were run for 100 years.
-One grid cell represents 1 ha and the model landscape comprised 1,000 x 1,000 ha; i.e., 10,000 square kilometers”.
-In contrast, the smallest grain achievable by Simplex is determined by slowest rate at which individuals can undergo BIDE (birth, immigration, death, emigration) processes.
-For example, under high residence times, an individual can move across 0.00001% of the x or y axis in one time step.
-Under low residence times, an individual can move across 10% or greater of the x or y axis in one time step.
+One grid cell represents 1 ha and the model landscape comprised 1,000 x 1,000 ha”.
+In Simplex, spatial grain is effectively determined by slowest rate at which individuals can traverse the landscape, which can vary from 0.00001% to 10% of the x or y axis in one time step.
 
@@
 *Core simulation process*.
-Simplex models begin simulation immediately after assembly from random combinations of state-variables and processes.
-Instead of operating by definitive time steps (i.e.
-days, generations), Simplex models advance turnover of the environmental matrix according to the initial rate of flow.
-If the initial rate of flow is 1.0, then the environmental matrix and inert particles would flow 1.0 units of distance.
-After each iteration of flow, each individual is given the chance to consume, grow, reproduce, starve, die, and to disperse towards resources and environmental optima.
+Simplex begin simulating immediately after assembling a model from random combinations of state-variables and processes.
+Simplex models advance turnover of the environmental matrix according to the rate of flow, which can be as low as 0.0 (i.e., an effectively static environment) or as high as 1.0 (i.e., where inert particles will flow 1.0 units of distance unless hindered by barriers).
+After each iteration of flow, every individual is given the chance to consume, grow, reproduce, starve, die, and to disperse towards environmental optima.
+However, each of these processes is probabilistic.
 
-*Duration: A run to mean reversion*.
-Once assembled, a Simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
-Mean reversion captures whether a system is fluctuating around a mean value.
-Once 100 generations have occurred within a model, Simplex examines whether a point of mean reversion has occured by conducting an Augmented Dickey-Fuller (ADF) Test, which is well-explained here: https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing.
-A model is stopped once mean reversion is determined to occurred and, unless the number of desired simulations (i.e.
-models) has been reached, Simplex simply constructs another model from random combinations of state-variables and processes, and then runs it to mean reversion.
+*Simulated life history*.\setlength\parindent{36pt} Simplex models simulate growth, reproduction, and death via weighted random sampling.
+This simulates the partly probabilistic and partly deterministic nature of environmental filtering and individual-level interactions.
+For example, individuals with a high cell quota have a low probability of dying and a high probability of reproducing.
+
+Inflow/Entrance.
+Resources and individuals enter from any point in the environment.
+Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among species in ecological communities (Hubbell 2001).
+Along with a species ID and species-specific maximum rates of resource uptake, active dispersal, resource efficiencies, cell maintenance, and environmental optima, each propagule was given a unique ID and a multi-resource cell quota that represent the state of internal resources.
+The average of these cell quotas determines the probability of reproduction.
+
+Dispersal.
+Individuals are allowed to actively move along environmental gradients and even against the direction of environmenta flow and towards their environmental optima.
+A better match to the environmental optima increases the chance of reproduction and an individual's ability to consume, grow, and reproduce.
+
+Consumption & growth.
+Sampled individuals consume resources according to their specific maximum rates of uptake and grow according to specific resource efficiencies and maintenance costs.
+However, can consumption can only happen in Simplex if individuals physically encounter resource particles.
+Uptake increases the individual cell quotas and decreases the size of the resource particle.
+Individual cell quotas are then decreased according to a specific physiological maintenance cost.
+
+Reproduction.
+Reproduction in Simplex is currently limited to clonal reproduction with the possibility of mutation.
+Individuals reproduce with a probability determined by the combination of mean cell quota and the proportional match to the environmental optima.
+The cell quota of each resource is evenly divided between the individual and its daughter.
+The daughter is then given a unique individual ID and the species ID of its mother, i.e., unless in the case of a random speciation event.
+
+Speciation.
+Speciation is simulated within Simplex as a discrete event (*sensu* Hubbell 2001 and Bell 2001) and is accompanied by mutations in the values of species-level state variables, much like that expected from a Wright-Fisher model (Hartl and Clark 1997).
+These random events allows for diversity to arise within the system, which the environmental filter can then select on.
+
+Death and emmigration.
+Individuals sampled at random will die if their smallest resource specific cell quota (i.e., N, C, P) is equal to or less than 0.
+Individuals, resource particles, and inert tracers are considered to have left or to have flowed out when they pass beyond edges of the environment.
+
+\setlength\parindent{0pt}
 
@@
 
-*Simulated life history*.\setlength\parindent{36pt} Simplex models simulate growth, reproduction, and death via weighted random sampling.
-This simulate the partly probabilistic and partly deterministic nature of environmental filtering and individual-level interactions.
-
-Inflow/Entrance.
-Resources and individuals enter from any point in the environment.
-Species identities of inflowing propagules are chosen at random from a log-series distribution, which often approximates the distribution of abundance among species in ecological communities (see Hubbell 2001).
-Along with a species ID and species-specific maximum rates of resource uptake, active dispersal, resource efficiencies, cell maintenance, and environmental optima, each propagule was given a unique ID and a multi-resource cell quota that represent the state of internal resources.
-The average of these cell quotas determine the probability of reproduction.
-
-Dispersal.
-Individuals are allowed to actively move along environmental gradients (sometimes against the direction of environmenta flow) and towards their optima.
-A better match to one's environmental optima increases the chance of reproduction and the individual's ability to perform (consume, grow).
-
-Consumption & growth.
-Sampled individuals consume resources according to their specific maximum rates of uptake and grow according to specific resource efficiencies and maintenance costs.
-Uptake increases the individual cell quotas and decreases ambient resources.
-Individual cell quotas are then decreased according to a specific physiological maintenance cost.
-
-Reproduction.
-Reproduction in Simplex is currently limited to clonal reproduction with the possibility of mutation.
-Individuals reproduce with a probability determined by the combination of mean cell quota and the proportional match to the environmental optima.
-The cell quota of each resource is evenly divided between the individual and its daughter.
-The daughter is given a unique individual ID and the species ID of its mother, unless in the case of speciation, but is allowed small mutations in individual-level state variables.
-
-Speciation.
-Speciation is simulated within Simplex as a discrete event (*sensu* Hubbell 2001 and Bell 2001) and is accompanied by mutations in the values of species-level state variables, much like that expected from a Wright-Fisher model (Hartl and Clark 1997).
-These random events allows for diversity to arise within the system, which the environmental filter can then select on.
-
-Death and emmigration.
-Individuals sampled at random will die if their smallest resource specific cell quota (i.e., N, C, P) is equal to or less than 0.
-Individuals, resource particles, and inert tracers are considered to have left or to have flowed out when they pass beyond edges of the environment.
-
-\setlength\parindent{0pt}
+*Duration: A run to mean reversion*.
+Once assembled, a Simplex model simulates ecological processes (birth, death, dispersal, growth, consumption, etc.) until the system reaches a point of mean reversion, i.e., the tendency of a system to reverse a directional change in, say, total abundance.
+Mean reversion captures whether a system is fluctuating around a mean value.
+Once 100 generations have occurred within a model, Simplex examines whether a point of mean reversion has occured by conducting an Augmented Dickey-Fuller (ADF) Test, which is well-explained here: https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing.
+A model is stopped once mean reversion is determined to occurred and, unless the number of desired simulations (i.e.
+models) has been reached, Simplex simply constructs another model from random combinations of state-variables and processes, and then runs it to mean reversion.
 
@@
 *Ecological complexity*.
-Simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of a particular state-variable, process, or combination thereof using univariate and multivariate tests.
+Simplex assembles models from random combinations of constraints (state-variables) and processes to generate output data that allow the user to test the general influence of state-variables and processes.
 
-*Nutrient limited growth*. All models assembled by Simplex employ the universal concept that individual growth and activity is fueled and limited by resources.
+*Process-based stochasticity*.
+Simplex was designed to incorporate random sampling and probability-based changes in as many ways as possible.
+This does not imply neutral dynamics or completely random dynamics, but instead makes all processes and dynamics (even strong environmental filtering) subject to a degree of uncertainty.
 
-*Resource diversity & heterogeneity*. Simplex allows the user to explore the influence of the number and abundances of different resources on ecological diversity and ecosystem processes.
+*Nutrient limited growth*.
+All models assembled by Simplex employ the universal concept that individual growth, reproduction, and activity is limited by resources within individuals (endogenous) rather than resouces simply within the environment (exogenous).
 
@@
 
-These data are stored in file as R-formatted data.frames. These files can be directly imported into an R or Python environment.
+These data are stored in .csv files. The file *SimData.csv* is stored an R-formatted data.frames. These files can be directly imported into an R or Python environment.
 
@@
 Droop (1968, 1983) gave a relationship between specific growth rate ($\mu$) and cell quota ($Q$): $$\mu = \mu'_{m} (1 - k_{q}/Q)$$
-where $k_{q}$ is the minimum cell quota needed for life, also referred to as the subsistence quota. $\mu'$ is the
+where $k_{q}$ is the minimum cell quota needed for life, also referred to as the subsistence quota. $\mu'_{m}$ is the theoretical maximum growth rate.
 
@@
 Simplex models use Pirt's concept of a constant maintenance requirement.
-Simplex also draws from Pirt's simple relation for substrate use: $$use(total) = use(maintenance) + use(growth)$$
+Simplex also draws from Pirt's simple relation for substrate use: $use(total) = use(maintenance) + use(growth)$
 
@@
 
-The Simplex platform assembles individual-based models from random combinations of state-variables and processes, to explore a degrees of variation in ecological conditions and complexity.
-In this way, Simplex allows ecologists to examine whether the influence of particular variables should be robust to a wide range of ecological and environmental conditions.
-In addition to its unique modeling aspects, Simplex includes analysis code that shows users how to import and analyze simulated datasets, including variance partitioning and fitting of species abundance models.
+The Simplex platform assembles individual-based models (IBMs) from random combinations of state-variables and processes, to simulate high degrees of variation and complexity.
+In this way, Simplex allows ecologists to examine whether the influence of particular variables should be robust to a wide range of conditions.
+For example, if a user wants to examine the influence of area on commonness and rarity, or the potential relationship between growth rate and physiological maintenance, then the user only needs to run Simplex, allow it to assemble and simulate a large number of models, and then analyze the output files.
+This is unlike many IBMs that are often built for a specific system or for individuals with highly specified behaviors and abilities (e.g.
+learning).
 
-Future developments of Simplex will have trophic dynamics including predator-prey, symbionts and mutualists, and producers and consumers.
+In addition to its unique modeling aspects, Simplex includes analysis code that shows users how to import and analyze simulated datasets, including simple univariate relationships, descriptive statistics, variance partitioning, and the fitting of species abundance models.
+By taking advantage of modern computational tools, version control, social coding, Simplex is well-positioned to continue development as both a research and education/instructional tool for ecology.
+
+Future developments and released versions of Simplex will have trophic dynamics including predator-prey, symbionts and mutualists, and producers and consumers.
 Individuals will also be attributed with additional behaviors (e.g., search patterns), sensing (e.g., of resources), and learning (e.g., predator avoidance).
@@
 
-By taking advantage of modern computing and version control and social coding tools (e.g., Python, R/RStudio, Knitr, Git, GitHub) Simplex is well-positioned to continue development as both a research and education/instructional tool for ecology.
-Future developments and released versions will include additional ecological dynamcis such as predator-prey, mutualism, and parasitism.
-Improvements to Simplex will also include an explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity.
-For example, given the chemical structure of phosphate and the size of a resource particle, the particle can be assigned *x* units of phosphorus as well as the biocomplexity value of phosphate (estimated as a form of Shannon's entropy).
-
+Future developments will also include explicit stoichiometry, where individual resource particles have ratios of Carbon, Nitrogen, and Phosphorus, as well as degrees of biocomplexity.
+For example, given the chemical structure of phosphate and the size of a resource particle, the particle can be assigned *x* units of phosphorus as well as the biocomplexity value of phosphate, which can be estimated as a form of entropy (Bertz 1983).
 Likewise, improvements and future versions of Simplex will provide increasing numbers of files for statistical analysis of Simplex output.
-These files will, in a sense, be able to be used as a complete analysis of some particular aspect of Simplex output (e.g., demographic, resource-related, spatial, multivariate, ecological model testing).
+These files will, in a sense, be able to be used as a complete analysis of some particular aspect of Simplex output (e.g., demographic, resource-related, spatial, diversity-related, multivariate, ecological model testing).
 These and all Simplex files are freely available and open source, allowing users to modify Simplex source code to suit their particular needs.
@@
 
-1. Black AJ, McKane AJ (2012) Stochastic formulation of ecological models and their applications. Trends in Ecology and Evolution. 27:337-345.
+Black AJ, McKane AJ (2012) Stochastic formulation of ecological models and their applications. Trends in Ecology and Evolution. 27:337-345.
 
-2. DeAngelis DL, Gross LJ (1992) Individual-based Models and Approaches in Ecology: Populations, Communities, and Ecosystems. Chapman & Hall. New York.
+Bertz, S (1983) On the complexity of graphs and molecules. Bulletin of Mathematical Biology 45:849–855.
 
-3. Droop MR (1983) Twenty-five years of algal growth kinetics, a personal view. Botanica Marina. 26: 99-112.
+DeAngelis DL, Gross LJ (1992) Individual-based Models and Approaches in Ecology: Populations, Communities, and Ecosystems. Chapman & Hall. New York.
 
-4. Gotelli NJ, Entsminger GL (2001) Swap and fill algorithms in null model analysis: rethinking the knight’s tour. Oecologia. 129:281–291
+Droop MR (1983) Twenty-five years of algal growth kinetics, a personal view. Botanica Marina. 26: 99-112.
 
-5.
+Gotelli NJ, Entsminger GL (2001) Swap and fill algorithms in null model analysis: rethinking the knight’s tour. Oecologia. 129:281–291
+
+Grimm V, Railsback SF (2005) Individual-based Modeling and Ecology. Princeton University Press, Princeton.
+
 Grimm V, Berger U, Bastiansen F, Eliassen S, Ginot V, Giske J, Goss-Custard J, Grand T, Heinz SK, Huse G, Huth A, Jepsen JU, Jørgensen C, Mooij WM, Mu¨ller B, Pe’er G, Pioub C, Railsback SF, Robbins AM, Robbins MM, Rossmanith E, Ru¨ger N, Strand E, Souissim S, Stillmane RA, Vabøg R, Visser U, DeAngelis DL (2006) A standard protocol for describing individual-based and agent-based models.
@@
 
-6. Grimm V, Railsback SF (2005) Individual-based Modeling and Ecology. Princeton University Press, Princeton.
+Hartl DL, Clark AG (1997) Principle of Population Genetics. Sinauer Associates, Inc. Publishers, Sunderland, MA.
 
-7. Hartl DL, Clark AG (1997) Principle of Population Genetics. Sinauer Associates, Inc. Publishers, Sunderland, MA.
+Hubbell SP (2001) The unified neutral theory of biodiversity and biogeography. Princeton University Press, Princeton, NJ.
 
-8. Hubbell SP (2001) The unified neutral theory of biodiversity and biogeography. Princeton University Press, Princeton, NJ.
+Leadbeater BSC (2006) The 'Droop equation' - Michael Droop and the legacy of the 'Cell-quota model' of phytoplankton growth. Protist. 157: 345-358.
 
-9. Leadbeater BSC (2006) The 'Droop equation' - Michael Droop and the legacy of the 'Cell-quota model' of phytoplankton growth. Protist. 157: 345-358.
+Locey K (2013) Synthesized ecological molecular neutral theory. figshare. http://dx.doi.org/10.6084/m9.figshare.640105.
 
-10. Locey K (2013) Synthesized ecological molecular neutral theory. figshare. http://dx.doi.org/10.6084/m9.figshare.640105.
+Locey KJ, Lennon JT. (2015) Scaling laws predict global microbial diversity. PeerJ PrePrints 3:e1808 https://dx.doi.org/10.7287/peerj.preprints.1451v1
 
-11. Locey KJ, Lennon JT. (2015) Scaling laws predict global microbial diversity. PeerJ PrePrints 3:e1808 https://dx.doi.org/10.7287/peerj.preprints.1451v1
+Matthews RB, Gilbert NG, Roach A, Polhill JG, Gotts NM (2007) Agent-based land-use models: a review of applications. Landscape Ecology. 22: 1447-1459.
 
-12. Matthews RB, Gilbert NG, Roach A, Polhill JG, Gotts NM (2007) Agent-based land-use models: a review of applications. Landscape Ecology. 22: 1447-1459.
+Pirt SJ (1965) The Maintenance Energy of Bacteria in Growing Cultures. Proceedings of the Royal Society of London. Series B, Biological Sciences, 163:224-231.
 
-13. Pirt SJ (1965) The Maintenance Energy of Bacteria in Growing Cultures. Proceedings of the Royal Society of London. Series B, Biological Sciences, 163:224-231.
+Rosindell J, Harmon LJ, Etienne RS (2015) Unifying ecology and macroevolution with individual-based theory. Ecology Letters. 18: 472-482.
 
-14. Rosindell J, Harmon LJ, Etienne RS (2015) Unifying ecology and macroevolution with individual-based theory. Ecology Letters. 18: 472-482.
-
-15. Succi S (2001) The Lattice Boltzmann Equation: For Fluid Dynamics and Beyond. Oxford University Press. New York.
+Succi S (2001) The Lattice Boltzmann Equation: For Fluid Dynamics and Beyond. Oxford University Press. New York.

### 2015-11-07 model/model.py
modified model.py, bide.py, and randparam.py to ensure that non-fluid movement is modeled -KL

@@ -29,8 +29,8 @@ import randparams as rp
 3.) change plot_system = 'no' to 'yes' on or near line 66
 
 Because generating animations requires computing time and memory, doing so can
-be computationally demanding. To quicken the process, comment out plt.show() on
-or near line 368.
+be computationally demanding. To quicken the process, use plot_system = 'no' on
+or near line 66.
 """
 
 # https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing
@@ -39,6 +39,7 @@ or near line 368.
 
 GenPath = mydir + 'GitHub/simplex/results/simulated_data/'
 
+"""
 ############## OPEN OUTPUT DATA FILE ###########################################
 OUT1 = open(GenPath + 'examples/SimData.csv','w')
 OUT2 = open(GenPath + 'examples/RADs.csv','w')
@@ -57,9 +58,7 @@ OUT3.close()
 OUT4.close()
 OUT5.close()
 OUT6.close()
-
-
-
+"""
 
 def nextFrame(arg):
 
@@ -68,7 +67,7 @@ def nextFrame(arg):
[…]

### 2015-11-07 tools/bide/bide.py

@@ -287,6 +287,101 @@ def fluid_movement(TypeOf, List, t_In, xAge, Xs, Ys, ux, uy, w, h, u0):
 
 
 
+def nonfluid_movement(TypeOf, motion, List, t_In, xAge, Xs, Ys, ux, uy, w, h, u0):
+
+    Type, IDs, ID, Vals = [], [], int(), []
+
+    if TypeOf == 'resource':
+        Type, IDs, ID, Vals = List
+    elif TypeOf == 'individual':
+        Type, IDs, ID, Vals, DispD, GrowthList, MaintList, N_RList, P_RList, C_RList, DispList = List
+    else:
+        IDs = List
+
+    if Xs == []:
+        if TypeOf == 'tracer':
+            return [IDs, Xs, Ys, xAge, t_In]
+        elif TypeOf == 'individual':
+            return [Type, Xs, Ys, xAge, IDs, ID, t_In, Vals, GrowthList, MaintList, N_RList, P_RList, C_RList, DispList]
+        elif TypeOf == 'resource':
+            return [Type, Xs, Ys, xAge, IDs, ID, t_In, Vals]
+
+
+    limit, distance, direction, pop = 0.1, 0, 0, 'no'
+    x, y = 0, 0
+
+    for i, val in enumerate(IDs):
+
+        # get distance
+        if TypeOf == 'individual':
+            distance = np.random.uniform(0, u0)
+        else:
+            distance = np.random.uniform(0, u0)
+
+        x, y = Xs[i], Ys[i]
+
+        # go up or down
+        #if TypeOf != 'resource':
+        if motion == 'unidirectional':
+            direction = 1
+        if motion == 'random_walk':
+            direction = choice([-1, 1])
+
+        y = y + (direction * distance)
+
+        # get distance
+        if TypeOf == 'individual':
+            distance = np.random.uniform(0, u0)
+        else:
+            distance = np.random.uniform(0, u0)
+
+        # go forward or backward
+        #if TypeOf != 'resource':
+        if motion == 'unidirectional':
+            direction = 1
+        if motion == 'random_walk':
+            direction = choice([-1, 1])
+
+        x = x + (direction * distance)
+
+        if x > w - limit or x < limit:
+            pop = 'yes'
+        elif y > h - limit or y < limit:
+            pop = 'yes'
+
+
+        if pop == 'no':
[…]

### 2015-11-07 tools/randparams/randparams.py

@@ -7,12 +7,12 @@ import os
 def get_rand_params():
     """ Get random model parameter values. Others are chosen in bide.py """
 
-    motion = choice(['fluid'])
+    motion = choice(['fluid', 'random_walk']) # 'fluid', 'unidirectional'
     D = choice([2, 2]) # number of spatial dimensions
 
-    width = choice([10, 15, 20, 30, 40, 50])
-    height = choice([5, 6, 7, 8, 9, 10])
-    barriers = choice([1,2,3,4])
+    width = choice([10, 20, 40, 80, 160])
+    height = choice([5, 10, 20, 40, 80])
+    barriers = choice([0, 0])
 
     pulse = choice([0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0])
     flux = choice(['yes'])
@@ -24,17 +24,17 @@ def get_rand_params():
     phase = choice([0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10])
     # 0 = in phase; 16 = entirely out of phase
 
-    disturb = choice([0.0001])#, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.1])
-    rates = np.array([1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.2, 0.1, 0.09, 0.08, 0.07, 0.06, 0.05, 0.04, 0.02, 0.01, 0.0075, 0.005, 0.001, 0.0005, 0.0001])  # inflow speeds
+    disturb = choice([0.00001])#, 0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.1])
+    rates = np.array([1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.09, 0.08, 0.07, 0.06, 0.05, 0.04, 0.02, 0.01, 0.0075, 0.005, 0.001, 0.0005, 0.0001])  # inflow speeds
 
-    alpha = np.random.uniform(0.99, 0.9999)
+    alpha = np.random.uniform(0.95, 0.999)
     reproduction = choice(['fission', 'sexual'])
     speciation = choice(['yes', 'no'])
 
-    seedCom = choice([100]) # size of starting community
-    m = choice([0.0, 0.0001, 0.0005, 0.001, 0.005, 0.01]) # m = probability of immigration
+    seedCom = 100 # size of starting community
[…]

### 2017-03-09 model/BIDE/bide.py
overhauled core modeling code -KL

@@ -1,0 +1,518 @@
+# -*- coding: utf-8 -*-
+from __future__ import division
+from random import randint, choice
+import numpy as np
+import sys
+from math import sqrt, pi
+
+
+def get_closest(RIDs, RX, RY, RZ, Rtypes, coords):
+    closest = 0
+    if len(RIDs) == 0:
+        return closest
+
+    x1, y1, z1 = coords
+    Try = min([20, len(RIDs)])
+    minDist, ct = 10**10, 0
+
+    while ct < Try:
+        ct += 1
+        j = randint(0, len(RIDs)-1)
+        x = RX[j]
+        y = RY[j]
+        z = RZ[j]
+
+        dist = sqrt((x1 - x)**2 + (y1 - y)**2 + (z1 - z)**2)
+
+        if dist < minDist:
+            minDist = dist
+            closest = RIDs[j]
+
+        return closest
+
+
+def decomposition(i, n):
+    gn = choice([[1/3, 1/3, 1/3], [0.5, 0.3, 0.2], [0.6, 0.2, 0.2],
+                [0.8, 0.1, 0.1], [0.99, 0.005, 0.005], [0.9, 0.05, 0.05]])
+    return gn
+
+
+def get_color(ID, colorD): # FUNCTION TO ASSIGN COLORS TO Sp_
+
+    r1 = lambda: randint(0,255)
+    r2 = lambda: randint(0,255)
+    r3 = lambda: randint(0,255)
+
+    color = '#%02X%02X%02X' % (r1(),r2(),r3())
+    colorD[ID] = color
+
+    return colorD
+
+
+
+def ResIn(RTypes, RVals, RX, RY, RZ, RID, RIDs, params, u1):
+
+    w, h, l, seed, m, r, nN, rmax, gmax, maintmax, dmax, amp, freq, phase, rates, pmax, dormlim, smax = params
+
+    for i in range(r):
+        x = np.random.binomial(1, u1)
+
+        if x == 1:
+            rval = np.random.uniform(1, rmax)
+            rtype = randint(0, nN-1)
+            RTypes.append(rtype)
+            RVals.append(rval)
+            RIDs.append(RID)
+            RID += 1
+
+            RX.append(float(np.random.uniform(0, 0.9*h)))
+            RY.append(float(np.random.uniform(0, 0.9*l)))
+            RZ.append(float(np.random.uniform(0, 0.9*w)))
[…]

### 2017-03-09 model/main_script/model_3D_Movie.py

@@ -1,0 +1,310 @@
+from __future__ import division
+
+import matplotlib.animation as animation
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+
+from random import choice, shuffle #, randint
+import numpy as np
+from numpy import sin, pi, mean
+import sys
+import os
+import time
+
+mydir = os.path.expanduser("~/")
+sys.path.append(mydir + "GitHub/residence-time/model/metrics")
+import metrics
+sys.path.append(mydir + "GitHub/residence-time/model/bide")
+import bide2 as bide
+sys.path.append(mydir + "GitHub/residence-time/model/randparams")
+import randparams as rp
+sys.path.append(mydir + "GitHub/residence-time/model/spatial")
+
+""" To generate movies:
+1.) uncomment line 'ani.save' on or near line 364
+2.) adjust the frames variable on or near line 364, to change movie length
+3.) change plot_system = 'no' to 'yes' on or near line 66
+
+Because generating animations requires computing time and memory, doing so can
+be computationally demanding. To quicken the process, use plot_system = 'no' on
+or near line 66.
+"""
+
+# https://www.quantstart.com/articles/Basics-of-Statistical-Mean-Reversion-Testing
+# http://statsmodels.sourceforge.net/0.5.0/generated/statsmodels.tsa.stattools.adfuller.html
+
+
+GenPath = mydir + 'GitHub/residence-time/results/simulated_data/'
+
+
+OUT = open(GenPath + 'SimData.csv','w+')
[…]

### 2017-03-09 tools/DiversityTools/FeasibleFunctions/feasible_functions.pyx

@@ -1,0 +1,333 @@
+from __future__ import division
+
+import sys
+import os
+from os import path, access, R_OK  # W_OK for write permission
+
+mydir = os.path.expanduser("~/GitHub/rare-bio/")
+mydir2 = os.path.expanduser("~/Desktop/")
+
+sys.path.append(mydir + "/tools/macroecotools")
+import macroecotools
+sys.path.append(mydir + "/tools/partitions")
+import partitions as parts
+
+import  matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid.inset_locator import inset_axes
+
+import numpy as np
+from scipy import stats
+from scipy.stats import gaussian_kde
+
+import random
+from random import choice
+import decimal
+import re
+
+import math
+
+########################################################################################################
+######   A Section devoted to evenness indices and descriptive statistical functions ###################
+
+"""
+    Modified from Locey and White (2013)
+    (insert reference)
+
+"""
+def min_max(n,s):
+
+    _min = int(math.floor(float(n)/float(s)))
+    if int(n%s) > 0:
+        _min +=1
+
+    return _min
+
+
+def get_modes(_list,which):
+
+    modes = []
+    for i in _list:
+        _mode = np.mode(i)
+        if which is 'high':
+            modes.append(int(max(_mode)))
+        if which is 'low':
+            modes.append(int(min(_mode)))
+
+    return modes
+
+
+def get_modal(_list):
+
+    """ Finds the mode from a kernel density function across a sample """
+    exp_mode = 0.0
+    density = gaussian_kde(_list)
+    n = len(_list)
+    xs = np.linspace(min(_list),max(_list),n)
+    density.covariance_factor = lambda : .001
+    density._compute_covariance()
+    D = [xs,density(xs)]
+    d = 0
+    maxd = 0.0
+    while d < len(D[1]):
+        if D[1][d] > maxd:
+            maxd = D[1][d]
+            exp_mode = D[0][d]
+        d += 1
+    return exp_mode
+
+def get_kdens_choose_kernel(_list,kernel):
+    """ Finds the kernel density function across a sample of SADs """
+    density = gaussian_kde(_list)
+    n = len(_list)
+    xs = np.linspace(min(_list),max(_list),n)
+    #xs = np.linspace(0.0,1.0,n)
[…]

### 2017-03-09 tools/DiversityTools/PredRadDATA/PredRadDATA_ALL.pyx

@@ -1,0 +1,129 @@
+from __future__ import division
+import sys
+import os
+import random
+import numpy as np
+from scipy import stats
+
+mydir = os.path.expanduser("~/GitHub/rare-bio/")
+mydir2 = os.path.expanduser("~/Desktop/")
+
+sys.path.append(mydir + "/tools/GenAnalysis/tools/")
+import macroecotools
+import macroeco_distributions
+import predRADs
+import mete
+import pln
+
+sys.path.append(mydir + "tools/feasible_functions")
+import feasible_functions as ff
+
+
+datasets = []
+for name in os.listdir(mydir2 +'data/micro'):
+        datasets.append([name, 'micro'])
+
+for name in os.listdir(mydir2 +'data/macro'):
+        datasets.append([name, 'macro'])
+
+
+ct = 0
+numMicros = 0
+numMacros = 0
+numEMP = 0
+numEMPopen = 0
+
+models = ['METE', 'MaxEntGeomSeries', 'IntegerPartitions', 'Zipf', 'PoissonLogNormal']
+
+for dataset in datasets:
+
+    name = dataset[0] # name of dataset
+    if name == '.DS_Store': continue
+
+    kind = dataset[1] # micro or macro
+
+    for model in models:
+        OUT = open(mydir2 + 'data/'+kind+'/'+name+'/'+name+'-'+model+'_SADMetricData_NoMicrobe1s.txt','w+')
+        #OUT = open(mydir2 + 'data/'+kind+'/'+name+'/'+name+'-'+model+'_SADMetricData.txt','w+')
+        RADs = []
+
+        if kind == 'macro':
+            path = kind
+            RADs = ff.get_SADs(mydir2 +'/data/'+path, name)
+            print 'macro', name, len(RADs)
+
+        if kind == 'micro':
+            path = kind
+            RADs = ff.GetSADsFromBiom_labeled(mydir2 +'data/'+path+'/EMPopen/', name)
+            print 'micro', name, len(RADs)
+
+        ct = 0
+        numRADs = len(RADs)
+        for RAD in RADs:
+
+            if kind == 'micro':
+                RAD = list([x for x in RAD if x > 1])
+
+            elif kind == 'macro':
+	            RAD = list([x for x in RAD if x > 0])
+
+
+            N = sum(RAD)
+            S = len(RAD)
+
+            if S < 2 or N < 10: continue
+
+            if model == 'MaxEntGeomSeries':
+                # Geometric Series for N and S
[…]

### 2017-03-09 tools/DiversityTools/PredRadDATA/PredRadDATA_EMP.py

@@ -1,0 +1,60 @@
+from __future__ import division
+import sys
+import os
+#import random
+import numpy as np
+from scipy import stats
+
+mydir = os.path.expanduser("~/Desktop/Repos/rare-bio/")
+sys.path.append(mydir + "tools/feasible_functions")
+import feasible_functions as ff
+
+mydir2 = os.path.expanduser("~/Desktop/")
+OUT = open(mydir + 'output/EMPopen-RADdata.txt','w+')
+
+ct = 0
+numEMP = 0
+numEMPopen = 0
+RADs = []
+    
+RADs = ff.GetSADsFromBiom_labeled(mydir2 +'data/micro/EMPopen', 'EMPopen')
+
+for RAD in RADs:
+    
+    RAD = list([x for x in RAD if x != 0])
+        
+    N = int(sum(RAD))
+    S = len(RAD)
+    
+    if S < 2: continue 
+    
+    Evar = ff.e_var(RAD)
+    ESimp = ff.simpsons_evenness(RAD)
+    ENee = ff.NHC_evenness(RAD)
+    EPielou = ff.pielous_evenness(RAD)
+            
+    EHeip = ff.Heips_evenness(RAD)
+    EQ = ff.EQ_evenness(RAD)
+            
+    BP = ff.Berger_Parker(RAD)
+    SimpDom = ff.simpsons_dom(RAD)
+            
+    rareRel = ff.rarityRel(RAD)
+    rareOnes = ff.rarityOnes(RAD)
+            
+    skew = float(np.log(stats.skew(RAD)))
+    
+    Var = np.var(np.array(RAD), ddof=1)
+    
+    ct += 1
+    
+    print>>OUT, N, S, Evar, ESimp, ENee, EHeip, EQ, EPielou, BP, SimpDom, rareRel, rareOnes, skew
+    print N, S, len(RADs) - ct
+    
+print 'micros:', len(RADs)
+    
+OUT.close()
+            
+            
+            
+            

### 2017-03-09 tools/DiversityTools/SSADdata/SSADdata/SSADdata.pyx

@@ -1,0 +1,156 @@
+from __future__ import division
+import sys
+import os
+#import random
+import numpy as np
+from scipy import stats
+
+
+mydir = os.path.expanduser("~/")
+sys.path.append(mydir + "tools/metrics")
+import metrics
+
+#OUT1 = open("/GitHub/rare-bio/output/EMPclosed-SSAD-ResultsTable.txt",'w+')
+
+
+def get_EMP_SSADs():
+
+    DATA = mydir + "data/micro/EMPopen/EMPopen-SSADdata.txt"
+
+    SSADdict = {}
+
+    with open(DATA) as f:
+
+        for d in f:
+            if d.strip():
+
+                d = d.split()
+                species = d[0]
+                #sample = d[1]
+                abundance = float(d[2])
+
+                if abundance > 0:
+                    if species in SSADdict:
+                        SSADdict[species].append(abundance)
+                    else:
+                        SSADdict[species] = [abundance]
+
+
+
+    SSADs = []
+    SSADlist = SSADdict.items()
+
+    for tup in SSADlist:
+
+        SSAD = tup[1]
+        if len(SSAD) >= 1:
+            SSAD.sort()
+            SSAD.reverse()
+            SSADs.append(SSAD)
+
+    return SSADs
+
+
+
+def get_Macro_SSADs():
+
+    for name in os.listdir('/data/macro'):
+
+        if name == 'BCI': continue
+        if name == '.DS_Store': continue
+
+        else:
+            print name
+
+            SSADdict = {}
+
+            path ='/data/macro/'+name+'/'+name
+            OUT = open(path + '-SSADMetricData.txt','w+')
+            DATA = path+'-data.txt'
+
+            with open(DATA) as f:
+
+                for d in f:
+                    if d.strip():
+                        d = d.split()
+
+                        if name == 'GENTRY':
+                            species = name + d[1] # Dataset name plus species identifier
+                            abundance = float(d[2])
+
+                        else:
+                            #site = d[0]
+                            #year = d[1]
+                            species = name + d[2] # Dataset name plus species identifier
+                            abundance = float(d[3])
+
+                        if abundance > 0:
+                            if species in SSADdict:
+                                SSADdict[species].append(abundance)
+                            else:
[…]

### 2017-03-09 tools/DiversityTools/SpeciesBySite/SpeciesBySite.pyx

@@ -1,0 +1,58 @@
+import os
+
+mydir = os.path.expanduser("~/data")
+
+OrC = 'closed'
+
+def GetSADsFromBiom_labeled(path, dataset):
+
+    minS = 2
+
+    IN = path + '/' + dataset + '-SSADdata.txt'
+    n = sum(1 for line in open(IN))
+
+    SpeciesDict = {}
+
+    print 'Starting build of SpeciesDict'
+    with open(IN) as f:
+
+        for d in f:
+
+            #print 'Reading in SSAD data. Lines left:', n
+            #n -= 1
+
+            if d.strip():
+
+                d = d.split()
+                species = d[0]
+                sample = d[1]
+                abundance = float(d[2])
+
+                if abundance > 0:
+                    if sample not in SpeciesDict:
+
+                        SpeciesDict[sample] = [species]
+
+                    else:
+                        SpeciesDict[sample].append(species)
+
+    print 'Finished building SpeciesDict'
+
+
+    OUT = open(path + '/' + dataset + '-SbyS.txt','w+')
+
+    SpeciesLists = SpeciesDict.items()
+    n = len(SpeciesLists)
+
+    for i, site in enumerate(SpeciesLists):
+
+        if len(site) >= minS:
+            print n - i
+            print >> OUT, site
+
+    OUT.close()
+
+    print 'Finished generating SbyS file'
+    return
+
+GetSADsFromBiom_labeled(mydir +'/micro/EMP'+OrC, 'EMP'+OrC)

### 2017-03-09 tools/DiversityTools/SpeciesBySite/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("SpeciesBySite.pyx"))

### 2017-03-09 tools/DiversityTools/expS_vs_obsN/expS_vs_obsN.pyx

@@ -1,0 +1,289 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+
+import numpy as np
+import random
+from random import randrange, sample
+import scipy as sc
+from scipy import stats
+
+import os
+import sys
+from scipy.stats.distributions import t
+
+import statsmodels.stats.api as sms
+import statsmodels.api as sm
+import statsmodels.formula.api as smf
+from statsmodels.sandbox.regression.predstd import wls_prediction_std
+from statsmodels.stats.outliers_influence import summary_table
+
+import itertools as it
+import pandas as pd
+from math import log10
+import linecache
+
+mydir = os.path.expanduser("~/Desktop/Repos/rare-bio/")
+mydir2 = os.path.expanduser("~/Desktop/")
+
+
+
+""" This code creates a figures of three subplots. First, there is a plot of
+observed species richness versus observed total abundance. Second, there is a
+plot estimated richness versus number of samples. Third, there is a subplot of
+expected richness versus dominance. """
+
+
+def EstimateS(SiteList):
+    m = len(SiteList)
+    m_inf = 0
+    SpDict = {}
+
+    for site in SiteList:
+        if min(site) <= 10: m_inf += 1
+
+        for sp in site:
+            if sp in SpDict:
+                SpDict[sp] += 1
+            else: SpDict[sp] = 1
+
+    IncVals = SpDict.values()
+    S = len(IncVals)
+    qs = [0]*10
+
+    for i, q in enumerate(qs):
+        qs[i] = IncVals.count(i+1)
+
+    # Chao2
+    q1 = qs[0]
+    q2 = qs[1]
+    chao2 = S + (((m-1)/m) * ((q1*(q1-1)) / (2*(q2+1))))
+
+    var = 'und'
+    if q1 > 0 and q2 > 0:
+        var = q2 * (0.5*(q1/q2)**2 + (q1/q2)**3 + 0.25*(q1/q2)**4)
+
+    # ICE
+    num = 0
+    n_inf = 0
+    for i, qk in enumerate(qs):
+        num += (i+1)*i*qk
+        n_inf += (i+1)*qk
[…]

### 2017-03-09 tools/DiversityTools/getRADs/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("getRADs.pyx"))

### 2017-03-09 tools/DiversityTools/macroeco_distributions/tests_macroeco_distributions.py

@@ -1,0 +1,87 @@
+"""Tests for the macroeco_distributions module"""
+
+from __future__ import division
+from macroeco_distributions import *
+import nose
+from nose.tools import assert_almost_equals, assert_equals
+from math import log
+
+dist_dic = {'pln': pln, 'trunc_logser': trunc_logser, 'trunc_expon': trunc_expon, 
+            'trunc_pareto': trunc_pareto, 'trunc_geom': trunc_geom, 'trunc_nbinom': nbinom_lower_trunc, 
+            'trunc_weibull': trunc_weibull, 'trunc_geom_zeros': trunc_geom_with_zeros}
+
+dist_solver_dic = {'pln': pln_solver, 'logser': logser_solver, 'trunc_logser': trunc_logser_solver, 
+                   'trunc_geom': trunc_geom_solver, 'trunc_expon': trunc_expon_solver, 
+                   'trunc_pareto': trunc_pareto_solver, 'zipf': zipf_solver,
+                   'trunc_nbinom': nbinom_lower_trunc_solver}
+
+# Testing SAD data from BBS (5-year total for one route)
+# Parameter values obtained by using the solvers in macroeco_distributions ver. 4e5fc52
+sad_1 = [109, 14, 4, 4, 680, 195, 13, 3, 123, 116, 1, 5, 105, 26, 14, 2, 9, 29, 15, 133, 5, 41, 45, 33, 
+         17, 27, 37, 11, 169, 1, 27, 7, 19, 23, 100, 4, 8, 5, 19, 1, 21, 12, 6, 1, 10, 2, 1, 94, 2, 4, 28, 1, 3, 
+         34, 3, 20, 72, 21, 1, 84, 10, 528, 18, 1, 1, 10, 10, 48, 7]
+
+sad_1_pars_dic = {'pln': (2.27027, 1.83484), 'logser': 0.99621, 'trunc_logser': 0.99621, 
+                   'trunc_geom': 0.021218, 'trunc_expon': 0.021678, 
+                   'trunc_pareto': 0.38849, 'zipf': 1.32303, 'trunc_nbinom': (0.12298, 0.0054890)}
+
+sad_2 = [71, 27, 88, 21, 1, 2, 6, 2, 54, 3, 4, 33, 21, 12, 36, 29, 11, 14, 3, 3, 1, 10, 1, 28, 2, 119, 235, 
[…]

### 2017-03-09 tools/DiversityTools/metrics/MetricTimeTest/metricsTimeTest.pyx

@@ -1,0 +1,153 @@
+from __future__ import division
+import sys
+import os
+import random
+import numpy as np
+from scipy import stats
+import time
+
+mydir = os.path.expanduser("~/")
+
+sys.path.append(mydir + "tools/metrics")
+import metrics
+sys.path.append(mydir + "tools/getRADs")
+import getRADs
+
+def getMetrics():
+    datasets = []
+    for name in os.listdir(mydir +'data/micro'):
+            datasets.append([name, 'micro'])
+    for name in os.listdir(mydir +'data/macro'):
+            datasets.append([name, 'macro'])
+
+
+    for dataset in datasets:
+
+        name = dataset[0] # name of dataset
+        kind = dataset[1] # micro or macro
+
+        print name
+        xnames = ['AGSOIL', '.DS_Store', 'CATLIN', 'CHU', 'EMPclosed']
+        if name in xnames: continue
+
+        RADs = []
+
+        if kind == 'macro':
+            RADs = getRADs.get_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+            print 'macro', name, len(RADs)
+
+
+        if kind == 'micro':
+            #RADs = getPhyloSADs.get_SADs(seqID, dataset[0], 'genus')
+
+            if name == 'EMPclosed' or name == 'EMPopen':
+                RADs = getRADs.EMP_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+
+            else:
+                RADs = getRADs.get_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+
+        ct = 0
+        numRADs = len(RADs)
+        for RAD in RADs:
+
+            if kind == 'micro':
+                RAD = list([x for x in RAD if x > 0])
+
+            elif kind == 'macro':
+                RAD = list([x for x in RAD if x > 0])
+
+
+            N = sum(RAD)
+            S = len(RAD)
+
+            if S < 10 or N < 11: continue
+
+            times = []
+            # Evenness
+            t = time.clock()
+            Evar = metrics.e_var(RAD)
+            t = time.clock() - t
+            times.append(['Evar', t])
+
+            t = time.clock()
+            ESimp = metrics.e_simpson(RAD)
+            t = time.clock() - t
+            times.append(['Simpsons Evenness', t])
+
+            t = time.clock()
+            EQ = metrics.EQ(RAD)
+            t = time.clock() - t
+            times.append(['EQ', t])
+
+            t = time.clock()
+            O = metrics.OE(RAD)
+            t = time.clock() - t
[…]

### 2017-03-09 tools/DiversityTools/metrics/metrics.pyx

@@ -1,0 +1,520 @@
+from __future__ import division
+import sys
+import numpy as np
+from scipy import stats
+from scipy.stats import gaussian_kde
+import mpmath as mpm
+from scipy.optimize import fsolve
+import re
+import random
+import os
+import math
+
+mydir = os.path.expanduser("~/GitHub/rare-bio")
+data = os.path.expanduser("~/data")
+
+################################################################################
+#### Section of evenness indices and descriptive statistical functions #########
+
+
+############### RARITY #########################################################
+def r_singletons(sad): # relative abundance of the least abundant taxon
+
+    """ percent of species represented by a single individual """
+
+    return 100 * sad.count(1)/len(sad)
+
+
+def p_ZPtOne(sad):
+
+    """ percent taxa with less than 0.1% N """
+    N = sum(sad)
+    S = len(sad)
+
+    sad = np.array(sad)/N
+    sad = sad*100
+    sad = sad.tolist()
+
+    numR = 0
+    for sp in sad:
+        if sp < 0.1:
+            numR += 1
+
+    return numR/S
+
+
+def Rlogskew(sad):
+    S = len(sad)
+
+    if S <= 2.0:
+        print 'S < 2, cannot compute log-skew'
+        sys.exit()
+
+    sad = np.log10(sad)
+    mu = np.mean(sad)
+
+    num = 0
+    denom = 0
+    for ni in sad:
+        num += ((ni - mu)**3.0)/S
+        denom += ((ni - mu)**2.0)/S
+
+    t1 = num/(denom**(3.0/2.0))
+    t2 = (S/(S - 2.0)) * np.sqrt((S - 1.0)/S)
+
+    return t1 * t2
+
+
+############### LOGNORMAL VARIABLES ############################################
+
+def Preston(sad):
+    N = sum(sad)
+    Nmax = max(sad)
+
+    left = (2 * N)/(np.sqrt(np.pi) * Nmax)
+
+    func = lambda a : left - (math.erf(np.log(2)/a) / a)
+
+    guess = 0.1 # alpha is often ~0.2, but appears to be lower for larger N
+    a = fsolve(func, guess)
+
+    expS = (np.sqrt(np.pi) / a) * np.exp( (np.log(2)/(2*a))**2 )
+
[…]

### 2017-03-09 tools/DiversityTools/partitions/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("partitions.pyx"))

### 2017-03-09 tools/DiversityTools/predRADs/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("predRADs.pyx"))

### 2017-03-09 tools/DiversityTools/radMetrics/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("radMetrics.pyx"))

### 2017-03-09 tools/SADfits/SADfits.py

@@ -1,0 +1,138 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+import sys
+import os
+from random import shuffle
+import numpy as np
+
+########### PATHS ##############################################################
+
+mydir = os.path.expanduser("~/GitHub/residence-time")
+tools = os.path.expanduser(mydir + "/tools")
+
+sys.path.append(tools + "/DiversityTools/macroeco_distributions")
+import macroeco_distributions as md
+sys.path.append(tools + "/DiversityTools/distributions")
+import distributions as dist
+sys.path.append(tools + "/DiversityTools/macroecotools")
+import macroecotools as mct
+sys.path.append(tools + "/metrics")
+import metrics
+sys.path.append(tools + "/DiversityTools/mete")
+import mete
+#sys.path.append(tools + "/pln")
+#import pln
+
+from scipy.stats.kde import gaussian_kde
+from macroeco_distributions import pln, pln_solver
+from numpy import empty
+
+
+def get_kdens_choose_kernel(_list,kernel):
+    """ Finds the kernel density function across a sample of SADs """
+    density = gaussian_kde(_list)
+    n = len(_list)
+    xs = np.linspace(min(_list),max(_list),n)
+    #xs = np.linspace(0.0,1.0,n)
+    density.covariance_factor = lambda : kernel
+    density._compute_covariance()
+    D = [xs,density(xs)]
+    return D
+
+
+
+def get_rad_pln(S, mu, sigma, lower_trunc = True):
+    """Obtain the predicted RAD from a Poisson lognormal distribution"""
+    abundance = list(empty([S]))
+    rank = range(1, int(S) + 1)
+    cdf_obs = [(rank[i]-0.5) / S for i in range(0, int(S))]
+    j = 0
+    cdf_cum = 0
+    i = 1
+    while j < S:
+        cdf_cum += pln.pmf(i, mu, sigma, lower_trunc)
+        while cdf_cum >= cdf_obs[j]:
+            abundance[j] = i
+            j += 1
+            if j == S:
+                abundance.reverse()
+                return abundance
+        i += 1
+
+
+
+def get_rad_from_obs(ab, dist):
[…]

### 2017-03-09 tools/bide/bide.py

@@ -3,11 +3,12 @@ from __future__ import division
 from random import randint, choice
 import numpy as np
 import sys
+import math
 #from math import modf
 #import decimal
 import time
 
-limit = 0.2
+limit = 0.1
 
 def coord(d):
     return float(np.random.uniform(0.1*d, 0.9*d))
@@ -17,14 +18,14 @@ def GetIndParam(means):
     vals = []
 
     if isinstance(means, float) or isinstance(means, int):
-        std = means/10.0
+        std = means/1000.0
         vals = np.random.normal(means, std)
         if vals < 0.00001:
             vals = 0.00001
 
     else:
         for val in means:
-            std = val/10.0
+            std = val/1000.0
             i = np.random.normal(val, std)
             if i < 0.00001:
                 i = 0.00001
@@ -37,6 +38,7 @@ def GetIndParam(means):
 def GetRAD(vector):
     RAD = []
     unique = list(set(vector))
+
     for val in unique:
         RAD.append(vector.count(val)) # the abundance of each Sp_
 
@@ -56,20 +58,29 @@ def get_color(ID, colorD): # FUNCTION TO ASSIGN COLORS TO Sp_
 
 
 
-def NewTracers(motion, IDs, Xs, Ys, t_In, w, h, u0):
+def NewTracers(numt, motion, IDs, Xs, Ys, t_In, w, h, u0):
 
-    x = np.random.binomial(1, u0)
-    if x == 1:
-        IDs.append(0)
-        t_In.append(0)
-        Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
-        Xs.append(float(np.random.uniform(0.1*w, 0.9*w)))
+    for t in range(numt):
+        x = np.random.binomial(1, u0/2)
+
+        if x == 1:
+            IDs.append(0)
+            t_In.append(0)
+
+            if motion == 'brown_noise' or motion == 'white_noise':
+                Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+                Xs.append(float(np.random.uniform(0.1*w, 0.9*w)))
+
+            else:
+                Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+                Xs.append(float(np.random.uniform(0.1*w, 0.15*w)))
 
     return [IDs, t_In, Xs, Ys]
 
 
 
[…]

### 2017-03-09 tools/metrics/metrics.py

@@ -1,11 +1,74 @@
 from __future__ import division
+import  matplotlib.pyplot as plt
 import sys
 import numpy as np
+from numpy import log10
 from scipy import stats
 from scipy.stats import gaussian_kde
 from scipy.optimize import fsolve
 import math
 
+
+
+def count_pts_within_radius(x, y, radius, logscale=0):
+    """Count the number of points within a fixed radius in 2D space"""
+    #TODO: see if we can improve performance using KDTree.query_ball_point
+    #http://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.KDTree.query_ball_point.html
+    #instead of doing the subset based on the circle
+    raw_data = np.array([x, y])
+    x = np.array(x)
+    y = np.array(y)
+    raw_data = raw_data.transpose()
+
+    # Get unique data points by adding each pair of points to a set
+    unique_points = set()
+    for xval, yval in raw_data:
+        unique_points.add((xval, yval))
+
+    count_data = []
+    for a, b in unique_points:
+        if logscale == 1:
+            num_neighbors = len(x[((log10(x) - log10(a)) ** 2 +
+                                   (log10(y) - log10(b)) ** 2) <= log10(radius) ** 2])
+        else:
+            num_neighbors = len(x[((x - a) ** 2 + (y - b) ** 2) <= radius ** 2])
+        count_data.append((a, b, num_neighbors))
+    return count_data
+
+
+
+def plot_color_by_pt_dens(x, y, radius, loglog=0, plot_obj=None):
+    """Plot bivariate relationships with large n using color for point density
+
+    Inputs:
+    x & y -- variables to be plotted
+    radius -- the linear distance within which to count points as neighbors
+    loglog -- a flag to indicate the use of a loglog plot (loglog = 1)
+
+    The color of each point in the plot is determined by the logarithm (base 10)
+    of the number of points that occur with a given radius of the focal point,
+    with hotter colors indicating more points. The number of neighboring points
+    is determined in linear space regardless of whether a loglog plot is
+    presented.
+
+    """
+    plot_data = count_pts_within_radius(x, y, radius, loglog)
+    sorted_plot_data = np.array(sorted(plot_data, key=lambda point: point[2]))
+
+    if plot_obj == None:
+        plot_obj = plt.axes()
+
[…]

### 2017-03-09 tools/analyses/Rmd/BasicAnalysis.Rmd
modified model.py, bide.py, and randparam.py to ensure that non-fluid movement is modeled -KL
added overhauled modeling code -KL

@@
 
-# Replacing 0's with 1's to allow log-transforms below
 simplex.dat <- simplex.dat[complete.cases(simplex.dat), ]
+simplex.dat$tau <- (simplex.dat(width)*simplex.dat(height))/simplex.dat(flow.rate)
 
@@
 phys.dat <- as.matrix(subset(simplex.dat, 
-                      select = c(flow.rate,
+                      select = c(tau
+                                 flow.rate,
                                  barriers,

### 2017-03-09 fig-scripts/ExaminePhi.py

@@ -1,0 +1,117 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import os
+import sys
+from scipy import stats
+
+
+mydir = os.path.expanduser('~/GitHub/residence-time')
+sys.path.append(mydir+'/tools')
+mydir2 = os.path.expanduser("~/")
+
+df = pd.read_csv(mydir + '/results/simulated_data/SimData.csv')
+
+df2 = pd.DataFrame({'width' : df['width']})
+df2['flow'] = df['flow.rate']
+df2['tau'] = (df2['width']**3)/df2['flow']
+df2['dil'] = np.log10(1/df2['tau'])
+
+df2['N'] = df['total.abundance']
+df2['S'] = df['species.richness']
+df2['Prod'] = df['ind.production']
+df2['E'] = np.log10(df['simpson.e'])
+df2['W'] = np.log10(df['Whittakers.turnover'])
+df2['Dorm'] = df['Percent.Dormant']
+
+df2['AvgG'] = df['avg.per.capita.growth']
+df2['AvgDisp'] = df['avg.per.capita.active.dispersal']
+df2['AvgRPF'] = df['avg.per.capita.RPF']
+df2['AvgE'] = df['avg.per.capita.N.efficiency']
+df2['AvgMaint'] = df['avg.per.capita.maint']
+df2['MF'] = df['avg.per.capita.MF']/np.mean(df['avg.per.capita.MF'])
+
+E = 0.04
+'''
+df2['P'] = df2['AvgMaint'] * df2['AvgRPF'] * df2['MF']
+df2['G'] = df2['AvgG'] * df2['AvgDisp'] * df2['AvgE']
+df2['phi'] = np.log10(df2['P'] / (df2['G'] + E))
+'''
+
+df2['P'] = (1/df2['AvgMaint']) * (1-df2['AvgRPF']) * df2['MF']
+df2['G'] = (df2['AvgG'] * df2['AvgDisp'])
+df2['phi'] = np.log10(df2['G'] / df2['P'])
+
+df2 = df2.replace([np.inf, -np.inf], np.nan).dropna()
+
+#exp_phi = np.mean(df2['phi']) # average of logs
+#exp_dil  = np.mean(df2['dil']) # log of the average
+
[…]

### 2017-03-09 fig-scripts/ExaminePhiFig.py

@@ -1,0 +1,104 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+import os
+import sys
+#from scipy import stats
+
+def obs_pred_rsquare(obs, pred):
+
+    """Determines the prop of variability in a data set accounted for by a model
+
+    In other words, this determines the proportion of variation explained by
+    the 1:1 line in an observed-predicted plot.
+
+    """
+    return 1 - sum((obs - pred) ** 2) / sum((obs - np.mean(obs)) ** 2)
+
+
+mydir = os.path.expanduser('~/GitHub/residence-time')
+sys.path.append(mydir+'/tools')
+mydir2 = os.path.expanduser("~/")
+
+df = pd.read_csv(mydir + '/results/simulated_data/SimData.csv')
+
+df2 = pd.DataFrame({'width' : df['width']})
+df2['flow'] = df['flow.rate']
+df2['tau'] = (df2['width']**3)/df2['flow']
+df2['dil'] = np.log10(1/df2['tau'])
+
+df2['N'] = df['total.abundance']
+df2['S'] = df['species.richness']
+df2['Prod'] = df['ind.production']
+df2['Active'] = (100 - df['Percent.Dormant'])/100
+
+df2['AvgG'] = df['active.avg.per.capita.growth']
+df2['AvgDisp'] = df['active.avg.per.capita.active.dispersal']
+df2['AvgRPF'] = df['dormant.avg.per.capita.RPF']
+df2['AvgE'] = 1 - df['active.avg.per.capita.N.efficiency']
+df2['AvgMaint'] = df['active.avg.per.capita.maint']
+df2['MF'] = df['dormant.avg.per.capita.MF']/np.max(df['dormant.avg.per.capita.MF'])
+
+E = 0.06
+
+
+
+fs = 8
+gd = 20
+mct = 1
+binz = 'log'
+fig = plt.figure()
+fig.add_subplot(3, 3, 1)
+
+df2['P'] = np.log10((df2['AvgRPF']) * df2['AvgMaint'] * 1/df2['MF'] * df2['Active'])
+df2 = df2.replace([np.inf, -np.inf], np.nan).dropna()
+
[…]

### 2017-03-09 fig-scripts/Fig1-2-line.py

@@ -1,0 +1,232 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+import pandas as pd
+from random import randint
+import numpy as np
+import os
+import statsmodels.api as sm
+import sys
+
+p = 1
+fr = 0.2
+_lw = 0.5
+w = 1
+sz = 5
+
+
+mydir = os.path.expanduser('~/GitHub/residence-time')
+df = pd.read_csv(mydir + '/results/simulated_data/SimData.csv')
+
+df2 = pd.DataFrame({'width' : df['width'].groupby(df['ct']).mean()})
+
+df2['sim'] = df['sim'].groupby(df['ct']).mean()
+df2['flow'] = df['flow.rate'].groupby(df['ct']).mean()
+df2['tau'] = np.log10(df2['width']**p/df2['flow'])
+
+df2['R'] = df['resource.particles'].groupby(df['ct']).mean()
+df2['N'] = df['total.abundance'].groupby(df['ct']).mean()
+df2['Prod'] = df['ind.production'].groupby(df['ct']).mean()
+df2['S'] = df['species.richness'].groupby(df['ct']).mean()
+df2['E'] = df['simpson.e'].groupby(df['ct']).mean()
+df2['W'] = df['Whittakers.turnover'].groupby(df['ct']).mean()
+df2['Dorm'] = df['Percent.Dormant'].groupby(df['ct']).mean()
+
+df2['Grow'] = df['active.avg.per.capita.growth'].groupby(df['ct']).mean()
+df2['Maint'] = np.log10(df['active.avg.per.capita.maint'].groupby(df['ct']).mean())
+df2['Disp'] = df['active.avg.per.capita.active.dispersal'].groupby(df['ct']).mean()
+df2['RPF'] = df['dormant.avg.per.capita.rpf'].groupby(df['ct']).mean()
+df2['Eff'] = df['active.avg.per.capita.efficiency'].groupby(df['ct']).mean()
+df2['MF'] = df['active.avg.per.capita.mf'].groupby(df['ct']).min()
+
+#df2 = df2.replace([np.inf, -np.inf], np.nan).dropna()
+
+#### plot figure ###############################################################
[…]

### 2017-03-09 model/LBM/LBM.py

@@ -1,0 +1,147 @@
+from __future__ import division
+import numpy as np
+
+""" This code was initially inspired and is almost entirely based on Dan
+Schroeder's http://physics.weber.edu/schroeder/fluids/, obtained in May 2014 """
+
+
+def SetLattice(u0, viscosity, width, height, lefts, bottoms, barriers): # Lattice Boltzmann PARAMETERS
+
+    four9ths = 4.0/9.0    # abbreviations for lattice-Boltzmann weight factors
+    one9th   = 1.0/9.0
+    one36th  = 1.0/36.0
+
+    ############## particle densities along 9 directions
+    n0 = four9ths * (np.ones((height,width)) - 1.5*u0**2)
+    nN = one9th * (np.ones((height,width)) - 1.5*u0**2)
+    nS = one9th * (np.ones((height,width)) - 1.5*u0**2)
+    nE = one9th * (np.ones((height,width)) + 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+    nW = one9th * (np.ones((height,width)) - 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+    nNE = one36th * (np.ones((height,width)) + 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+    nSE = one36th * (np.ones((height,width)) + 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+    nNW = one36th * (np.ones((height,width)) - 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+    nSW = one36th * (np.ones((height,width)) - 3*u0 + 4.5*u0**2 - 1.5*u0**2)
+
+    rho = n0 + nN + nS + nE + nW + nNE + nSE + nNW + nSW   # macroscopic density
+    ux = (nE + nNE + nSE - nW - nNW - nSW) / rho	# macroscopic x velocity
+    uy = (nN + nNE + nNW - nS - nSE - nSW) / rho	# macroscopic y velocity
+
+    barrier = np.zeros((height, width), bool)  # Initialize barriers
+                                               # True wherever there's a barrier
+
+    BarrierHeight = 0.1
+    BarrierWidth = 0.1
+
+    for i in range(barriers):
+
+        left = lefts[i]
+        bottom = bottoms[i]
+
[…]

### 2017-03-09 model/metrics/metrics.py

@@ -1,0 +1,815 @@
+from __future__ import division
+import matplotlib.pyplot as plt
+import sys
+import numpy as np
+from numpy import log10
+from scipy import stats
+from scipy.stats import gaussian_kde
+from scipy.optimize import fsolve
+import math
+
+
+def GetRAD(vector):
+    RAD = []
+    unique = list(set(vector))
+
+    for val in unique:
+        RAD.append(vector.count(val)) # the abundance of each Sp_
+
+    return RAD, unique # the rad and the specieslist
+
+
+def separateCom(lists):
+
+    SpeciesIDs, IndIDs, Qs, GrowthList, MaintList, MFDList, RPFList, N_RList, DispList, DormList, ADList, SizeList = lists
+    a_SpeciesIDs, a_IndIDs, a_Qs, a_GrowthList, a_MaintList, a_MFDList, a_RPFList, a_N_RList, a_DispList, a_DormList, a_SizeList = [list([]) for _ in xrange(11)]
+    d_SpeciesIDs, d_IndIDs, d_Qs, d_GrowthList, d_MaintList, d_MFDList, d_RPFList, d_N_RList, d_DispList, d_DormList, d_SizeList = [list([]) for _ in xrange(11)]
+
+    for i, ind in enumerate(ADList):
+        if ind == 'd':
+
+            d_SpeciesIDs.append(SpeciesIDs[i])
+            d_IndIDs.append(IndIDs[i])
+            d_Qs.append(Qs[i])
+            d_GrowthList.append(GrowthList[i])
+            d_MaintList.append(MaintList[i])
+            d_MFDList.append(MFDList[i])
+            d_RPFList.append(RPFList[i])
+            d_N_RList.append(np.var(N_RList[i]))
+            d_DispList.append(DispList[i])
+            d_DormList.append(DormList[i])
+            d_SizeList.append(SizeList[i])
+
+        elif ind == 'a':
+
+            a_SpeciesIDs.append(SpeciesIDs[i])
+            a_IndIDs.append(IndIDs[i])
+            a_Qs.append(Qs[i])
+            a_GrowthList.append(GrowthList[i])
+            a_MaintList.append(MaintList[i])
+            a_MFDList.append(MFDList[i])
+            a_RPFList.append(RPFList[i])
+            a_N_RList.append(np.var(N_RList[i]))
+            a_DispList.append(DispList[i])
[…]

### 2017-03-09 model/randparams/randparams.py

@@ -1,0 +1,42 @@
+from __future__ import division
+from random import randint
+import numpy as np
+import sys
+
+def get_rand_params(width=0):
+    """ Get random model parameter values. Others are chosen in bide.py """
+
+    seedCom = 1000 # size of starting community
+
+    if width == 0:
+        width = 4
+    elif width < 10:
+        width += 1
+    elif width >= 10:
+        width = 4
+
+    height = float(width)
+    length = float(height)
+
+    low = np.random.uniform(4, 4)
+    rates = np.linspace(0, -low, 40)
+    rates = 10**rates
+    rates = rates.tolist()
+
+    nN = randint(3, 3)
+    amp = np.random.uniform(10**-1, 10**-1)
+    freq = np.random.uniform(10**-1, 10**-1)
+    phase = np.random.uniform(10**-1, 10**-1)
+    m = np.random.uniform(0.1, 0.1)
+
+    r = randint(20, 20)
+    rmax = np.random.uniform(1000, 1000)
+
+    dormlim = np.random.uniform(0.1, 0.1)
+    gmax = np.random.uniform(0.2, 0.2)
+    dmax = np.random.uniform(0.2, 0.2)
+    pmax = np.random.uniform(0.1, 0.1)
+    mmax = np.random.uniform(0.1, 0.1)
+    smax = np.random.uniform(100, 100)
+
+    return [width, height, length, seedCom, m, r, nN, rmax, gmax, mmax, dmax, amp, freq, phase, rates, pmax, dormlim, smax]

### 2017-03-09 model/spatial/spatial.py

@@ -1,0 +1,112 @@
+from __future__ import division
+#import sys
+from random import choice, randint
+import numpy as np
+import math
+import sys
+
+
+def SAR(Xs, Ys, Zs, SpIDs, w):
+    
+    boxes = [list([]) for _ in xrange(w**2)]
+    
+    index = 0
+    for i, val in enumerate(SpIDs):
+        x = int(round(Xs[i]))
+        y = int(round(Ys[i]))
+
+        index = int(round(x + (y * w)))
+
+        if index > len(boxes) - 1:
+            index = len(boxes) - 1
+        elif index < 0:
+            index = 0
+            
+        boxes[index].append(val)
+    
+    
+    q = []
+    sar = [] 
+    while boxes:
+        i = randint(0, len(boxes)-1)
+        box = boxes.pop(i)
+        q.extend(box)
+        sar.append(len(list(set(q))))
+        
+    return sar
+
+    
+    
+
+def distance(p0, p1):
+
+    """ take two (x, y) tuples as parameters
+    http://stackoverflow.com/questions/5407969/distance-formula-between-two-points-in-a-list"""
+
+    return math.sqrt((p0[0] - p1[0])**2 + (p0[1] - p1[1])**2)
+
+
+
+def nearest_neighbor(xlist1, xlist2, ylist1, ylist2, q=1):
+
+    """ xlist1 and ylist1 are assumed to be the lists of individual organisms
+    xlist2 and ylist2 are assumed to be the lists of whatever the individual
+    organisms are being measured with respect to their distance to """
+
+    n = len(xlist1)
+    r = len(xlist2)
+    r = min([10, r])
+
+    refpoints = min([10, n])
+    DistList = []
+
+    for ref in range(refpoints):
+
+        i = randint(0, n-1)
+        x1 = xlist1[i]
+        y1 = ylist1[i]
+        MinDist = 10000
+
+        for j in range(r):
+
+            x2 = xlist2[j]
+            y2 = ylist2[j]
+            dist = distance((x1, y1), (x2, y2))
+
+            if dist < MinDist:
+                MinDist = dist
+
+        DistList.append(MinDist)
+
+    return np.mean(DistList)
+
+
+
+
+
+def avg_dist(xlist1, xlist2, ylist1, ylist2, q=1):
+
+    """ xlist1 and ylist1 are assumed to be the lists of individual organisms
+    xlist2 and ylist2 are assumed to be the lists of whatever the individual
[…]

### 2017-03-09 tools/BIDE/bide.py

@@ -1,0 +1,1226 @@
+# -*- coding: utf-8 -*-
+from __future__ import division
+from random import randint, choice
+import numpy as np
+import sys
+import math
+#from math import modf
+#import decimal
+import time
+
+limit = 0.1
+
+def coord(d):
+    return float(np.random.uniform(0.1*d, 0.9*d))
+
+
+def GetIndParam(means):
+    vals = []
+
+    if isinstance(means, float) or isinstance(means, int):
+        std = means/1000.0
+        vals = np.random.normal(means, std)
+        if vals < 0.00001:
+            vals = 0.00001
+
+    else:
+        for val in means:
+            std = val/1000.0
+            i = np.random.normal(val, std)
+            if i < 0.00001:
+                i = 0.00001
+            vals.append(i)
+
+    return vals
+
+
+
+def GetRAD(vector):
+    RAD = []
+    unique = list(set(vector))
+
+    for val in unique:
+        RAD.append(vector.count(val)) # the abundance of each Sp_
+
+    return RAD, unique # the rad and the specieslist
+
+
+def get_color(ID, colorD): # FUNCTION TO ASSIGN COLORS TO Sp_
+
+    r1 = lambda: randint(0,255)
+    r2 = lambda: randint(0,255)
+    r3 = lambda: randint(0,255)
+
+    color = '#%02X%02X%02X' % (r1(),r2(),r3())
+    colorD[ID] = color
+
+    return colorD
+
+
+
+def NewTracers(numt, motion, IDs, Xs, Ys, t_In, w, h, u0):
+
+    for t in range(numt):
+        x = np.random.binomial(1, u0/2)
+
+        if x == 1:
+            IDs.append(0)
+            t_In.append(0)
+
+            if motion == 'brown_noise' or motion == 'white_noise':
+                Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+                Xs.append(float(np.random.uniform(0.1*w, 0.9*w)))
+
+            else:
+                Ys.append(float(np.random.uniform(0.1*h, 0.9*h)))
+                Xs.append(float(np.random.uniform(0.1*w, 0.15*w)))
+
+    return [IDs, t_In, Xs, Ys]
+
+
+
+def ResIn(motion, Type, Vals, Xs, Ys, ID, IDs, t_In, numr, rmax, nN, nP, nC, w, h, u0):
+
+
+    for r in range(numr):
+        x = np.random.binomial(1, u0/2)
+
+        if x == 1:
[…]

### 2017-03-09 tools/DiversityTools/ObsPred/ObsPred.pyx

@@ -1,0 +1,175 @@
+from __future__ import division
+import sys
+import os
+import random
+import numpy as np
+from scipy import stats
+
+import statsmodels.api as sm
+import statsmodels.formula.api as smf
+import pandas as pd
+
+########### PATHS ##############################################################
+
+tools = os.path.expanduser("~/tools")
+sys.path.append(tools + "/macroeco_distributions")
+import macroeco_distributions as md
+sys.path.append(tools + "/distributions")
+import distributions as dist
+sys.path.append(tools + "/macroecotools")
+import macroecotools
+sys.path.append(tools + "/partitions")
+import partitions as parts
+sys.path.append(tools + "/FeasibleFunctions")
+import feasible_functions as ff
+sys.path.append(tools + "/metrics")
+import metrics
+sys.path.append(tools + "/mete")
+import mete
+sys.path.append(tools + "/pln")
+import pln
+
+data = os.path.expanduser("~/data")
+mydir = os.path.expanduser("~/GitHub/rare-bio")
+#macroeco_dir = os.path.expanduser("~/GitHub/macroecotools")
+
+########### END PATHS ##########################################################
+
+
+########### FUNCTIONS ##########################################################
+
+def import_obs_pred_data(input_filename):
+    # TAKEN FROM THE mete_sads.py script used for White et al. (2012)
+    data = np.genfromtxt(input_filename, dtype = "S15, f8, f8",
+    names = ['site','obs','pred'], delimiter = " ")
+    return data
+
+########### END FUNCTIONS ######################################################
+
+
+########### GET DATA ###########################################################
+
+datasets = []
+for name in os.listdir(data +'/micro'):
+        datasets.append([name, 'micro'])
+
+for name in os.listdir(data +'/macro'):
+        datasets.append([name, 'macro'])
+
+########### END GET DATA #######################################################
+
+
+
+########### DECLARE VARIABLES ##################################################
+
+ct = 0
+numMicros = 0
+numMacros = 0
+numEMP = 0
+numEMPopen = 0
+
+models = ['METE', 'MaxEntGeomD']
+#models = ['IntPartsNS', 'IntPartsN']
+########### END DECLARE VARIABLES ##############################################
+
+
[…]

### 2017-03-09 tools/DiversityTools/ObsPred/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("ObsPred.pyx"))

### 2017-03-09 tools/DiversityTools/PredRadDATA/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("PredRadDATA_ALL.pyx"))

### 2017-03-09 tools/DiversityTools/SSADdata/SSADdata/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("SSADdata.pyx"))

### 2017-03-09 tools/DiversityTools/SSADdata/SsadDATA_ALL.py

@@ -1,0 +1,102 @@
+from __future__ import division
+import sys
+import os
+import random
+import numpy as np
+from scipy import stats
+
+mydir = os.path.expanduser("~/Desktop/Repos/rare-bio/")
+mydir2 = os.path.expanduser("~/Desktop/")
+
+
+sys.path.append(mydir + "tools/feasible_functions")
+import feasible_functions as ff
+
+
+
+datasets = []
+seqID = '99'
+
+"""
+for name in os.listdir('/Users/lisalocey/Desktop/RareBio/data/micro9599/'+seqID+'data'): 
+    index = name.find('-')
+    name = name[0:index]
+    datasets.append([name, 'micro'])        
+"""
+
+for name in os.listdir(mydir2 +'data/micro'):
+        datasets.append([name, 'micro'])   
+
+for name in os.listdir(mydir2 +'data/macro'):
+        datasets.append([name, 'macro'])   
+
+
+
+ct = 0
+numMicros = 0
+numMacros = 0
+numEMP = 0
+numEMPopen = 0
+    
+for dataset in datasets:
+    
+    name = dataset[0] # name of dataset
+    if name == '.DS_Store': continue
+    
+    kind = dataset[1] # micro or macro
+    OUT = open(mydir2 + 'data/'+kind+'/'+name+'/'+name+'-SSADMetricData.txt','w+')
+    RADs = []
+    
+    if kind == 'macro':
+        path = kind
+        RADs = ff.get_SADs(mydir2 +'/data/'+path, name)
+
+        print 'macro', name, len(RADs)
+        
+        
+    if kind == 'micro':
+        #RADs = getPhyloSADs.get_SADs(seqID, dataset[0], 'genus')    
+
+        path = kind
+        if name == 'EMPclosed': continue
+        if name == 'EMPopen':
+            RADs = ff.GetSADsFromBiom_labeled(mydir2 +'data/'+path+'/EMPopen/', name)
+        
+        else: RADs = ff.get_SADs(mydir2 +'data/'+path, name)
+                                                                                                        
+    ct = 0
+    numRADs = len(RADs)
+    for RAD in RADs:
+        
+        RAD = list([x for x in RAD if x != 0])
+        
+        N = sum(RAD)
+        S = len(RAD)
+        
+        if S < 2: continue 
+        
+        Evar = ff.e_var(RAD)
[…]

### 2017-03-09 tools/DiversityTools/macroecotools/macroecotools.py

@@ -1,0 +1,430 @@
+"""Module for conducting standard macroecological plots and analyses"""
+
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+import colorsys
+from numpy import log10
+import pandas
+import numpy
+import math
+
+def AIC(k, L):
+    """Computes the Akaike Information Criterion.
+
+    Keyword arguments:
+    L  --  log likelihood value of given distribution.
+    k  --  number of fitted parameters.
+
+    """
+    AIC = 2 * k - 2 * L
+    return AIC
+
+
+def AICc(k, L, n):
+    """Computes the corrected Akaike Information Criterion. 
+    
+    Keyword arguments:
+    L  --  log likelihood value of given distribution.
+    k  --  number of fitted parameters.
+    n  --  number of observations.
+       
+    """
+    AICc = 2 * k - 2 * L + 2 * k * (k + 1) / (n - k - 1)
+    return AICc
+
+def aic_weight(AICc_list, n, cutoff = 4):
+    """Computes Akaike weight for one model relative to others
+    
+    Based on information from Burnham and Anderson (2002).
+    
+    Keyword arguments:
+    n           --  number of observations.
+    cutoff      --  minimum number of observations required to generate a weight.
+    AICc_list   --  list of AICc values for each model
+    
+    """
+    if n < cutoff:
+        AICc_weights = None
+        
+    else:
+        AICc_min = min(AICc_list) # Minimum AICc value for the entire list
+        relative_likelihoods = []
+        
+        for AICc in AICc_list:
+            delta_AICc = AICc - AICc_min
+            relative_likelihood = np.exp(-(delta_AICc)/2)
+            relative_likelihoods.append(relative_likelihood)
+            
+        relative_likelihoods = np.array(relative_likelihoods)
+        
+        AICc_weights = relative_likelihoods / sum(relative_likelihoods)
+            
+        return(AICc_weights)   
+
+def get_pred_iterative(cdf_obs, dist, *pars):
+    """Function to get predicted abundances (reverse-sorted) for distributions with no analytical ppf."""
+    cdf_obs = np.sort(cdf_obs)
+    abundance  = list(np.empty([len(cdf_obs)]))
+    j = 0
+    cdf_cum = 0
+    i = 1
+    while j < len(cdf_obs):
[…]

### 2017-03-09 tools/DiversityTools/macroecotools/test_macroecotools.py

@@ -1,0 +1,97 @@
+"""Test suite for macroecotools"""
+
+from pandas import Series, DataFrame
+from numpy import array, array_equal
+from macroecotools import *
+
+comp_data = DataFrame({'site': [1, 1, 2, 3, 3, 3, 4],
+                       'year': [1, 2, 1, 1, 1, 2, 2],
+                       'genus': ['a', 'a', 'a', 'a', 'a', 'd', 'f'],
+                       'species': ['b', 'b', 'b', 'b', 'c', 'e', 'g'],
+                       'spid': ['ab', 'ab', 'ab', 'ab', 'ac', 'de', 'fg'],
+                       'counts': [1, 2, 5, 5, 4, 3, 10]})
+
+def test_combined_spID_multiple_str():
+    """Test combining multiple strings in a single species identifiers"""
+    assert combined_spID("fam", "gen", "sp") == "famgensp"
+
+def test_combined_spID_str():
+    """Test that a single string remains unchanged as a species identifier"""
+    assert combined_spID("Dipodomys spectabilis") == "Dipodomys spectabilis"
+
+def test_combined_spID_single_series():
+    """Test that a single column data frame/series returns the same"""
+    ids = Series(['gen1 sp1', 'gen1 sp2', 'gen2 sp3'])
+    assert ids.equals(combined_spID(ids))
+
+def test_combined_spID_multiple_series():
+    """Test that multiple series return the right combined series"""
+    ids1 = Series(['gen1', 'gen1', 'gen2', 'gen2'])
+    ids2 = Series(['sp1', 'sp1', 'sp2', 'sp3'])
+    combined_ids = Series(['gen1sp1', 'gen1sp1', 'gen2sp2', 'gen2sp3'])
+    assert combined_ids.equals(combined_spID(ids1, ids2))
+
+def test_combined_spID_lists():
+    """Test that multiple lists return the right combined list"""
+    ids1 = ['gen1', 'gen1', 'gen2', 'gen2']
+    ids2 = ['sp1', 'sp1', 'sp2', 'sp3']
+    ids3 = ['subsp1', 'subsp2', 'subsp3', 'subsp4']
+    combined_ids = ['gen1sp1subsp1', 'gen1sp1subsp2', 'gen2sp2subsp3', 'gen2sp3subsp4']
+    assert combined_ids == combined_spID(ids1, ids2, ids3)
+
+def test_combined_spID_arrays():
[…]

### 2017-03-09 tools/DiversityTools/mete/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("mete.pyx"))

### 2017-03-09 tools/DiversityTools/metrics/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+from Cython.Build import cythonize
+
+setup(ext_modules = cythonize("metrics.pyx"))

### 2017-03-09 tools/DiversityTools/predRADs/predRADs.pyx

@@ -1,0 +1,158 @@
+from __future__ import division
+import sys
+
+mydir = os.path.expanduser("~/GitHub/rare-bio")
+mydir2 = os.path.expanduser("~/Desktop")
+
+sys.path.append(mydir + "/tools/partitions")
+import partitions as parts
+sys.path.append(mydir + "/tools/FeasibleFunctions")
+import feasible_functions as ff
+
+sys.path.append(mydir + "/tools/metrics")
+import metrics
+sys.path.append(mydir + "/tools/macroecotools")
+import macroecotools
+sys.path.append(mydir + "/tools/macroeco_distributions")
+import macroeco_distributions as md
+sys.path.append(mydir + "/tools/predRADs")
+import predRADs
+sys.path.append(mydir + "/tools/mete")
+import mete
+sys.path.append(mydir + "/tools/pln")
+import pln
+
+sys.path.append("/Users/lisalocey/Desktop/global/GenAnalysis/tools/")
+import macroeco_distributions as md
+import macroecotools as mt
+import feasible_functions as ff
+import mete
+import pln
+
+import numpy as np
+import random
+import math
+
+
+def truncate(RAD, maxn):
+
+    while max(RAD) > maxn:
+        m = max(RAD)
+        i = RAD.index(m)
+        n1 = RAD[i]
+        n2 = random.randint(1,n1-1)
+        RAD[i] -= n2
+
+        i = random.randint(0, len(RAD)-1)
+        RAD[i] += n2
+
+    return RAD
+
+
+def get_GeomSeries(N,S,zeros=False):
+
+    rank = range(1,S+1)
+    cdf = [(S-i+0.5)/S for i in rank]
+    SNratio = S/N
+    if zeros == False:
+        abd = md.trunc_geom.ppf(np.array(cdf), SNratio, N)
+
+    return abd
+
+
+def RandComp(Q, N, maxn=False):
+
+    composition = []
+
+    indices = []
+    while len(indices) < N-1:
+        index = random.randint(1, Q)
+        if index in indices: continue
+        else: indices.append(index)
+
+    indices.sort()
+    Qlist = [1]*Q
+
+    partsOF = [Qlist[i:j] for i, j in zip([0]+indices, indices+[None])]
+    for p in partsOF:
+        composition.append(sum(p))
+
+    RAD = truncate(composition, maxn)
+    RAD.sort()
+    RAD.reverse()
+
+    return composition
+
+
+
[…]

### 2017-03-09 tools/DiversityTools/radMetrics/radMetrics.py

@@ -1,0 +1,118 @@
+from __future__ import division
+import sys
+import os
+import random
+import numpy as np
+from scipy import stats
+
+
+mydir = os.path.expanduser("~/")
+
+sys.path.append(mydir + "tools/metrics")
+import metrics
+sys.path.append(mydir + "tools/getRADs")
+import getRADs
+
+def getMetrics():
+    datasets = []
+    for name in os.listdir(mydir +'data/micro'):
+            datasets.append([name, 'micro'])
+    for name in os.listdir(mydir +'data/macro'):
+            datasets.append([name, 'macro'])
+
+
+    for dataset in datasets:
+
+        name = dataset[0] # name of dataset
+        kind = dataset[1] # micro or macro
+
+        #print name, kind
+        #xnames = ['AGSOIL', '.DS_Store', 'SLUDGE']
+        #xnames = ['HMP']
+
+        if name == '.DS_Store': continue
+        if name in ['MGRAST'] : pass
+        else: continue
+
+        #OUT = open(mydir + 'data/'+kind+'/'+name+'/'+name+'-SADMetricData_NoMicrobe1s.txt','w+')
+        OUT = open(mydir + 'data/'+kind+'/'+name+'/'+name+'-SADMetricData.txt','w+')
+        RADs = []
+
+        if kind == 'macro':
+            RADs = getRADs.get_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+            print 'macro', name, len(RADs)
+
+
+        if kind == 'micro':
+            #RADs = getPhyloSADs.get_SADs(seqID, dataset[0], 'genus')
+
+            if name == 'EMPclosed' or name == 'EMPopen':
+                RADs = getRADs.EMP_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+
+            else:
+                RADs = getRADs.get_SADs(mydir +'data/'+kind+'/'+name+'/', name)
+
+            print 'micro', name, len(RADs)
+
+        ct = 0
+        numRADs = len(RADs)
+        for RAD in RADs:
+
+            if kind == 'micro':
+                RAD = list([x for x in RAD if x > 0]) # greater than 1 means singletons are excluded
+
+            elif kind == 'macro':
+                RAD = list([x for x in RAD if x > 0])
+
+
+            N = sum(RAD)
+            S = len(RAD)
+
+            if S < 10: continue
[…]

### 2017-03-09 tools/Rbin/metrics.R

@@ -1,0 +1,51 @@
+#### A function to generate observed richness
+S.obs <- function(x = ""){ rowSums(x > 0) * 1}
+
+
+sp.turnover <- function(site1, site2){
+  
+  site1 <- site1[site1 != '']
+  site1 <- site1[!is.na(site1)]
+  site2 <- site2[site2 != '']
+  site2 <- site2[!is.na(site2)]
+  #if(length(site1) | length(site2) == 0){
+  #  return -1
+  #  }
+  gamma = union(site1, site2)         # Gamma species pool
+  s     = length(gamma)                                   # Gamma richness
+  a.bar = mean(c(length(site1), length(site2)))   # Mean sample richness
+  b.w   = round(s/a.bar - 1, 4)
+  return(b.w)
+  }
+
+
+
+get.vectors <- function(inputFile, type){
+  vectors <- c()
+  con  <- file(inputFile, open = "r")
+  
+  while (length(oneLine <- readLines(con, n = 1)) > 0) {
+    myLine <- unlist((strsplit(oneLine, ",")))
+    myLine[myLine == '[]'] <- -1
+    myLine[is.na(myLine)] <- -1
+    myLine <- myLine[which(myLine!=-1)]
+    myLine <- myLine[which(myLine!="")]
+    
+    if (length(myLine > 0)){
+      #if(type == 'char'){
+      #  myLine <- as.vector(myLine)
+      #}
+      if(type == 'number'){
+        myLine <- as.numeric(myLine)
+      }
+      
+      i <- length(vectors)+1
+      vectors[[i]] <- myLine
+    }
+  }
+  close(con)
+  return(vectors)
+}
+
+
+sem <- function(x) return(sqrt(var(x)/length(x)))

### 2017-03-09 tools/Rbin/multiplot.R

@@ -1,0 +1,133 @@
+################################################################################
+#                                                                              #
+# Functions for calculating metrics of diversity, evenness, rarity, etc.       #
+# These are not included in other diversity packages, e.g., Vegan              #
+#                                                                              #
+################################################################################
+#                                                                              #
+# Written by: Ken Locey                                                        #
+# Obtained from:                                                               #
+# http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/      #
+#                                                                              #
+################################################################################
+#                                                                              #
+# Recent Changes:                                                              #
+#                                                                              #
+# Future Changes (To-Do List):                                                 #
+#         1.                                                                   #
+#         2. Add warnings                                                      #
+#                                                                              #
+################################################################################
+
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  
+  numPlots = length(plots)
+  
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                     ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+  
+  if (numPlots==1) {
[…]

### 2017-03-09 tools/SADfits/SAD-Models.py

@@ -1,0 +1,429 @@
+from __future__ import division
+import os, collections, math
+from scipy import stats, optimize
+import statsmodels.formula.api as smf
+import statsmodels.api as sm
+from scipy.stats import nbinom
+import numpy as np
+import pandas as pd
+from macroeco_distributions import pln, pln_solver, negbin_solver, trunc_geom
+from scipy.optimize import fsolve
+import scipy.optimize as opt
+from numpy import log, log2, exp, sqrt, log10
+from math import erf, pi
+from sklearn.grid_search import GridSearchCV
+from sklearn.neighbors import KernelDensity
+
+
+"""This code was written using MIT liscenced code from the following Weecology
+repos: METE (https://github.com/weecology/METE) and macroecotools
+(https://github.com/weecology/macroecotools). """
+
+
+
+class predictS:
+    def __init__(self, N, Nmax, predictNmax = True):
+        self.N = N
+        self.Nmax = Nmax
+        self.predictNmax = predictNmax
+
+    def alpha(self, a, Nmax, Nmin=1):
+
+        """Numerically solve for Preston's a. Needed to estimate S using the lognormal"""
+
+        y = sqrt(pi*Nmin*Nmax)/(2.0*a) * exp((a * log2(sqrt(Nmax/Nmin)))**2.0)
+        y = y * exp((log(2.0)/(2.0*a))**2.0)
+        y = y * erf(a * log2(sqrt(Nmax/Nmin)) - log(2.0)/(2.0*a))
+        y += erf(a * log2(sqrt(Nmax/Nmin)) + log(2.0)/(2.0*a))
+        y -= self.N
+
+        return y # find alpha
+
+    def s(self, a, Nmax, Nmin=1):
+
+        """Predict S from the lognormal using Nmax as the only empirical input"""
+
+        return sqrt(pi)/a * exp( (a * log2(sqrt(Nmax/Nmin)))**2) # Using equation 10
+
+
+    def getNmax(self, b=0.6148, slope=0.942904468437):
+
+        """Predict Nmax using N and the scaling law of Nmax with N predicted by the lognormal"""
+
+        #NmaxCalc = 10 ** (b + slope*(log10(self.N)))
+        NmaxCalc = (self.N **  slope) * b
+        return int(round(NmaxCalc))
+
+
+    def getS(self, predictNmax=True):
+
+        guess = 0.1 # initial guess for Pre ston's alpha
+        Nmin = 1
+
[…]

### 2017-03-09 tools/SimpleIBMs/WhiteBeltModel_Heat.py

@@ -1,0 +1,110 @@
+from __future__ import division
+from random import choice, randrange
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy as sc
+from scipy import stats
+
+
+######################### COMMUNITY SIMULATION FUNCTION ########################
+def simCOM(gens, dorm, disp, env, lgp = 0.92):
+    COM = list(np.random.logseries(lgp, 2000)) # populate the community
+
+    for i, microbe in enumerate(COM):
+        x, y = np.random.uniform(0, 1, 2)
+        COM[i] = [microbe, choice([0, 1]), x, y] # 0 is active; 1 is dormant
+
+    for g in range(gens):
+
+        i = randrange(len(COM))          # Choose an individual at random to die
+	x = np.random.binomial(1, dorm)
+	if x == 0 or COM[i][1] == 0: COM.pop(i)
+
+        i = randrange(len(COM))             # Choose an individual to transition
+        diff = np.abs(min([COM[i][2], COM[i][3]]) - env)
+        COM[i][1] = np.random.binomial(1, diff)
+
+        i = randrange(len(COM))               # Choose an individual to disperse
+        x = np.random.normal(loc=COM[i][2], scale=disp)
+	y = np.random.normal(loc=COM[i][3], scale=disp)
+
+	if x > 1: x = 1 # keeping the individual within bounds
+	elif x < 0: x = 0
+	if y > 1: y = 1
+	elif y < 1: y = 1
+	COM[i][2], COM[i][3] = [x, y]
+
+	for i, val in enumerate(COM):
+            if val[1] == 0: # Choose an individual at random to reproduce
+                COM.append(val)
+                break
+
+        print g, len(COM)
+    return COM
+
+################ GET SLOPES AND INTERCEPTS ####### Regression Approach #########
+def getVals(Zvals, Yints, COM):
+    Slist, Alist, grain = [ [], [], 2]
+
+    for i in range(1, 10):
+        slist, grain = [[], grain/2] # grain on the log2 scale
+
+        for ind in COM:
+            if min([ind[2], ind[3]]) <= grain: slist.append(ind[0])
+
+        if len(list(set(slist))) > 0:
+            Slist.append(len(list(set(slist))))
+            Alist.append(grain)
+
[…]

### 2017-03-09 tools/metrics/metrics.py

@@ -1,0 +1,767 @@
+from __future__ import division
+import  matplotlib.pyplot as plt
+import sys
+import numpy as np
+from numpy import log10
+from scipy import stats
+from scipy.stats import gaussian_kde
+from scipy.optimize import fsolve
+import math
+
+
+
+def count_pts_within_radius(x, y, radius, logscale=0):
+    """Count the number of points within a fixed radius in 2D space"""
+    #TODO: see if we can improve performance using KDTree.query_ball_point
+    #http://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.KDTree.query_ball_point.html
+    #instead of doing the subset based on the circle
+    raw_data = np.array([x, y])
+    x = np.array(x)
+    y = np.array(y)
+    raw_data = raw_data.transpose()
+
+    # Get unique data points by adding each pair of points to a set
+    unique_points = set()
+    for xval, yval in raw_data:
+        unique_points.add((xval, yval))
+
+    count_data = []
+    for a, b in unique_points:
+        if logscale == 1:
+            num_neighbors = len(x[((log10(x) - log10(a)) ** 2 +
+                                   (log10(y) - log10(b)) ** 2) <= log10(radius) ** 2])
+        else:
+            num_neighbors = len(x[((x - a) ** 2 + (y - b) ** 2) <= radius ** 2])
+        count_data.append((a, b, num_neighbors))
+    return count_data
+
+
+
+def plot_color_by_pt_dens(x, y, radius, loglog=0, plot_obj=None):
+    """Plot bivariate relationships with large n using color for point density
+
+    Inputs:
+    x & y -- variables to be plotted
+    radius -- the linear distance within which to count points as neighbors
+    loglog -- a flag to indicate the use of a loglog plot (loglog = 1)
+
+    The color of each point in the plot is determined by the logarithm (base 10)
+    of the number of points that occur with a given radius of the focal point,
[…]

### 2017-03-09 tools/randparams/randparams.py

@@ -1,0 +1,129 @@
+from __future__ import division
+from random import choice, randint
+import numpy as np
+import sys
+import os
+
+def get_rand_params(fixed):
+    """ Get random model parameter values. Others are chosen in bide.py """
+
+    envgrads = []
+    seedCom = 100 # size of starting community
+    rates = []
+
+    if fixed is True:
+
+        rates = np.array([1.0, 0.5, 0.1, 0.05, 0.01, 0.005, 0.001])
+
+        #motion = 'white_noise'
+        #motion = 'brown_noise'
+        motion = 'fluid'
+
+        #width = choice([5, 20, 5, 20])
+        width = 10
+        #height = int(width)
+        height = choice([5])
+
+        num_envgrads = 2
+        for i in range(num_envgrads):
+            x = np.random.uniform(0, width)
+            y = np.random.uniform(0, height)
+            envgrads.append([x, y])
+
+        nNi = 3 # max number of Nitrogen types
+        nP = 3 # max number of Phosphorus types
+        nC = 3 # max number of Carbon types
+
+        amp = 0.001
+        freq = 0.001
+        phase = 0.0
+        pulse = 0.001
+        flux = 'yes'
+
+        disturb = 0.00001
+        m = 0.01
+        speciation = 0.01
+        maintmax = 0.001
+
+        reproduction = 'fission'
+        alpha = 0.99
+        barriers = 0
+
+        #r = choice([1, 10, 100])#resource particles flowing in per time step
+        r = 100
+        gmax = 0.9
+        rmax = 1000
+        dmax = 0.1
+        #p_of_going_active = np.random.uniform(0.0001, 0.01)
+        #maint_factor = np.random.uniform(1, 100)
+
+
+    elif fixed is False:
+
+        #motion = choice(['fluid', 'white_noise']) # 'fluid', 'unidirectional'
+        motion = 'fluid'
+        #motion = 'white_noise'
+
+        if motion == 'white_noise':
+            rates = np.array([0.00001])
+        else:
+            rates = [choice([1.0, 0.8, 0.6, 0.4, 0.2, 0.08, 0.06, 0.04, 0.02, 0.008, 0.006, 0.002])]
+
+            #rates = np.array([1.0, 0.9, 0.8, 0.7, 0.6, 0.4, 0.2, 0.08, 0.06, 0.04, 0.02, 0.008, 0.006, 0.002])
+            #rates = np.array([1.0, 0.7, 0.4, 0.1, 0.07, 0.04, 0.01, 0.007])
[…]

### 2017-03-09 tools/spatial/spatial.py

@@ -1,0 +1,102 @@
+from __future__ import division
+#import sys
+from random import choice, randint
+import numpy as np
+#from scipy import stats
+#from scipy.stats import gaussian_kde
+#from scipy.optimize import fsolve
+import math
+
+
+def distance(p0, p1):
+
+    """ take two (x, y) tuples as parameters
+    http://stackoverflow.com/questions/5407969/distance-formula-between-two-points-in-a-list"""
+
+    return math.sqrt((p0[0] - p1[0])**2 + (p0[1] - p1[1])**2)
+
+
+
+def nearest_neighbor(xlist1, xlist2, ylist1, ylist2, q=1):
+
+    """ xlist1 and ylist1 are assumed to be the lists of individual organisms
+    xlist2 and ylist2 are assumed to be the lists of whatever the individual
+    organisms are being measured with respect to their distance to """
+
+    n = len(xlist1)
+    r = len(xlist2)
+    r = min([10, r])
+
+    refpoints = min([10, n])
+    DistList = []
+
+    for ref in range(refpoints):
+
+        i = randint(0, n-1)
+        x1 = xlist1[i]
+        y1 = ylist1[i]
+        MinDist = 10000
+
+        for j in range(r):
+
+            x2 = xlist2[j]
+            y2 = ylist2[j]
+            dist = distance((x1, y1), (x2, y2))
+
+            if dist < MinDist:
+                MinDist = dist
+
+        DistList.append(MinDist)
+
+    return np.mean(DistList)
+
+
+
+
+
+def avg_dist(xlist1, xlist2, ylist1, ylist2, q=1):
+
+    """ xlist1 and ylist1 are assumed to be the lists of individual organisms
+    xlist2 and ylist2 are assumed to be the lists of whatever the individual
+    organisms are being measured with respect to their distance to """
+
+    nmax = len(xlist1)
+    rmax = len(xlist2)
+
+    refpoints = min([100, nmax])
+    dist = []
+
+    for n in range(refpoints):
+        for j in range(q):
+
+            i1 = choice(range(nmax))
+            x1 = xlist1[i1]
+            y1 = ylist1[i1]
+
+            i2 = choice(range(rmax))
+            x2 = xlist2[i2]
+            y2 = ylist2[i2]
+
+            dist.append(distance((x1, y1), (x2, y2)))
+
+    return np.mean(dist)
+
+
+"""
[…]

### 2017-03-09 tools/unit_tests/test_metrics.py

@@ -1,0 +1,123 @@
+from __future__ import division
+import os
+import sys
+
+mydir = os.path.expanduser("~/")
+sys.path.append(mydir + "GitHub/simplex/tools/metrics")
+import metrics
+
+
+def test_percent_ones():
+
+    """ tests for the percent of species in a list called sad,
+    represented by a single individual
+
+    where the function returns 100 * sad.count(1)/len(sad) """
+
+    assert metrics.percent_ones([1,10,1,11,1.1,0.1,1,1,111,1.1]) == 100*4.0/10.0
+    assert metrics.percent_ones([1]) == 100
+    assert metrics.percent_ones([2, 3, 4, 11, 111, 0.1, 1.1]) == 0.0
+
+
+def test_percent_pt_one():
+
+    """ test for percent taxa with less than 0.1 percent relative abundance
+    in a list called sad """
+
+    assert metrics.percent_pt_one([1]) == 0
+    assert metrics.percent_pt_one([1000, 1]) == 50
+    assert metrics.percent_pt_one([999, 1]) == 0
+
+
+def test_Rlogskew():
+
+    """ test for log-skew of a frequency distribution """
+
+    assert metrics.Rlogskew([1,1]) == 'NaN'
+    assert metrics.Rlogskew([1.0,123]) == 'NaN'
+    assert metrics.Rlogskew([5,5,5,5,5]) == 'NaN'
+    assert metrics.Rlogskew([5,5,5,5,5,5,5,5,1]) == -3.0
+
+
+
+def test_Preston():
+
+    """ test for the value of alpha for Preston's lognormal distribution """
+    assert metrics.Preston([0]) == 'NaN'
+    assert metrics.Preston([1]) == (0.72823274985908792, 3.0525896286614604)
+    assert metrics.Preston([1,2,3,4,5,6,7,8,9,10]) == (0.16113216807474443, 1123.4019840427286)
+
+
+def test_Berger_Parker():
+
+    """ test for the value of the Berger-Parker index of dominance, i.e.,
+    relative abundance of the most abundant species """
+
+    assert metrics.Berger_Parker([100]) == 1.0
+    assert metrics.Berger_Parker([1,1,1,1,1,1,1,1,1,1]) == 0.1
+    assert metrics.Berger_Parker([0]) == 'NaN'
[…]
